### 2018-10-03 DryDep.tex
Intitial commit of paper documents.

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[gmd, manuscript]{copernicus}

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% Fossil Record (fr)
% Geographica Helvetica (gh)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\begin{document}

\title{Update of the ozone dry deposition in the OsloCTM3 v1.0}

% \Author[affil]{given_name}{surname}

\Author[1]{Stefanie}{Falk}
\Author[2,a]{Amund}{S{\o}vde Haslerud}
%\Author[]{}{}

\affil[1]{Department of Geosciences, University of Oslo, Oslo, Norway}
\affil[2]{CICERO Center for International Climate Research, Oslo, Norway}
\affil[a]{Kjeller Vindteknikk, Kjeller, Norway}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

\runningtitle{Update of ozone dry deposition in OsloCTM3}

\runningauthor{Falk}

\correspondence{Stefanie Falk (stefanie.falk@geo.uio.no)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
  Since the industrial revolution, tropospheric background ozone concentrations have been increasing strongly mainly in the northern hemisphere. In recent years, the number of episodes of peak concentrations especially in summertime has been decreasing in North America and Europe due to the implementation of air quality regulations. At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution. High concentrations of ozone in ambient air are hazardous not only to humans but to the ecosystem in general. The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future. Most chemistry-transport models (CTMs) and chemistry-climate models (CCMs) tend to overestimate surface ozone values. This may be partly due to uncertainties in the distribution and amount of precursor species and stratosphere-troposphere exchange as well as scavenging processes removing ozone from the atmosphere. Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition. The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone. Dry deposition velocities are typically calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these conceptual dry deposition velocities are hard to measure explicitly. Following the European Monitoring and Evaluation Programme (EMEP) MSC-W chemistry transport model, we compute canopy resistance from stomatal conductance. In this paper, we present an update on the dry deposition scheme implemented in the OsloCTM3v1.0 and evaluate the model with respect to ozone and different parameters in the stomatal conductance parameterization, e.g., photosynthetic photon flux density (PPFD) and phenology. Furthermore, we investigate the impact on modeled ozone concentrations, both at the surface and in the free troposphere. We show that ozone dry deposition in the new version (v.1.1) decreases with respect to the previous version (v1.0), leading to an increase in surface ozone of up to 40\,\unit{\%}. We also put our results in context to recent multi-model studies as well as satellite data.
\end{abstract}

%\copyrightstatement{TEXT}
\introduction  %% \introduction[modified heading if necessary]
Ozone is an important trace gas for all lifeforms on Earth. Depending on the place of its occurrence it has either a positive or negative connotation. In the stratosphere, ozone absorbs most of the ultraviolet (UV)-light from the sun in the range of 100--315\,\unit{nm}, thus shielding the Earth's surface from the most harmful UV-radiation. In addition, ozone is a potent greenhouse gas in both, stratosphere and troposphere. With a radiative forcing of $0.40 \pm 0.20\,\unit{Wm^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}. In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant. Continuously high concentrations of ambient air ozone are hazardous to the whole ecosystem. It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about 3--15\,\unit{\%} \citep{PJ:Ainsworth2017} as well as 7\,\unit{\%} loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}. The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}. Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days \citep{JGR:Stevenson2005,ACP:Young2013}. It is mainly produced photochemically in situ from precursor gases such as hydrocarbons (e.g., \chem{CH_4}) in the presents of nitrogen oxides (\chem{NO_x}), carbon monoxide (\chem{CO}):
\begin{reaction}
  ^\bullet\chem{OH} + \chem{CO} \rightarrow {^\bullet\chem{HOCO}}\\
  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{H_2O} + \chem{CO_2}
 \end{reaction}
\begin{reaction}
  ^\bullet\chem{HO_2} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
  \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
  ^\bullet\chem{O(^3P)} + \chem{O_2} \rightarrow \chem{O_3}
\end{reaction}
This leads to a net production via:
\begin{reaction}
  \chem{CO} + 2\chem{O_2} + h\nu \rightarrow \chem{CO_2} + \chem{O_3}
\end{reaction}
A part from these, additional reaction pathways involving volatile organic compounds (VOCs) exist. Another possible source of tropospheric ozone transportation down from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}. Major sinks are photolytical destruction, de-nitration via
\begin{reaction}
  \chem{NO} + \chem{O_3} \rightarrow \chem{NO_2} + \chem{O_2}
\end{reaction}
and
\begin{reaction}
  \chem{NO_2} + \chem{O_3} \rightarrow \chem{NO_3} + \chem{O_2},
\end{reaction}
and dry deposition processes. The global abundance of tropospheric ozone is highly variable and it changes, e.g., with time of the day ($\sim$15:00 local time), season (mid-June), altitude, and location \citep{ACP:Schnell2015}. Typical concentrations of surface ozone reach from 10\,\unit{ppb} over the tropical Pacific to 100\,\unit{ppb} in the surrounding areas of high emitting sources \citep[Chapter 8]{IPCC2013}. This variability poses a challenge on both trend analysis from observation as well as validation and intercomparison of models. From the observational side, there is only a limited number of long-term ozone observations, mainly restricted to European sites. From these only three display a statistical significant trend, indicating a doubling of tropospheric ozone since the onset of the industrialization \citep[Chapter 2]{IPCC2013}. But especially the the very low pre-industrial ozone values cannot be reproduced by the likes of most models \citep[Chapter 8]{IPCC2013}.

, relation to weather \citep{ACP:Otero2018}
Dry deposition in detail, conductance sum \citep{AE:Wesely1989}, stomatal conductance \citep{PP:Buckley2017}, snow/ice \citep{ACP:Helmig2007,AE:Oltmans2008} and water \citep{JGR:Helmig2012,ACP:Luhar2017}.

What is the status and what are the challenges in the dry deposition of ozone - we do not include bromine explosion in the polar regions which are tremendously reducing surface ozone in polar spring \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
simple plant function type (PFT)-based scheme. In this scheme, an average canopy resistance is computed from stomatal conductance of fractional PFTs in a grid cell.

Convention on Long-Range Transboundary Air Pollution (CLRTAP) -> European Monitoring and Evaluation Programme (EMEP),

Ozone accumulated over long periods of time inflicts damage on vegetation which leads to loss of primary production in agriculture and forestry.

Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)

\section{Model description}
General model description followed by detailed dry deposition scheme \citep{ACP:Simpson2012,ICP:MappingManual2017}. This is followed by the necessary pre-processing of files. Old scheme \citep{AE:Wesely1989, JGR:Hough1991}
\citet{GMD:Sovde2012,GMD:Lund2018}
\subsection{Ozone dry deposition scheme}
\citet{ACP:Simpson2012} \citet{ICP:MappingManual2017}
\subsection{Greening season and photoactive radiation}

\section{Evaluation}

\subsection{Sensitivity tests}
Include table of all sensitivity runs.

\subsection{Comparison with recent multi-model evaluation}
Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}.

\subsection{Comparison with observations}
The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (http://bodekerscientific.com/).

\section{Discussion}

%\subsection{HEADING}
%\subsubsection{HEADING}

\conclusions  %% \conclusions[modified heading if necessary]
What did we gain?

%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.

%\codeavailability{TEXT} %% use this section when having only software code available

%\dataavailability{TEXT} %% use this section when having only data sets available

\codedataavailability{TEXT} %% use this section when having data sets and software code available

%\sampleavailability{TEXT} %% use this section when having geoscientific samples available

%\appendix
%\section{}    %% Appendix A

%\subsection{}     %% Appendix A1, A2, etc.

%\noappendix       %% use this to mark the end of the appendix section

%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:

%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
%% They will be correctly named automatically.

%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:

%\appendixfigures  %% needs to be added in front of appendix figures

%\appendixtables   %% needs to be added in front of appendix tables

%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.

\authorcontribution{Stefanie Falk wrote the paper, did the data analysis, and finalized the implementation of the stomatal conductance in the dry deposition scheme. Amund S{\o}vde Haslerud did most of the implementation and documentation of the updated dry deposition scheme.} %% it is strongly recommended to make use of this section

\competinginterests{The authors declare that they have no conflict of interest.} %% this section is mandatory even if you declare that no competing interests are present

%\disclaimer{TEXT} %% optional section

\begin{acknowledgements}
  This work was supported by the Norwegian Research Council (NRC) through the project The double punch: Ozone and climate stresses on vegetation (OzoNorClim).\\
  We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
  The used Leaf Area Index (LAI) and roughness length ($Z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A. (\doi{10.3334/ORNLDAAC/970}).
\end{acknowledgements}

%% REFERENCES

%% The reference list is compiled as follows:

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{DryDep.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES

%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
%% between bibliography and first table and/or figure as well as between each table and/or figure.

%% ONE-COLUMN FIGURES

%%f
%\begin{figure}[t]
%\includegraphics[width=8.3cm]{FILE NAME}
%\caption{TEXT}
%\end{figure}
%
%%% TWO-COLUMN FIGURES
%
%%f
%\begin{figure*}[t]
%\includegraphics[width=12cm]{FILE NAME}
%\caption{TEXT}
%\end{figure*}
%
%
%%% TABLES
%%%
%%% The different columns must be seperated with a & command and should
%%% end with \\ to identify the column brake.
%
%%% ONE-COLUMN TABLE
%
%%t
%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}
%
%%% TWO-COLUMN TABLE
%
%%t
%\begin{table*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table*}
%
%%% LANDSCAPE TABLE
%
%%t
%\begin{sidewaystable*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{sidewaystable*}
%
%
%%% MATHEMATICAL EXPRESSIONS
%
%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
%%%
%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
%%% Vectors are identified in bold italic font using \vec{x}
%%% Matrices are identified in bold roman font
%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
%%% The character * should not be applied as mutliplication sign
%
%
%%% EQUATIONS
%
%%% Single-row equation
%
%\begin{equation}
%
%\end{equation}
%
%%% Multiline equation
%
%\begin{align}
%& 3 + 5 = 8\\
%& 3 + 5 = 8\\
%& 3 + 5 = 8
%\end{align}
%
%
%%% MATRICES
%
%\begin{matrix}
%x & y & z\\
%x & y & z\\
%x & y & z\\
%\end{matrix}
%
%
%%% ALGORITHM
%
%\begin{algorithm}
%\caption{...}
%\label{a1}
%\begin{algorithmic}
%...
%\end{algorithmic}
%\end{algorithm}
%
%
%%% CHEMICAL FORMULAS AND REACTIONS
%
%%% For formulas embedded in the text, please use \chem{}
%
%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
%
%\begin{reaction}
%%% \rightarrow should be used for normal (one-way) chemical reactions
%%% \rightleftharpoons should be used for equilibria
%%% \leftrightarrow should be used for resonance structures
%\end{reaction}
%
%
%%% PHYSICAL UNITS
%%%
%%% Please use \unit{} and apply the exponential notation

\end{document}

### 2018-10-04 DryDep.tex
Almost finished introduction. Pulled in some more paper sources.

@@
 Depending on the place of its occurrence it has either a positive or negative connotation.
-In the stratosphere, ozone absorbs most of the ultraviolet (UV)-light from the sun in the range of 100--315\,\unit{nm}, thus shielding the Earth's surface from the most harmful UV-radiation.
+In the stratosphere, ozone absorbs most of the ultraviolet (UV)-light from the sun within the range of 100--315\,\unit{nm}, thus shielding the Earth's surface from the most harmful UV-radiation.
 In addition, ozone is a potent greenhouse gas in both, stratosphere and troposphere.
-With a radiative forcing of $0.40 \pm 0.20\,\unit{Wm^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
+With a radiative forcing of $0.40 \pm 0.20\,\unit{Wm^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
 In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
@@
 It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about 3--15\,\unit{\%} \citep{PJ:Ainsworth2017} as well as 7\,\unit{\%} loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
-The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.
-Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days \citep{JGR:Stevenson2005,ACP:Young2013}.
-It is mainly produced photochemically in situ from precursor gases such as hydrocarbons (e.g., \chem{CH_4}) in the presents of nitrogen oxides (\chem{NO_x}), carbon monoxide (\chem{CO}):
+The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
+Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs) and hydrocarbons (e.g., \chem{CH_4}) in the presents of nitrogen oxides (\chem{NO_x}).
+A typical reaction cylce for \chem{CO} is sketched below.
+In a sequence of rapid reactions a peroxy radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}
 \begin{reaction}
   ^\bullet\chem{OH} + \chem{CO} \rightarrow {^\bullet\chem{HOCO}}\\
-  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{H_2O} + \chem{CO_2}
+  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{HO_2^\bullet} + \chem{CO_2}.
 \end{reaction}
+Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolysed. The resulting atomic oxygen reacts then with \chem{O_2} to form an ozone molecule
 \begin{reaction}
-  ^\bullet\chem{HO_2} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
+  \chem{HO_2^\bullet} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
   \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
-  ^\bullet\chem{O(^3P)} + \chem{O_2} \rightarrow \chem{O_3}
+  ^\bullet\chem{O(^3P)} + \chem{O_2} \rightarrow \chem{O_3}.
 \end{reaction}
-This leads to a net production via:
+Such a cycle leads to a net production via:
 \begin{reaction}
-  \chem{CO} + 2\chem{O_2} + h\nu \rightarrow \chem{CO_2} + \chem{O_3}
+  \chem{CO} + 2\chem{O_2} + h\nu \rightarrow \chem{CO_2} + \chem{O_3}.
 \end{reaction}
-A part from these, additional reaction pathways involving volatile organic compounds (VOCs) exist.
-Another possible source of tropospheric ozone transportation down from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
-Major sinks are photolytical destruction, de-nitration via
+A part from this, additional reaction pathways involving volatile organic compounds (VOCs) exist.
+Another possible source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
+Major sinks are photolysis to \chem{O({^1{D}})} followed by a reaction with water vapor, reaction with \chem{HO_2} \citep{ACP:Seinfeld2006}, and de-nitration via
 \begin{reaction}
@@
 \end{reaction}
-and dry deposition processes.
-The global abundance of tropospheric ozone is highly variable and it changes, e.g., with time of the day ($\sim$15:00 local time), season (mid-June), altitude, and location \citep{ACP:Schnell2015}.
-Typical concentrations of surface ozone reach from 10\,\unit{ppb} over the tropical Pacific to 100\,\unit{ppb} in the surrounding areas of high emitting sources \citep[Chapter 8]{IPCC2013}.
-This variability poses a challenge on both trend analysis from observation as well as validation and intercomparison of models.
-From the observational side, there is only a limited number of long-term ozone observations, mainly restricted to European sites.
-From these only three display a statistical significant trend, indicating a doubling of tropospheric ozone since the onset of the industrialization \citep[Chapter 2]{IPCC2013}.
-But especially the the very low pre-industrial ozone values cannot be reproduced by the likes of most models \citep[Chapter 8]{IPCC2013}.
+and scavenging processes, in particular dry deposition.
+A very special sink which is linked to dry deposition are so called ozone depleting events.
+They occure in spring-time in the polar boundary layer where an outburst of bromine monoxid \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
+Various schemes have been sucessfully applied to different types of atmospheric models \citep[e.g.][]{ACP:Yang2010,ACP:Toyota2011,ACP:Toyota2014a,GMD:Falk2018}.
+Although these events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes, yet.\\
+Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
+The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, and location \citep{ACP:Schnell2015} or weather conditions in general \citep{ACP:Otero2018}.
+Typical concentrations of surface ozone range from 10\,\unit{ppb} over the tropical Pacific to 100\,\unit{ppb} in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
+This variability poses a challenge on both, trend analysis from observation as well as validation and intercomparison of models.
+At the observational side, there is only a limited number of long-term ozone observations, mainly restricted to European sites.
+Amongst these, only three display a statistical significant trend, indicating a doubling of tropospheric ozone since the onset of the industrialization \citep[Chapter 2]{IPCC2013}.
+But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models, while there is a general tendency to underestimate tropospheric ozone among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP) \citep[Chapter 8]{IPCC2013}.\\
+Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
+The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
+Dry deposition velocities $v^i_d$ (also refered to as contuctances $G^i$) for any gaseous specie $i$ are typically calculated following a resistance-analogous approach
+\begin{equation}
+  v^i_d = \frac{1}{R_a + R^i_b + R^i_c},
+\end{equation}
+wherein aerodynamic $R_a$, quasi laminar $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
+For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gaseous specie to specie and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
+In its original formulation, fixed seasonal average dry deposition resistances have been used for each landuse type.
+For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years, e.g., dry deposition of ozone to ocean \citep{ACP:Luhar2017} and stomatal conductance as part of canopy conductance \citep{PP:Buckley2017} or updates on the parameters have been provided from model validation with observation \citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
 
-, relation to weather \citep{ACP:Otero2018}
-Dry deposition in detail, conductance sum \citep{AE:Wesely1989}, stomatal conductance \citep{PP:Buckley2017}, snow/ice \citep{ACP:Helmig2007,AE:Oltmans2008} and water \citep{JGR:Helmig2012,ACP:Luhar2017}.
-
-What is the status and what are the challenges in the dry deposition of ozone - we do not include bromine explosion in the polar regions which are tremendously reducing surface ozone in polar spring \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
-simple plant function type (PFT)-based scheme. In this scheme, an average canopy resistance is computed from stomatal conductance of fractional PFTs in a grid cell.
-
-Convention on Long-Range Transboundary Air Pollution (CLRTAP) -> European Monitoring and Evaluation Programme (EMEP),
-
-Ozone accumulated over long periods of time inflicts damage on vegetation which leads to loss of primary production in agriculture and forestry.
+What is the status and what are the challenges in the dry deposition of ozone.
 
@@
 Old scheme \citep{AE:Wesely1989, JGR:Hough1991}
+
+historical anthropogenic and biomass burning emissions \citep{Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}
+
 \citet{GMD:Sovde2012,GMD:Lund2018}
 \subsection{Ozone dry deposition scheme}
-\citet{ACP:Simpson2012} \citet{ICP:MappingManual2017}
+\label{subsec:DryDep}
+Convention on Long-Range Transboundary Air Pollution (CLRTAP) -> European Monitoring and Evaluation Programme (EMEP),
+simple plant function type (PFT)-based scheme. In this scheme, an average canopy resistance is computed from stomatal conductance of fractional PFTs in a grid cell.
+\citep{ACP:Simpson2012} \citet{ICP:MappingManual2017}
 \subsection{Greening season and photoactive radiation}
@@
 Include table of all sensitivity runs.
+\citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}
+dry dep velocity over water \citep{JGR:Helmig2012}
 
@@
 \subsection{Comparison with observations}
-The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (http://bodekerscientific.com/).
+The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}).
 
@@
   The used Leaf Area Index (LAI) and roughness length ($Z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A.
-(\doi{10.3334/ORNLDAAC/970}).
+(\doi{10.3334/ORNLDAAC/970}).\\
+  Community Emission Data System (CEDS) historical emission inventory is provided by the Joint Global Research Institute project (\url{http://www.globalchange.umd.edu/ceds/}.)
 \end{acknowledgements}

### 2018-10-05 DryDep.tex
Finished first draft of introduction.

@@
 Furthermore, we investigate the impact on modeled ozone concentrations, both at the surface and in the free troposphere.
-We show that ozone dry deposition in the new version (v.1.1) decreases with respect to the previous version (v1.0), leading to an increase in surface ozone of up to 40\,\unit{\%}.
+We show that ozone dry deposition in the new version (v.1.1) decreases with respect to the previous version (v1.0), leading to an increase in surface ozone of up to ??\,\unit{\%}.
 We also put our results in context to recent multi-model studies as well as satellite data.
+{\bf TODO: Finsish the results section as soon as there are results.}
 \end{abstract}
@@
 The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs) and hydrocarbons (e.g., \chem{CH_4}) in the presents of nitrogen oxides (\chem{NO_x}).
-A typical reaction cylce for \chem{CO} is sketched below.
-In a sequence of rapid reactions a peroxy radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}
+Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs - also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
+A typical reaction mechanism for \chem{CO} is sketched below.
+In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}
 \begin{reaction}
@@
 \end{reaction}
-Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolysed. The resulting atomic oxygen reacts then with \chem{O_2} to form an ozone molecule
+Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolysed.
+The resulting atomic oxygen reacts then with \chem{O_2} (and any other available co-reactant $M$) to form an ozone molecule
 \begin{reaction}
@@
   \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
-  ^\bullet\chem{O(^3P)} + \chem{O_2} \rightarrow \chem{O_3}.
+  ^\bullet\chem{O(^3P)} + \chem{O_2} + M \rightarrow \chem{O_3} + M.
 \end{reaction}
@@
 \end{reaction}
-A part from this, additional reaction pathways involving volatile organic compounds (VOCs) exist.
-Another possible source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
-Major sinks are photolysis to \chem{O({^1{D}})} followed by a reaction with water vapor, reaction with \chem{HO_2} \citep{ACP:Seinfeld2006}, and de-nitration via
+Similar reaction mechanisms involving volatile organic compounds (VOCs) exist \citep{ACP:Monks2015}.
+Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
+Based on observations, STE might only amount to roughly 10\,\unit{\%} ($550 \pm 140$\,\unit{Tg a^{-1}}) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be 5000\,\unit{Tg a^{-1}} \citep{ACP:Monks2015}.
+Ozone is removed from the atmosphere by photochemical reactions or scavenging processes.
+Major sinks are photolysis followed by a reaction with water vapor to from \chem{OH}
 \begin{reaction}
-  \chem{NO} + \chem{O_3} \rightarrow \chem{NO_2} + \chem{O_2}
+  \chem{O_3} + h\nu \rightarrow \chem{O(^1D)} + \chem{O_2})\\
+  \chem{O(^1D)} + \chem{H_2O} \rightarrow 2\chem{OH},
 \end{reaction}
-and
+reactions with \chem{HO_2} \citep{ACP:Seinfeld2006}
 \begin{reaction}
-  \chem{NO_2} + \chem{O_3} \rightarrow \chem{NO_3} + \chem{O_2},
+  {\bf TODO: Reaction},
 \end{reaction}
-and scavenging processes, in particular dry deposition.
-A very special sink which is linked to dry deposition are so called ozone depleting events.
-They occure in spring-time in the polar boundary layer where an outburst of bromine monoxid \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
-Various schemes have been sucessfully applied to different types of atmospheric models \citep[e.g.][]{ACP:Yang2010,ACP:Toyota2011,ACP:Toyota2014a,GMD:Falk2018}.
-Although these events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes, yet.\\
+or de-nitration reaction via
+\begin{reaction}
+  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2}.
+\end{reaction}
+A minor sink which is related to dry deposition are so called ozone depleting events.
+They occur in spring-time in the polar boundary layer where an outburst of bromine monoxide \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
+Various schemes ranging from bulk-snow parameterizations \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models.
+Although these events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3.\\
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
-The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, and location \citep{ACP:Schnell2015} or weather conditions in general \citep{ACP:Otero2018}.
+The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
 Typical concentrations of surface ozone range from 10\,\unit{ppb} over the tropical Pacific to 100\,\unit{ppb} in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
@@
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
-Dry deposition velocities $v^i_d$ (also refered to as contuctances $G^i$) for any gaseous specie $i$ are typically calculated following a resistance-analogous approach
+It is estimated that about $1000 \pm 200$\,\unit{Tg a^{-1}} of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
+Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](h_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
+Dry deposition velocities $v^i_d$ (also referred to as conductance $G^i$) for any gaseous specie $i$ are typically calculated following a resistance-analogous approach
 \begin{equation}
@@
 For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gaseous specie to specie and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
-In its original formulation, fixed seasonal average dry deposition resistances have been used for each landuse type.
-For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years, e.g., dry deposition of ozone to ocean \citep{ACP:Luhar2017} and stomatal conductance as part of canopy conductance \citep{PP:Buckley2017} or updates on the parameters have been provided from model validation with observation \citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
+Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each landuse type.
+For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years, e.g., dry deposition of ozone to ocean \citep{ACP:Luhar2017} and stomatal conductance as part of canopy conductance \citep{PP:Buckley2017} or updates on the parameters have been provided from model validation with observation \citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
+The variaty of differing formulations leads to a wide spread in model intercomparisons \citet{ACP:Hardacre2015} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition.\\
 
-What is the status and what are the challenges in the dry deposition of ozone.
-
-Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
+In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the updates of the dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present the pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season and photosynthetic photon flux density (Section~\ref{subsec:pre-pro}).
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the drydeposition scheme and validate our results with results from the modelintercomparison of \citet{ACP:Hardacre2015} and to observations.
+{\bf TODO: Outlook on paper.}
 
 \section{Model description}
+\label{sec:model_des}
 General model description followed by detailed dry deposition scheme \citep{ACP:Simpson2012,ICP:MappingManual2017}.
@@
 \citet{GMD:Sovde2012,GMD:Lund2018}
+
 \subsection{Ozone dry deposition scheme}
@@
 \citep{ACP:Simpson2012} \citet{ICP:MappingManual2017}
-\subsection{Greening season and photoactive radiation}
+
+\subsection{Greening season and photosynthetic photon flux density}
+\label{subsec:pre-pro}
 
 \section{Evaluation}
+\label{sec:eval}
 
 \subsection{Sensitivity tests}
+\label{subsec:sens}
 Include table of all sensitivity runs.
 \citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}
-dry dep velocity over water \citep{JGR:Helmig2012}
+dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
 
 \subsection{Comparison with recent multi-model evaluation}
+\label{subsec:model}
 Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}.
@@
 \subsection{Comparison with observations}
+\label{subsec:obs}
 The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}).

### 2018-10-08 DryDep.tex
Removed some parts of the abstract, e.g., reference to EMEP. Needs more work at some point.

@@
 Dry deposition velocities are typically calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these conceptual dry deposition velocities are hard to measure explicitly.
-Following the European Monitoring and Evaluation Programme (EMEP) MSC-W chemistry transport model, we compute canopy resistance from stomatal conductance.
-In this paper, we present an update on the dry deposition scheme implemented in the OsloCTM3v1.0 and evaluate the model with respect to ozone and different parameters in the stomatal conductance parameterization, e.g., photosynthetic photon flux density (PPFD) and phenology.
-Furthermore, we investigate the impact on modeled ozone concentrations, both at the surface and in the free troposphere.
-We show that ozone dry deposition in the new version (v.1.1) decreases with respect to the previous version (v1.0), leading to an increase in surface ozone of up to ??\,\unit{\%}.
-We also put our results in context to recent multi-model studies as well as satellite data.
+In this paper, we present an update on the general dry deposition scheme implemented in the OsloCTM3v1.0.
+We evaluate the resulting modelled ozone dry deposition with respect to obervations and multi-model studies.
+We will also estimate the impact of a set of parameters on the modeled ozone concentrations, both at the surface and in the free troposphere.
+We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to ??\,\unit{\%}.
 {\bf TODO: Finsish the results section as soon as there are results.}

### 2018-10-09 DryDep.tex
Finished revision of first draft of introduction. May be tagged.

@@
 \begin{abstract}
-  Since the industrial revolution, tropospheric background ozone concentrations have been increasing strongly mainly in the northern hemisphere.
-In recent years, the number of episodes of peak concentrations especially in summertime has been decreasing in North America and Europe due to the implementation of air quality regulations.
+  Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere.
+In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations.
 At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
 High concentrations of ozone in ambient air are hazardous not only to humans but to the ecosystem in general.
-The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future.
-Most chemistry-transport models (CTMs) and chemistry-climate models (CCMs) tend to overestimate surface ozone values.
-This may be partly due to uncertainties in the distribution and amount of precursor species and stratosphere-troposphere exchange as well as scavenging processes removing ozone from the atmosphere.
+The impact of ozone damage on vegetation and agricultural plants in combination with advancing climate change may affect food security in the future.
+While the future scenarios in themselves are uncertain, there are limiting factors constraining the accuracy of surface ozone modeling also at present: The distribution and amount of precursor/decomposing species, the stratosphere-troposphere exchange as well as scavenging processes.
 Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
-Dry deposition velocities are typically calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these conceptual dry deposition velocities are hard to measure explicitly.
-In this paper, we present an update on the general dry deposition scheme implemented in the OsloCTM3v1.0.
-We evaluate the resulting modelled ozone dry deposition with respect to obervations and multi-model studies.
+The conceptual dry deposition velocities are calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these are hard to measure explicitly.
+In this paper, we present the update from an empirical dry deposition parameterization to a more process based implementation in the OsloCTM3.
+Our focus lies mainly on a stomatal conductance-based description of the canopy resistance.
+We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies.
 We will also estimate the impact of a set of parameters on the modeled ozone concentrations, both at the surface and in the free troposphere.
 We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to ??\,\unit{\%}.
-{\bf TODO: Finsish the results section as soon as there are results.}
+{\bf TODO: Finish the results section as soon as there are results.}
 \end{abstract}
@@
 It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about 3--15\,\unit{\%} \citep{PJ:Ainsworth2017} as well as 7\,\unit{\%} loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
-The impact of ozone damage on vegetation and agricultural plants in particular may even affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
+The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
 Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs - also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
@@
 \end{reaction}
-Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolysed.
+Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
 The resulting atomic oxygen reacts then with \chem{O_2} (and any other available co-reactant $M$) to form an ozone molecule
@@
 \end{reaction}
-Similar reaction mechanisms involving volatile organic compounds (VOCs) exist \citep{ACP:Monks2015}.
+Similar cycles involving volatile organic compounds (VOCs) exist \citep{ACP:Monks2015}.
 Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
@@
 \end{reaction}
-or de-nitration reaction via
+de-nitration reactions via
 \begin{reaction}
-  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2}.
+  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2},
 \end{reaction}
-A minor sink which is related to dry deposition are so called ozone depleting events.
+and dry deposition.
+We will come back to the latter later in this section and cover the implemented scheme in more detail in Section~\ref{subsec:DryDep}.
+A dry deposition related sink limited to Arctic regions are so called ozone depleting events.
 They occur in spring-time in the polar boundary layer where an outburst of bromine monoxide \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
-Various schemes ranging from bulk-snow parameterizations \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models.
-Although these events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3.\\
+Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.
+Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3 as of now.\\
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
@@
 Amongst these, only three display a statistical significant trend, indicating a doubling of tropospheric ozone since the onset of the industrialization \citep[Chapter 2]{IPCC2013}.
-But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models, while there is a general tendency to underestimate tropospheric ozone among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP) \citep[Chapter 8]{IPCC2013}.\\
+But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models.
+Among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP), there is a general tendency to underestimate tropospheric ozone burden (e.g., $10-20$\,\unit{\%} negative bias at 250\,\unit{hPa} in the southern hemisphere (SH) tropical region) \citep[Chapter 8]{IPCC2013}.
+With respect to surface ozone, \citet{ACP:Schnell2015} conclude that all ACCMIP models, which reported hourly surface ozone, tend to overestimate surface ozone values in North America and Europe in comparison with available observations.
+A key to fathom these slightly contradicting results may lie in the used dry deposition schemes.\\
 Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
@@
 wherein aerodynamic $R_a$, quasi laminar $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
-For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gaseous specie to specie and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
+For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
 Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each landuse type.
-For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years, e.g., dry deposition of ozone to ocean \citep{ACP:Luhar2017} and stomatal conductance as part of canopy conductance \citep{PP:Buckley2017} or updates on the parameters have been provided from model validation with observation \citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
-The variaty of differing formulations leads to a wide spread in model intercomparisons \citet{ACP:Hardacre2015} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition.\\
+For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years.
+\citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
+An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
+Canopy conductance is formulated at the single-leave level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
+But progress has also been made on process oriented modeling of stomatal conductance \citep{PP:Buckley2017}.
+The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
 
-In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the updates of the dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present the pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season and photosynthetic photon flux density (Section~\ref{subsec:pre-pro}).
-In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the drydeposition scheme and validate our results with results from the modelintercomparison of \citet{ACP:Hardacre2015} and to observations.
-{\bf TODO: Outlook on paper.}
+In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}).
+In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
+{\bf TODO: Precise outlook on paper when writing is done.}
 
@@
 
-historical anthropogenic and biomass burning emissions \citep{Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}
+historical anthropogenic and biomass burning emissions \citep{ACP:Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}
 
@@
 
-\subsection{Comparison with recent multi-model evaluation}
+\subsection{Comparison with recent multi-model comparison}
 \label{subsec:model}
@@
 \section{Discussion}
+\label{sec:disc}
 
@@
 \conclusions  %% \conclusions[modified heading if necessary]
+\label{sec:conc}
 What did we gain?

### 2018-10-09 DryDep.tex
Started to write the model and scheme decription.

@@
 \label{sec:model_des}
-General model description followed by detailed dry deposition scheme \citep{ACP:Simpson2012,ICP:MappingManual2017}.
-This is followed by the necessary pre-processing of files.
-Old scheme \citep{AE:Wesely1989, JGR:Hough1991}
-
+The OsloCTM3 is a three dimensional global chemistry transort model (CTM).
+The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}.
+A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
+\emph{The OsloCTM3 can also be coupled to MEGAN \citep{ACP:Guenther2006}.
+A publication focussing on this is planed.}
+Technical: Partly parallelized (low number of effective cores 16, huge memory consumption if run on T159 resolution, code has been ported to fortran90 keeping the dynamical core UCI-CTM mainly intact.)
 historical anthropogenic and biomass burning emissions \citep{ACP:Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}
-
-\citet{GMD:Sovde2012,GMD:Lund2018}
 
@@
 \label{subsec:DryDep}
-Convention on Long-Range Transboundary Air Pollution (CLRTAP) -> European Monitoring and Evaluation Programme (EMEP),
-simple plant function type (PFT)-based scheme. In this scheme, an average canopy resistance is computed from stomatal conductance of fractional PFTs in a grid cell.
-\citep{ACP:Simpson2012} \citet{ICP:MappingManual2017}
+
+{\bf TODO: Some words about the old scheme \citep{AE:Wesely1989, JGR:Hough1991}...}
+
+We follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{ACP:Simpson2012,ICP:MappingManual2017}.
+It is a more physical approach compared to \citet{AE:Wesely1989, JGR:Hough1991}.
+The scheme is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+Since \chem{CO} has a very small uptake and is not included in the EMEP scheme, the old OsloCTM2 parameterization is kept.
+In adition to the gaseous species, some of the aerosol deposition rates have also been modified, namely black carbon (BC) and organic carbon (OC), sulphuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
 
@@
 \label{subsec:pre-pro}
+      {\bf TODO: The neccessary pre-processing routines...}
 

### 2018-10-10 DryDep.tex
Add drawing of mapping of different landuse categories. Copied all text about Ra and almost all about Rb from manual. Fixed bug in make. TODO: Complete bibliography.

@@
 A typical reaction mechanism for \chem{CO} is sketched below.
-In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}
-\begin{reaction}
-  ^\bullet\chem{OH} + \chem{CO} \rightarrow {^\bullet\chem{HOCO}}\\
-  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{HO_2^\bullet} + \chem{CO_2}.
-\end{reaction}
+In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
+%\begin{reaction}
+%  ^\bullet\chem{OH} + \chem{CO} \rightarrow {^\bullet\chem{HOCO}}\\
+%  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{HO_2^\bullet} + \chem{CO_2}.
+%\end{reaction}
 Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
-The resulting atomic oxygen reacts then with \chem{O_2} (and any other available co-reactant $M$) to form an ozone molecule
-\begin{reaction}
-  \chem{HO_2^\bullet} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
-  \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
-  ^\bullet\chem{O(^3P)} + \chem{O_2} + M \rightarrow \chem{O_3} + M.
-\end{reaction}
+The resulting atomic oxygen reacts then with \chem{O_2} (and any other available co-reactant $M$) to form an ozone molecule.
+%\begin{reaction}
+%  \chem{HO_2^\bullet} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
+%  \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
+%^\bullet\chem{O(^3P)} + \chem{O_2} + M \rightarrow \chem{O_3} + M.
+%\end{reaction}
 Such a cycle leads to a net production via:
@@
 Ozone is removed from the atmosphere by photochemical reactions or scavenging processes.
-Major sinks are photolysis followed by a reaction with water vapor to from \chem{OH}
-\begin{reaction}
-  \chem{O_3} + h\nu \rightarrow \chem{O(^1D)} + \chem{O_2})\\
-  \chem{O(^1D)} + \chem{H_2O} \rightarrow 2\chem{OH},
-\end{reaction}
-reactions with \chem{HO_2} \citep{ACP:Seinfeld2006}
-\begin{reaction}
-  {\bf TODO: Reaction},
-\end{reaction}
-de-nitration reactions via
-\begin{reaction}
-  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2},
-\end{reaction}
+Major sinks are photolysis followed by a reaction with water vapor to from \chem{OH},
+%\begin{reaction}
+%  \chem{O_3} + h\nu \rightarrow \chem{O(^1D)} + \chem{O_2})\\
+%  \chem{O(^1D)} + \chem{H_2O} \rightarrow 2\chem{OH},
+%\end{reaction}
+reactions with \chem{HO_2} \citep{ACP:Seinfeld2006},
+%\begin{reaction}
+%  {\bf TODO: Reaction},
+%\end{reaction}
+de-nitration reactions,
+%\begin{reaction}
+%  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2},
+%\end{reaction}
 and dry deposition.
@@
   v^i_d = \frac{1}{R_a + R^i_b + R^i_c},
+  \label{eq:drydep_velo}
 \end{equation}
-wherein aerodynamic $R_a$, quasi laminar $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
+wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
 For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
@@
 \label{sec:model_des}
-The OsloCTM3 is a three dimensional global chemistry transort model (CTM).
+The OsloCTM3 is a three dimensional global chemistry transport model (CTM).
 The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}.
@@
 We follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{ACP:Simpson2012,ICP:MappingManual2017}.
-It is a more physical approach compared to \citet{AE:Wesely1989, JGR:Hough1991}.
-The scheme is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
-Since \chem{CO} has a very small uptake and is not included in the EMEP scheme, the old OsloCTM2 parameterization is kept.
+It is a more physical approach compared to the previous used Wesely scheme.
+The EMEP scheme is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+Since \chem{CO} has a very small uptake and is not included in the EMEP scheme, the old OsloCTM2 \citep{OsloCTM2} parameterization is kept.
 In adition to the gaseous species, some of the aerosol deposition rates have also been modified, namely black carbon (BC) and organic carbon (OC), sulphuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
+In Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is splitted into three different parts which will be addressed in the following.
+
+\subsubsection{Aerodynamic resistance}
+\label{subsubsec:Ra}
+The aerodynamical resistance $R_a$ as described in \citet{Simpson2003,ACP:Simpson2012}
+\begin{equation}
+  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{z_0}\right)-\Psi_m\left(\frac{z_0}{L}\right)}\right],
+\end{equation}
+with $\kappa$ the K\'{a}rm\'{a}n constant, $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
+For certain values of $z$, $z_0$, and $L$, this may result in unphysical (negative) values for $R_a$.
+For this reason, we diverge from the EMEP scheme at this point and follow the method of \citet{Monteith1973}.
+The sensible heat flux $\Phi_{Q_\text{sensible}}$ can be written as
+\begin{equation}
+\Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z R_a},
+\end{equation}
+wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at reference height $T_z$, and $\partial_z R_a$ the diffusion resistance to sensible heat between surface and the reference height.
+From the eddy diffusion theorem, a similar formulation can be derived
+\begin{equation}
+  \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z u} \cdot u_*^2.
+\end{equation}
+From this and assuming $\partial_z R_a \rightarrow R_a $ and $\partial_z u \rightarrow u_z $ for a finite $z$, we find 
+\begin{equation}
+  R_a = \frac{u_z}{u_*^2}.
+\end{equation}
+As reference height $z$ we use the height at midlevel of the lowermost model level (roughly 8\,\unit{m}) and $u_z$ and $u_*$ are available from the meteorological input fields.
+
+\subsubsection{Quasi-laminar layer resistance}
+\label{subsubsec:Rb}
+$R_b^i$ is species specific, as already mentioned, and its definition differs over land and ocean.
+In case a gridbox contains both land and ocean, a weighted mean is calculated using the land and ocean conductances, respectively.\\
+
+Over land we use
+\begin{equation}
+  R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
+\end{equation}
+wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$.
+Assuming a constant kinetic viscosity of air $\nu_\text{air}$, $\nu_\text{air}$ can be calculated from the Schmidt number for water $\text{Sc}_\chem{H_2O}=0.6$ and a molecular diffusivity of $D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$
+\begin{equation}
+  \text{Sc}_i = \frac{\nu}{D_i} = \text{Sc}_\chem{H_2O}\cdot\frac{D_\chem{H_2O}}{D_i},
+\end{equation}
+with $D_i$ the molecular diffusivity for a gas $i$. The ration $D_\chem{H_2O}/D_i$ is given in Table~\ref{tab:diffusivity}.\\
+%
+\begin{table}[t]
+  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$.}
+  \begin{tabular}{lc}
+    \tophline
+     & $D_\chem{H_2O}/D_i$\\
+    \middlehline
+    \chem{O_3} & 1.6\\
+    \chem{SO_2} & 1.9\\
+    \chem{NO_2} & 1.6\\
+    \chem{H_2O_2} & 1.4\\
+    \chem{HCHO} & 1.3\\
+    \chem{NO} & 1.3\\
+    \chem{CH_3CHO} & 1.6\\
+    \bottomhline
+  \end{tabular}
+  %\belowtable{} % Table Footnotes
+  \label{tab:diffusivity}
+\end{table}
+
+Over ocean we use
+\begin{equation}
+  R_b^i = \frac{1}{\kappa u_*}\cdot\ln\left({\frac{z_0}{D_i}\cdot \kappa u_*}\right)
+\end{equation}
+with a minimum auf 10\,\unit{ms^{-1}} and a maximum of 1000\,\unit{ms^{-1}}.
+The computation of roughneth length $z_0$ over ocean is devided into a \emph{calm} and a \emph{rough} ocean case, where windspeeds above 3\,\unit{ms^{-1}} are considered as boundary between the cases.
+Calm sea follows \citet{Hinze1975,Garrat1992}:
+\begin{equation}
+  z_0^\text{calm water} = \text{min}\left(2\cdot10^{-3}, 0.135 \frac{\nu}{u_*}\right),
+\end{equation}
+where $\nu$ is the kinematic viscosity of air, computed from surface pressure $P_0$, 2\,\unit{m} temperature $T\text{2M}$, and universal gas constant for air $R_\text{air}$
+\begin{equation}
+  \nu = \frac{6.2\cdot 10^{-8} \cdot T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}}.
+\end{equation}
+\subsubsection{Surface resistance}
+\label{subsubsec:Rc}
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=8.3cm]{pictures/pft_categories.pdf}\\
+%  \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
+%\caption{Plant functional types (PFT). (a) CML2 dynamic landuse 0.5x0.5\,\unit{^\circ}}
+%\label{fig:pft_landuse}
+%\end{figure}
+Mapping of pft landuse types from either ISLSCP2 MODIS or CML2 categories to 9 classical biosphere types used in the EMEP scheme.
+EMEP categories: Forests, Mediterranean scrub, Crops, Moorland (savanna++), Grassland, Wetlands, Tundra, Desert, Water, Urban
+Weighting of conductance with landuse fraction and sum. Inverse of this is gridbox average resistance.
 

### 2018-10-11 DryDep.tex
Update of bibliography. Finished writing about Rb... some questions remain... where does the absolute viscosity come from? And where do the tabulated values come from?

@@
 
-{\bf TODO: Some words about the old scheme \citep{AE:Wesely1989, JGR:Hough1991}...}
+The original dry deposition scheme of the OsloCTM2 and OsloCTM3 was not based on actual calculation of the deposition velocities from equations taking the state of the atmosphere into account, but rather paramerterized following the original work of \citet{AE:Wesely1989} with paramerter updates from \citet{JGR:Hough1991}.
+This means that seasonal day and night average deposition velocities for different land use types (water, forest, grass, tundra/desert, and ice and snow) had been used.
+Day was distiquished from night by solar zenith angles below 90\,\unit{^\circ}.
+Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses.
+Since winter and summer parameters for ocean are equal in this parameterization, no distinquct treatment was needed for ocean gridboxes.
+In addition, reduced uptake due to snow cover above 1\,\unit{m} and 10\,\unit{cm} for forest and grass/tundra, respectively, was taken into account.
+In the following, we will refer to this static parameterization as \emph{Wesley scheme}.\\
 
-We follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{ACP:Simpson2012,ICP:MappingManual2017}.
-It is a more physical approach compared to the previous used Wesely scheme.
-The EMEP scheme is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
-Since \chem{CO} has a very small uptake and is not included in the EMEP scheme, the old OsloCTM2 \citep{OsloCTM2} parameterization is kept.
+Getting away from the Wesely schme, we follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{Simpson2003,ACP:Simpson2012}.
+We will refere to this as \emph{EMEP scheme} thoughout the rest of the paper.
+The EMEP scheme is a more physical approach compared to the previous used Wesely scheme.
+It is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+Since \chem{CO} has a very small uptake and is not included in \citet{Simpson2003,ACP:Simpson2012}, the old OsloCTM2 \citep{OsloCTM2} parameterization is kept.
 In adition to the gaseous species, some of the aerosol deposition rates have also been modified, namely black carbon (BC) and organic carbon (OC), sulphuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
-In Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is splitted into three different parts which will be addressed in the following.
+As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically splitted into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which will be addressed seperately in the following.
 
@@
 \label{subsubsec:Ra}
-The aerodynamical resistance $R_a$ as described in \citet{Simpson2003,ACP:Simpson2012}
+The aerodynamical resistance $R_a$ in \citet{Simpson2003,ACP:Simpson2012} is described as
 \begin{equation}
@@
 \end{equation}
-wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at reference height $T_z$, and $\partial_z R_a$ the diffusion resistance to sensible heat between surface and the reference height.
+wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at reference height $T_z$, and $\partial_z R_a$ the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$.
 From the eddy diffusion theorem, a similar formulation can be derived
@@
 $R_b^i$ is species specific, as already mentioned, and its definition differs over land and ocean.
-In case a gridbox contains both land and ocean, a weighted mean is calculated using the land and ocean conductances, respectively.\\
-
+In case a gridbox contains both land and ocean, a weighted mean is calculated using the land and ocean conductances, respectively.
+\subsection*{Land}
 Over land we use
@@
 wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$.
-Assuming a constant kinetic viscosity of air $\nu_\text{air}$, $\nu_\text{air}$ can be calculated from the Schmidt number for water $\text{Sc}_\chem{H_2O}=0.6$ and a molecular diffusivity of $D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$
+From the definition of $\text{Sc}_i = \nu/D_i$, with the kinematic viscosity of air $\nu$, $\text{Sc}_i $ can be derived as water equivalent:
 \begin{equation}
@@
 \end{equation}
-with $D_i$ the molecular diffusivity for a gas $i$. The ration $D_\chem{H_2O}/D_i$ is given in Table~\ref{tab:diffusivity}.\\
+with the molecular diffusivity for any gas $D_i$, the Schmidt number of water $\text{Sc}_\chem{H_2O}=0.6$ and a its molecular diffusivity $D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$.
+The ratios $D_\chem{H_2O}/D_i$ are given in Table~\ref{tab:diffusivity}.
 %
@@
 \end{table}
-
+%
+\subsection*{Ocean}
 Over ocean we use
@@
 \end{equation}
-with a minimum auf 10\,\unit{ms^{-1}} and a maximum of 1000\,\unit{ms^{-1}}.
-The computation of roughneth length $z_0$ over ocean is devided into a \emph{calm} and a \emph{rough} ocean case, where windspeeds above 3\,\unit{ms^{-1}} are considered as boundary between the cases.
-Calm sea follows \citet{Hinze1975,Garrat1992}:
+with a lower threashold of 10\,\unit{ms^{-1}} and an upper limit of 1000\,\unit{ms^{-1}}.
+The computation of roughness length $z_0$ over ocean is devided into a \emph{calm} and a \emph{rough} sea case, where windspeeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
+Calm sea follows \citet{Hinze1975,Garrat1992} with a slightly increased coefficient:
 \begin{equation}
-  z_0^\text{calm water} = \text{min}\left(2\cdot10^{-3}, 0.135 \frac{\nu}{u_*}\right),
+  z_0^\text{calm} = \text{min}\left(2\cdot10^{-3}, 0.135 \frac{\nu}{u_*}\right).
 \end{equation}
-where $\nu$ is the kinematic viscosity of air, computed from surface pressure $P_0$, 2\,\unit{m} temperature $T\text{2M}$, and universal gas constant for air $R_\text{air}$
+The kinematic viscosity of air $\nu = \mu/\rho$ can be computed from the absolute viscosity of air $\mu$, and substituting $\rho$ uising the gas equation by surface pressure $P_0$, 2\,\unit{m} temperature $T_\text{2M}$, and the universal gas constant for air $R_\text{air}$
 \begin{equation}
-  \nu = \frac{6.2\cdot 10^{-8} \cdot T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}}.
+  \nu = \frac{6.2\cdot 10^{-8}\,\unit{kg \cdot m^{-1}\cdot s^{-1} \cdot T^{-1}} \cdot T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}}.
 \end{equation}
+The rough sea case follows the method of \citet{Charnock1955,Wu1980}:
+\begin{equation}
+  z_0^\text{rough} = \text{min}\left(2\cdot10^{-3}, 0.018 \frac{u^2_*}{g}\right).
+\end{equation}
+The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
+Since the $z_0$ computed with this paramterization is very small, the resulting $R^i_b$ is often defined by its imposed lower limit of 10\,\unit{sm^{-1}}.
+
 \subsubsection{Surface resistance}
@@
 Weighting of conductance with landuse fraction and sum. Inverse of this is gridbox average resistance.
+-> \citep{ICP:MappingManual2017}
 

### 2018-10-12 DryDep.tex
Continued writing on Section 2. Started with stomatal and non-stomatal conductance. Included comments by Frode.

@@
 \begin{abstract}
-  Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere.
-In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations.
-At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
  High concentrations of ozone in ambient air are hazardous not only to humans but to the ecosystem in general.
@@
 In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
+Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere.
+In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations.
+At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
 Continuously high concentrations of ambient air ozone are hazardous to the whole ecosystem.
@@
 A key to fathom these slightly contradicting results may lie in the used dry deposition schemes.\\
-Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
+Removal of any substance through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
@@
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](h_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
-Dry deposition velocities $v^i_d$ (also referred to as conductance $G^i$) for any gaseous specie $i$ are typically calculated following a resistance-analogous approach
+Dry deposition velocities $v^i_d$ (also referred to as conductance $G^i$) for any gaseous species $i$ are typically calculated following a resistance-analogous approach
 \begin{equation}
@@
 wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
-For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on landuse types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
-Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each landuse type.
+For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
+Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type.
 For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years.
@@
 \emph{The OsloCTM3 can also be coupled to MEGAN \citep{ACP:Guenther2006}.
-A publication focussing on this is planed.}
+A publication focusing on this is planed.}
 Technical: Partly parallelized (low number of effective cores 16, huge memory consumption if run on T159 resolution, code has been ported to fortran90 keeping the dynamical core UCI-CTM mainly intact.)
@@
 
-The original dry deposition scheme of the OsloCTM2 and OsloCTM3 was not based on actual calculation of the deposition velocities from equations taking the state of the atmosphere into account, but rather paramerterized following the original work of \citet{AE:Wesely1989} with paramerter updates from \citet{JGR:Hough1991}.
-This means that seasonal day and night average deposition velocities for different land use types (water, forest, grass, tundra/desert, and ice and snow) had been used.
-Day was distiquished from night by solar zenith angles below 90\,\unit{^\circ}.
+In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account.
+Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}.
+This means that seasonal day and night average deposition velocities for different land use types (water, forest, grass, tundra/desert, and ice and snow) were in use.
+Day was distinguished from night by solar zenith angles below 90\,\unit{^\circ}.
 Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses.
-Since winter and summer parameters for ocean are equal in this parameterization, no distinquct treatment was needed for ocean gridboxes.
-In addition, reduced uptake due to snow cover above 1\,\unit{m} and 10\,\unit{cm} for forest and grass/tundra, respectively, was taken into account.
-In the following, we will refer to this static parameterization as \emph{Wesley scheme}.\\
+Since winter and summer parameters for ocean are equal in this parameterization, no distinctive treatment was needed for ocean gridboxes.
+In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account.
+In the following, we will refer to this static parameterization as \emph{Wesely scheme}.\\
 
-Getting away from the Wesely schme, we follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{Simpson2003,ACP:Simpson2012}.
-We will refere to this as \emph{EMEP scheme} thoughout the rest of the paper.
-The EMEP scheme is a more physical approach compared to the previous used Wesely scheme.
-It is used for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
-Since \chem{CO} has a very small uptake and is not included in \citet{Simpson2003,ACP:Simpson2012}, the old OsloCTM2 \citep{OsloCTM2} parameterization is kept.
-In adition to the gaseous species, some of the aerosol deposition rates have also been modified, namely black carbon (BC) and organic carbon (OC), sulphuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
-As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically splitted into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which will be addressed seperately in the following.
+For the new dry deposition scheme, we follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{WASP:Simpson2003,ACP:Simpson2012}.
+We will refer to this as \emph{EMEP scheme} throughout the rest of the paper.
+The EMEP scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature) as well as dynamics (e.g., wind stress) of the atmosphere into account and also acknowledges to a certain degree the global variety of plants and the variability throughout the seasons.
+It is implemented for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
+In addition to the gaseous species, some of the aerosol deposition rates are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
+As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which will be addressed separately in the following.
 
@@
 \label{subsubsec:Ra}
-The aerodynamical resistance $R_a$ in \citet{Simpson2003,ACP:Simpson2012} is described as
+The aerodynamic resistance $R_a$ in \citet{WASP:Simpson2003,ACP:Simpson2012} is described as
 \begin{equation}
@@
 with $\kappa$ the K\'{a}rm\'{a}n constant, $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
-For certain values of $z$, $z_0$, and $L$, this may result in unphysical (negative) values for $R_a$.
+For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
 For this reason, we diverge from the EMEP scheme at this point and follow the method of \citet{Monteith1973}.
@@
   \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z R_a},
+  \label{eq:sens_heat}
 \end{equation}
@@
   \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z u} \cdot u_*^2.
+  \label{eq:eddy_theo}
 \end{equation}
-From this and assuming $\partial_z R_a \rightarrow R_a $ and $\partial_z u \rightarrow u_z $ for a finite $z$, we find 
+Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $\partial_z R_a \rightarrow R_a $ and $\partial_z u \rightarrow u_z $ for a finite $z$, we find 
 \begin{equation}
@@
 \end{equation}
-As reference height $z$ we use the height at midlevel of the lowermost model level (roughly 8\,\unit{m}) and $u_z$ and $u_*$ are available from the meteorological input fields.
+As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields.
 
@@
 \label{subsubsec:Rb}
+
 $R_b^i$ is species specific, as already mentioned, and its definition differs over land and ocean.
@@
 \subsection*{Land}
-Over land we use
+Over land we use the definition \citep[Eq.~(53),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \begin{table}[t]
-  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$.}
-  \begin{tabular}{lc}
+  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
+  \begin{tabular}{lcrr}
     \tophline
-     & $D_\chem{H_2O}/D_i$\\
+    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & $H^i_*$\\
+    & & & (\unit{M\cdot atm^{-1}})\\
     \middlehline
-    \chem{O_3} & 1.6\\
-    \chem{SO_2} & 1.9\\
-    \chem{NO_2} & 1.6\\
-    \chem{H_2O_2} & 1.4\\
-    \chem{HCHO} & 1.3\\
-    \chem{NO} & 1.3\\
-    \chem{CH_3CHO} & 1.6\\
+    \chem{O_3}     & 1.6 & 1.0 & $10^{-2}$\\
+    \chem{SO_2}    & 1.9 & 0.0 & $10^5$\\
+    \chem{NO_2}    & 1.6 & 0.1 & $10^{-2}$\\
+    \chem{H_2O_2}  & 1.4 & 1.0 & $10^5$\\
+    \chem{HCHO}    & 1.3 & 0.0 & $6 \cdot 10^3$\\
+    \chem{NO}      & 1.3 & 0.0 & $2 \cdot 10^{-3}$\\
+    \chem{CH_3CHO} & 1.6 & 15.0 & $0.0$\\
     \bottomhline
@@
 \subsection*{Ocean}
-Over ocean we use
+Over ocean we use the definition \citep[Eq.~(54),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \end{equation}
-with a lower threashold of 10\,\unit{ms^{-1}} and an upper limit of 1000\,\unit{ms^{-1}}.
-The computation of roughness length $z_0$ over ocean is devided into a \emph{calm} and a \emph{rough} sea case, where windspeeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
-Calm sea follows \citet{Hinze1975,Garrat1992} with a slightly increased coefficient:
+with an imposed a lower threshold of 10\,\unit{sm^{-1}} and an upper limit of 1000\,\unit{sm^{-1}}.
+The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
+Calm sea follows \citet{Hinze1975,Garratt1992} with a slightly increased coefficient:
 \begin{equation}
-  z_0^\text{calm} = \text{min}\left(2\cdot10^{-3}, 0.135 \frac{\nu}{u_*}\right).
+  z_0^\text{calm} = \text{min}\left(2\cdot10^{-3}, 0.135 \cdot \frac{\nu}{u_*}\right).
 \end{equation}
-The kinematic viscosity of air $\nu = \mu/\rho$ can be computed from the absolute viscosity of air $\mu$, and substituting $\rho$ uising the gas equation by surface pressure $P_0$, 2\,\unit{m} temperature $T_\text{2M}$, and the universal gas constant for air $R_\text{air}$
+The kinematic viscosity of air $\nu = \mu/\rho$ can be computed from the absolute viscosity of air $\mu$, and substituting $\rho$ with surface pressure $P_0$ using the ideal gas law, 2\,\unit{m} temperature $T_\text{2M}$, and the universal gas constant for air $R_\text{air}$
 \begin{equation}
-  \nu = \frac{6.2\cdot 10^{-8}\,\unit{kg \cdot m^{-1}\cdot s^{-1} \cdot T^{-1}} \cdot T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}}.
+  \nu = \frac{T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}} \cdot 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}}.
 \end{equation}
-The rough sea case follows the method of \citet{Charnock1955,Wu1980}:
+The rough sea case follows the method of \citet{QJRMS:Charnock1955,JPO:Wu1980}:
 \begin{equation}
-  z_0^\text{rough} = \text{min}\left(2\cdot10^{-3}, 0.018 \frac{u^2_*}{g}\right).
+  z_0^\text{rough} = \text{min}\left(2\cdot10^{-3}, 0.018 \cdot \frac{u^2_*}{g}\right).
 \end{equation}
 The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
-Since the $z_0$ computed with this paramterization is very small, the resulting $R^i_b$ is often defined by its imposed lower limit of 10\,\unit{sm^{-1}}.
+Since the $z_0$ computed with this parameterization is very small, the resulting $R^i_b$ is often defined by its imposed lower limit of 10\,\unit{sm^{-1}}.
 
@@
 \label{subsubsec:Rc}
-%\begin{figure}[t]
-%  \centering
-%  \includegraphics[width=8.3cm]{pictures/pft_categories.pdf}\\
-%  \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
-%\caption{Plant functional types (PFT). (a) CML2 dynamic landuse 0.5x0.5\,\unit{^\circ}}
-%\label{fig:pft_landuse}
-%\end{figure}
-Mapping of pft landuse types from either ISLSCP2 MODIS or CML2 categories to 9 classical biosphere types used in the EMEP scheme.
-EMEP categories: Forests, Mediterranean scrub, Crops, Moorland (savanna++), Grassland, Wetlands, Tundra, Desert, Water, Urban
-Weighting of conductance with landuse fraction and sum. Inverse of this is gridbox average resistance.
+
+The surface resistance $R_c^i$ consists of both, stomatal and non-stomatal resistances. When computing gridbox averages, we use the sum of conductances (deposition velocities)
+\begin{equation}
+  G_c^i = \text{LAI} \cdot G_\text{sto}^i + G_\text{ns}^i, 
+\end{equation}
+wherein $\text{LAI}$ is the leaf area index (zero for non-vegetated surfaces), $G_\text{sto}^i$ the stomatal conductance, and $G_\text{ns}^i$ the non-stomatal conductance.
+
+{\bf TODO: Mapping of pft land surface types from either ISLSCP2 MODIS or CML2 categories to 9 classical biosphere types used in the EMEP scheme.
+EMEP categories: Forests, Mediterranean scrub, Crops, Moorland (savanna++), Grassland, Wetlands, Tundra, Desert, Water, Urban.
+Weighting of conductance with land surface fraction and sum.
+Inverse of this is gridbox average resistance.}
+
+\subsubsection*{Stomatal conductance}
 -> \citep{ICP:MappingManual2017}
+{\bf TODO: Write about latitude dependent vegetation height and EMEP algorithm in general.}
 
-\subsection{Greening season and photosynthetic photon flux density}
+\subsubsection*{Non-stomatal conductance}
+In the EMEP scheme, non-stomatal conductances are explicitly calculated for \chem{O_3}, \chem{SO_2}, \chem{HNO_3}, and \chem{NH_3}.
+For all other species, an interpolation between \chem{O3} and \chem{SO_2} values is carried out.
+
+The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending in the soil/surface. For a land surface type $N$, we can write
+\begin{equation}
+  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
+\end{equation}
+$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces.
+The external leaf resistance $r_\text{ext} = 2000\,\unit{sm^{-1}}\cdot F_T$.
+Below $-1\,\unit{^\circ C}$,s a temperature correction factor $F_T$ is used
+\begin{equation}
+  F_T = \exp{(-2\cdot(1+T_0))}.
+\end{equation}
+It ranges between 1--2.
+\subsection{Pre-processing}
 \label{subsec:pre-pro}
-      {\bf TODO: The neccessary pre-processing routines...}
+      {\bf TODO: The necessary pre-processing routines...}
+
+      \subsubsection{Greening season}
+
+      \subsubsection{Photosynthetic photon flux density (PPFD)}
 
@@
 \label{subsec:model}
-Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}.
+Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
 
@@
 \label{subsec:obs}
-The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}).
+The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}). \citep{AE:Hu2017} -> need to look into this one...
 
@@
 
-%\appendix
-%\section{}    %% Appendix A
-
-%\subsection{}     %% Appendix A1, A2, etc.
+\appendix
+\section{Land use and plant function types (PFT)}    %% Appendix A
+\appendixfigures
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=8.3cm]{pictures/pft_categories.pdf}\\
+  \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
+  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface 0.5x0.5\,\unit{^\circ}}
+  \label{fig:pft_landsurface}
+\end{figure}
+\subsection{}     %% Appendix A1, A2, etc.
 

### 2018-10-15 DryDep.tex
Continued writing about non-stomatal conductance. Next is to write about vegetation height and snow cover.

@@
 The impact of ozone damage on vegetation and agricultural plants in combination with advancing climate change may affect food security in the future.
-While the future scenarios in themselves are uncertain, there are limiting factors constraining the accuracy of surface ozone modeling also at present: The distribution and amount of precursor/decomposing species, the stratosphere-troposphere exchange as well as scavenging processes.
-Removal of any gas through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
+While the future scenarios in themselves are uncertain, there are limiting factors constraining the accuracy of surface ozone modeling also at present: The distribution and amount of ozone precursor/depleting substances, the stratosphere-troposphere exchange as well as scavenging processes.
+Removal of any substance through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
 The conceptual dry deposition velocities are calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these are hard to measure explicitly.
-In this paper, we present the update from an empirical dry deposition parameterization to a more process based implementation in the OsloCTM3.
-Our focus lies mainly on a stomatal conductance-based description of the canopy resistance.
+In this paper, we present an update of the dry deposition scheme implemented in the OsloCTM3.
+We change from an empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere into account.
+Examining the sensitivity of the scheme to certain parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
 We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies.
-We will also estimate the impact of a set of parameters on the modeled ozone concentrations, both at the surface and in the free troposphere.
+We will also estimate the impact on the modeled ozone concentrations, both at the surface and in the free troposphere.
 We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to ??\,\unit{\%}.
@@
 In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
-Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere.
-In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations.
+Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
+In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ACP:Wilson2012}.
 At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
@@
 Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs - also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
-A typical reaction mechanism for \chem{CO} is sketched below.
+A typical reaction mechanism for \chem{CO} is sketched in the following.
 In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
@@
 \end{reaction}
-Similar cycles involving volatile organic compounds (VOCs) exist \citep{ACP:Monks2015}.
+Similar cycles involving VOCs exist \citep{ACP:Monks2015}.
 Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
@@
 %\end{reaction}
-reactions with \chem{HO_2} \citep{ACP:Seinfeld2006},
+reactions with \chem{HO_2},
 %\begin{reaction}
@@
 Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.
-Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3 as of now.\\
+Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3 as of now.
+We also do not consider the contribution of very short-lived ozone depleting substances (VSLS), that affect stratospheric ozone \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017}, to the tropospheric ozone abundance.\\
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
@@
 At the observational side, there is only a limited number of long-term ozone observations, mainly restricted to European sites.
-Amongst these, only three display a statistical significant trend, indicating a doubling of tropospheric ozone since the onset of the industrialization \citep[Chapter 2]{IPCC2013}.
+Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
 But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models.
@@
 It is estimated that about $1000 \pm 200$\,\unit{Tg a^{-1}} of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
-Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](h_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
-Dry deposition velocities $v^i_d$ (also referred to as conductance $G^i$) for any gaseous species $i$ are typically calculated following a resistance-analogous approach
+Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
+Dry deposition velocities $v^i_d$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$ are typically calculated following a resistance-analogous approach
 \begin{equation}
@@
 \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
-An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}.
-Canopy conductance is formulated at the single-leave level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
-But progress has also been made on process oriented modeling of stomatal conductance \citep{PP:Buckley2017}.
-The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
+An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$]{ACP:Helmig2007}.
+Canopy conductance is parameterized at the single-leave level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
+But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
+The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
 
@@
 {\bf TODO: Precise outlook on paper when writing is done.}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Model description}
@@
 A publication focusing on this is planed.}
-Technical: Partly parallelized (low number of effective cores 16, huge memory consumption if run on T159 resolution, code has been ported to fortran90 keeping the dynamical core UCI-CTM mainly intact.)
-historical anthropogenic and biomass burning emissions \citep{ACP:Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}
-
+{\bf TODO: Technical: Partly parallelized (low number of effective cores 16, huge memory consumption if run on T159 resolution, code has been ported to fortran90 keeping the dynamical core UCI-CTM mainly intact.)
+historical anthropogenic and biomass burning emissions \citep{ACP:Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Ozone dry deposition scheme}
 \label{subsec:DryDep}
-
 In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account.
@@
 The EMEP scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature) as well as dynamics (e.g., wind stress) of the atmosphere into account and also acknowledges to a certain degree the global variety of plants and the variability throughout the seasons.
-It is implemented for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+We implement it for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
 In addition to the gaseous species, some of the aerosol deposition rates are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
-As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which will be addressed separately in the following.
+As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which we will recap in the following.
 
@@
 \end{equation}
-with $\kappa$ the K\'{a}rm\'{a}n constant, $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
+with $\kappa$ the K\'{a}rm\'{a}n constant (about $0.40$), $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
-For this reason, we diverge from the EMEP scheme at this point and follow the method of \citet{Monteith1973}.
+For this reason, we diverge from the EMEP scheme at this point and fall back to the method of \citet{Monteith1973}.
 The sensible heat flux $\Phi_{Q_\text{sensible}}$ can be written as
@@
 As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Quasi-laminar layer resistance}
 \label{subsubsec:Rb}
-
-$R_b^i$ is species specific, as already mentioned, and its definition differs over land and ocean.
-In case a gridbox contains both land and ocean, a weighted mean is calculated using the land and ocean conductances, respectively.
-\subsection*{Land}
-Over land we use the definition \citep[Eq.~(53),][]{ACP:Simpson2012}
+The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces.
+In case a gridbox contains both, a weighted mean is calculated using both conductances, respectively.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Land}
+Over land we use \citep[Eq.~(53),][]{ACP:Simpson2012}
 \begin{equation}
   R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
+  \label{eq:Rb}
 \end{equation}
 wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$.
-From the definition of $\text{Sc}_i = \nu/D_i$, with the kinematic viscosity of air $\nu$, $\text{Sc}_i $ can be derived as water equivalent:
+Eq.~(\ref{eq:Rb}) differs from a similar formulation in \citet{ACP:Seinfeld2006} by a factor of roughly $1.25$.
+From $\text{Sc}_i = \nu/D_i$, with the kinematic viscosity of air $\nu$, we derive a Schmidt number in water equivalent:
 \begin{equation}
-  \text{Sc}_i = \frac{\nu}{D_i} = \text{Sc}_\chem{H_2O}\cdot\frac{D_\chem{H_2O}}{D_i},
+  \text{Sc}_i = \frac{D_\chem{H_2O}}{D_i} \cdot \text{Sc}_\chem{H_2O},
 \end{equation}
@@
   \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
-  \begin{tabular}{lcrr}
+  \begin{tabular}{lcrc}
     \tophline
@@
 \end{table}
-%
-\subsection*{Ocean}
-Over ocean we use the definition \citep[Eq.~(54),][]{ACP:Simpson2012}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Ocean}
+Over ocean we use \citep[Eq.~(54),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \end{equation}
-with an imposed a lower threshold of 10\,\unit{sm^{-1}} and an upper limit of 1000\,\unit{sm^{-1}}.
-The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
-Calm sea follows \citet{Hinze1975,Garratt1992} with a slightly increased coefficient:
+with an imposed lower threshold of 10\,\unit{sm^{-1}} and an upper limit of 1000\,\unit{sm^{-1}}.
+The computation of roughness length $z_0$ over ocean is divided into \emph{calm} and a \emph{rough} sea, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
+For calm sea, we apply the limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
 \begin{equation}
@@
 \end{equation}
-The kinematic viscosity of air $\nu = \mu/\rho$ can be computed from the absolute viscosity of air $\mu$, and substituting $\rho$ with surface pressure $P_0$ using the ideal gas law, 2\,\unit{m} temperature $T_\text{2M}$, and the universal gas constant for air $R_\text{air}$
+The kinematic viscosity of air $\nu$ herein can be computed from 
 \begin{equation}
-  \nu = \frac{T_\text{2M}}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}} \cdot 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}}.
+  \nu = \frac{\mu}{\rho} = \frac{\mu(T)}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}},
 \end{equation}
+with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}} \cdot T_\text{2M}$ and substituting the air density $\rho$ using the ideal gas law.
+$P_0$ is the surface pressure, $T_\text{2M}$ the 2\,\unit{m} temperature and $R_\text{air}$ the universal gas constant for air.
 The rough sea case follows the method of \citet{QJRMS:Charnock1955,JPO:Wu1980}:
@@
 Since the $z_0$ computed with this parameterization is very small, the resulting $R^i_b$ is often defined by its imposed lower limit of 10\,\unit{sm^{-1}}.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Surface resistance}
 \label{subsubsec:Rc}
-
 The surface resistance $R_c^i$ consists of both, stomatal and non-stomatal resistances. When computing gridbox averages, we use the sum of conductances (deposition velocities)
@@
 Inverse of this is gridbox average resistance.}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection*{Stomatal conductance}
@@
 {\bf TODO: Write about latitude dependent vegetation height and EMEP algorithm in general.}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection*{Non-stomatal conductance}
 In the EMEP scheme, non-stomatal conductances are explicitly calculated for \chem{O_3}, \chem{SO_2}, \chem{HNO_3}, and \chem{NH_3}.
-For all other species, an interpolation between \chem{O3} and \chem{SO_2} values is carried out.
+For all other species, an interpolation between \chem{O_3} and \chem{SO_2} values is carried out.
 
@@
 \end{equation}
-$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces.
-The external leaf resistance $r_\text{ext} = 2000\,\unit{sm^{-1}}\cdot F_T$.
-Below $-1\,\unit{^\circ C}$,s a temperature correction factor $F_T$ is used
+$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
 \begin{equation}
-  F_T = \exp{(-2\cdot(1+T_0))}.
+  r_\text{ext} = 2000\cdot F_T\,\unit{sm^{-1}}.
 \end{equation}
-It ranges between 1--2.
+Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ ranging between $1-2$
+\begin{equation}
+  F_T = \exp{(-2\cdot(1+\theta_0))}\text{, for }1 \leq F_T \leq 2.
+\end{equation}
+$\theta_0$ is the surface or 2\,\unit{m} temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
+%
+\begin{equation}
+  \text{SAI} = 
+  \begin{cases}
+    \text{LAI}+1 & \text{\emph{forest / wetland}},\\
+    \text{LAI}\cdot 5/3.5 & \text{\emph{cropland, I. part of growing season}},\\
+    \text{LAI} + 1.5 & \text{\emph{cropland, II. part of growing season}},\\
+    0 & \text{\emph{cropland, winter}}.
+  \end{cases}
+\end{equation}
+%
+Extending the EMEP scheme to the southern hemisphere, the growing season for crops is defined as:
+\begin{itemize}
+\item[] NH: I. day 90--140; II. day 141--270
+\item[] SH: I. day 272--322; II. day 323--452 (day 87)
+\end{itemize}
+
+In this way, vegetation affects the conductance also by being there not only by uptake through the stomata. The in canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
+\begin{equation}
+  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N)}{u_*},
+\end{equation}
+where $h(N)$ is the canopy height and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
+According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.
+$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow $f_\text{snow}$:
+\begin{equation}
+  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
+\end{equation}
+$f_\text{snow}$ ranges between $1-2$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Pre-processing}
@@
       {\bf TODO: The necessary pre-processing routines...}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
       \subsubsection{Greening season}
-
+      \label{subsubsection:greening}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
       \subsubsection{Photosynthetic photon flux density (PPFD)}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%      
 \section{Evaluation}
 \label{sec:eval}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Sensitivity tests}
@@
 dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Comparison with recent multi-model comparison}
@@
 Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Comparison with observations}
@@
 The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}). \citep{AE:Hu2017} -> need to look into this one...
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}

### 2018-10-16 DryDep.tex
Finished non-stomatal resistance part and began to work on preprocessing, still need to write about stomatal resistance, though - Added supplement figure - made pft_landuse_mapping look more fancy - updated latex_setup files.

@@
 A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
-\emph{The OsloCTM3 can also be coupled to MEGAN \citep{ACP:Guenther2006}.
-A publication focusing on this is planed.}
-{\bf TODO: Technical: Partly parallelized (low number of effective cores 16, huge memory consumption if run on T159 resolution, code has been ported to fortran90 keeping the dynamical core UCI-CTM mainly intact.)
-historical anthropogenic and biomass burning emissions \citep{ACP:Lamarque2010}, meteorological driver \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}, Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018}}
+The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN) \citep{ACP:Guenther2006}.
+A publication focusing on this is planed.
+Parallelization of the OsloCTM3 has not been optimized as of now.
+It is possible to run the model in resolution of at least T159N80L60, but it is not very time and memory efficient.
+Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$.
+In the following, we will give a detailed account of the new dry deposition scheme.
+Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summerize them, non the less.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}.
-This means that seasonal day and night average deposition velocities for different land use types (water, forest, grass, tundra/desert, and ice and snow) were in use.
+This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use.
 Day was distinguished from night by solar zenith angles below 90\,\unit{^\circ}.
 Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses.
-Since winter and summer parameters for ocean are equal in this parameterization, no distinctive treatment was needed for ocean gridboxes.
+For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes.
 In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account.
-In the following, we will refer to this static parameterization as \emph{Wesely scheme}.\\
+In the following, we will refer to this parameterization as \emph{Wesely scheme}.\\
 
-For the new dry deposition scheme, we follow the Convention on Long-Range Transboundary Air Pollution (CLRTAP) implemented in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{WASP:Simpson2003,ACP:Simpson2012}.
+Regarding the new dry deposition scheme, we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{WASP:Simpson2003,ACP:Simpson2012}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
 We will refer to this as \emph{EMEP scheme} throughout the rest of the paper.
-The EMEP scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature) as well as dynamics (e.g., wind stress) of the atmosphere into account and also acknowledges to a certain degree the global variety of plants and the variability throughout the seasons.
-We implement it for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+The EMEP scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature)  of the atmosphere as well as dynamics (e.g., wind stress) of the boundary layer into account.
+To a certain degree, the global variety of plants and their variability throughout the seasons is also acknowledged.
+The EMEP scheme is implemented for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
@@
 \end{equation}
-wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at reference height $T_z$, and $\partial_z R_a$ the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$.
+wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$.
+$\partial_z R_a$ is the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$.
+$\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
 From the eddy diffusion theorem, a similar formulation can be derived
@@
 \label{subsubsec:Rb}
-The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces.
-In case a gridbox contains both, a weighted mean is calculated using both conductances, respectively.
+The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces. In case a gridbox contains both, a weighted mean is calculated.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-with the molecular diffusivity for any gas $D_i$, the Schmidt number of water $\text{Sc}_\chem{H_2O}=0.6$ and a its molecular diffusivity $D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$.
+with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$).
 The ratios $D_\chem{H_2O}/D_i$ are given in Table~\ref{tab:diffusivity}.
@@
   \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
-  \begin{tabular}{lcrc}
+  \begin{tabular}{lcrr@{$\cdot$}l}
     \tophline
-    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & $H^i_*$\\
-    & & & (\unit{M\cdot atm^{-1}})\\
+    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
+    & & & \multicolumn{2}{l}{(\unit{M\cdot atm^{-1}})}\\
     \middlehline
-    \chem{O_3}     & 1.6 & 1.0 & $10^{-2}$\\
-    \chem{SO_2}    & 1.9 & 0.0 & $10^5$\\
-    \chem{NO_2}    & 1.6 & 0.1 & $10^{-2}$\\
-    \chem{H_2O_2}  & 1.4 & 1.0 & $10^5$\\
-    \chem{HCHO}    & 1.3 & 0.0 & $6 \cdot 10^3$\\
-    \chem{NO}      & 1.3 & 0.0 & $2 \cdot 10^{-3}$\\
-    \chem{CH_3CHO} & 1.6 & 15.0 & $0.0$\\
+    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
+    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
+    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
+    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
+    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
+    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
+    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
     \bottomhline
@@
 with an imposed lower threshold of 10\,\unit{sm^{-1}} and an upper limit of 1000\,\unit{sm^{-1}}.
-The computation of roughness length $z_0$ over ocean is divided into \emph{calm} and a \emph{rough} sea, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
-For calm sea, we apply the limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
+The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
+For calm sea, we apply the following limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
 \begin{equation}
@@
 \begin{equation}
-  \nu = \frac{\mu}{\rho} = \frac{\mu(T)}{\frac{P_0}{T_\text{2M}\cdot R_\text{air}}},
+  \nu = \frac{\mu}{\rho} = \frac{\mu(T)}{\frac{P_0}{T\cdot R_\text{air}}},
 \end{equation}
-with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}} \cdot T_\text{2M}$ and substituting the air density $\rho$ using the ideal gas law.
-$P_0$ is the surface pressure, $T_\text{2M}$ the 2\,\unit{m} temperature and $R_\text{air}$ the universal gas constant for air.
+with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}} \cdot T$ and substituting the air density $\rho$ using the ideal gas law.
+$P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ the universal gas constant for air.
 The rough sea case follows the method of \citet{QJRMS:Charnock1955,JPO:Wu1980}:
@@
 
-{\bf TODO: Mapping of pft land surface types from either ISLSCP2 MODIS or CML2 categories to 9 classical biosphere types used in the EMEP scheme.
-EMEP categories: Forests, Mediterranean scrub, Crops, Moorland (savanna++), Grassland, Wetlands, Tundra, Desert, Water, Urban.
-Weighting of conductance with land surface fraction and sum.
-Inverse of this is gridbox average resistance.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 -> \citep{ICP:MappingManual2017}
-{\bf TODO: Write about latitude dependent vegetation height and EMEP algorithm in general.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-Extending the EMEP scheme to the southern hemisphere, the growing season for crops is defined as:
-\begin{itemize}
-\item[] NH: I. day 90--140; II. day 141--270
-\item[] SH: I. day 272--322; II. day 323--452 (day 87)
-\end{itemize}
+Extending the EMEP scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
+\begin{table}[t]
+  \caption{Definition of growing season for crops used in the OsloCTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
+  \begin{tabular}{lrr}
+    \tophline
+    & I. part & II. part \\
+    & (days) & (days) \\
+    \middlehline
+    NH & 90--140 & 141--270 \\
+    SH & 272--322 & 323--452 \\
+    \bottomhline
+  \end{tabular}
+  %\belowtable{} % Table Footnotes
+  \label{tab:growing_season}
+\end{table}
+%
 
@@
 \begin{equation}
-  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N)}{u_*},
+  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
 \end{equation}
-where $h(N)$ is the canopy height and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
-According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.
-$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow $f_\text{snow}$:
+where $h(N, \text{lat})$ is the latitutde dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
+According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
+$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
 \begin{equation}
@@
 \end{equation}
-$f_\text{snow}$ ranges between $1-2$.
+As initially mentioned, the neccessary depth of snow to cover a certain type of vegetation, differs.
+Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.\\
+
+The total conductance $G_\text{ns}^\chem{O_3}$ in each gridbox is the sum of all $G_\text{ns}^\chem{O_3}(N)$ weighted by the land surface type fraction $f_L(N)$
+\begin{equation}
+  G_\text{ns}^\chem{O_3} = \sum_{N=0}^{N_\text{max}}{G_\text{ns}^\chem{O_3}(N) \cdot f_L(N)}.
+\end{equation}
+
+For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
+Since the dry deposition rate for \chem{CO} is very small, we use a fixed dry deposition velocity $v_d^\chem{CO}$ of 0.03\,\unit{cms^{-1}}.
+
+The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ N}.
+To adapt this to a global model, we need to make a few additional assumptions.
+The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes (40--60\,\unit{^\circ}).
+Towards the pol, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
+At a latitude of 74\,\unit{^\circ} a minimum height of $3/10 \cdot h(N)$ is reached and kept constant.
+Towards the equator, we increase the height linearly so that at a latitude of 10\,\unit{^\circ} a maximum height of $2 \cdot h(N)$ is achieved which is then held constant.
+We also assume symmetry in both hemispheres.
+Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not irrealistic.
+For four example PFTs, results are shown in the Supplement (Fig.~S.1).
+Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
+For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+From MODIS category 'Barren or sparsely vegetated', everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.
+%
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{pictures/pft_landuse_mapping.pdf}
+  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM) 2 to the nine land surface types used in the EMAP scheme.
+Water bodies and snow and ice categories of MODIS are actually not mapped.
+For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+From MODIS category 'Barren or sparsely vegetated', everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.}
+  \label{fig:pft_mapping}
+\end{figure*}
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsec:pre-pro}
-      {\bf TODO: The necessary pre-processing routines...}
+As mentioned in the previous section, there are two variables needed for computing the stomatal conductance which are not directly available from the meteorological input data.
+The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the varying amount of light plants can use to photosynthesize.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsubsection:greening}
+The EMEP scheme is making extensive use of the time elapsed since the onset of the greening season (GDAY) and the total length of the greening season (GLEN).
+There is a latitude dependent parametrization for the greening season that may work well for all but high northern latitudes, especially in Europe, and is not taking meteorology into account.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Photosynthetic photon flux density (PPFD)}
+From OpenIFS an accumulated photoactive radiation (PAR) is available, but needs to be deaccumulated.
+Strictly speaking, since this field is already integrated over the whole spectrum, it is not the actual PAR but PPFD.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 \label{sec:eval}
+For meteorological input, we use \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}.
+CEDS historical emission inventory is used for emissions.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-16 DryDep_supplement.tex
Finished non-stomatal resistance part and began to work on preprocessing, still need to write about stomatal resistance, though - Added supplement figure - made pft_landuse_mapping look more fancy - updated latex_setup files.

\documentclass[manuscript]{copernicus}
\graphicspath{{pictures/}}          % graphics
\begin{document}
\clearpage
\setcounter{page}{1}

\section*{Supplement}
\subsection*{S.1 Latitude dependent vegetation height}
\begin{center}
  \includegraphics[height=0.45\textheight]{pictures/tree_height.pdf}
\end{center}

\end{document}

### 2018-10-17 DryDep.tex
Finished first version of Others and aerosols. Inserted equations for stomatal conductance computation. No spell-check done. Update of bibliography.

@@
 In the following, we will give a detailed account of the new dry deposition scheme.
-Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summerize them, non the less.
+Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, non the less.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
-In addition to the gaseous species, some of the aerosol deposition rates are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
+In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
 As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which we will recap in the following.
@@
 \end{equation}
-with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg m^{-1} s^{-1} K^{-1}} \cdot T$ and substituting the air density $\rho$ using the ideal gas law.
+with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{mol \, m^{-2} s^{-1} K^{-1}} \cdot T$ \citep{PBL} ($\{T \in \mathbb{R} | (233 < T < 303)\,\unit{K}\}$) and substituting the air density $\rho$ using the ideal gas law.
 $P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ the universal gas constant for air.
@@
 
+\begin{equation}
+  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{(f_\text{min}, f_\text{T} \cdot f_\text{D} \cdot f_\text{SW})}
+\end{equation}
+
+\begin{equation}
+  f_\text{T} = \frac{T_\text{2m}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{2m}}{T_{\text{max}}-T_\text{opt}}\right)^\beta, 
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$
+
+\begin{equation}
+  f_\text{D} = f_{\text{min}}+(1-f_{\text{min}}) \frac{D_{\text{min}} - D}{D_{\text{min}}-D_{\text{max}}}
+\end{equation}
+
+\begin{equation}
+  f_\text{SW} =
+  \begin{cases}
+    1 & \quad \text{if  SWVL1} \geq 0.5\\
+    2 \cdot \text{SWVL1}  & \quad \text{if SWVL1} < 0.5
+  \end{cases}
+\end{equation}
+
+\begin{equation}
+  f_\text{phen} =
+  \begin{cases}
+    0  & \quad \text{if  not growing season} \\
+    \max{(\phi_a, \phi_c)}  & \quad \text{if growing season} 
+  \end{cases}
+\end{equation}
+
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+\end{equation}
+
+\begin{equation}
+  g^\text{m}_{\text{max}} = g_{\text{max}} \cdot R \cdot \frac{T_\text{surf}}{P_\text{surf}}
+\end{equation} 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ ranging between $1-2$
+Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ and $\{F_T \in \mathbb{R} | (1 \leq F_T \leq 2)\}$
 \begin{equation}
-  F_T = \exp{(-2\cdot(1+\theta_0))}\text{, for }1 \leq F_T \leq 2.
+  F_T = \exp{(-2\cdot(1+\theta_0))}.
 \end{equation}
@@
 \end{equation}
-where $h(N, \text{lat})$ is the latitutde dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
+where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
 According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
@@
 \end{equation}
-As initially mentioned, the neccessary depth of snow to cover a certain type of vegetation, differs.
+As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs.
 Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.\\
@@
 
-For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
-Since the dry deposition rate for \chem{CO} is very small, we use a fixed dry deposition velocity $v_d^\chem{CO}$ of 0.03\,\unit{cms^{-1}}.
-
+\subsubsection*{Latitude dependent vegetation height}
 The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ N}.
@@
 We also assume symmetry in both hemispheres.
-Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not irrealistic.
+Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic.
 For four example PFTs, results are shown in the Supplement (Fig.~S.1).
@@
 From MODIS category 'Barren or sparsely vegetated', everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.
+
+\subsection{Dry deposition scheme for other gases and aerosols}
+For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
+Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{cms^{-1}}.
+
+The EMEP scheme for aerosols (BC, OC, \chem{SO_4}, \chem{MSA}, SOA) defines the surface deposition velocity $V_\text{ds}$ as
+\begin{equation}
+  \frac{V_\text{ds}}{u_*} =
+  \begin{cases}
+    a_1 &  L \ge 0\,\unit{m}, \\
+    a_1 \cdot F_N \left[ 1 + \left(\frac{a_2}{25}\right)^{2/3}\right]& -25\,\unit{m} < L < 0\,\unit{m}, \\
+    a_1 \cdot F_N \left[ 1 + \left(\frac{-a_2}{L}\right)^{2/3}\right]& L \le -25\,\unit{m}, \\
+  \end{cases}
+\end{equation}
+wherein $F_N = 3$ for fine nitrate and ammonium and $F_N = 1$ for all other species and $L$ is again the Obukhov length.
+$a_2 = 300\,\unit{m}$, while $a_1$ differes for forest and non-forest.
+To account for hydrophilic and hydrophobic BC/OC aerosols on dry and wet surfaces, we diverge slightly from the EMAP scheme in the definition of the parameter $a_1$.
+From pre-studies with an aerosol microphysic model, we find
+\begin{equation}
+  a_1^\text{hydrophob.} = 0.025\,\unit{cms^{-1}} \cdot \overline{u_*}
+\end{equation}
+and
+\begin{equation} 
+  a_1^\text{hydrophil.} = 0.2\,\unit{cms^{-1}} \cdot \overline{u_*},
+\end{equation}
+with the annual mean friction velocity $\overline{u_*}$. Our definition of $a_1$ is then:
+\begin{equation}
+  a_1 =
+  \begin{cases}
+    \emph{land} &
+    \begin{cases}
+      \emph{non-forest} &  a_1^\text{hydrophob.}\\
+      \emph{forest} &
+      \begin{cases}
+        \emph{dry} & \max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
+        \emph{wet} & 1.5\cdot\max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
+    \end{cases}\\
+    \end{cases}\\
+    \emph{ocean} &
+    \begin{cases}
+      a_1^\text{hydrophil.}\\
+      a_1^\text{hydrophob.}\\
+    \end{cases}\\
+    \emph{ice} &
+    \begin{cases}
+      \theta \le 0\,\unit{^\circ C} & a_1^\text{hydrophob.}\\
+      \theta > 0\,\unit{^\circ C} &
+      \begin{cases}
+        a_1^\text{hydrophil.}\\
+        a_1^\text{hydrophob.}\\
+      \end{cases}\\
+    \end{cases}\\
+  \end{cases}
+  \label{eq:dd_aerosol_a1}
+\end{equation}
+%
+The total aerosol surface deposition is then
+\begin{equation}
+ V_\text{ds}^\text{tot} = \sum_{N=0}^{N_\text{max}} V_\text{ds}(N) \cdot f_L(N).
+\end{equation}
 %
@@
 The EMEP scheme is making extensive use of the time elapsed since the onset of the greening season (GDAY) and the total length of the greening season (GLEN).
-There is a latitude dependent parametrization for the greening season that may work well for all but high northern latitudes, especially in Europe, and is not taking meteorology into account.
+There is a latitude dependent parameterization for the greening season that may work well for all but high northern latitudes, especially in Europe, and is not taking meteorology into account.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-18 DryDep.tex
Updated tree_heights.pdf. Spell checked text. Not much more progress, since I have to figure out impact of viscosity parameterization...

@@
 \end{equation}
-with a temperature dependent absolute viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{mol \, m^{-2} s^{-1} K^{-1}} \cdot T$ \citep{PBL} ($\{T \in \mathbb{R} | (233 < T < 303)\,\unit{K}\}$) and substituting the air density $\rho$ using the ideal gas law.
+with a temperature dependent dynamic viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$ \citep{PBL} ($\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$) and substituting the air density $\rho$ using the ideal gas law.
 $P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ the universal gas constant for air.
@@
 wherein $F_N = 3$ for fine nitrate and ammonium and $F_N = 1$ for all other species and $L$ is again the Obukhov length.
-$a_2 = 300\,\unit{m}$, while $a_1$ differes for forest and non-forest.
+$a_2 = 300\,\unit{m}$, while $a_1$ differs for forest and non-forest.
 To account for hydrophilic and hydrophobic BC/OC aerosols on dry and wet surfaces, we diverge slightly from the EMAP scheme in the definition of the parameter $a_1$.

### 2018-10-19 DryDep.tex
Little progress in terms of writing. Checked the dynamical viscosity.

@@
 \end{equation}
-As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields.
+As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are computed directly from the meteorological input fields.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{equation}
-  \nu = \frac{\mu}{\rho} = \frac{\mu(T)}{\frac{P_0}{T\cdot R_\text{air}}},
+  \nu = \frac{\mu}{\rho} = \frac{\mu(T)}{\frac{P_0}{T\cdot R_\text{air}}}.
+  \label{eq:kinematic_viscosity}
 \end{equation}
-with a temperature dependent dynamic viscosity of air $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$ \citep{PBL} ($\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$) and substituting the air density $\rho$ using the ideal gas law.
-$P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ the universal gas constant for air.
+For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Southerland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$.
+In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
+$P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ is the universal gas constant for air.
 The rough sea case follows the method of \citet{QJRMS:Charnock1955,JPO:Wu1980}:
@@
 The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
-Since the $z_0$ computed with this parameterization is very small, the resulting $R^i_b$ is often defined by its imposed lower limit of 10\,\unit{sm^{-1}}.
+Since the $z_0$ computed with this parameterization is rather small ($0 < z_0 < 1\cdot 10^{-4}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{sm^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
+For this reason, the choice of $\mu(T)$ has no large effect on $<R_b>$ (1.4\,\unit{\%}).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (Fig.~S.1).
+For four example PFTs, results are shown in the Supplement (Fig.~S.2).
+
+\subsubsection*{Mapping of land surface types}
 Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).

### 2018-10-22 DryDep.tex
Finally reached the conclusion about the dynamic viscosity. Included some supplementory material and started formulationg the stomatal conductance section.

@@
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
-Dry deposition velocities $v^i_d$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$ are typically calculated following a resistance-analogous approach
+Dry deposition velocities $v^i_d$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
 \begin{equation}
@@
 \end{equation}
-wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989}.
-For all gases $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
+wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989,ACP:Seinfeld2006}.
+For all gases, $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.).
 Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type.
@@
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$]{ACP:Helmig2007}.
-Canopy conductance is parameterized at the single-leave level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
+Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
 But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
@@
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
-For this reason, we diverge from the EMEP scheme at this point and fall back to the method of \citet{Monteith1973}.
-The sensible heat flux $\Phi_{Q_\text{sensible}}$ can be written as
+For this reason, we diverge from the EMEP scheme at this point and fall back to the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
 \begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z R_a},
+  \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z R_a}.
   \label{eq:sens_heat}
 \end{equation}
-wherein $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$.
-$\partial_z R_a$ is the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$.
-$\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
+Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z R_a$ is the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$, and $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
 From the eddy diffusion theorem, a similar formulation can be derived
@@
 with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$).
-The ratios $D_\chem{H_2O}/D_i$ are given in Table~\ref{tab:diffusivity}.
+The used ratios $D_\chem{H_2O}/D_i$ are displayed in Table~\ref{tab:diffusivity}.
 %
@@
 The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
-For calm sea, we apply the following limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
+For calm sea, we apply the following upper limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
 \begin{equation}
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Southerland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$.
+But despite its rough acuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$.
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
-Since the $z_0$ computed with this parameterization is rather small ($0 < z_0 < 1\cdot 10^{-4}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{sm^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
-For this reason, the choice of $\mu(T)$ has no large effect on $<R_b>$ (1.4\,\unit{\%}).
+Since the $z_0$ computed with this parameterization is rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{sm^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection*{Stomatal conductance}
--> \citep{ICP:MappingManual2017}
-
+The stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
+%
 \begin{equation}
-  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{(f_\text{min}, f_\text{T} \cdot f_\text{D} \cdot f_\text{SW})}
+  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{(f_\text{min}, f_\text{T} \cdot f_\text{D} \cdot f_\text{SW})}.
 \end{equation}
-
+%
+The factors herein vary within the range $0-1$ and account for leaf phenology ($f_\text{phen}$), photoactive radiation ($f_\text{light}$), temperature ($f_\text{T}$), water vapor pressure deficit ($f_\text{D}$), and soil water content ($f_\text{SW}$).
+The temperature ajustment is computed from
+%
 \begin{equation}
@@
 \end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$
-
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$.
+The parameters $T_\text{min}$ and $T_\text{opt}$ are tabulated for various species of plants and can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
+%
 \begin{equation}

### 2018-10-22 DryDep_supplement.tex
Little progress in terms of writing. Checked the dynamical viscosity.
Finally reached the conclusion about the dynamic viscosity. Included some supplementory material and started formulationg the stomatal conductance section.

@@
 \section*{Supplement}
-\subsection*{S.1 Latitude dependent vegetation height}
+\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
+\subsubsection*{S.1.1 Sutherland's law and linear fits}
+\appendixfigures
 \begin{center}
-  \includegraphics[height=0.45\textheight]{pictures/tree_height.pdf}
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
+\end{center}
+\subsubsection*{S.1.2 Comparison of $z_o$ and $R_b$ for linear fit through origin and Sutherland's law}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
+\end{center}
+\subsubsection*{S.1.3 $R_b$ for different species}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_1.pdf}
+  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_2.pdf}
+\end{center}
+
+\subsection*{S.2 Tabulated parameters for stomatal conductance}
+\appendixtables
+\begin{sidewaystable}[t]
+  \begin{tabular}{lccccccccccccccccc}
+    \tophline
+    Code & gmax & fmin & phia & phib & phic & phid & phie & phif & phiAS & phiAE & flight & Tmin & Topt & Tmax & Dmax & Dmin & Dcrit \\
+    \middlehline
+    CF & 140 & 0.1 & 0.8 & 0.8 & 0.8 & 0.8 & 1 & 1 & 0 & 0 & 0.006 & 0 & 18 & 36 & 0.5 & 3 & 1000 \\
+    DF & 150 & 0.1 & 0 & 0 & 1 & 0 & 20 & 30 & 0 & 0 & 0.006 & 0 & 20 & 35 & 1 & 3.25 & 1000 \\
+    NF & 200 & 0.1 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.013 & 8 & 25 & 38 & 1 & 3.2 & 1000 \\
+    BF & 200 & 0.02 & 1 & 1 & 0.3 & 1 & 130 & 60 & 80 & 35 & 0.009 & 1 & 23 & 39 & 2.2 & 4 & 1000 \\
+    TC & 300 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0105 & 0.01 & 12 & 26 & 40 & 1.2 & 3.2 & 8 \\
+    MC & 300 & 0.019 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0048 & 0 & 25 & 51 & 1 & 2.5 & 1000 \\
+    RC & 360 & 0.02 & 0.2 & 0.2 & 1 & 0.2 & 20 & 45 & 0 & 0 & 0.0023 & 8 & 24 & 50 & 0.31 & 2.7 & 10 \\
+    SNL & 60 & 0.01 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0.009 & 1 & 18 & 36 & 1.3 & 3 & 1000 \\
+    GR & 270 & 0.01 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0.009 & 12 & 26 & 40 & 1.3 & 3 & 1000 \\
+    MS & 200 & 0.01 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.012 & 4 & 20 & 37 & 1.3 & 3.2 & 1000 \\
+    WE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    TU & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    DE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    W & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    ICE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    U & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 & 1 & 0 & 1000 \\
+    \bottomhline
+  \end{tabular}
+  \vspace{2\baselineskip}
+  \begin{tabular}{lcccccccccccc}
+    \tophline
+    Code & RSO & RO3 & height & dSGS & dEGS & ddSGS & ddEGS & LAImin & LAImax & LS & LE \\
+    \middlehline
+    CF & 0 & 200 & 20 & 0 & 366 & 0 & 0 & 5 & 5 & 1 & 1\\
+    DF & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 & 0 & 4 & 20 & 30\\
+    NF & 0 & 200 & 8 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
+    BF & 0 & 200 & 15 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
+    TC & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57 & 0 & 3.5 & 70 & 20\\
+    MC & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57 & 0 & 3 & 70 & 44\\
+    RC & 0 & 200 & 1 & 130 & 250 & 0 & 0 & 0 & 4.2 & 35 & 65\\
+    SNL & 0 & 400 & 0.5 & 0 & 366 & 0 & 0 & 2 & 3 & 192 & 96\\
+    GR & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0 & 2 & 3.5 & 140 & 135\\
+    MS & 0 & 200 & 2 & 0 & 366 & 0 & 0 & 2.5 & 2.5 & 1 & 1\\
+    WE & 50 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    TU & 500 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    DE & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    W 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    ICE & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    U & 400 & 400 & 10 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    \bottomhline
+  \end{tabular}
+\end{sidewaystable}
+\subsection*{S.3 Latitude dependent vegetation height}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
 \end{center}

### 2018-10-23 DryDep.tex
Continued writing on section 2, stomatal conductance. Added plots to vis PAR de-accumulation and phenology. Updated supplement accordingly. TODO: Finish phenology scetch.

@@
 A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
+All relevant information and papers can also be found on the \href{http://folk.uio.no/mariantl/osloctm3/}{OsloCTM3 webpage}.
 The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN) \citep{ACP:Guenther2006}.
@@
 In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account.
-In the following, we will refer to this parameterization as \emph{Wesely scheme}.\\
+We will refer to this parameterization as \emph{Wesely scheme}.\\
 
@@
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
-In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}), and secondary organic aerosols (SOA).
-As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$) which we will recap in the following.
+In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA).
+As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different resistance contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy $R^i_c$) which we will recap in the following.
 
@@
 \label{subsubsec:Ra}
-The aerodynamic resistance $R_a$ in \citet{WASP:Simpson2003,ACP:Simpson2012} is described as
+The aerodynamic resistance in \citet{WASP:Simpson2003,ACP:Simpson2012} is described as
 \begin{equation}
@@
 \end{equation}
-with $\kappa$ the K\'{a}rm\'{a}n constant (about $0.40$), $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
+with $\kappa = 0.40$ the K\'{a}rm\'{a}n constant, $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
@@
 \label{subsubsec:Rc}
-The surface resistance $R_c^i$ consists of both, stomatal and non-stomatal resistances. When computing gridbox averages, we use the sum of conductances (deposition velocities)
+The surface resistance consists of both, stomatal and non-stomatal resistances. When computing gridbox averages, we use the sum of conductances (deposition velocities)
 \begin{equation}
@@
 \subsubsection*{Stomatal conductance}
-The stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
+The stomatal conductance is a measure of the rate of \chem{CO_2} entering or evapotranspiration through the stomata of a leaf.
+There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respirationn (e.g., light, available water, etc.).
+Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}.
+To reflect part of the underlying mechanism, the stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
 %
@@
 %
-The factors herein vary within the range $0-1$ and account for leaf phenology ($f_\text{phen}$), photoactive radiation ($f_\text{light}$), temperature ($f_\text{T}$), water vapor pressure deficit ($f_\text{D}$), and soil water content ($f_\text{SW}$).
-The temperature ajustment is computed from
+The factors herein are normalized and vary within the range $0-1$.
+They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\text{T}$), water vapor pressure deficit ($f_\text{D}$), and soil water content ($f_\text{SW}$).
+The temperature adjustment is computed from
 %
@@
 with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$.
-The parameters $T_\text{min}$ and $T_\text{opt}$ are tabulated for various species of plants and can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
+The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various landsurface types and can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
+Since $f_\text{T}$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+The water vapor deficite (VPD) is proportional to the satuation partial pressure of water ($P_\chem{H_2O}^s$) and relative humidity ($\text{RH}$)
 %
 \begin{equation}
-  f_\text{D} = f_{\text{min}}+(1-f_{\text{min}}) \frac{D_{\text{min}} - D}{D_{\text{min}}-D_{\text{max}}}
+  \text{VPD} = P_\chem{H_2O}^s \cdot (1 - \text{RH}/100).
 \end{equation}
-
+%
+Using tabulated values of $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ (Supplement \citet{ACP:Simpson2012} or Supplement~S.2), the water vapor pressure deficit penalty factor can be computed:
+%
+\begin{equation}
+  f_\text{D} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}.
+\end{equation}
+%
+The penalty factor with respect to available water in soil $\text{SW}$ is defined as 
+%
 \begin{equation}
@@
   \begin{cases}
-    1 & \quad \text{if  SWVL1} \geq 0.5\\
-    2 \cdot \text{SWVL1}  & \quad \text{if SWVL1} < 0.5
+    1 & \quad \text{SW} \geq 0.5,\\
+    2 \cdot \text{SW}  & \quad \text{SW} < 0.5.
   \end{cases}
 \end{equation}
-
+\citet{ACP:Simpson2012} suggest using SW in soil depths below 1\,\unit{m} (SWVL4 in OpenIFS).
+Throughout most of our simulations (Section~\ref{sec:eval}), we have instead used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance much more sensitive to precipitation than it would be otherwise.
+The phenology of a plant typically describes its lifecycle though out a year, e.g., in mid latitudes and for decideous species, it starts with the emergence of leafs in spring and ends in fall.
+In the EMEP scheme phenology is parametrized with respect to the begin of the greening season (GDAY) and its length (GLEN).
+Details about our choice of GDAY and GLEN are given in Section~\ref{subsubsec:greening}.
+In summary, the $f_\text{phen}$ parameterization reads as follows:
+%
 \begin{equation}
@@
   \begin{cases}
-    0  & \quad \text{if  not growing season} \\
-    \max{(\phi_a, \phi_c)}  & \quad \text{if growing season} 
+    \text{GDAY = 0}  & \quad 0 \\
+    \text{(GLEN $\ge$ 365} & \quad 1 \quad \emph{explicitly excluding Tropics}) \\
+    \emph{else}  & \quad 
+    \begin{cases}
+      \text{GDAY} \le \phi_\text{AS}  & \quad \phi_a \\
+      \text{GDAY} \le \phi_\text{AS} + \phi_e  & \quad \phi_b+(\phi_c-\phi_b)\cdot(\text{GDAY}-\phi_\text{AS})/\phi_e\\
+      \text{GDAY} \le \text{GLEN} - \phi_\text{AE} - \phi_f  & \quad \phi_c\\
+      \text{GDAY} \le \text{GLEN} - \phi_\text{AE}  & \quad \phi_d+(\phi_c-\phi_d)\cdot(\text{GLEN}-\phi_\text{AE}-\text{GDAY})/\phi_f\\
+      \emph{else} & \quad \phi_d
+    \end{cases}
   \end{cases}
 \end{equation}
-
+%
+The parameters $\phi_a$, $\phi_b$, $\phi_c$, $\phi_d$, $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ can as well be found in Supplement \citet{ACP:Simpson2012} or Supplement~S.2.
+If GDAY is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}$.
+Since the EMEP scheme is tuned to NH mid latitudes, the phenology does not apply to the tropics, hence, we later decided to set $f_\text{phen}$ to one if GLEN is greater or equal 365 which may be the case in the tropics and subtropics.
+The five phases described be the third case may dependent on surface type take a saddle or an trapezoidal shape (Fig.~\ref{fig:phenology}).
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/phenology.pdf}
+  \caption{Scatch of the five different phases described in $f_\text{phen}$.}
+  \label{fig:phenology}
+\end{figure}
+%
 \begin{equation}
@@
 \end{equation}
-
+%
 \begin{equation}
@@
     \text{LAI}+1 & \text{\emph{forest / wetland}},\\
-    \text{LAI}\cdot 5/3.5 & \text{\emph{cropland, I. part of growing season}},\\
-    \text{LAI} + 1.5 & \text{\emph{cropland, II. part of growing season}},\\
-    0 & \text{\emph{cropland, winter}}.
+    \text{LAI}\cdot 5/3.5 & \text{\emph{cropland}, I. part of growing season},\\
+    \text{LAI} + 1.5 & \text{\emph{cropland}, II. part of growing season},\\
+    0 & \text{\emph{cropland}, winter}.
   \end{cases}
@@
 \subsubsection{Greening season}
-\label{subsubsection:greening}
-The EMEP scheme is making extensive use of the time elapsed since the onset of the greening season (GDAY) and the total length of the greening season (GLEN).
+\label{subsubsec:greening}
+The EMEP scheme is making extensive use of the time elapsed since the onset of the greening season and the total length of the greening season.
 There is a latitude dependent parameterization for the greening season that may work well for all but high northern latitudes, especially in Europe, and is not taking meteorology into account.

### 2018-10-23 DryDep_supplement.tex
Continued writing on section 2, stomatal conductance. Added plots to vis PAR de-accumulation and phenology. Updated supplement accordingly. TODO: Finish phenology scetch.

--- 2018-10-16 DryDep_supplement.tex
@@
 \section*{Supplement}
-\subsection*{S.1 Latitude dependent vegetation height}
+\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
+\subsubsection*{S.1.1 Sutherland's law and linear fits}
+\appendixfigures
 \begin{center}
-  \includegraphics[height=0.45\textheight]{pictures/tree_height.pdf}
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
+\end{center}
+\subsubsection*{S.1.2 Comparison of $z_o$ and $R_b$ for $\mu(T) = m\cdot T$ and Sutherland's law: \chem{H_2O}}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
+\end{center}
+\subsubsection*{S.1.3 $R_b$ for different species}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_1.pdf}
+  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_2.pdf}
+\end{center}
+
+%\begin{minipage}[t]{1.\textwidth}
+\subsection*{S.2 Tabulated parameters for stomatal conductance}
+\appendixtables
+\begin{sidewaystable}[!htbp]
+  %\caption{S.2 Tabulated parameters for stomatal conductance}
+  \begin{tabular}{lcccccccccccccc}
+    \tophline
+    Code & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $f_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
+    &&&&&&&&&&&&&&\\
+    \middlehline
+    CF & 140 & 0.1 & 0.8 & 0.8 & 0.8 & 0.8 & 1 & 1 & 0 & 0 & 0.006 & 0 & 18 & 36  \\
+    DF & 150 & 0.1 & 0 & 0 & 1 & 0 & 20 & 30 & 0 & 0 & 0.006 & 0 & 20 & 35  \\
+    NF & 200 & 0.1 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.013 & 8 & 25 & 38 \\
+    BF & 200 & 0.02 & 1 & 1 & 0.3 & 1 & 130 & 60 & 80 & 35 & 0.009 & 1 & 23 & 39 \\
+    TC & 300 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0105 & 0.01 & 12 & 26 & 40 \\
+    MC & 300 & 0.019 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0048 & 0 & 25 & 51  \\
+    RC & 360 & 0.02 & 0.2 & 0.2 & 1 & 0.2 & 20 & 45 & 0 & 0 & 0.0023 & 8 & 24 & 50  \\
+    SNL & 60 & 0.01 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0.009 & 1 & 18 & 36 \\
+    GR & 270 & 0.01 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0.009 & 12 & 26 & 40 \\
+    MS & 200 & 0.01 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.012 & 4 & 20 & 37 \\
+    WE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+    TU & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
+    DE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
+    W & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+    ICE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+    U & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+    \bottomhline
+  \end{tabular}
+  \\
+  \vspace{2\baselineskip}
+  \begin{tabular}{lcccccccccccccc}
+    \tophline
+    Code & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & height & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$ & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
+    &&&&&&&&&&&&&& \\
+    \middlehline
+    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0 & 5 & 5 & 1 & 1\\
+    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 & 0 & 4 & 20 & 30\\
+    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
+    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
+    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57 & 0 & 3.5 & 70 & 20\\
+    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57 & 0 & 3 & 70 & 44\\
+    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0 & 0 & 4.2 & 35 & 65\\
+    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0 & 2 & 3 & 192 & 96\\
+    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0 & 2 & 3.5 & 140 & 135\\
+    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0 & 2.5 & 2.5 & 1 & 1\\
+    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    \bottomhline
+  \end{tabular}
+\end{sidewaystable}
+%\end{minipage}
+
+\subsection*{S.3 Latitude dependent vegetation height}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
+\end{center}
+
+\subsection*{S.4 De-accumulation of photoactive radiation (PAR) from OpenIFS: January 1st 2005}
+\subsubsection*{S.4.1 Output from OpenIFS: Accumulated PAR}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/PAR_as_is.png}
+\end{center}
+\subsubsection*{S.4.2 Partly de-accumulated}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/PAR_partly_deaccumulated.png}
+\end{center}
+\subsubsection*{S.4.3 De-accumulation of hour 00}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/PAR_deaccumulating_00h.png}
 \end{center}

### 2018-10-24 DryDep.tex
Finished describtion of dry deposition. Updated plot and added two more for supplement.

@@
 \label{subsubsec:Ra}
-The aerodynamic resistance in \citet{WASP:Simpson2003,ACP:Simpson2012} is described as
+The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. In \citet{WASP:Simpson2003,ACP:Simpson2012} it is described as
 \begin{equation}
@@
 \subsubsection*{Land}
-Over land we use \citep[Eq.~(53),][]{ACP:Simpson2012}
+Over land, we use \citep[Eq.~(53),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \subsubsection*{Ocean}
-Over ocean we use \citep[Eq.~(54),][]{ACP:Simpson2012}
+Over ocean, we use \citep[Eq.~(54),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \end{equation}
-For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Southerland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$.
-But despite its rough acuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$.
+For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$.
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$ (Supplement S.1).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 \begin{equation}
-  G_c^i = \text{LAI} \cdot G_\text{sto}^i + G_\text{ns}^i, 
+  G_c = \text{LAI} \cdot G_\text{sto} + G_\text{ns}, 
 \end{equation}
-wherein $\text{LAI}$ is the leaf area index (zero for non-vegetated surfaces), $G_\text{sto}^i$ the stomatal conductance, and $G_\text{ns}^i$ the non-stomatal conductance.
+wherein $\text{LAI}$ is the leaf area index (zero for non-vegetated surfaces), $G_\text{sto}$ the stomatal conductance, and $G_\text{ns}$ the non-stomatal conductance.
 
@@
 The stomatal conductance is a measure of the rate of \chem{CO_2} entering or evapotranspiration through the stomata of a leaf.
-There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respirationn (e.g., light, available water, etc.).
+There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.).
 Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}.
@@
   g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{(f_\text{min}, f_\text{T} \cdot f_\text{D} \cdot f_\text{SW})}.
+  \label{eq:stomatal}
 \end{equation}
@@
 with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$.
-The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various landsurface types and can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
+The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various land surface types.
+All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
 Since $f_\text{T}$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
-The water vapor deficite (VPD) is proportional to the satuation partial pressure of water ($P_\chem{H_2O}^s$) and relative humidity ($\text{RH}$)
+The water vapor deficit (VPD) is proportional to the saturation partial pressure of water ($P_\chem{H_2O}^s$) and relative humidity ($\text{RH}$)
 %
@@
 %
-Using tabulated values of $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ (Supplement \citet{ACP:Simpson2012} or Supplement~S.2), the water vapor pressure deficit penalty factor can be computed:
+Using tabulated values of $f_\text{min}$, $D_\text{min}$, $D_\text{max}$, the water vapor pressure deficit penalty factor can be computed:
 %
@@
 %
-The penalty factor with respect to available water in soil $\text{SW}$ is defined as 
+The penalty factor with respect to available soil water ($\text{SW}$) is defined as 
 %
@@
   \begin{cases}
-    1 & \quad \text{SW} \geq 0.5,\\
-    2 \cdot \text{SW}  & \quad \text{SW} < 0.5.
+    1 & \quad \emph{if} \quad \text{SW} \geq 0.5,\\
+    2 \cdot \text{SW}  & \quad \emph{if} \quad \text{SW} < 0.5.
   \end{cases}
@@
 \citet{ACP:Simpson2012} suggest using SW in soil depths below 1\,\unit{m} (SWVL4 in OpenIFS).
-Throughout most of our simulations (Section~\ref{sec:eval}), we have instead used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance much more sensitive to precipitation than it would be otherwise.
-The phenology of a plant typically describes its lifecycle though out a year, e.g., in mid latitudes and for decideous species, it starts with the emergence of leafs in spring and ends in fall.
-In the EMEP scheme phenology is parametrized with respect to the begin of the greening season (GDAY) and its length (GLEN).
+Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance much more sensitive to precipitation than it would be otherwise.
+The phenology of a plant typically describes its lifecycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
+In the EMEP scheme, phenology is parameterized with respect to the begin of the greening season (GDAY) and its length (GLEN).
 Details about our choice of GDAY and GLEN are given in Section~\ref{subsubsec:greening}.
@@
   \begin{cases}
-    \text{GDAY = 0}  & \quad 0 \\
-    \text{(GLEN $\ge$ 365} & \quad 1 \quad \emph{explicitly excluding Tropics}) \\
+    \emph{if} \quad \text{GDAY = 0}  & \quad 0 \\
+    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding Tropics)} \\
     \emph{else}  & \quad 
     \begin{cases}
-      \text{GDAY} \le \phi_\text{AS}  & \quad \phi_a \\
-      \text{GDAY} \le \phi_\text{AS} + \phi_e  & \quad \phi_b+(\phi_c-\phi_b)\cdot(\text{GDAY}-\phi_\text{AS})/\phi_e\\
-      \text{GDAY} \le \text{GLEN} - \phi_\text{AE} - \phi_f  & \quad \phi_c\\
-      \text{GDAY} \le \text{GLEN} - \phi_\text{AE}  & \quad \phi_d+(\phi_c-\phi_d)\cdot(\text{GLEN}-\phi_\text{AE}-\text{GDAY})/\phi_f\\
+      \emph{if} \quad \text{GDAY} \le \phi_\text{AS}  & \quad \phi_a \\
+      \emph{if} \quad \text{GDAY} \le \phi_\text{AS} + \phi_e  & \quad \phi_b+(\phi_c-\phi_b)\cdot(\text{GDAY}-\phi_\text{AS})/\phi_e\\
+      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE} - \phi_f  & \quad \phi_c\\
+      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE}  & \quad \phi_d+(\phi_c-\phi_d)\cdot(\text{GLEN}-\phi_\text{AE}-\text{GDAY})/\phi_f\\
       \emph{else} & \quad \phi_d
@@
 %
-The parameters $\phi_a$, $\phi_b$, $\phi_c$, $\phi_d$, $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ can as well be found in Supplement \citet{ACP:Simpson2012} or Supplement~S.2.
-If GDAY is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}$.
-Since the EMEP scheme is tuned to NH mid latitudes, the phenology does not apply to the tropics, hence, we later decided to set $f_\text{phen}$ to one if GLEN is greater or equal 365 which may be the case in the tropics and subtropics.
-The five phases described be the third case may dependent on surface type take a saddle or an trapezoidal shape (Fig.~\ref{fig:phenology}).
+The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}).
+If GDAY is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
+Since the EMEP scheme is tuned to NH mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal 365 which is the case in the tropics.
 %
@@
   \includegraphics[width=8.3cm]{pictures/phenology.pdf}
-  \caption{Scatch of the five different phases described in $f_\text{phen}$.}
+  \caption{Sketch of the five different phases described in $f_\text{phen}$.}
   \label{fig:phenology}
@@
 %
+Light in the wavelength band $(400-700)$\,\unit{nm} to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
+In the EMEP parameterization, an integrated photon flux -- photosynthetic photon flux density (PPFD) -- is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
 \begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD}).
 \end{equation}
 %
+Since $g_\text{max}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,\,m^{-2}}, a unit conversion is done:
 \begin{equation}
-  g^\text{m}_{\text{max}} = g_{\text{max}} \cdot R \cdot \frac{T_\text{surf}}{P_\text{surf}}
+  g^\text{m}_{\text{sto}}(N) = g_{\text{sto}}(N) \cdot R \cdot \frac{T_0}{P_0}.
+\end{equation}
+Herein $R$ is the universal gas constant. With this the total gridbox stomatal conductance $G_\text{sto}$ is computed weighted by the land surface type fraction $f_L(N)$:
+\begin{equation}
+  G_\text{sto} = \sum_{N=0}^{N_\text{max}} f_L(N)\cdot g^\text{m}_{\text{sto}}(N)
 \end{equation}
@@
   \begin{cases}
-    \text{LAI}+1 & \text{\emph{forest / wetland}},\\
-    \text{LAI}\cdot 5/3.5 & \text{\emph{cropland}, I. part of growing season},\\
-    \text{LAI} + 1.5 & \text{\emph{cropland}, II. part of growing season},\\
-    0 & \text{\emph{cropland}, winter}.
+    \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
+    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, I. part of growing season},\\
+    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, II. part of growing season},\\
+    0 & \emph{if} \quad \text{\emph{cropland}, winter}.
   \end{cases}
@@
 For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
-Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{cms^{-1}}.
+Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{cm\,s^{-1}}.
 
@@
   \begin{cases}
-    a_1 &  L \ge 0\,\unit{m}, \\
-    a_1 \cdot F_N \left[ 1 + \left(\frac{a_2}{25}\right)^{2/3}\right]& -25\,\unit{m} < L < 0\,\unit{m}, \\
-    a_1 \cdot F_N \left[ 1 + \left(\frac{-a_2}{L}\right)^{2/3}\right]& L \le -25\,\unit{m}, \\
+    a_1 &  \emph{if} \quad L \ge 0\,\unit{m}, \\
+    a_1 \cdot F_N \left[ 1 + \left(\frac{a_2}{25}\right)^{2/3}\right] & \emph{if} \quad -25\,\unit{m} < L < 0\,\unit{m}, \\
+    a_1 \cdot F_N \left[ 1 + \left(\frac{-a_2}{L}\right)^{2/3}\right] & \emph{if} \quad L \le -25\,\unit{m}, \\
   \end{cases}
@@
 \begin{equation}
-  a_1^\text{hydrophob.} = 0.025\,\unit{cms^{-1}} \cdot \overline{u_*}
+  a_1^\text{hydrophob.} = 0.025\,\unit{cm\,s^{-1}} \cdot \overline{u_*}
 \end{equation}
@@
 \begin{equation} 
-  a_1^\text{hydrophil.} = 0.2\,\unit{cms^{-1}} \cdot \overline{u_*},
+  a_1^\text{hydrophil.} = 0.2\,\unit{cm\,s^{-1}} \cdot \overline{u_*},
 \end{equation}
@@
   \begin{cases}
-    \emph{land} &
+    \emph{if} \quad \emph{land} &
     \begin{cases}
-      \emph{non-forest} &  a_1^\text{hydrophob.}\\
-      \emph{forest} &
+      \emph{if} \quad \emph{non-forest} &  a_1^\text{hydrophob.}\\
+      \emph{if} \quad \emph{forest} &
       \begin{cases}
-        \emph{dry} & \max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
-        \emph{wet} & 1.5\cdot\max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
+        \emph{if} \quad \emph{dry} & \max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
+        \emph{if} \quad \emph{wet} & 1.5\cdot\max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
     \end{cases}\\
     \end{cases}\\
-    \emph{ocean} &
+    \emph{if} \quad \emph{ocean} &
     \begin{cases}
@@
     \end{cases}\\
-    \emph{ice} &
+    \emph{if} \quad \emph{ice} &
     \begin{cases}
-      \theta \le 0\,\unit{^\circ C} & a_1^\text{hydrophob.}\\
-      \theta > 0\,\unit{^\circ C} &
+      \emph{if} \quad \theta \le 0\,\unit{^\circ C} & a_1^\text{hydrophob.}\\
+      \emph{if} \quad \theta > 0\,\unit{^\circ C} &
       \begin{cases}
@@
 As mentioned in the previous section, there are two variables needed for computing the stomatal conductance which are not directly available from the meteorological input data.
-The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the varying amount of light plants can use to photosynthesize.
+The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the amount of light plants need to photosynthesize.
+In the following, we present the necessary pre-processing of the variables.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsubsec:greening}
-The EMEP scheme is making extensive use of the time elapsed since the onset of the greening season and the total length of the greening season.
-There is a latitude dependent parameterization for the greening season that may work well for all but high northern latitudes, especially in Europe, and is not taking meteorology into account.
+The original EMEP scheme uses fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ together with leapsrates ($d_\text{SGS}$, $d_\text{EGS}$, $\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$ in Supplement~S.2) to define phenology and dry deposition over agricultural areas.
+However this parameterization fails in a global model, therefor another latitude dependent parameterization is needed.
+First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS) model.
+SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses leapsrates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
+This parameterization works well, if the focus of the scientific work lies on mid latitudes and North America.
+In Europe, however, most of northern Scandinavia has no vegetation period according to this parameterization.
+
+In agriculture, there are differing empirical rules to estimate the SGS and EGS.
+The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above 5\,\unit{^\circ C} and vice versa for EGS.
+If, within these 5 days, the temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset.
+With respect to model performance, we decided to preprocess our meteorological input date using this 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH and $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
+In all other cases and where the 5\,\unit{^\circ C}-days criterion fails, we use the SMOKE-BEIS parameterization.
+
+Instead of SGS and EGS, we use the derived parameters day of greening season (GDAY), which is the time elapsed starting at the SGS, and the total length of the greening season (GLEN), which is the time between EGS and SGS.
+Example maps of the computed GLEN are shown in Supplement~S.4.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Photosynthetic photon flux density (PPFD)}
-From OpenIFS an accumulated photoactive radiation (PAR) is available, but needs to be deaccumulated.
+From OpenIFS an accumulated photosynthetic active radiation (PAR) is available, but needs to be deaccumulated.
 Strictly speaking, since this field is already integrated over the whole spectrum, it is not the actual PAR but PPFD.
@@
 Include table of all sensitivity runs.
-\citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm s^{-1}}$,][]{ACP:Helmig2007}
+\citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
 dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)

### 2018-10-24 DryDep_supplement.tex
Finished describtion of dry deposition. Updated plot and added two more for supplement.

@@
     Code & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $f_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
-    &&&&&&&&&&&&&&\\
+    & (\unit{mmol\,s^{-1}\,\,m^{-2}}) &&&&&&&&&&&&&\\
     \middlehline
@@
 
-\subsection*{S.4 De-accumulation of photoactive radiation (PAR) from OpenIFS: January 1st 2005}
-\subsubsection*{S.4.1 Output from OpenIFS: Accumulated PAR}
+\subsection*{S.4 Temperature dependent greening season 2005}
+\subsubsection*{S.4.1 GLEN: NH mid--high latitudes only}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/GLEN_2005.pdf}
+\end{center}
+\subsubsection*{S.4.1 GLEN: NH and SH mid--high latitudes}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight]{pictures/GLEN_2005_she.pdf}
+\end{center}
+
+\subsection*{S.5 De-accumulation of photoactive radiation (PAR) from OpenIFS: January 1st 2005}
+\subsubsection*{S.5.1 Output from OpenIFS: Accumulated PAR}
 \appendixfigures
@@
 \end{center}
-\subsubsection*{S.4.2 Partly de-accumulated}
+\subsubsection*{S.5.2 Partly de-accumulated}
 \appendixfigures
@@
 \end{center}
-\subsubsection*{S.4.3 De-accumulation of hour 00}
+\subsubsection*{S.5.3 De-accumulation of hour 00}
 \appendixfigures

### 2018-10-25 DryDep.tex
Finished Section 2. Needs to be read one more time before tagging and shippment to co-authors.

@@
   \end{cases}
+  \label{eq:fphen}
 \end{equation}
@@
 Light in the wavelength band $(400-700)$\,\unit{nm} to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
-In the EMEP parameterization, an integrated photon flux -- photosynthetic photon flux density (PPFD) -- is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
+In the EMEP parameterization, an integrated photon flux is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
 \begin{equation}
   f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD}).
+  \label{eq:flight}
 \end{equation}
@@
   G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
+  \label{eq:non-stomata-cond}
 \end{equation}
@@
   \end{cases}
+  \label{eq:SAI}
 \end{equation}
@@
 \label{subsubsec:greening}
-The original EMEP scheme uses fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ together with leapsrates ($d_\text{SGS}$, $d_\text{EGS}$, $\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$ in Supplement~S.2) to define phenology and dry deposition over agricultural areas.
-However this parameterization fails in a global model, therefor another latitude dependent parameterization is needed.
-First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS) model.
-SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses leapsrates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
-This parameterization works well, if the focus of the scientific work lies on mid latitudes and North America.
-In Europe, however, most of northern Scandinavia has no vegetation period according to this parameterization.
+In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ together with leaprates ($d_\text{SGS}$, $d_\text{EGS}$, $\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
+For the growing season of crops in the computation of non-stomatal conductances, we use fixed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
+Since the eurocentric EMEP parameterization of SGS and EGS is not applicable in a global model, another latitude dependent parameterization is needed.
+First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).
+SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses leaprates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
+As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.
 
-In agriculture, there are differing empirical rules to estimate the SGS and EGS.
+In agriculture, there are different empirical rules to estimate the SGS and EGS.
 The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above 5\,\unit{^\circ C} and vice versa for EGS.
-If, within these 5 days, the temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset.
-With respect to model performance, we decided to preprocess our meteorological input date using this 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH and $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
-In all other cases and where the 5\,\unit{^\circ C}-days criterion fails, we use the SMOKE-BEIS parameterization.
+Based on this rule of thumb, we have pre-processed our meteorological input data offline.
+If, within these 5 days, the temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively.
+First, we used the 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
+In all other cases and where the 5\,\unit{^\circ C}-days criteria fails, we use the SMOKE-BEIS parameterization.
+The described algorithm written in python 2.7 has been included as Supplement~S.4.1.
+An example map of the computed GLEN using the 5\,\unit{^\circ C}-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005_she.pdf}
+  \caption{Pre-processing of greening season from meteorological surface temperature fields.
+Shown is the total length of the greening season (GLEN) for the year 2005.
+The 5\,\unit{^\circ C}-days criteria has been used in both hemispheres mid--high latitudes.
+Ocean has been shaded to indicate that greening season will only affect land.}
+  \label{fig:glen_2015_she}
+\end{figure}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Photosynthetic photon flux density}
+From OpenIFS an accumulated photosynthetic active radiation is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
+Since this field is already integrated over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
+Therefor, we will refer to the deaccumulated field as PPFD.
+An example PAR for January 2nd 2005 is shown in Supplement~S.5.1.
+Deaccumulation of hour 3 till 21, simply means
+\begin{equation}
+  \text{PPFD}(t_i) = \text{PAR}(t_{i+1})-\text{PAR}(t_i).
+\end{equation}
+If everything was right in the OpenIFS output, we would not need to do any further processing, however the first timestep of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately deaccumulate this time step.
+The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
+\begin{equation}
+  \text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right].
+\end{equation}
 
-Instead of SGS and EGS, we use the derived parameters day of greening season (GDAY), which is the time elapsed starting at the SGS, and the total length of the greening season (GLEN), which is the time between EGS and SGS.
-Example maps of the computed GLEN are shown in Supplement~S.4.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Photosynthetic photon flux density (PPFD)}
-From OpenIFS an accumulated photosynthetic active radiation (PAR) is available, but needs to be deaccumulated.
-Strictly speaking, since this field is already integrated over the whole spectrum, it is not the actual PAR but PPFD.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
   \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
-  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface 0.5x0.5\,\unit{^\circ}}
+  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface (0.5x0.5)\,\unit{^\circ}; (b) Zonal distribution of land surface types.}
   \label{fig:pft_landsurface}

### 2018-10-25 DryDep_supplement.tex
Finished Section 2. Needs to be read one more time before tagging and shippment to co-authors.

@@
 
-\subsection*{S.4 Temperature dependent greening season 2005}
-\subsubsection*{S.4.1 GLEN: NH mid--high latitudes only}
-\appendixfigures
-\begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/GLEN_2005.pdf}
-\end{center}
-\subsubsection*{S.4.1 GLEN: NH and SH mid--high latitudes}
-\appendixfigures
-\begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/GLEN_2005_she.pdf}
-\end{center}
+\subsection*{S.4 Temperature dependent greening season}
+\subsubsection*{S.4.1 Python code snippets}
+\begin{verbatim}
+#----------------------------------------------------------------------------------
+import numpy as np
+import pandas as pd
+import xarray as xr
+#----------------------------------------------------------------------------------
+def start_growing_season_fixed(lat, **kwargs):
+    '''
+    Begin of growing season parameterization adapted from SMOKE-BEIS.
+    '''
+    bLeap = kwargs.pop('leap', False)
+    if (bLeap):
+       MAY31 = 152
+       NOV1  = 306
+       DEC31 = 366
+    else:
+       MAY31 = 151
+       NOV1  = 305
+       DEC31 = 365
 
-\subsection*{S.5 De-accumulation of photoactive radiation (PAR) from OpenIFS: January 1st 2005}
+    if lat < -65:
+        return((0,))
+    elif lat < -23:
+        return((NOV1,1))
+    elif lat <= 23:
+        return((1,))
+    elif lat < 65:
+        return(((lat-23)*4.5,))
+    else:
+        return((0,))
+#----------------------------------------------------------------------------------
+def end_growing_season_fixed(lat, **kwargs):
+    '''
+    End of growing season parameterization adapted from SMOKE-BEIS.
+    '''
+    bLeap = kwargs.pop('leap', False)
+    if (bLeap):
+       MAY31 = 152
+       NOV1  = 306
+       DEC31 = 366
+    else:
+       MAY31 = 151
+       NOV1  = 305
+       DEC31 = 365
+
+    if lat < -65:
+        return((0,))
+    elif lat < -23:
+        return((DEC31,MAY31))
+    elif lat <= 23:
+        return((DEC31,))
+    elif lat < 65:
+        return((DEC31-(lat-23)*3.3,))
+    else:
+        return((0,))
+#----------------------------------------------------------------------------------
+def growing_season(temperature,**kwargs):
+    '''
+    Agricultural rule of thumb definition of growing season:
+    5 consecutive days above 5 degree Celsius 
+    and vice versa for end of growing season.
+    '''
+    # Shift start day of evaluation.
+    s_shift = kwargs.pop('s_shift',365/2)
+    # Number of days that need to fulfill temperature criteria
+    degree_days_crit = kwargs.pop('ddc', 5)
+    # Temperature criteria
+    temperature_crit = kwargs.pop('tc', 5)
+    # Switch to southern hemisphere evaluation
+    sh = kwargs.pop('sh', False)
+    # Activate verbose
+    verbose = kwargs.pop('verbose', False)
+    # Counters
+    count_gdd = 0
+    count_days = 0
+    start_gs = (0,False)
+    end_gs = (0,False)
+
+    if sh:
+    # Shift the temperature by halve a year - shifted back later.
+        temp = temperature.roll(time=s_shift)
+    else:
+        temp = temperature
+    for itemp in temp:
+        count_days += 1
+        if not start_gs[1]:
+            if itemp > temperature_crit:
+                count_gdd += 1
+            else:
+                count_gdd = 0
+            if count_gdd == degree_days_crit:
+                if sh:
+                    start_gs = (count_days+s_shift, True)
+                else:
+                    start_gs = (count_days, True)
+                count_gdd = 0
+        elif start_gs[1] and not end_gs[1] and count_days>s_shift:
+            if itemp <= temperature_crit:
+                count_gdd += 1
+                if verbose:
+                    print(count_days, count_gdd)
+            else:
+                count_gdd = 0
+            if count_gdd == degree_days_crit:
+                if sh:
+                    end_gs = (count_days-s_shift, True)
+                else:
+                    end_gs = (count_days, True)
+                count_gdd = 0
+
+    return({'sgs':start_gs,'egs':end_gs})
+#----------------------------------------------------------------------------------
+def growing_season_stadyn(xr_temp):
+    '''
+    Preprocess the greening season for OsloCTM3.
+    Setting using the 5deg-5days criteria and in case this fails
+    or is out of the defined bounds,
+    fall back to SMOKE-BEIS parameterization.
+    '''
+    # Fetch leap year from input data
+    bLeap = False
+    if xr_temp.time.size==366:
+       bLeap = True
+    # Use 5deg-5days criteria   
+    if (xr_temp.lat > 45 and xr_temp.lat < 85):
+        gs = growing_season(xr_temp)
+        sgs = (gs['sgs'][0],)
+        egs = (gs['egs'][0],)
+        if ( not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
+            # Check for failing 5deg-5days criteria => Fall back to fixed
+            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+    elif (xr_temp.lat < -35 and xr_temp.lat >= -65):
+        gs = growing_season(xr_temp, sh=True)
+        sgs = (gs['sgs'][0],1)
+        egs = (xr_temp.size, gs['egs'][0])
+        # Check for failing 5deg-5days criteria => Fall back to fixed
+        if (not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
+            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+    # Use SMOKE-BEIS parameterization
+    else:
+        sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+        egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
+    # Generate GDAY and GLEN fields
+    # in case no growing season has been allocated
+    if (sgs[0]==egs[0]==0):
+        gday = np.repeat(0,len(xr_temp.time))
+        glen = 0
+    # in normal cases
+    else:
+        # handle northern hemisphere
+        if len(sgs)==1:
+            gday = np.concatenate((np.repeat(0,int(sgs[0])-1),
+                                   np.arange(1,int(egs[0])-int(sgs[0])+1),
+                                   np.repeat(0,len(xr_temp.time)-
+                                   int(egs[0])+1)))
+            glen = int(egs[0])-int(sgs[0])+1
+        # handle southern hemisphere
+        else:
+            gday = np.concatenate((np.arange(1,int(egs[1])+1)+len(xr_temp.time)
+                                   -int(sgs[0])+1,
+                                   np.repeat(0,int(sgs[0])-int(egs[1])),
+                                   np.arange(1,len(xr_temp.time)-int(sgs[0])+1)))
+            glen = len(xr_temp.time)-(int(sgs[0])-int(egs[1]))+1
+    # This should not happen, but in case it does, print info.
+    if not(len(gday)==len(xr_temp.time)):
+        print(xr_temp, sgs, egs)
+    # Output
+    data_gday = xr.DataArray(gday.astype(int),[('time', xr_temp.time.data)])
+    return(data_gday, (int)(glen))
+#----------------------------------------------------------------------------------
+\end{verbatim}
+%\subsubsection*{S.4.2 GLEN: NH mid--high latitudes only}
+%\appendixfigures
+%\begin{center}
+%  \includegraphics[height=0.4\textheight, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005.pdf}
+%\end{center}
+
+\subsection*{S.5 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS: January 2nd 2005}
 \subsubsection*{S.5.1 Output from OpenIFS: Accumulated PAR}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/PAR_as_is.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_as_is.png}
 \end{center}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/PAR_partly_deaccumulated.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_partly_deaccumulated.png}
 \end{center}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/PAR_deaccumulating_00h.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_deaccumulating_00h.png}
 \end{center}

### 2018-10-29 DryDep.tex
Finished first draft of section 2.

@@
 In addition, ozone is a potent greenhouse gas in both, stratosphere and troposphere.
-With a radiative forcing of $0.40 \pm 0.20\,\unit{Wm^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
+With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
 In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
@@
 Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
-Based on observations, STE might only amount to roughly 10\,\unit{\%} ($550 \pm 140$\,\unit{Tg a^{-1}}) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be 5000\,\unit{Tg a^{-1}} \citep{ACP:Monks2015}.
+Based on observations, STE might only amount to roughly 10\,\unit{\%} ($550 \pm 140$\,\unit{Tg\,a^{-1}}) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be 5000\,\unit{Tg\,a^{-1}} \citep{ACP:Monks2015}.
 Ozone is removed from the atmosphere by photochemical reactions or scavenging processes.
@@
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
-It is estimated that about $1000 \pm 200$\,\unit{Tg a^{-1}} of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
+It is estimated that about $1000 \pm 200$\,\unit{Tg\,a^{-1}} of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
@@
 \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
-An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm s^{-1}}$]{ACP:Helmig2007}.
+An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$]{ACP:Helmig2007}.
 Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
@@
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
-In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA).
+In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA).
 As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different resistance contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy $R^i_c$) which we will recap in the following.
@@
 \begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z R_a}.
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial_z T}{\partial_z R_a}.
   \label{eq:sens_heat}
@@
 \begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air} c_P \cdot \frac{\partial_z T}{\partial_z u} \cdot u_*^2.
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial_z T}{\partial_z u} \cdot u_*^2.
   \label{eq:eddy_theo}
@@
 \end{equation}
-As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are computed directly from the meteorological input fields.
+As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with $\tau^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2s^{-1}}$).
+with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
 The used ratios $D_\chem{H_2O}/D_i$ are displayed in Table~\ref{tab:diffusivity}.
@@
     & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
-    & & & \multicolumn{2}{l}{(\unit{M\cdot atm^{-1}})}\\
+    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
     \middlehline
@@
 \end{equation}
-with an imposed lower threshold of 10\,\unit{sm^{-1}} and an upper limit of 1000\,\unit{sm^{-1}}.
-The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{ms^{-1}} are considered as boundary between both cases.
+with an imposed lower threshold of 10\,\unit{s\,m^{-1}} and an upper limit of 1000\,\unit{s\,m^{-1}}.
+The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{m\,s^{-1}} are considered as boundary between both cases.
 For calm sea, we apply the following upper limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
 \begin{equation}
-  z_0^\text{calm} = \text{min}\left(2\cdot10^{-3}, 0.135 \cdot \frac{\nu}{u_*}\right).
+  z_0^\text{calm} = \text{min}\left\{2\cdot10^{-3}, 0.135 \cdot \frac{\nu}{u_*}\right\}.
 \end{equation}
@@
 \end{equation}
-For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg \, m^{-1} s^{-1} K^{-1}} \cdot T$.
+For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
 But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$ (Supplement S.1).
@@
 \begin{equation}
-  z_0^\text{rough} = \text{min}\left(2\cdot10^{-3}, 0.018 \cdot \frac{u^2_*}{g}\right).
+  z_0^\text{rough} = \text{min}\left\{2\cdot10^{-3}, 0.018 \cdot \frac{u^2_*}{g}\right\}
 \end{equation}
+with a gravitational acceleration $g = 9.836\,\unit{m\,s^{-2}}$.
 The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
-Since the $z_0$ computed with this parameterization is rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{sm^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{s\,m^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{equation}
-  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{(f_\text{min}, f_\text{T} \cdot f_\text{D} \cdot f_\text{SW})}.
+  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
   \label{eq:stomatal}
@@
 The factors herein are normalized and vary within the range $0-1$.
-They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\text{T}$), water vapor pressure deficit ($f_\text{D}$), and soil water content ($f_\text{SW}$).
-The temperature adjustment is computed from
+They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
+
+The temperature adjustment $f_T$ of plants is computed from
 %
 \begin{equation}
-  f_\text{T} = \frac{T_\text{2m}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{2m}}{T_{\text{max}}-T_\text{opt}}\right)^\beta, 
+  f_T = \frac{T_\text{2m}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{2m}}{T_{\text{max}}-T_\text{opt}}\right)^\beta, 
 \end{equation}
@@
 All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
-Since $f_\text{T}$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+Since $f_T$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+
 The water vapor deficit (VPD) is proportional to the saturation partial pressure of water ($P_\chem{H_2O}^s$) and relative humidity ($\text{RH}$)
@@
 %
-Using tabulated values of $f_\text{min}$, $D_\text{min}$, $D_\text{max}$, the water vapor pressure deficit penalty factor can be computed:
+Using tabulated values of $f_\text{min}$, $D_\text{min}$, $D_\text{max}$, the water vapor pressure deficit penalty factor $f_D$ can be computed:
 %
 \begin{equation}
-  f_\text{D} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}.
+  f_D = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}.
 \end{equation}
-%
-The penalty factor with respect to available soil water ($\text{SW}$) is defined as 
+
+The penalty factor with respect to available soil water ($\text{SW}$) $f_\text{SW}$ is defined as 
 %
@@
 \citet{ACP:Simpson2012} suggest using SW in soil depths below 1\,\unit{m} (SWVL4 in OpenIFS).
-Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance much more sensitive to precipitation than it would be otherwise.
+Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance more sensitive to precipitation than it would be otherwise.
+
 The phenology of a plant typically describes its lifecycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
-In the EMEP scheme, phenology is parameterized with respect to the begin of the greening season (GDAY) and its length (GLEN).
-Details about our choice of GDAY and GLEN are given in Section~\ref{subsubsec:greening}.
-In summary, the $f_\text{phen}$ parameterization reads as follows:
+In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
+Details about our treatment of these are given in Section~\ref{subsubsec:greening}.
+In summary, our adaption of the $f_\text{phen}$ parameterization reads as follows:
 %
@@
 %
+Herein, we use the SGS and EGS derived parameters day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
 The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}).
-If GDAY is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
-Since the EMEP scheme is tuned to NH mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal 365 which is the case in the tropics.
+If GLEN is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
+Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$.
+Since the EMEP scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
 %
@@
 \end{figure}
-%
+
 Light in the wavelength band $(400-700)$\,\unit{nm} to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
-In the EMEP parameterization, an integrated photon flux is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
+The integral of PAR over these wavelengths is the photosynthetic photon flux denisty (PPFD).
+In the EMEP parameterization, this integrated photon flux is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
 \begin{equation}
@@
 \end{equation}
-%
-Since $g_\text{max}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,\,m^{-2}}, a unit conversion is done:
+
+Since $g_\text{max}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,m^{-2}}, a unit conversion is done:
 \begin{equation}
@@
 \end{equation}
-Herein $R$ is the universal gas constant. With this the total gridbox stomatal conductance $G_\text{sto}$ is computed weighted by the land surface type fraction $f_L(N)$:
+Herein, $R$ is the universal gas constant. With these, the total gridbox stomatal conductance $G_\text{sto}$ is computed weighted by the land surface type fraction $f_L(N)$:
 \begin{equation}
@@
 \end{equation}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending in the soil/surface. For a land surface type $N$, we can write
+The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
 \begin{equation}
@@
 \begin{equation}
-  r_\text{ext} = 2000\cdot F_T\,\unit{sm^{-1}}.
+  r_\text{ext} = 2000\,\unit{s\,m^{-1}}\cdot F_T.
 \end{equation}
@@
 As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs.
-Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.\\
+Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.
 
@@
 \subsubsection*{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ N}.
+The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ north}.
 To adapt this to a global model, we need to make a few additional assumptions.
@@
 For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
-From MODIS category 'Barren or sparsely vegetated', everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.
+From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.
 
@@
 \end{equation}
-wherein $F_N = 3$ for fine nitrate and ammonium and $F_N = 1$ for all other species and $L$ is again the Obukhov length.
+wherein $F_N = 3$ for fine nitrate and ammonium and $F_N = 1$ for all other species.
+$L$ is again the Obukhov length.
 $a_2 = 300\,\unit{m}$, while $a_1$ differs for forest and non-forest.
@@
       \begin{cases}
-        \emph{if} \quad \emph{dry} & \max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
-        \emph{if} \quad \emph{wet} & 1.5\cdot\max{(a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10})}\\
+        \emph{if} \quad \emph{dry} & \max{\left\{a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10}\right\}}\\
+        \emph{if} \quad \emph{wet} & 1.5\cdot\max{\left\{a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10}\right\}}\\
     \end{cases}\\
@@
 \label{subsubsec:greening}
-In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ together with leaprates ($d_\text{SGS}$, $d_\text{EGS}$, $\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
+In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with leaprates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
 For the growing season of crops in the computation of non-stomatal conductances, we use fixed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated photosynthetic active radiation is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
+From OpenIFS an accumulated PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
 Since this field is already integrated over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
@@
 
+%\clearpage
+
 \appendix
@@
 \appendixfigures
-\begin{figure}[t]
+\begin{figure*}[!htbp]
   \centering
@@
   \label{fig:pft_landsurface}
-\end{figure}
-\subsection{}     %% Appendix A1, A2, etc.
+\end{figure*}
+%\subsection{}     %% Appendix A1, A2, etc.
 

### 2018-10-29 DryDep_supplement.tex
Finished first draft of section 2.

@@
 \appendixtables
-\begin{sidewaystable}[!htbp]
+\begin{sidewaystable*}[!htbp]
   %\caption{S.2 Tabulated parameters for stomatal conductance}
@@
     \tophline
-    Code & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $f_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
-    & (\unit{mmol\,s^{-1}\,\,m^{-2}}) &&&&&&&&&&&&&\\
+    Code$^*$ & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $\alpha_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
+    & (\unit{mmol\,s^{-1}\,m^{-2}}) &&&&&&(\unit{days})&(\unit{days})&(\unit{days})&(\unit{days})&&(\unit{^\circ C})&(\unit{^\circ C})&(\unit{^\circ C})\\
     \middlehline
@@
     \tophline
-    Code & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & height & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$ & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
-    &&&&&&&&&&&&&& \\
+    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$ & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
+    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\
     \middlehline
@@
   \end{tabular}
-\end{sidewaystable}
+  \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean scrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
+\end{sidewaystable*}
 %\end{minipage}

### 2018-11-06 DryDep.tex
Started to include the actual resuls into the documents and updated some plots.

@@
 Include table of all sensitivity runs.
-\citep[e.g., $v^\chem{O_3}_{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
+\citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
 dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
+  \caption{Relative difference between the sensitivity simulation emep\_psgsssh and the Wesely scheme in (a) surface ozone concentration; (b) ozone dry deposition velosity; (c) amount of ozone removed from the atmosphere by dry deposition.}
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep}\\
+  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo}
+  \caption{Comparison of various model runs with the OsloCTM3 and the multi-model-mean results by \citet{ACP:Hardacre2015}.
+(a) total annual ozone drydeposition; (b) annual average ozone dry depositon velocity.}
+\end{figure*}
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_hem}
+  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition seperated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
+Comparison with results from \citet{ACP:Hardacre2015}.}
+\end{figure}
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
+  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
+\end{figure}
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
+  \caption{Average ozone dry deposition velocitys seperated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}). \citep{AE:Hu2017} -> need to look into this one...
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-monthly_ozone_drydep_stations_True}\\
+  \includegraphics[width=8.3cm]{pictures/final-monthly_drydep_divergence_True}
+  \caption{Ozone dry deposition velocities at different observation sites.
+Comparison between Weseley scheme and average result from all sensitivity tests to observations and multi-model-mean from \citet{ACP:Hardacre2015}.}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-11-07 DryDep.tex
Included table with simulation specifications and wrote some words about it. Need to make the diff plot, though.

@@
 \label{sec:eval}
-For meteorological input, we use \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}.
-CEDS historical emission inventory is used for emissions.
+In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme.
+We focus on the change in total ozone dry deposition, ozone dry deposition velocities, and surface ozone concentration related to the stomatal conductance part of the surface resistance.
+We evaluate the results in the context of the multi-model comparison and observations
+in\citet{ACP:Hardacre2015} (Section~\ref{subsec:model}-\ref{subsec:obs}).
+The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}.
+CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
+In the following, we will discuss the various model sensitivity studies.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsec:sens}
-Include table of all sensitivity runs.
+A summary of specifications of all simulations discussed in this section is given in Table~\ref{tab:simsum}.
+For simplicity, we only list the tested parameters.
+An x in the table denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.
+With respect to the sensitivity on light (PPFD), $f_\text{light}\equiv 1$ has been tested.
+Regarding the phenology, we have tested three different options: No phenology ($f_\text{phen}\equiv 1$), phenology including / excluding the Tropics (TR).
+Diverging from \citet{ACP:Simpson2012}, the simulations inititally ran with the uppermost soil water level (SWVL1).
+One sensitivity test was conducted using SWVL4.
+The meteorological year used in all simulations is 2005.
+Most of the simulations were run with CEDS emissions for 2014.
+One test has been conducted to look at the impact of the emissions.
+\begin{table*}[t]
+  \caption{Summary of specifications of all simulations discussed in this section.
+For simplicity, only the tested parameters are listed.
+An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
+  \begin{tabular}{lccccccccc}
+    \tophline
+    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
+    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{cm\,s^{-1}}) & (year)\\
+    \middlehline
+    Wesely type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
+    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
+    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
+    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
+    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
+    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
+    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
+    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
+    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
+    \bottomhline
+  \end{tabular}
+  \label{tab:simsum}
+  %\belowtable{} % Table Footnotes
+\end{table*}
 \citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
@@
 %
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_hem}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
   \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition seperated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
 Comparison with results from \citet{ACP:Hardacre2015}.}
-\end{figure}
+\end{figure*}
 %
@@
   This work was supported by the Norwegian Research Council (NRC) through the project The double punch: Ozone and climate stresses on vegetation (OzoNorClim).\\
-  We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
+  %We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
   The used Leaf Area Index (LAI) and roughness length ($Z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A.
@@
   Community Emission Data System (CEDS) historical emission inventory is provided by the Joint Global Research Institute project (\url{http://www.globalchange.umd.edu/ceds/}.)
+  Randerson, J.T., G.R.
+van der Werf, L.
+Giglio, G.J.
+Collatz, and P.S.
+Kasibhatla.
+2018.
+Global Fire Emissions Database, Version 4, (GFEDv4).
+ORNL DAAC, Oak Ridge, Tennessee, USA.
+https://doi.org/10.3334/ORNLDAAC/1293
 \end{acknowledgements}

### 2018-11-13 DryDep.tex
A lot about sensitivity studies and some new plots.

@@
 A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
-All relevant information and papers can also be found on the \href{http://folk.uio.no/mariantl/osloctm3/}{OsloCTM3 webpage}.
-The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN) \citep{ACP:Guenther2006}.
+The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}.
 A publication focusing on this is planed.
 Parallelization of the OsloCTM3 has not been optimized as of now.
-It is possible to run the model in resolution of at least T159N80L60, but it is not very time and memory efficient.
+While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, it is very time and memory consuming to run the CTM with full chemistry at this resolution.
 Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$.
 In the following, we will give a detailed account of the new dry deposition scheme.
-Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, non the less.
+Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, nonetheless.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with $\tau^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).
+As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with the turbulent surface stress $\tau_0^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection*{Stomatal conductance}
-The stomatal conductance is a measure of the rate of \chem{CO_2} entering or evapotranspiration through the stomata of a leaf.
+The stomatal conductance is a measure of the rate of \chem{CO_2} evapotranspiration through the stomata of a leaf.
 There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.).
@@
 Extending the EMEP scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
+%
 \begin{table}[t]
@@
 
-In this way, vegetation affects the conductance also by being there not only by uptake through the stomata. The in canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
+In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata. The in-canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
 \begin{equation}
@@
 The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes (40--60\,\unit{^\circ}).
-Towards the pol, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
+Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
 At a latitude of 74\,\unit{^\circ} a minimum height of $3/10 \cdot h(N)$ is reached and kept constant.
@@
 We also assume symmetry in both hemispheres.
-Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic.
+Presuming a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic.
 For four example PFTs, results are shown in the Supplement (Fig.~S.2).
@@
 Based on this rule of thumb, we have pre-processed our meteorological input data offline.
-If, within these 5 days, the temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively.
+If, within these 5 days, the average temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively.
 First, we used the 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
@@
 From OpenIFS an accumulated PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
-Since this field is already integrated over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
+Since this field is already integrated assumingly over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
 Therefor, we will refer to the deaccumulated field as PPFD.
@@
 The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
+%
 \begin{equation}
@@
 \end{equation}
-
+%
+The resulting PPFD fields are still accumulated over a time period of three hours and should be devided by 3.
+A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about 30\,\unit{\%} below observations.
+To counter this, we decided to refrain from the devision at this stage.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme.
-We focus on the change in total ozone dry deposition, ozone dry deposition velocities, and surface ozone concentration related to the stomatal conductance part of the surface resistance.
-We evaluate the results in the context of the multi-model comparison and observations
-in\citet{ACP:Hardacre2015} (Section~\ref{subsec:model}-\ref{subsec:obs}).
-The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}.
+We focus on the changes in ozone, e.g., total dry deposition, dry deposition velocities, and surface concentration, related to the stomatal conductance parameterization.
+We evaluate our results with respect to \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
+The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}.
 CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sensitivity tests}
+\subsection{Sensitivity studies}
 \label{subsec:sens}
-A summary of specifications of all simulations discussed in this section is given in Table~\ref{tab:simsum}.
+As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme hase been revised majorly.
+Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of the changes.
+We have conducted 9 different model simulations to assess the parameter space of the EMEP scheme.
+A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+For all model integrations, the meteorological reference year is 2005.
+This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years.
+After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}.
+As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively.
+During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
+Especially in Europe, basically resulting in the subpression of canopy resistance in these regions.
+The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the borthern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}).
+Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3.
+
+focussion on the details of the canopy resistance parameterization. 
  For simplicity, we only list the tested parameters.
@@
 One test has been conducted to look at the impact of the emissions.
+All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
 \begin{table*}[t]
@@
     EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
+    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
     EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
     EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
-    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
     \bottomhline

### 2018-11-13 DryDep_supplement.tex
Started to include the actual resuls into the documents and updated some plots.
A lot about sensitivity studies and some new plots.

@@
 \begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_as_is.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.7cm}]{pictures/PAR_as_is.png}
 \end{center}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_partly_deaccumulated.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.7cm}]{pictures/PAR_partly_deaccumulated.png}
 \end{center}
@@
 
+\subsection*{S.6 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-total_ozone_drydepvelo_monthly_pft_sh}
+\end{center}
 \end{document}

### 2018-11-14 DryDep.tex
Finish section 3.1 and corrected plot.

@@
 In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme.
-We focus on the changes in ozone, e.g., total dry deposition, dry deposition velocities, and surface concentration, related to the stomatal conductance parameterization.
+We focus on the changes in ozone, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ respective atmospheric surface ozone burden in units of \unit{kg}, related to the stomatal conductance parameterization.
 We evaluate our results with respect to \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
@@
 \label{subsec:sens}
-As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme hase been revised majorly.
-Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of the changes.
-We have conducted 9 different model simulations to assess the parameter space of the EMEP scheme.
+As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme has undergone a major revision.
+Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change.
+Therefor, we have conducted 9 different model simulations to mainly assess the parameter space of the stomatal conductance within EMEP scheme.
 A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
@@
 During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
-Especially in Europe, basically resulting in the subpression of canopy resistance in these regions.
-The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the borthern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}).
+Especially in Europe, this basically results in the subpression of canopy resistance in these regions.
+The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}).
 Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3.
-
-focussion on the details of the canopy resistance parameterization. 
- For simplicity, we only list the tested parameters.
+Accidently, we used forcings for the year 2014 instead of our meteorological year 2005 and used this opportunnity to assess the impact on ozone regarding differing emissions.
+As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+A final sensityivity study was conducted changing to deep soil water content (\emph{EMEP\_SWVL4}).
+All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
 An x in the table denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.
-With respect to the sensitivity on light (PPFD), $f_\text{light}\equiv 1$ has been tested.
-Regarding the phenology, we have tested three different options: No phenology ($f_\text{phen}\equiv 1$), phenology including / excluding the Tropics (TR).
-Diverging from \citet{ACP:Simpson2012}, the simulations inititally ran with the uppermost soil water level (SWVL1).
-One sensitivity test was conducted using SWVL4.
-The meteorological year used in all simulations is 2005.
-Most of the simulations were run with CEDS emissions for 2014.
-One test has been conducted to look at the impact of the emissions.
-All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
+
 \begin{table*}[t]
@@
 \end{table*}
-\citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
-dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
-
+%
+In Fig.~\ref{fig:diff_maps}, we show the relative difference between \emph{EMEP\_ppgsssh} and the \emph{Wesely type} by means of surface ozone burden, dry deposition velocity and removal of ozone due to dry deposition.
+Surface ozone increases globally except for regions covered by tropical forest.
+Especially in desert regions in Africa and Asia, surface ozone increases by up to 100\,\unit{\%}.
+Consistently, dry deposition velocities decrease globally and up to 100\,\unit{\%} in desert regions, while they increase over tropical forest.
+With respect to total dry deposition, the picture is a bit less clear.
+We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only a small change.
+%
 \begin{figure}[t]
   \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
-  \caption{Relative difference between the sensitivity simulation emep\_psgsssh and the Wesely scheme in (a) surface ozone concentration; (b) ozone dry deposition velosity; (c) amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \label{fig:diff_maps}
 \end{figure}
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
+
 \begin{figure*}[t]
@@
 (a) total annual ozone drydeposition; (b) annual average ozone dry depositon velocity.}
+  \label{fig:mmm_drydep}
 \end{figure*}
@@
 Comparison with results from \citet{ACP:Hardacre2015}.}
+  \label{fig:mmm_drydep_hem}
 \end{figure*}
@@
   \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
+  \label{fig:mmm_drydep_season}
 \end{figure}
@@
   \caption{Average ozone dry deposition velocitys seperated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
+  \label{fig:mmm_drydep_season_pft}
 \end{figure}
@@
 Comparison between Weseley scheme and average result from all sensitivity tests to observations and multi-model-mean from \citet{ACP:Hardacre2015}.}
+  \label{fig:mmm_drydep_stations}
 \end{figure}
@@
 \label{sec:disc}
+dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017})
 

### 2018-11-15 DryDep.tex
Update of some figures, correction of introduction text to Section 3. Tried to figure out about dry deposition onto deserts...

@@
 \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
-An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$]{ACP:Helmig2007}.
+An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.
 Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
@@
 For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
-Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{cm\,s^{-1}}.
+Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{m\,s^{-1}}.
 
@@
 \begin{equation}
-  a_1^\text{hydrophob.} = 0.025\,\unit{cm\,s^{-1}} \cdot \overline{u_*}
+  a_1^\text{hydrophob.} = 0.025\,\unit{m\,s^{-1}} \cdot \overline{u_*}
 \end{equation}
@@
 \begin{equation} 
-  a_1^\text{hydrophil.} = 0.2\,\unit{cm\,s^{-1}} \cdot \overline{u_*},
+  a_1^\text{hydrophil.} = 0.2\,\unit{m\,s^{-1}} \cdot \overline{u_*},
 \end{equation}
@@
 \label{sec:eval}
-In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme.
-We focus on the changes in ozone, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ respective atmospheric surface ozone burden in units of \unit{kg}, related to the stomatal conductance parameterization.
-We evaluate our results with respect to \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
+In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme.
+We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ or surface ozone burden in units of \unit{kg}, respectivly.
+We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
 The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}.
-CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
-In the following, we will discuss the various model sensitivity studies.
+CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
+Biogenic emissions are taken from MEGAN output.
+In the following, we will present the various model sensitivity studies.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change.
-Therefor, we have conducted 9 different model simulations to mainly assess the parameter space of the stomatal conductance within EMEP scheme.
-A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+Therefor, the conducted sensitivity studies (9 in total) mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
+A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
 For all model integrations, the meteorological reference year is 2005.
@@
 Especially in Europe, this basically results in the subpression of canopy resistance in these regions.
-The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}).
-Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3.
-Accidently, we used forcings for the year 2014 instead of our meteorological year 2005 and used this opportunnity to assess the impact on ozone regarding differing emissions.
+The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
+Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}).
+Accidently, we have used forcings for the year 2014 instead of 2005.
+Though, we take this as opportunnity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
 As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
-A final sensityivity study was conducted changing to deep soil water content (\emph{EMEP\_SWVL4}).
+A final sensitivity study was conducted changing to soil water level at depths deeper than 1\,\unit{m} (\emph{EMEP\_SWVL4}).
 All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
@@
     Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
-    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{cm\,s^{-1}}) & (year)\\
+    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{m\,s^{-1}}) & (year)\\
     \middlehline
-    Wesely type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
+    Wesely\_type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
     EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the relative difference between \emph{EMEP\_ppgsssh} and the \emph{Wesely type} by means of surface ozone burden, dry deposition velocity and removal of ozone due to dry deposition.
+In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone due to dry deposition.
 Surface ozone increases globally except for regions covered by tropical forest.
 Especially in desert regions in Africa and Asia, surface ozone increases by up to 100\,\unit{\%}.
-Consistently, dry deposition velocities decrease globally and up to 100\,\unit{\%} in desert regions, while they increase over tropical forest.
+Consistently, dry deposition velocities decrease globally (up to 100\,\unit{\%}) in desert regions, while they increase over tropical forest.
 With respect to total dry deposition, the picture is a bit less clear.
-We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only a small change.
+We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
 %
@@
   \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
-  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
   \label{fig:diff_maps}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with recent multi-model comparison}
+\subsection{Comparison with multi-model evaluation}
 \label{subsec:model}
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep}\\
-  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo}
-  \caption{Comparison of various model runs with the OsloCTM3 and the multi-model-mean results by \citet{ACP:Hardacre2015}.
-(a) total annual ozone drydeposition; (b) annual average ozone dry depositon velocity.}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
+  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) surface ozone concentrations; and in comparison with the multi-model evaluation by \citet{ACP:Hardacre2015}: (b) Annual average ozone dry depositon velocity; (c) Total annual ozone dry deposition.
+The different colors indicate sets of associated simulation.}
   \label{fig:mmm_drydep}
-\end{figure*}
+\end{figure}
 %

### 2018-11-16 DryDep.tex
Updates some more plots. Started writing the next section.

@@
 We change from an empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere into account.
-Examining the sensitivity of the scheme to certain parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
-We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies.
-We will also estimate the impact on the modeled ozone concentrations, both at the surface and in the free troposphere.
-We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to ??\,\unit{\%}.
+Examining the sensitivity of the scheme to various parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
+We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies and also estimate the impact on the modeled ozone concentrations, both at the surface and in the free troposphere.
+We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to $100\,\unit{\%}$.
 {\bf TODO: Finish the results section as soon as there are results.}
@@
 Continuously high concentrations of ambient air ozone are hazardous to the whole ecosystem.
-It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about 3--15\,\unit{\%} \citep{PJ:Ainsworth2017} as well as 7\,\unit{\%} loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
+It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} as well as $7\,\unit{\%}$ loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
 The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs - also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
+Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
 A typical reaction mechanism for \chem{CO} is sketched in the following.
@@
 Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
-Based on observations, STE might only amount to roughly 10\,\unit{\%} ($550 \pm 140$\,\unit{Tg\,a^{-1}}) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be 5000\,\unit{Tg\,a^{-1}} \citep{ACP:Monks2015}.
+Based on observations, STE might only amount to roughly $10\,\unit{\%}$ ($550 \pm 140\,\unit{Tg\,a^{-1}}$) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be $5000\,\unit{Tg\,a^{-1}}$ \citep{ACP:Monks2015}.
 Ozone is removed from the atmosphere by photochemical reactions or scavenging processes.
@@
 We also do not consider the contribution of very short-lived ozone depleting substances (VSLS), that affect stratospheric ozone \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017}, to the tropospheric ozone abundance.\\
-Since ozone is highly reactive, its global mean life-time in the troposphere is roughly 22 days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
+Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
 The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
-Typical concentrations of surface ozone range from 10\,\unit{ppb} over the tropical Pacific to 100\,\unit{ppb} in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
+Typical concentrations of surface ozone range from $10\,\unit{ppb}$ over the tropical Pacific to $100\,\unit{ppb}$ in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
 This variability poses a challenge on both, trend analysis from observation as well as validation and intercomparison of models.
@@
 But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models.
-Among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP), there is a general tendency to underestimate tropospheric ozone burden (e.g., $10-20$\,\unit{\%} negative bias at 250\,\unit{hPa} in the southern hemisphere (SH) tropical region) \citep[Chapter 8]{IPCC2013}.
+Among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP), there is a general tendency to underestimate tropospheric ozone burden (e.g., $10-20\,\unit{\%}$ negative bias at $250\,\unit{hPa}$ in the southern hemisphere (SH) tropical region) \citep[Chapter 8]{IPCC2013}.
 With respect to surface ozone, \citet{ACP:Schnell2015} conclude that all ACCMIP models, which reported hourly surface ozone, tend to overestimate surface ozone values in North America and Europe in comparison with available observations.
@@
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
-It is estimated that about $1000 \pm 200$\,\unit{Tg\,a^{-1}} of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
-Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_d$.
-Dry deposition velocities $v^i_d$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
-\begin{equation}
-  v^i_d = \frac{1}{R_a + R^i_b + R^i_c},
+It is estimated that about $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
+Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
+Dry deposition velocities $v^i_\text{DD}$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
+\begin{equation}
+  v^i_\text{DD} = \frac{1}{R_a + R^i_b + R^i_c},
   \label{eq:drydep_velo}
@@
 But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
-The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
+The variety of differing formulations and choices of parameters leads to a wide spread of results in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about $20\,\unit{\%}$ uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
 
 In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
-In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}).
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}).
 In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
@@
 This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use.
-Day was distinguished from night by solar zenith angles below 90\,\unit{^\circ}.
-Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses.
+Day was distinguished from night by solar zenith angles below $90\,\unit{^\circ}$.
+Winter was defined by temperatures below $273.15\,\unit{K}$ for gridboxes containing land masses.
 For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes.
-In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account.
+In addition, a reduced uptake due to snow cover above $1\,\unit{m}$ for forest and $10\,\unit{cm}$ for grass/tundra, respectively, was taken into account.
 We will refer to this parameterization as \emph{Wesely scheme}.\\
@@
 \end{equation}
-with $\kappa = 0.40$ the K\'{a}rm\'{a}n constant, $u_*$ the friction velocity, $\Psi_m$ the integrated stability equations for momentum, $d$ a constant (typically 0.7\,\unit{m}), and $L$ the Obukhov length.
+with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a constant $d$ (typically $0.7\,\unit{m}$), and the Obukhov length $L$.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
@@
 \end{equation}
-As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with the turbulent surface stress $\tau_0^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).
+As reference height $z$, we use the height at mid-level of the lowermost model level (roughly $8\,\unit{m}$).
+From the meteorological input fields $u_z$ is directly available, while $u_*$ can be computed via
+\begin{equation}
+  u_* = \sqrt{\frac{\tau_0}{\rho}},
+\end{equation}
+with the turbulent surface stress $\tau_0^2 = \text{NSSS}^2 + \text{EWSS}^2$, and northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-with an imposed lower threshold of 10\,\unit{s\,m^{-1}} and an upper limit of 1000\,\unit{s\,m^{-1}}.
-The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, where wind speeds above 3\,\unit{m\,s^{-1}} are considered as boundary between both cases.
+with an imposed lower threshold of $10\,\unit{s\,m^{-1}}$ and an upper limit of $1000\,\unit{s\,m^{-1}}$.
+The computation of roughness length $z_0$ over ocean is divided into a \emph{calm} and a \emph{rough} sea case, with a threshold of $3\,\unit{m\,s^{-1}}$.
 For calm sea, we apply the following upper limit \citep[][with a slightly higher coefficient of $0.135$]{Hinze1975,Garratt1992}
@@
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
-$P_0$ is the surface pressure, as $T$ the 2\,\unit{m} temperature is chosen, and $R_\text{air}$ is the universal gas constant for air.
+$P_0$ is the surface pressure, as $T$ the $2\,\unit{m}$ temperature is chosen, and $R_\text{air}$ is the universal gas constant for air.
 The rough sea case follows the method of \citet{QJRMS:Charnock1955,JPO:Wu1980}:
@@
 with a gravitational acceleration $g = 9.836\,\unit{m\,s^{-2}}$.
-The allowed maximum roughness length in both cases is set to 2\,\unit{mm}.
-Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of 10\,\unit{s\,m^{-1}} in about 91\,\unit{\%} of all cases (see Supplement S.1).
+The allowed maximum roughness length in both cases is set to $2\,\unit{mm}$.
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement S.1).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection*{Stomatal conductance}
-The stomatal conductance is a measure of the rate of \chem{CO_2} evapotranspiration through the stomata of a leaf.
+The stomatal conductance is a measure of the rate of \chem{CO_2} exchange and evapotranspiration through the stomata of a leaf.
 There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.).
@@
 \end{equation}
-\citet{ACP:Simpson2012} suggest using SW in soil depths below 1\,\unit{m} (SWVL4 in OpenIFS).
+\citet{ACP:Simpson2012} suggest using SW in soil depths below $1\,\unit{m}$ (SWVL4 in OpenIFS).
 Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance more sensitive to precipitation than it would be otherwise.
@@
 
-Light in the wavelength band $(400-700)$\,\unit{nm} to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
+Light in the wavelength band $(400-700)\,\unit{nm}$ to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
 The integral of PAR over these wavelengths is the photosynthetic photon flux denisty (PPFD).
@@
 \begin{equation}
-  F_T = \exp{(-2\cdot(1+\theta_0))}.
-\end{equation}
-$\theta_0$ is the surface or 2\,\unit{m} temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
+  F_T = \exp{(-0.2\cdot(1+\theta_0))}.
+\end{equation}
+$\theta_0$ is the surface or $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
 %
@@
 As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs.
-Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.
+Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.
 
@@
 \subsubsection*{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ north}.
+The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between $60$ and $74\,\unit{^\circ north}$.
 To adapt this to a global model, we need to make a few additional assumptions.
-The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes (40--60\,\unit{^\circ}).
+The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes ($40-60\,\unit{^\circ}$).
 Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
-At a latitude of 74\,\unit{^\circ} a minimum height of $3/10 \cdot h(N)$ is reached and kept constant.
-Towards the equator, we increase the height linearly so that at a latitude of 10\,\unit{^\circ} a maximum height of $2 \cdot h(N)$ is achieved which is then held constant.
+At a latitude of $74\,\unit{^\circ}$ a minimum height of $3/10 \cdot h(N)$ is reached and kept constant.
+Towards the equator, we increase the height linearly so that at a latitude of $10\,\unit{^\circ}$ a maximum height of $2 \cdot h(N)$ is achieved which is then held constant.
 We also assume symmetry in both hemispheres.
-Presuming a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic.
+Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
 For four example PFTs, results are shown in the Supplement (Fig.~S.2).
@@
 For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
-From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.
+From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.
+This mapping differs from the one used in the Wesely scheme.
 
@@
 For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
-Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{m\,s^{-1}}.
+Since the dry deposition velocity for \chem{CO} is very small, we use a prescribed value for $v_\text{DD}^\chem{CO}$ of $0.03\,\unit{cm\,s^{-1}}$.
 
@@
 \begin{equation}
-  a_1^\text{hydrophob.} = 0.025\,\unit{m\,s^{-1}} \cdot \overline{u_*}
+  a_1^\text{hydrophob.} = 0.025\,\unit{cm\,s^{-1}} \cdot \overline{u_*}
 \end{equation}
@@
 \begin{equation} 
-  a_1^\text{hydrophil.} = 0.2\,\unit{m\,s^{-1}} \cdot \overline{u_*},
+  a_1^\text{hydrophil.} = 0.2\,\unit{cm\,s^{-1}} \cdot \overline{u_*},
 \end{equation}
@@
 For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
-From MODIS category 'Barren or sparsely vegetated', everything polward from 60\,\unit{^\circ} is defined as tundra, while everything equatorward is categorized as desert.}
+From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
   \label{fig:pft_mapping}
@@
 \label{subsubsec:greening}
-In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use fixed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with leaprates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
-For the growing season of crops in the computation of non-stomatal conductances, we use fixed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
+In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use prescribed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with leaprates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
+For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
 Since the eurocentric EMEP parameterization of SGS and EGS is not applicable in a global model, another latitude dependent parameterization is needed.
@@
 In agriculture, there are different empirical rules to estimate the SGS and EGS.
-The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above 5\,\unit{^\circ C} and vice versa for EGS.
+The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above $5\,\unit{^\circ C}$ and vice versa for EGS.
 Based on this rule of thumb, we have pre-processed our meteorological input data offline.
-If, within these 5 days, the average temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively.
-First, we used the 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
-In all other cases and where the 5\,\unit{^\circ C}-days criteria fails, we use the SMOKE-BEIS parameterization.
+If, within these 5 days, the average temperature drops below / rises above $5\,\unit{^\circ C}$, the counters are reset, respectively.
+First, we used the $5\,\unit{^\circ C}$-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
+In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we use the SMOKE-BEIS parameterization.
 The described algorithm written in python 2.7 has been included as Supplement~S.4.1.
-An example map of the computed GLEN using the 5\,\unit{^\circ C}-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
+An example map of the computed GLEN using the $5\,\unit{^\circ C}$-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
 %
@@
 Shown is the total length of the greening season (GLEN) for the year 2005.
-The 5\,\unit{^\circ C}-days criteria has been used in both hemispheres mid--high latitudes.
+The $5\,\unit{^\circ C}$-days criteria has been used in both hemispheres mid--high latitudes.
 Ocean has been shaded to indicate that greening season will only affect land.}
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
+From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
 Since this field is already integrated assumingly over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
 Therefor, we will refer to the deaccumulated field as PPFD.
-An example PAR for January 2nd 2005 is shown in Supplement~S.5.1.
+An example PAR for January 2 2005 is shown in Supplement~S.5.1.
 Deaccumulation of hour 3 till 21, simply means
@@
 %
-The resulting PPFD fields are still accumulated over a time period of three hours and should be devided by 3.
-A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about 30\,\unit{\%} below observations.
+The resulting PPFD fields are still accumulated over a time period of three hours and should be devided by $3$.
+A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations.
 To counter this, we decided to refrain from the devision at this stage.
@@
 In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme.
-We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ or surface ozone burden in units of \unit{kg}, respectivly.
+We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface ozone burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectivly.
 We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
@@
 Accidently, we have used forcings for the year 2014 instead of 2005.
-Though, we take this as opportunnity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
+Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
 As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
-A final sensitivity study was conducted changing to soil water level at depths deeper than 1\,\unit{m} (\emph{EMEP\_SWVL4}).
+A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).
 All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
@@
     Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
-    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{m\,s^{-1}}) & (year)\\
+    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
     \middlehline
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone due to dry deposition.
-Surface ozone increases globally except for regions covered by tropical forest.
-Especially in desert regions in Africa and Asia, surface ozone increases by up to 100\,\unit{\%}.
-Consistently, dry deposition velocities decrease globally (up to 100\,\unit{\%}) in desert regions, while they increase over tropical forest.
-With respect to total dry deposition, the picture is a bit less clear.
-We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-%
 \begin{figure}[t]
@@
 %
+In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone due to dry deposition.
+Surface ozone increases globally except for regions covered by tropical forest.
+Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$.
+Consistently, dry deposition velocities decrease globally (up to $100\,\unit{\%}$) in desert regions, while they increase over tropical forest.
+With respect to total dry deposition, the picture is a bit less clear.
+We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
+A possible explanation is the huge difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities in the EMEP scheme.
+We cover this in more detail in the following section.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsec:model}
-Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
+Compare to \citet{ACP:Hardacre2015} as well as . \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
+
+\subsubsection{Zonal averages}
 
@@
 %
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
+  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
+  \label{fig:mmm_drydep_season}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Average seasonal cycles}
+%
 \begin{figure*}[t]
@@
 %
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
-  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
-  \label{fig:mmm_drydep_season}
-\end{figure}
+As suggested by \citet{ACP:Hardacre2015}, we retrospectively seperate ozone dry deposition velocities with respect to surface types.
+Based on CML2 dynamic land surface, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from $22\,\unit{\%}$ only for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean.
+Since we have not perform a full unfolding on the data, the results should be regarded with suspicion.
 %
@@
 \end{figure}
+%
+The Wesely scheme uses $v^\chem{O_3}_\text{tundra/desert} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{desert} = 0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$.
+From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}.
+This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme.
+Similarly, the dry deposition over water $v^\chem{O_3}_\text{water}=0.07\,\unit{cm\,s^{-1}}$ in the Wesely scheme, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{water} = 0.05\,\unit{cm\,s^{-1}}$.
+$v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ \citep{JGR:Helmig2012}.
+\citet{ACP:Luhar2017} found $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
+%
+\begin{table*}[t]
+\caption{Total ozone dry deposition seperated into land, ocean, and ice contributions.}
+\begin{tabular}{lcccccccccccc}
+\tophline
+& & Ocean & & & Ice & & & Land & & & Total\\
+& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
+& & (Tg a-1) & & & (Tg a-1) & & & (Tg a-1) & & & (Tg a-1)\\
+\middlehline
+88.7 110.3 199.0 OsloCTM3: Wesely type\\
+67.1 84.7 151.8 OsloCTM3: EMEP\_full\\
+66.6 84.1 150.7 OsloCTM3: EMEP\_offLight\\
+67.1 84.6 151.7 OsloCTM3: EMEP\_offPhen\\
+68.1 86.1 154.2 OsloCTM3: EMEP\_SWVL4\\
+67.1 84.6 151.7 OsloCTM3: EMEP\_ppgs\\
+67.0 84.5 151.6 OsloCTM3: EMEP\_ppgssh\\
+66.5 84.8 151.3 OsloCTM3: EMEP\_ppgssh\_ice\\
+66.8 83.5 150.4 OsloCTM3: EMEP\_ppgs\_2005\\
+
+0.7 5.0 5.7 OsloCTM3: Wesely type\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_full\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_offLight\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_offPhen\\
+0.6 4.5 5.1 OsloCTM3: EMEP\_SWVL4\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_ppgs\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_ppgssh\\
+0.2 1.3 1.4 OsloCTM3: EMEP\_ppgssh\_ice\\
+0.6 4.4 5.0 OsloCTM3: EMEP\_ppgs\_2005\\
+
+346.1 153.5 499.6 OsloCTM3: Wesely type\\
+236.0 112.3 348.2 OsloCTM3: EMEP\_full\\
+267.8 129.8 397.6 OsloCTM3: EMEP\_offLight\\
+239.9 116.0 355.9 OsloCTM3: EMEP\_offPhen\\
+241.8 115.0 356.8 OsloCTM3: EMEP\_SWVL4\\
+238.7 112.6 351.3 OsloCTM3: EMEP\_ppgs\\
+243.9 119.4 363.3 OsloCTM3: EMEP\_ppgssh\\
+240.3 120.4 360.7 OsloCTM3: EMEP\_ppgssh\_ice\\
+237.8 110.0 347.8 OsloCTM3: EMEP\_ppgs\_2005\\
+
+612.4 347.9 960.2 OsloCTM3: Wesely type\\
+408.6 245.6 654.2 OsloCTM3: EMEP\_full\\
+451.3 267.2 718.6 OsloCTM3: EMEP\_offLight\\
+413.9 249.9 663.7 OsloCTM3: EMEP\_offPhen\\
+417.2 250.7 667.9 OsloCTM3: EMEP\_SWVL4\\
+411.4 246.0 657.4 OsloCTM3: EMEP\_ppgs\\
+419.4 253.6 673.0 OsloCTM3: EMEP\_ppgssh\\
+413.0 251.1 664.1 OsloCTM3: EMEP\_ppgssh\_ice\\
+409.2 240.7 649.9 OsloCTM3: EMEP\_ppgs\_2005\\
+\bottomhline
+\end{tabular}
+%\belowtable{} % Table Footnotes
+\end{table*}
+%
 
@@
   \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
-  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface (0.5x0.5)\,\unit{^\circ}; (b) Zonal distribution of land surface types.}
+  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface $(0.5\times0.5)\,\unit{^\circ}$; (b) Zonal distribution of land surface types.}
   \label{fig:pft_landsurface}
@@
   %We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
-  The used Leaf Area Index (LAI) and roughness length ($Z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A.
+  The used Leaf Area Index (LAI) and roughness length ($z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A.
 (\doi{10.3334/ORNLDAAC/970}).\\

### 2018-11-16 DryDep_supplement.tex
Updates some more plots. Started writing the next section.

@@
 \begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.7cm}]{pictures/PAR_as_is.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_as_is.png}
 \end{center}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.7cm}]{pictures/PAR_partly_deaccumulated.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_partly_deaccumulated.png}
 \end{center}
@@
 \begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.75cm}]{pictures/PAR_deaccumulating_00h.png}
+  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.65cm}]{pictures/PAR_deaccumulating_00h.png}
 \end{center}

### 2018-11-19 DryDep.tex
Continued writing: Section zonal distributions, and some spell-checking.

@@
 
-The phenology of a plant typically describes its lifecycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
+The phenology of a plant typically describes its life-cycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
 In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
@@
 Light in the wavelength band $(400-700)\,\unit{nm}$ to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
-The integral of PAR over these wavelengths is the photosynthetic photon flux denisty (PPFD).
+The integral of PAR over these wavelengths is the photosynthetic photon flux density (PPFD).
 In the EMEP parameterization, this integrated photon flux is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated.
-Since this field is already integrated assumingly over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
-Therefor, we will refer to the deaccumulated field as PPFD.
+From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be de-accumulated.
+Since this field is already integrated presumably over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
+Therefor, we will refer to the de-accumulated field as PPFD.
 An example PAR for January 2 2005 is shown in Supplement~S.5.1.
-Deaccumulation of hour 3 till 21, simply means
+De-Accumulation of hour 3 till 21, simply means
 \begin{equation}
@@
 \end{equation}
-If everything was right in the OpenIFS output, we would not need to do any further processing, however the first timestep of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately deaccumulate this time step.
+If everything was right in the OpenIFS output, we would not need to do any further processing, however the first time step of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately de-accumulate this time step.
 The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
@@
 %
-The resulting PPFD fields are still accumulated over a time period of three hours and should be devided by $3$.
+The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$.
 A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations.
-To counter this, we decided to refrain from the devision at this stage.
+To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS versions.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme.
-We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface ozone burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectivly.
+We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively.
 We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
@@
 Therefor, the conducted sensitivity studies (9 in total) mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
-A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
 For all model integrations, the meteorological reference year is 2005.
@@
 During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
-Especially in Europe, this basically results in the subpression of canopy resistance in these regions.
+Especially in Europe, this basically results in the suppression of canopy resistance in these regions.
 The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
 Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}).
-Accidently, we have used forcings for the year 2014 instead of 2005.
+Accidentally, we have used forcings for the year 2014 instead of 2005.
 Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
-As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
 A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).
 All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
-An x in the table denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.
+An x in the table denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.
 
@@
 For simplicity, only the tested parameters are listed.
-An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
+An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
   \begin{tabular}{lccccccccc}
     \tophline
-    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
+    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
     & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
     \middlehline
-    Wesely\_type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
+    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
     EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
@@
   \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
-  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
   \label{fig:diff_maps}
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone due to dry deposition.
+In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition.
 Surface ozone increases globally except for regions covered by tropical forest.
@@
 We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-A possible explanation is the huge difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities in the EMEP scheme.
+A possible explanation to these changes especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme.
 We cover this in more detail in the following section.
@@
 \label{subsec:model}
-Compare to \citet{ACP:Hardacre2015} as well as . \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
-
-\subsubsection{Zonal averages}
+%
+In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}.
+For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$.
+First (Section~\ref{subsubsec:zonal}), we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations.
+In Section~\ref{subsubsec:seasons}, we study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated.
+From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
+%
+\subsubsection{Zonal distribution}
+\label{subsubsec:zonal}
+%
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts.
+In the deep Tropics, the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
+On average, we find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere.
+Among the senistivity studies, there is only a low variance.
+Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres.
+Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) generally decreases the ozone concentration by about $2\,\unit{ppb}$, while taking deeper soil water content into consideration (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
+Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference.
+The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
+An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.
+
+In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ration between surface concentration and total dry deposition since we have no output of these.
+$v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme even smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
+In the remaining latitude bands, except for EMEP\_ppgssh\_ice, the EMEP scheme is constantly above the multi-model-mean.
+This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard.
 
@@
   \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
-  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) surface ozone concentrations; and in comparison with the multi-model evaluation by \citet{ACP:Hardacre2015}: (b) Annual average ozone dry depositon velocity; (c) Total annual ozone dry deposition.
-The different colors indicate sets of associated simulation.}
+  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
+The different colors indicate sets of simulation with similar baselines.
+The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
   \label{fig:mmm_drydep}
@@
   \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
-  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
+  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocities.}
   \label{fig:mmm_drydep_season}
@@
 \subsubsection{Average seasonal cycles}
+\label{subsubsec:seasons}
 %
@@
   \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
-  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition seperated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
+  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
 Comparison with results from \citet{ACP:Hardacre2015}.}
@@
 %
-As suggested by \citet{ACP:Hardacre2015}, we retrospectively seperate ozone dry deposition velocities with respect to surface types.
+As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
 Based on CML2 dynamic land surface, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from $22\,\unit{\%}$ only for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean.
@@
 %
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
-  \caption{Average ozone dry deposition velocitys seperated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
+  \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
   \label{fig:mmm_drydep_season_pft}
-\end{figure}
+\end{figure*}
 %
@@
 \begin{table*}[t]
-\caption{Total ozone dry deposition seperated into land, ocean, and ice contributions.}
+\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
 \begin{tabular}{lcccccccccccc}
 \tophline
-& & Ocean & & & Ice & & & Land & & & Total\\
+\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total}\\
 & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
-& & (Tg a-1) & & & (Tg a-1) & & & (Tg a-1) & & & (Tg a-1)\\
+& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})}\\
 \middlehline
-88.7 110.3 199.0 OsloCTM3: Wesely type\\
-67.1 84.7 151.8 OsloCTM3: EMEP\_full\\
-66.6 84.1 150.7 OsloCTM3: EMEP\_offLight\\
-67.1 84.6 151.7 OsloCTM3: EMEP\_offPhen\\
-68.1 86.1 154.2 OsloCTM3: EMEP\_SWVL4\\
-67.1 84.6 151.7 OsloCTM3: EMEP\_ppgs\\
-67.0 84.5 151.6 OsloCTM3: EMEP\_ppgssh\\
-66.5 84.8 151.3 OsloCTM3: EMEP\_ppgssh\_ice\\
-66.8 83.5 150.4 OsloCTM3: EMEP\_ppgs\_2005\\
-
-0.7 5.0 5.7 OsloCTM3: Wesely type\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_full\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_offLight\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_offPhen\\
-0.6 4.5 5.1 OsloCTM3: EMEP\_SWVL4\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_ppgs\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_ppgssh\\
-0.2 1.3 1.4 OsloCTM3: EMEP\_ppgssh\_ice\\
-0.6 4.4 5.0 OsloCTM3: EMEP\_ppgs\_2005\\
-
-346.1 153.5 499.6 OsloCTM3: Wesely type\\
-236.0 112.3 348.2 OsloCTM3: EMEP\_full\\
-267.8 129.8 397.6 OsloCTM3: EMEP\_offLight\\
-239.9 116.0 355.9 OsloCTM3: EMEP\_offPhen\\
-241.8 115.0 356.8 OsloCTM3: EMEP\_SWVL4\\
-238.7 112.6 351.3 OsloCTM3: EMEP\_ppgs\\
-243.9 119.4 363.3 OsloCTM3: EMEP\_ppgssh\\
-240.3 120.4 360.7 OsloCTM3: EMEP\_ppgssh\_ice\\
-237.8 110.0 347.8 OsloCTM3: EMEP\_ppgs\_2005\\
-
-612.4 347.9 960.2 OsloCTM3: Wesely type\\
-408.6 245.6 654.2 OsloCTM3: EMEP\_full\\
-451.3 267.2 718.6 OsloCTM3: EMEP\_offLight\\
-413.9 249.9 663.7 OsloCTM3: EMEP\_offPhen\\
-417.2 250.7 667.9 OsloCTM3: EMEP\_SWVL4\\
-411.4 246.0 657.4 OsloCTM3: EMEP\_ppgs\\
-419.4 253.6 673.0 OsloCTM3: EMEP\_ppgssh\\
-413.0 251.1 664.1 OsloCTM3: EMEP\_ppgssh\_ice\\
-409.2 240.7 649.9 OsloCTM3: EMEP\_ppgs\_2005\\
+Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 \\
+EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 \\
+EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 \\
+EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 \\
+EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 \\
+EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 \\
+EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 \\
+EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 \\
+EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 \\
 \bottomhline
@@
 \label{subsec:obs}
-The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/}). \citep{AE:Hu2017} -> need to look into this one...
+The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/})? \citep{AE:Hu2017} -> need to look into this one...
 
@@
   \caption{Ozone dry deposition velocities at different observation sites.
-Comparison between Weseley scheme and average result from all sensitivity tests to observations and multi-model-mean from \citet{ACP:Hardacre2015}.}
+Comparison between Wesely scheme and average result from all sensitivity tests to observations and multi-model-mean from \citet{ACP:Hardacre2015}.}
   \label{fig:mmm_drydep_stations}
@@
 dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017})
-
+\citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
 %\subsection{HEADING}

### 2018-11-20 DryDep.tex
Finished section about zonal distribution, started with seasonal cycles. Had to update a few plots. No spell-check so far...

@@
     \emph{if} \quad \text{GDAY = 0}  & \quad 0 \\
-    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding Tropics)} \\
+    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding tropics)} \\
     \emph{else}  & \quad 
@@
 A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations.
-To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS versions.
+To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS cycles.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
-The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}.
+The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1nc4}.
 CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
@@
 \label{subsec:sens}
-As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme has undergone a major revision.
-Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change.
-Therefor, the conducted sensitivity studies (9 in total) mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
+Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change.
+Therefor, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
 A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
@@
 During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
-Especially in Europe, this basically results in the suppression of canopy resistance in these regions.
+Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia.
 The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
@@
 For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$.
-First (Section~\ref{subsubsec:zonal}), we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations.
-In Section~\ref{subsubsec:seasons}, we study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated.
+In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}).
 From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts.
-In the deep Tropics, the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
-On average, we find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere.
-Among the senistivity studies, there is only a low variance.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts.
+In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
+We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
+Among the sensitivity studies, there is only a low variance.
 Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres.
-Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) generally decreases the ozone concentration by about $2\,\unit{ppb}$, while taking deeper soil water content into consideration (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
+Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
 Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference.
@@
 
-In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ration between surface concentration and total dry deposition since we have no output of these.
-$v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme even smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
-In the remaining latitude bands, except for EMEP\_ppgssh\_ice, the EMEP scheme is constantly above the multi-model-mean.
-This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard.
+In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ratio between surface concentration and total dry deposition since we have no direct output of these.
+The $v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme and about a factor of $3$ smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
+In the remaining latitude bands, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean.
+This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3.
+The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
+The closer to $1$ these factores are the higher do the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
+That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation.
+Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (see figure in Supplement~S.6.1).
+The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions.
+
+The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b).
+In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in any of the sensitivity studies.
+The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$.
+Consulting the old parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$.
+Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$).
+These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative change we see between the Wesely and the EMEP scheme.
+We further elaborate on this in the following section.
 
@@
 %
+In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month.
+Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
+There are two major phases apparent, NH and SH greening season.
+Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity.
+In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ.
+The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b).
+In general, the seasonal patters are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
+What differs most is the molding and intensitity of the NH peak dry deposition.Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme.
+Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer into October in the Wesely scheme.
+Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
+%
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
-  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocities.}
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
+  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
+  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
+The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
   \label{fig:mmm_drydep_season}
@@
 %
+To disentangle the contributions of different regions to the global ozone budget further, we will briefly look at different projections of seasonal cycles.
+
+In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown.
+As reference, we have again added the multi-model-mean by \citet{ACP:Hardacre2015}.
+As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics and in the SH.
+The highest ozone dry deposition is found in the tropics and amounts on average to the  peak levelof dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme.
+The average tropical ozone dry deposition in the Wesely scheme diverges by about $5\,\unit{Tg}$ in comparison to its NH maximum.
+The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH.
+While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the verious realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively.
+This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).
+
+%
 \begin{figure*}[t]
   \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
-  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
-Comparison with results from \citet{ACP:Hardacre2015}.}
+  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
+The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
   \label{fig:mmm_drydep_hem}

### 2018-11-20 DryDep_supplement.tex
Finished section about zonal distribution, started with seasonal cycles. Had to update a few plots. No spell-check so far...

@@
 
-\subsection*{S.6 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
+\subsection*{S.6 }
+\subsubsection*{S.6.1 Normed zonal average ozone dry deposition velocity}
+\appendixfigures
+\begin{center}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
+\end{center}
+
+\subsubsection*{S.6.2 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
 \appendixfigures

### 2018-11-21 DryDep.tex
Found some mistake in the analysis. Re-did the plots, but haven't had the nerve to go through the text again...

@@
 Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$.
-Consistently, dry deposition velocities decrease globally (up to $100\,\unit{\%}$) in desert regions, while they increase over tropical forest.
+Consistently, dry deposition velocities decrease globally by yhe same order of magnitude in these regions, while they increase over tropical forest.
 With respect to total dry deposition, the picture is a bit less clear.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with multi-model evaluation}
+\subsection{Comparison with modeling results}
 \label{subsec:model}
@@
 Among the sensitivity studies, there is only a low variance.
-Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres.
+Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
 Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
-Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference.
+Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference.
 The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
@@
 
-In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ratio between surface concentration and total dry deposition since we have no direct output of these.
-The $v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme and about a factor of $3$ smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
-In the remaining latitude bands, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean.
+Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
+\begin{equation}
+  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
+\end{equation}
+with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
+This way of computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
+We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
+The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
+The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
+In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean.
 This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3.
 The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
-The closer to $1$ these factores are the higher do the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
+The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (see figure in Supplement~S.6.1).
-The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions.
+Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (for surface tupe partitioning see figure in Supplement~S.6.1).
+The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions, and its counterpart in the northern hemisphere.
 
 The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b).
-In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in any of the sensitivity studies.
+In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in all sensitivity studies.
 The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$.
-Consulting the old parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$.
-Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$).
-These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative change we see between the Wesely and the EMEP scheme.
+Consulting the Wsely scheme parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
+Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
+These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative difference between the Wesely and the EMEP scheme.
 We further elaborate on this in the following section.
@@
 The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b).
-In general, the seasonal patters are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
-What differs most is the molding and intensitity of the NH peak dry deposition.Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme.
-Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer into October in the Wesely scheme.
+In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
+What differs most is the molding and intensitity of the NH peak dry deposition.
+Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme.
+Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme.
 Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
@@
 %
-To disentangle the contributions of different regions to the global ozone budget further, we will briefly look at different projections of seasonal cycles.
+To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.
 
 In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown.
-As reference, we have again added the multi-model-mean by \citet{ACP:Hardacre2015}.
-As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics and in the SH.
-The highest ozone dry deposition is found in the tropics and amounts on average to the  peak levelof dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme.
-The average tropical ozone dry deposition in the Wesely scheme diverges by about $5\,\unit{Tg}$ in comparison to its NH maximum.
+As reference, we have added the multi-model-mean by \citet{ACP:Hardacre2015}.
+As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface).
+The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme.
+In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum.
 The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH.
-While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the verious realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively.
+While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the various realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively.
 This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).
@@
 As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on CML2 dynamic land surface, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from $22\,\unit{\%}$ only for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean.
-Since we have not perform a full unfolding on the data, the results should be regarded with suspicion.
+Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $22\,\unit{\%}$ for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean.
+From this CLM~2 based map, it was not possible to select gridboxes with $100\,\unit{\%}$ for each surface type.
+Since we have not perform a full unfolding on the data, the results should be treated with slight caution.
+In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories.
+The patterns and absolute numbers differ between the schemes and the multi-model-mean.
+Exceptually large is the divergence for desert regions.
+The multi-model-mean displays a very pronounced seasonal cycle ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
+The Wesely scheme as well as the EMEP scheme do not indicate any seasonal cycle.
+The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition,  about one order of magnitude lower than in the EMEP scheme than in the Wesely scheme.
+From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}.
+This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme.
+Similarly, the dry deposition velocities over water differ.
+From measurements over the ocean a value of $v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}.
+In a comparison of various dry deposition mechanisms over ocean waters implemented in their model, \citet{ACP:Luhar2017} give a range of $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
 %
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
+  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
   \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
@@
 %
-The Wesely scheme uses $v^\chem{O_3}_\text{tundra/desert} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{desert} = 0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$.
-From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}.
-This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme.
-Similarly, the dry deposition over water $v^\chem{O_3}_\text{water}=0.07\,\unit{cm\,s^{-1}}$ in the Wesely scheme, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{water} = 0.05\,\unit{cm\,s^{-1}}$.
-$v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ \citep{JGR:Helmig2012}.
-\citet{ACP:Luhar2017} found $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
+
 %

### 2018-11-21 DryDep_supplement.tex
Found some mistake in the analysis. Re-did the plots, but haven't had the nerve to go through the text again...

@@
 
-\subsection*{S.6 }
-\subsubsection*{S.6.1 Normed zonal average ozone dry deposition velocity}
+\subsection*{S.6 Normed zonal average ozone dry deposition velocity}
+%\subsubsection*{S.6.1 }
 \appendixfigures
@@
 
-\subsubsection*{S.6.2 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
-\appendixfigures
-\begin{center}
-  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-total_ozone_drydepvelo_monthly_pft_sh}
-\end{center}
 \end{document}

### 2018-11-22 DryDep.tex
Almost finished comparison with mmm. Updated plot again.

@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$ (Supplement S.1).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement S.1).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$.
+This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
 In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
@@
 Among the sensitivity studies, there is only a low variance.
-Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
+Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
 Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
-Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference.
+Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference.
 The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
@@
 with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
-This way of computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
+This way of retrospective computing of dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
 We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
@@
 The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
-That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (for surface tupe partitioning see figure in Supplement~S.6.1).
-The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions, and its counterpart in the northern hemisphere.
+That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation.
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1).
+The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere.
+However, it is not clear whether these roughly $10\,\unit{\%}$ of desert in the SH in comparison with $80\,\unit{\%}$ being covered by ocean in the same zonal band could have such an impact.
 
 The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b).
-In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in all sensitivity studies.
-The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$.
-Consulting the Wsely scheme parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
+In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of te amount given by \emph{Wesely\_type}.
+The occurence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$.
+Consulting the parameter file used in the Wesely scheme, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
-These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative difference between the Wesely and the EMEP scheme.
+These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP schemebut may not sufficiently.
 We further elaborate on this in the following section.
@@
 Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity.
-In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ.
+In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ.
 The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b).
@@
 Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
+In comparison, the available two month of the multi-model-mean display a different pattern.
+Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
 %
@@
 
-In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown.
-As reference, we have added the multi-model-mean by \citet{ACP:Hardacre2015}.
+In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown.
+We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference.
 As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface).
@@
 The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH.
-While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the various realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively.
-This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).
+The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realisations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
+This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
 
@@
 As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $22\,\unit{\%}$ for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean.
-From this CLM~2 based map, it was not possible to select gridboxes with $100\,\unit{\%}$ for each surface type.
-Since we have not perform a full unfolding on the data, the results should be treated with slight caution.
+Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+From this CLM~2 based map, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type.
+Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
 In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories.
-The patterns and absolute numbers differ between the schemes and the multi-model-mean.
-Exceptually large is the divergence for desert regions.
-The multi-model-mean displays a very pronounced seasonal cycle ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
-The Wesely scheme as well as the EMEP scheme do not indicate any seasonal cycle.
-The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition,  about one order of magnitude lower than in the EMEP scheme than in the Wesely scheme.
-From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}.
-This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme.
+The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean.
+In our sensitivity study, the divergence of the average dry deposition velocities in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
+In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a very pronounced seasonal cycle in desert regions ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
+The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme.
+From measurements of ozone fluxes in the Sahara desert \citep{AE:Gusten1995}, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$ have been deduced.
+This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme.
 Similarly, the dry deposition velocities over water differ.
-From measurements over the ocean a value of $v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}.
-In a comparison of various dry deposition mechanisms over ocean waters implemented in their model, \citet{ACP:Luhar2017} give a range of $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
+From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}.
+In a model study of dry deposition mechanisms to ocean waters by means of prescribed $v_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
+With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
+With respect to forests, we might be able to improve the model performance further by allowing for the full range of generic tree types of the EMEP scheme, instead of using a broad average.
 %
@@
   \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
-  \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
+  \caption{Average seasonal cycles of ozone dry deposition velocities separated by land use type.
+Results from \citep{ACP:Hardacre2015} are shown as a reference.
+We refrain from showing the full extend of \emph{EMEP\_offLight} here, since it is an extreme scenario and has been discussed already.}
   \label{fig:mmm_drydep_season_pft}
@@
 
+Finally, we have a look at the different global as well as hemispheric sinks for ozone.
+Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink, while ice and snow account for far less than $1\,\unit{\%}$.
+The remainder is desposited to various other land surfaces of which deserts might be the most neglected in process modelling.
+The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}.
+But also the results of ~\citet{ACP:Luhar2017} yield a lower ozone dry deposition than the models participating in the model intercomparison.
 %
@@
 \label{sec:disc}
+
+{\bf Go through the finished text and resume/continue the started discussions in the evaluation section.}
 dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017})
 \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
+
+Come back to the dicussion of the multi-model-mean - ther seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
 %\subsection{HEADING}

### 2018-11-23 DryDep.tex
Finishes model-model comparison, started writing model-obs comparison - should probably get some satellite data...

@@
 Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$.
-Consistently, dry deposition velocities decrease globally by yhe same order of magnitude in these regions, while they increase over tropical forest.
+Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest.
 With respect to total dry deposition, the picture is a bit less clear.
@@
 For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$.
-In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}).
+In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}).
 From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
@@
 with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
-This way of retrospective computing of dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
+This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
 We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
@@
 Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1).
-The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere.
-However, it is not clear whether these roughly $10\,\unit{\%}$ of desert in the SH in comparison with $80\,\unit{\%}$ being covered by ocean in the same zonal band could have such an impact.
-
-The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b).
+The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
+However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.
+
+The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c).
 In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of te amount given by \emph{Wesely\_type}.
@@
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
-These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP schemebut may not sufficiently.
+These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP scheme.
 We further elaborate on this in the following section.
@@
 Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
-From this CLM~2 based map, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type.
+Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type.
 Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
@@
 The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean.
-In our sensitivity study, the divergence of the average dry deposition velocities in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
-In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a very pronounced seasonal cycle in desert regions ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
+The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
+In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
 The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme.
-From measurements of ozone fluxes in the Sahara desert \citep{AE:Gusten1995}, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$ have been deduced.
+From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
 This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme.
@@
 From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}.
-In a model study of dry deposition mechanisms to ocean waters by means of prescribed $v_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
+In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
 With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
-With respect to forests, we might be able to improve the model performance further by allowing for the full range of generic tree types of the EMEP scheme, instead of using a broad average.
+With respect to forests, we might be able to improve the model performance further by allowing more PFTs, instead of using a broad average.
 %
@@
 Finally, we have a look at the different global as well as hemispheric sinks for ozone.
-Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink, while ice and snow account for far less than $1\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
 The remainder is desposited to various other land surfaces of which deserts might be the most neglected in process modelling.
 The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}.
-But also the results of ~\citet{ACP:Luhar2017} yield a lower ozone dry deposition than the models participating in the model intercomparison.
+But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
 %
@@
 \label{subsec:obs}
-The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/})? \citep{AE:Hu2017} -> need to look into this one...
+In this section, we compare our results to observations. For all comparisons with observations, we use the original resolution of the OsloCTM3 instead of the re-grided.
+
+In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown.
+We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band.
+The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}.
+Although it is quite interesting to compare to the multi-model-mean, our focus lies on the performance of the OsloCTM3 with respect to the observational averages.
+At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
+We use a $\chi^2$-test to assess this subjectiv in a more objective way:
+\begin{equation}
+  \frac{\chi^2}{\text{NDF}} = \frac{\sum (\overline{\chem{O_3}}_\text{DD}^\text{sim}-\overline{\chem{O_3}}_\text{DD}^\text{obs})^2}{\text{NDF}}.
+\end{equation}
+The smaller the number, the better does the simulation represent the observation.
+The results shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}.
+The $\chi^2$-test reveals that in $5$ of $6$ cases the EMEP scheme performs better than then Wesely scheme, although a statisfying result is only achieved for two sites.
+With only one full year of simulation, the uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Many different aspects
+
+The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/})?
 
@@
   \caption{Ozone dry deposition velocities at different observation sites.
-Comparison between Wesely scheme and average result from all sensitivity tests to observations and multi-model-mean from \citet{ACP:Hardacre2015}.}
+Comparison between Wesely scheme and average result from all sensitivity studies to observations and multi-model-mean taken from \citet{ACP:Hardacre2015}.
+The model uncertainty of the OsloCTM3 is given as $1\,\sigma$ error band.
+The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.}
   \label{fig:mmm_drydep_stations}
@@
 {\bf Go through the finished text and resume/continue the started discussions in the evaluation section.}
-dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017})
+dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean?
+Although it is a low number there is $\frac{2}{3}$ ocean on earth...
+(Think, I read this in Hardacre paper...
+-> \citet{ACP:Luhar2017}).
+Well, most of it is already answered in the last part of the model to model evaluation.
 \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
@@
 Come back to the dicussion of the multi-model-mean - ther seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
+May also include \citep{AE:Hu2017} in this part of the discussion.
 %\subsection{HEADING}

### 2018-11-26 DryDep.tex
Finished section 3. May need to add comparison with satellite data. Updates plots.

@@
 We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively.
-We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
+We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
 The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1nc4}.
@@
 Biogenic emissions are taken from MEGAN output.
-In the following, we will present the various model sensitivity studies.
+In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-A possible explanation to these changes especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme.
+A possible explanation to the divergence especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme.
 We cover this in more detail in the following section.
@@
 %
-In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}.
-For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$.
-In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}).
-From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
+In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}.
+For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
+In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}).
+From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
 %
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
@@
 We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
-Among the sensitivity studies, there is only a low variance.
+Among the sensitivity studies, there is only a low absolute variance.
 Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
 Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
-Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference.
+Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics).
+Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$.
 The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
-An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.
+An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.
 
@@
 The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
-In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean.
+In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean.
 This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3.
 The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
-The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
+The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation.
@@
 The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c).
-In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of te amount given by \emph{Wesely\_type}.
-The occurence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$.
+In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}.
+The occurrence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$.
 Consulting the parameter file used in the Wesely scheme, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
@@
 These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP scheme.
-We further elaborate on this in the following section.
+We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).
 
@@
 %
-In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month.
+In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month.
 Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
-There are two major phases apparent, NH and SH greening season.
+There are two major phases apparent: NH and SH greening season.
 Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity.
@@
 In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
-What differs most is the molding and intensitity of the NH peak dry deposition.
-Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme.
+What differs most is the molding and intensity of the NH peak dry deposition.
+Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme.
 Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme.
@@
 
-In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown.
+In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown.
 We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference.
@@
 The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH.
-The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realisations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
+The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
 This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
@@
 The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
-In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$).
+In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
 The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme.
@@
 Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
-The remainder is desposited to various other land surfaces of which deserts might be the most neglected in process modelling.
+The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling.
 The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}.
@@
 \label{subsec:obs}
-In this section, we compare our results to observations. For all comparisons with observations, we use the original resolution of the OsloCTM3 instead of the re-grided.
+In this section, we compare our results to observations. For all comparisons with observations, we use the original resolution of the OsloCTM3 instead of the re-gridded.
 
@@
 At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
-We use a $\chi^2$-test to assess this subjectiv in a more objective way:
-\begin{equation}
-  \frac{\chi^2}{\text{NDF}} = \frac{\sum (\overline{\chem{O_3}}_\text{DD}^\text{sim}-\overline{\chem{O_3}}_\text{DD}^\text{obs})^2}{\text{NDF}}.
-\end{equation}
-The smaller the number, the better does the simulation represent the observation.
-The results shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}.
-The $\chi^2$-test reveals that in $5$ of $6$ cases the EMEP scheme performs better than then Wesely scheme, although a statisfying result is only achieved for two sites.
-With only one full year of simulation, the uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Many different aspects
-
-The sites used in \citet{ACP:Hardacre2015} and additional ozone zonal data from satellites and sondes (\url{http://bodekerscientific.com/})?
+We use a $\chi^2$-test
+\begin{equation}
+  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
+\end{equation}
+with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
+The closer to $1$ this test scores, the better does the simulation represent the observation.
+A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
+The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}.
+The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+In the case of Castel Porziano, Auchencoth Moss, and Citrus Orchard, there may be a hint to an underlying model bias of a too small dry deposition of about $2.8\,\unit{mmol\,m^{-2}\,s^{-1}}$.
+With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed.
+Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data.
+Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role.
+In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
+
+{\bf Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
 
@@
   \includegraphics[width=8.3cm]{pictures/final-monthly_drydep_divergence_True}
-  \caption{Ozone dry deposition velocities at different observation sites.
-Comparison between Wesely scheme and average result from all sensitivity studies to observations and multi-model-mean taken from \citet{ACP:Hardacre2015}.
+  \caption{Ozone dry deposition fluxes at different observation sites.
+(a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
 The model uncertainty of the OsloCTM3 is given as $1\,\sigma$ error band.
-The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.}
+The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
+(b) Model divergence from observation and $\chi^2$-test results.}
   \label{fig:mmm_drydep_stations}
@@
 
-Come back to the dicussion of the multi-model-mean - ther seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
+Come back to the discussion of the multi-model-mean - there seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
 May also include \citep{AE:Hu2017} in this part of the discussion.

### 2018-11-27 DryDep.tex
Wrote the conclusions, but not sure what should go to discussions... corrected some typos.

@@
 Parallelization of the OsloCTM3 has not been optimized as of now.
-While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, it is very time and memory consuming to run the CTM with full chemistry at this resolution.
+While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, with the highest model level at $0.1\,\unit{hPa}$, it is very time and memory consuming to run the CTM with full chemistry at this resolution.
 Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$.
@@
 Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
-For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
 From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.
@@
 Water bodies and snow and ice categories of MODIS are actually not mapped.
-For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
 From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
@@
 In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}.
-The occurrence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$.
-Consulting the parameter file used in the Wesely scheme, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
+The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a change in $v_\text{desert}^\chem{O_3}$.
+Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
@@
 Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity.
-In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ.
+In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ.
 The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b).
@@
 Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
-Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type.
+Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
 Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
@@
 The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean.
-The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
+The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
 In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
@@
 \label{subsec:obs}
-In this section, we compare our results to observations. For all comparisons with observations, we use the original resolution of the OsloCTM3 instead of the re-gridded.
+In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 instead of the re-gridded.
 
@@
 The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}.
-Although it is quite interesting to compare to the multi-model-mean, our focus lies on the performance of the OsloCTM3 with respect to the observational averages.
 At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
@@
 
-{\bf Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
+{\bf TODO: Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
 
@@
 \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
-
+Resolution and representativeness of site
 Come back to the discussion of the multi-model-mean - there seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
@@
 \label{sec:conc}
-What did we gain?
+We have presented an update of the dry deposition scheme in the OsloCTM3 from a purely prescribed parameterization based on the works of \citet{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.
+In general, this update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
+Most of this decrease is due to changes in the baseline dry deposition velocities at the order of one magnitude in case of desert and grasslands or the accumulative effect of small changes over the ocean in the southern hemisphere.
+With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the OsloCTM3 may slightly underestimate ozone dry deposition in the former case and overestimate it in the latter.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile studying dry deposition in these regimes at a more process-based level (e.g., 2-layer gas exchange with ocean waters).
+Although ice and snow only amounts to less than $1\,\unit{\%}$ to the global annual ozone dry deposition, a change in dry deposition velocities not only almost doubles the surface ozone concentrations in the high Arctics but also affects surface ozone down to latitudes at about $50\,\unit{^\circ}$.
+In comparison to multi-model-mean results, we conclude that it is important to use an updated ozone dry deposition to ice and snow \citep{ACP:Helmig2007}.
+We have extensively studied the parameter space of the stomatal conductance parameterization and found the largest sensitivity of total surface ozone in the tropics.
+This may indicate that changes in solar radiation due to changes cloud and aerosol may also strongly influence surface ozone in the tropics.
+In the most extreme test scenario, allowing for an unlimited photosynthesis, the increase in surface ozone amounts to up to $16\,\unit{\%}$ compared to the reference experiment.
+Switching off the phenology in the tropics only increases ozone there by up to $10\,\unit{\%}$, while the complete switch-off amounts only to a maximum increase of $5\,\unit{\%}$.
+Total surface ozone increases by about $7.7\,\unit{\%}$ in the northern hemisphere ($(50-75)\,\unit{^\circ N}$) if a greening season based on a $5\,\unit{^\circ C}$-days criteria is used.
+An important factor in the global ozone budget are also emissions of precursor substances.
+In a comparison between 2005 and 2014 emissions using the same meteorology, we found a decrease in total surface ozone up to $5\,\unit{\%}$ in the tropics.
+We also evaluated the model with observed dry deposition velocities at six sites in the northern hemisphere and found that the updated dry deposition scheme performs better than the old one, but is not able to reproduce the measurements at most sites well.
+This may be due to several reasons.
+The rather low model resolution in both horizontal and vertical affects the distribution and transport (e.g., long-range, stratosphere--troposphere exchange) as well as the regional production of ozone.
+Depending on the location of the observation site and its respective representativeness for a larger area ozone dry deposition and ozone concentrations will be over- or underestimated.
+In addition, a comparison of very few years of measurement to only one specific year of simulation may reflect the year to year variability more than the actual model performance.
+For further diagnostics, we shall make dry deposition velocities directly available in the model output.
 
@@
 
-\codedataavailability{TEXT} %% use this section when having data sets and software code available
+\codedataavailability{The OsloCTM3 is made publicly available on git-hub under a MIT license.
+Model results can be made available under request.} %% use this section when having data sets and software code available
 
@@
   \centering
-  \includegraphics[width=8.3cm]{pictures/pft_categories.pdf}\\
-  \includegraphics[width=8.3cm]{pictures/pft_categories_zonal.pdf}
-  \caption{Plant functional types (PFT). (a) CML2 dynamic land surface $(0.5\times0.5)\,\unit{^\circ}$; (b) Zonal distribution of land surface types.}
+  \includegraphics[width=0.8\textwidth]{pictures/pft_categories.pdf}\\
+  \includegraphics[width=0.8\textwidth]{pictures/pft_categories_zonal.pdf}
+  \caption{Partitioning of land surface types. (a) CML2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}
   \label{fig:pft_landsurface}
@@
 
-\authorcontribution{Stefanie Falk wrote the paper, did the data analysis, and finalized the implementation of the stomatal conductance in the dry deposition scheme.
-Amund S{\o}vde Haslerud did most of the implementation and documentation of the updated dry deposition scheme.} %% it is strongly recommended to make use of this section
+\authorcontribution{Stefanie Falk has finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the OsloCTM3, conducted the tests, and evaluated the results.
+Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the documentation.
+Both authors contributed to the writing of the paper.} %% it is strongly recommended to make use of this section
 

### 2018-11-28 DryDep.tex
Finished discussion and clonclusions. Decided to pull these two section together.

@@
 In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}).
-In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
+In Section\ref{sec:conc}, we will summarize and discuss our results and draw conclusions for further development of the model.
 {\bf TODO: Precise outlook on paper when writing is done.}
@@
 Finally, we have a look at the different global as well as hemispheric sinks for ozone.
-Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
 The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling.
@@
 In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
-
-{\bf TODO: Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
-
+%
 \begin{figure}[t]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{sec:disc}
-
-{\bf Go through the finished text and resume/continue the started discussions in the evaluation section.}
-dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean?
-Although it is a low number there is $\frac{2}{3}$ ocean on earth...
-(Think, I read this in Hardacre paper...
--> \citet{ACP:Luhar2017}).
-Well, most of it is already answered in the last part of the model to model evaluation.
-\citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
-Resolution and representativeness of site
-Come back to the discussion of the multi-model-mean - there seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones.
-May also include \citep{AE:Hu2017} in this part of the discussion.
-%\subsection{HEADING}
-%\subsubsection{HEADING}
-
-\conclusions  %% \conclusions[modified heading if necessary]
+{\bf TODO: Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from a purely prescribed parameterization based on the works of \citet{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.
-In general, this update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
-Most of this decrease is due to changes in the baseline dry deposition velocities at the order of one magnitude in case of desert and grasslands or the accumulative effect of small changes over the ocean in the southern hemisphere.
-With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the OsloCTM3 may slightly underestimate ozone dry deposition in the former case and overestimate it in the latter.
-Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile studying dry deposition in these regimes at a more process-based level (e.g., 2-layer gas exchange with ocean waters).
-Although ice and snow only amounts to less than $1\,\unit{\%}$ to the global annual ozone dry deposition, a change in dry deposition velocities not only almost doubles the surface ozone concentrations in the high Arctics but also affects surface ozone down to latitudes at about $50\,\unit{^\circ}$.
-In comparison to multi-model-mean results, we conclude that it is important to use an updated ozone dry deposition to ice and snow \citep{ACP:Helmig2007}.
-We have extensively studied the parameter space of the stomatal conductance parameterization and found the largest sensitivity of total surface ozone in the tropics.
-This may indicate that changes in solar radiation due to changes cloud and aerosol may also strongly influence surface ozone in the tropics.
-In the most extreme test scenario, allowing for an unlimited photosynthesis, the increase in surface ozone amounts to up to $16\,\unit{\%}$ compared to the reference experiment.
-Switching off the phenology in the tropics only increases ozone there by up to $10\,\unit{\%}$, while the complete switch-off amounts only to a maximum increase of $5\,\unit{\%}$.
-Total surface ozone increases by about $7.7\,\unit{\%}$ in the northern hemisphere ($(50-75)\,\unit{^\circ N}$) if a greening season based on a $5\,\unit{^\circ C}$-days criteria is used.
-An important factor in the global ozone budget are also emissions of precursor substances.
-In a comparison between 2005 and 2014 emissions using the same meteorology, we found a decrease in total surface ozone up to $5\,\unit{\%}$ in the tropics.
-We also evaluated the model with observed dry deposition velocities at six sites in the northern hemisphere and found that the updated dry deposition scheme performs better than the old one, but is not able to reproduce the measurements at most sites well.
+We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
+Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
+However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models are decreasing ozone dry deposition \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+
+Most of this decrease is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of small changes over the ocean which is especially important in the southern hemisphere.
+With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition over the former and overestimate it over the latter.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust and wind drag.
+
+Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$.
+In a comparison with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model.
+
+We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants.
+This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
+Tropospheric total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increases.
+The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.
+
+In the northern hemisphere mid and high latitudes ($(50-75)\,\unit{^\circ N}$), total surface ozone increases by about $7.7\,\unit{\%}$ if the beginning and end of the vegetation period is estimated based on a $5\,\unit{^\circ C}$-days criteria instead of prescribed.
+This is very important for any study focusing on ozone in the boreal and subarctic regions.
+
+An important factor in the global ozone budget are emissions of precursor substances.
+We cover this by using the same meteorology with different years of CEDS emissions.
+We chose the years 2005 and 2014 for our comparison.
+In 2014, surface ozone burden is higher in the southern hemisphere and in the tropics ($\sim 5\,\unit{\%}$) compared to 2005, while it is lower in the northern hemisphere ($\sim 2\,\unit{\%}$).
+This is most likely reflecting the ongoing industrialization process of countries in the southern hemisphere and the commitment and implementation of air quality regulations of industrialized nations in the northern hemisphere.
+
+We also evaluated the model with respect to observed dry deposition velocities at six sites in the northern hemisphere and found that the updated dry deposition scheme performs better than the old one, but is not able to reproduce the measurements at most sites quantitatively.
 This may be due to several reasons.
-The rather low model resolution in both horizontal and vertical affects the distribution and transport (e.g., long-range, stratosphere--troposphere exchange) as well as the regional production of ozone.
-Depending on the location of the observation site and its respective representativeness for a larger area ozone dry deposition and ozone concentrations will be over- or underestimated.
+The rather low model resolution in both horizontal ($2.25^\circ\times 2.25^\circ$) and vertical (L60, $P_\text{max}=0.1\,\unit{hPa}$) affects the distribution and transport (e.g., long-range, convection, and stratosphere--troposphere exchange) of ozone and its precursors.
+Depending on the location of the observation site and its respective representativeness for a larger area, ozone dry deposition and ozone concentrations are expected to be over- or underestimated in the model.
 In addition, a comparison of very few years of measurement to only one specific year of simulation may reflect the year to year variability more than the actual model performance.
-For further diagnostics, we shall make dry deposition velocities directly available in the model output.
+
+Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
+The model performance could also be improved by allowing for more plant fictionally types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
+The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
 

### 2018-11-29 DryDep.tex
Fixed some formulations to make things clearer.

@@
 \subsubsection*{Mapping of land surface types}
-Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
+Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
@@
   \includegraphics[width=12cm]{pictures/pft_landuse_mapping.pdf}
-  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM) 2 to the nine land surface types used in the EMAP scheme.
+  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 to the nine land surface types used in the EMAP scheme.
 Water bodies and snow and ice categories of MODIS are actually not mapped.
@@
 In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}.
-For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
+For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
 In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}).
@@
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
+We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.
+The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
@@
 
-Most of this decrease is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of small changes over the ocean which is especially important in the southern hemisphere.
-With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition over the former and overestimate it over the latter.
-Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust and wind drag.
+Most of this decrease in the OsloCTM3 is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of a small change in dry deposition to the ocean which is especially amplified in the southern hemisphere.
+With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate it over the latter.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust.
 
@@
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
-Tropospheric total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increases.
+Total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increase.
 The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.
@@
 Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
-The model performance could also be improved by allowing for more plant fictionally types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
+The model performance could also be improved by allowing for more plant fuctional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
 The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.

### 2018-12-03 DryDep.tex
Suggestions by Frode - Part I.

@@
 
+As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
+Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
+Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
+
 \subsection{Dry deposition scheme for other gases and aerosols}
@@
   \includegraphics[width=12cm]{pictures/pft_landuse_mapping.pdf}
-  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 to the nine land surface types used in the EMAP scheme.
+  \caption{Mapping of land surface categories.
+Either land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types used in the EMAP scheme.
 Water bodies and snow and ice categories of MODIS are actually not mapped.
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
-From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
+From MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
   \label{fig:pft_mapping}
@@
 Since this field is already integrated presumably over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
-Therefor, we will refer to the de-accumulated field as PPFD.
+Therefore, we will refer to the de-accumulated field as PPFD.
 An example PAR for January 2 2005 is shown in Supplement~S.5.1.
@@
 A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations.
-To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS cycles.
+To counter this, we decided to refrain from the division at this stage, but need to bear this in mind for later OpenIFS cycles.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change.
-Therefor, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
+Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
 A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
@@
 Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia.
-The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
+The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
 Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}).
-Accidentally, we have used forcings for the year 2014 instead of 2005.
+Accidentally, we have used emissions for the year 2014 instead of 2005.
 Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
@@
     \tophline
-    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
+    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Emissions\\
     & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
@@
 From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
+
+Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
+\begin{equation}
+  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
+\end{equation}
+with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
+This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
+We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
 %
@@
 
-Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
-\begin{equation}
-  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
-\end{equation}
-with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
-This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
-We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
-The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
+The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
 The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
@@
 This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3.
-The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
+The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
 The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
@@
 There are two major phases apparent: NH and SH greening season.
-Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity.
+Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer).
+Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity.
 In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ.
@@
 %
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
-  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
-The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
-  \label{fig:mmm_drydep_season}
-\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
-This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
+This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
 
@@
 %
-As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
-Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
-Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
 In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories.
@@
 
-{\bf TODO: Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.
+We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
+The dry deposition to vegetation is now senisitive to parameters such as photosyntheiscal active radiation, soil water content, moisture, and plant phenology.
 The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
@@
 \appendix
-\section{Land use and plant function types (PFT)}    %% Appendix A
+\section{Figures}    %% Appendix A
 \appendixfigures
@@
   \includegraphics[width=0.8\textwidth]{pictures/pft_categories_zonal.pdf}
-  \caption{Partitioning of land surface types. (a) CML2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}
+  \caption{Partitioning of land surface types. (a) CLM~2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}
   \label{fig:pft_landsurface}
 \end{figure*}
+
+\appendixfigures
+\begin{figure}[!htbp]
+  \includegraphics[width=0.8\textwidth]{pictures/final-total_ozone_drydepvelo_monthly}\\
+  \includegraphics[width=0.8\textwidth]{pictures/final-total_ozone_drydep_monthly}\\
+  \includegraphics[width=0.6\textwidth]{pictures/legend}
+  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
+The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
+  \label{fig:mmm_drydep_season}
+\end{figure}
 %\subsection{}     %% Appendix A1, A2, etc.
 
-%\noappendix       %% use this to mark the end of the appendix section
+\noappendix       %% use this to mark the end of the appendix section
 

### 2018-12-04 DryDep.tex
Finished Part II of comments. But need to think about drydep velocity computation again.

@@
 
-\title{Update of the ozone dry deposition in the OsloCTM3 v1.0}
+\title{Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0}
 
@@
 The impact of ozone damage on vegetation and agricultural plants in combination with advancing climate change may affect food security in the future.
-While the future scenarios in themselves are uncertain, there are limiting factors constraining the accuracy of surface ozone modeling also at present: The distribution and amount of ozone precursor/depleting substances, the stratosphere-troposphere exchange as well as scavenging processes.
+While the future scenarios in themselves are uncertain, there are limiting factors constraining the accuracy of surface ozone modeling also at present: The distribution and amount of ozone precursors and ozone depleting substances, the stratosphere-troposphere exchange as well as scavenging processes.
 Removal of any substance through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
@@
 Examining the sensitivity of the scheme to various parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
-We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies and also estimate the impact on the modeled ozone concentrations, both at the surface and in the free troposphere.
-We show that ozone dry deposition decreases with respect to the previous version, leading to an increase in surface ozone of up to $100\,\unit{\%}$.
-{\bf TODO: Finish the results section as soon as there are results.}
+We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies and also estimate the impact on the modeled ozone concentrations at the surface.
+We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-47\,\unit{\%}$), leading to an increase in surface ozone of up to $100\,\unit{\%}$.
+While high sensitivity to changes in dry deposition to vegetation is found in the tropics, the largest impact on global scales is associated to changes in dry deposition to the ocean and deserts.
 \end{abstract}
@@
 
-As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
-Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
-Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
-
 \subsection{Dry deposition scheme for other gases and aerosols}
@@
 The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
-This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
-
+This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in long-wave radiative transfer (e.g., due changes in cloud cover or aerosol).
+
+As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
+Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
+Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
 %
@@
 \caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
-\begin{tabular}{lcccccccccccc}
+\begin{tabular}{lccccccccccccr}
 \tophline
-\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total}\\
+\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
 & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
-& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})}\\
+& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
 \middlehline
-Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 \\
-EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 \\
-EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 \\
-EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 \\
-EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 \\
-EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 \\
-EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 \\
-EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 \\
-EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 \\
+Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
+EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
+EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
+EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
+EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
+EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
+EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
+EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
+EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
 \bottomhline
 \end{tabular}
-%\belowtable{} % Table Footnotes
+\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.} % Table Footnotes
 \end{table*}
@@
 We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
-The dry deposition to vegetation is now senisitive to parameters such as photosyntheiscal active radiation, soil water content, moisture, and plant phenology.
+The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology.
 The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
-However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models are decreasing ozone dry deposition \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
-
-Most of this decrease in the OsloCTM3 is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of a small change in dry deposition to the ocean which is especially amplified in the southern hemisphere.
-With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate it over the latter.
+However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+
+Most of the decrease in ozone dry deposition in the OsloCTM3 is due to changes in dry deposition velocities over the ocean and deserts.
+In case of desert and grasslands this change is at the order of one magnitude.
+Over the ocean it is due to the accumulative effect of a small absolute change in dry deposition which is especially amplified in the southern hemisphere.
+With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
 Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust.
 
-Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$.
-In a comparison with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model.
+Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
+By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model.
 
@@
 Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
-The model performance could also be improved by allowing for more plant fuctional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
+The model performance could also be improved by allowing for more plant functional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
 The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
@@
 
-%\clearpage
+\clearpage
 
@@
   \includegraphics[width=0.8\textwidth]{pictures/final-total_ozone_drydep_monthly}\\
-  \includegraphics[width=0.6\textwidth]{pictures/legend}
+  \includegraphics[width=0.7\textwidth]{pictures/legend}
   \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.

### 2018-12-05 DryDep.tex
Updated dry deposition velocity calculation and corresponding number in the text. Need to go through the whole of it again.

@@
 
-Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
-\begin{equation}
-  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
-\end{equation}
-with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$.
-This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}.
-We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
+For each gridbox, average dry deposition velocities have been reconstructed from the ratio between the monthly averaged  surface ozone concentration and total ozone dry deposition 
+\begin{equation}
+  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month}.
+\end{equation}
+Herein $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$ with the average height of the lowermost model level in each gridbox $\Delta h_\text{month}$, which is changing with temperature and surface pressure, and the respective number of seconds in a month $s_\text{month}$.
 %
@@
 Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics).
-Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$.
+Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $(54-70)\,\unit{^\circ N}$.
 The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
@@
 The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
-The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$.
+The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ smaller than in the Wesely scheme.
 In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean.
@@
 The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
-That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation.
+That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
 Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1).
@@
 The different colors indicate sets of simulation with similar baselines.
-The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
+The multi-model mean from the evaluation of TF~HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
   \label{fig:mmm_drydep}
@@
 %
-In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month.
+In Appendix Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month.
 Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
@@
 Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer).
-Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity.
+Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity and a slight increase of $v_\text{DD}^\chem{O_3}$ in the region $(20-40)\,\unit{^\circ S}$.
 In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ.
-The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b).
+The corresponding total monthly ozone dry deposition is shown in Appendix Fig.~\ref{fig:mmm_drydep_season}b).
 In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
@@
 The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean.
-The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$).
+The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$).
 In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
@@
 From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
-This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme.
+This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme.
 Similarly, the dry deposition velocities over water differ.
@@
 In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
-With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
-With respect to forests, we might be able to improve the model performance further by allowing more PFTs, instead of using a broad average.
+With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
+With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs, especially in regions covered by tropical or coniferous forest.
 %
@@
 
-Finally, we have a look at the different global as well as hemispheric sinks for ozone.
+Finally, we take a look at the different global as well as hemispheric sinks for ozone.
 Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
@@
 The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
-In the case of Castel Porziano, Auchencoth Moss, and Citrus Orchard, there may be a hint to an underlying model bias of a too small dry deposition of about $2.8\,\unit{mmol\,m^{-2}\,s^{-1}}$.
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed.
@@
 In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
+
 %
@@
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
-However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
 
@@
   \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
-The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
+The multi-model mean from the evaluation of TF~HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
   \label{fig:mmm_drydep_season}

### 2018-12-06 DryDep.tex
Rearangement of supplement. Minor changes in the text. Almost finished proofreading.

@@
 In this paper, we present an update of the dry deposition scheme implemented in the OsloCTM3.
-We change from an empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere into account.
+We change from an empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere and vegetation into account.
 Examining the sensitivity of the scheme to various parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
@@
 %\end{reaction}
-and dry deposition.
-We will come back to the latter later in this section and cover the implemented scheme in more detail in Section~\ref{subsec:DryDep}.
+and dry deposition. We will come back to the latter later in this section and cover the implemented scheme in more detail in Section~\ref{subsec:DryDep}.
+
 A dry deposition related sink limited to Arctic regions are so called ozone depleting events.
@@
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
-Dry deposition velocities $v^i_\text{DD}$, which later on are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
+Species dependent dry deposition velocities $v^i_\text{DD}$, which are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
 \begin{equation}
@@
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.
-Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
+Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as for single plant species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
 But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
@@
 In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
-In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}).
-In Section\ref{sec:conc}, we will summarize and discuss our results and draw conclusions for further development of the model.
-{\bf TODO: Precise outlook on paper when writing is done.}
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations at the surface (Section~\ref{subsec:obs}).
+In Section~\ref{sec:conc}, we will summarize and discuss our results and draw conclusions for further development of the model.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Parallelization of the OsloCTM3 has not been optimized as of now.
-While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, with the highest model level at $0.1\,\unit{hPa}$, it is very time and memory consuming to run the CTM with full chemistry at this resolution.
-Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$.
+While the meteorological data driving the OsloCTM3 is given in a resolution of T159N80L60, with the highest model level at $0.1\,\unit{hPa}$, it is very time and memory consuming to run the OsloCTM3 with full chemistry at this resolution.
+Therefore, we reduced the horizontal resolution to $2.25^\circ\times2.25^\circ$ in our experiments.
 In the following, we will give a detailed account of the new dry deposition scheme.
@@
 In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA).
+
 As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different resistance contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy $R^i_c$) which we will recap in the following.
@@
 with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
-The used ratios $D_\chem{H_2O}/D_i$ are displayed in Table~\ref{tab:diffusivity}.
-%
-\begin{table}[t]
-  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
-  \begin{tabular}{lcrr@{$\cdot$}l}
-    \tophline
-    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
-    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
-    \middlehline
-    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
-    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
-    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
-    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
-    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
-    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
-    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
-    \bottomhline
-  \end{tabular}
-  %\belowtable{} % Table Footnotes
-  \label{tab:diffusivity}
-\end{table}
+The used ratios $D_\chem{H_2O}/D_i$ are tabulated in Supplement~S.1.
+%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement S.1).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.2: Figs.~1--2).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 The allowed maximum roughness length in both cases is set to $2\,\unit{mm}$.
-Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement S.1).
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.2: Fig.~3).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various land surface types.
-All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or in Supplement~S.2.
+All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or tabulated in Supplement~S.3.
 Since $f_T$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
@@
 \end{equation}
-\citet{ACP:Simpson2012} suggest using SW in soil depths below $1\,\unit{m}$ (SWVL4 in OpenIFS).
+\citet{ACP:Simpson2012} suggest using SW of soil depths below $1\,\unit{m}$ (SWVL4 in OpenIFS).
 Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance more sensitive to precipitation than it would be otherwise.
@@
 The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}).
-If GLEN is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
+If GLEN is zero we are, e.g., in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
 Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$.
@@
 Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (Fig.~S.2).
+For four example PFTs, results are shown in the Supplement (S.4: Fig.4).
 
 \subsubsection*{Mapping of land surface types}
-Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
+The OsloCTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
@@
 In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we use the SMOKE-BEIS parameterization.
-The described algorithm written in python 2.7 has been included as Supplement~S.4.1.
+The described algorithm written in python~2.7 has been included as Supplement~S.5.
 An example map of the computed GLEN using the $5\,\unit{^\circ C}$-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
@@
 Therefore, we will refer to the de-accumulated field as PPFD.
-An example PAR for January 2 2005 is shown in Supplement~S.5.1.
-De-Accumulation of hour 3 till 21, simply means
+An example PAR for January 2nd 2005 is shown in Supplement~S.6 (Fig.~5).
+De-Accumulation of hour 3 till 21, simply means computing
 \begin{equation}
@@
 \end{equation}
-If everything was right in the OpenIFS output, we would not need to do any further processing, however the first time step of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately de-accumulate this time step.
-The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
+If everything was right in the OpenIFS output, we would not need to do any further processing, however the first time step of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (Supplement~S.6: Fig.~6), so it is impossible to accurately de-accumulate this time step.
+The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.6: Fig.~7)
 %
@@
 Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
-As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+As mentioned previously, the simulations were initially conducted with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
 A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).
@@
 The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
-The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ smaller than in the Wesely scheme.
+The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ lower than in the Wesely scheme.
 In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean.
@@
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1).
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.7: Fig.~8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
@@
 
-Finally, we take a look at the different global as well as hemispheric sinks for ozone.
+Finally, we take a look at the different global as well as hemispheric sinks for ozone (Table~\ref{tab:ozone_sinks}).
 Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
@@
 \end{tabular}
-\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.} % Table Footnotes
+\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
+\label{tab:ozone_sinks}
 \end{table*}

### 2018-12-06 DryDep_supplement.tex
Suggestions by Frode - Part I.
Rearangement of supplement. Minor changes in the text. Almost finished proofreading.

@@
 \documentclass[manuscript]{copernicus}
+\usepackage{graphicx}
+\usepackage{subcaption}
+%\usepackage{subfigure}
 \graphicspath{{pictures/}}          % graphics
@@
 \setcounter{page}{1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Supplement}
+\subsection*{S.1 Aerodynamical resistance}
+\begin{table}[!htbp]
+  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
+  \begin{tabular}{lcrr@{$\cdot$}l}
+    \tophline
+    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
+    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
+    \middlehline
+    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
+    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
+    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
+    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
+    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
+    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
+    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
+    \bottomhline
+  \end{tabular}
+  %\belowtable{} % Table Footnotes
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
+  \caption{Sutherland's law and divergence linear fits with one or two degrees of freedom.
+The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3.
+Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is neglicible.}
+\end{figure}
+%
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
+  \caption{Comparison of resulting $z_o$ and $R_b$ for $\mu(T) = m\cdot T$  (left panel) and Sutherland's law (right panel).
+The example displays the distributions for \chem{H_2O} after one day of model integration for both, rough and calm sea case.}
+\end{figure}
 
-\section*{Supplement}
-\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
-\subsubsection*{S.1.1 Sutherland's law and linear fits}
-\appendixfigures
-\begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
-\end{center}
-\subsubsection*{S.1.2 Comparison of $z_o$ and $R_b$ for $\mu(T) = m\cdot T$ and Sutherland's law: \chem{H_2O}}
-\appendixfigures
-\begin{center}
-  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
-\end{center}
-\subsubsection*{S.1.3 $R_b$ for different species}
-\appendixfigures
-\begin{center}
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}[b]{1.\textwidth}
     \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_1.pdf}
+    \caption{}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{1.\textwidth}
     \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_2.pdf}
-\end{center}
-
-%\begin{minipage}[t]{1.\textwidth}
-\subsection*{S.2 Tabulated parameters for stomatal conductance}
-\appendixtables
-\begin{sidewaystable*}[!htbp]
-  %\caption{S.2 Tabulated parameters for stomatal conductance}
+    \caption{}
+  \end{subfigure}
+  \caption{Resulting $R_b$ for different species.
+Shown are results of one day model integrations for both, Southerland's law and $\mu(T) = m\cdot T$.
+Shown are also the percentages above / below the given threashold.}
+\end{figure}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{S.3 Stomatal conductance}
+\begin{table*}[!htbp]
+  \caption{Tabulated parameters for stomatal conductance computation in the EMEP scheme}
   \begin{tabular}{lcccccccccccccc}
@@
   \vspace{2\baselineskip}
-  \begin{tabular}{lcccccccccccccc}
+  \begin{tabular}{lcccccccccc}
     \tophline
-    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$ & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
-    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\
+    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$\\% & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
+    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})\\%&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\
     \middlehline
-    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0 & 5 & 5 & 1 & 1\\
-    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 & 0 & 4 & 20 & 30\\
-    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
-    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
-    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57 & 0 & 3.5 & 70 & 20\\
-    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57 & 0 & 3 & 70 & 44\\
-    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0 & 0 & 4.2 & 35 & 65\\
-    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0 & 2 & 3 & 192 & 96\\
-    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0 & 2 & 3.5 & 140 & 135\\
-    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0 & 2.5 & 2.5 & 1 & 1\\
-    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
-    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
-    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
-    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
-    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
-    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
+    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0\\% & 5 & 5 & 1 & 1\\
+    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 \\%& 0 & 4 & 20 & 30\\
+    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0\\% & 4 & 4 & 1 & 1\\
+    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 \\%& 4 & 4 & 1 & 1\\
+    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57\\% & 0 & 3.5 & 70 & 20\\
+    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57\\% & 0 & 3 & 70 & 44\\
+    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0\\% & 0 & 4.2 & 35 & 65\\
+    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 2 & 3 & 192 & 96\\
+    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0\\% & 2 & 3.5 & 140 & 135\\
+    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0\\% & 2.5 & 2.5 & 1 & 1\\
+    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
     \bottomhline
@@
   \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean scrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
-\end{sidewaystable*}
-%\end{minipage}
+\end{table*}
 
-\subsection*{S.3 Latitude dependent vegetation height}
-\appendixfigures
-\begin{center}
+\clearpage
+\subsection*{S.4 Latitude dependent vegetation height}
+\begin{figure}[!htbp]
+  \centering
   \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
-\end{center}
+  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global describtion.
+Example heights are shown for coniferous, decideous, needleleaf, and broadleaf forests as well as Mediterranean shrub.
+As reference the original range of EMEP is added.}
+\end{figure}
 
-\subsection*{S.4 Temperature dependent greening season}
-\subsubsection*{S.4.1 Python code snippets}
+\subsection*{S.5 Temperature dependent greening season (python 2.7)}
+{\scriptsize
 \begin{verbatim}
@@
 \end{verbatim}
-%\subsubsection*{S.4.2 GLEN: NH mid--high latitudes only}
-%\appendixfigures
-%\begin{center}
-%  \includegraphics[height=0.4\textheight, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005.pdf}
-%\end{center}
+}
+\clearpage
 
-\subsection*{S.5 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS: January 2nd 2005}
-\subsubsection*{S.5.1 Output from OpenIFS: Accumulated PAR}
-\appendixfigures
-\begin{center}
+\subsection*{S.6 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
+\begin{figure}[!htbp]
+  \centering
   \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_as_is.png}
-\end{center}
-\subsubsection*{S.5.2 Partly de-accumulated}
-\appendixfigures
-\begin{center}
+  \caption{Example output from OpenIFS for January 2nd 2005. PAR is accumulated over the period of one day.}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
   \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_partly_deaccumulated.png}
-\end{center}
-\subsubsection*{S.5.3 De-accumulation of hour 00}
-\appendixfigures
-\begin{center}
+  \caption{Partly de-accumulated output from OpenIFS. $\text{PPFD}(t_i) = \text{PAR}(t_{i+1})-\text{PAR}(t_i)$.}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
   \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.65cm}]{pictures/PAR_deaccumulating_00h.png}
-\end{center}
+  \caption{De-accumulation of hour 00 in the partly de-accumulated fields.
+$\text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right]$.}
+\end{figure}
+%\clearpage
 
-\subsection*{S.6 Normed zonal average ozone dry deposition velocity}
-%\subsubsection*{S.6.1 }
-\appendixfigures
-\begin{center}
+\subsection*{S.7 Normalized zonal average ozone dry deposition velocity}
+\begin{figure}[!htbp]
+  \centering
   \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
-\end{center}
+  \caption{Normalized zonal average ozone dry deposition velocity.}
+\end{figure}
 

### 2018-12-07 DryDep.tex
Spell-checked, corrected version with respect to velocity computation.

@@
 Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as for single plant species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}.
-But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
+But progress has also been made on process-oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}.
 The variety of differing formulations and choices of parameters leads to a wide spread of results in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about $20\,\unit{\%}$ uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\
@@
 
+There seems to be a discrepancy between the OsloCTM3 response and the multi-model-mean, since the Wesely scheme is closer to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
+Without knowing the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this discrepancy.
+
 \begin{figure}[t]
@@
 Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
-The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling.
+The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process-modeling.
 The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}.
@@
 \label{subsec:obs}
-In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 instead of the re-gridded.
+In this section, we compare our results to observations.
+For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
+Although we do not assess the impacts of differing resolution in models on compatibility with observation, we can assume that both, high and low biases, may occur.
 
@@
 A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
-The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}.
+The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b).
 The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed.
-Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data.
-Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role.
+Further more, the observational averages comprise at most nine years worth of data.
+Statistically, these data may still subject to interannual variability.
+Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
 In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
@@
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
+We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
 The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology.
-The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
+Due to changes in dry deposition velocities $v_\text{DD}^\chem{O_3}$, the annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
 However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+We found the response of the OsloCTM3 to the changes in dry deposition velocities consistent between the old and the new dry deposition scheme.
+A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in surface ozone concentration $[\chem{O_3}(z_0)]$.
+However, there is an apparent discrepancy when comparing with the multi-model-mean.
+Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved.
 
@@
 With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
-Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}.
 

### 2018-12-07 DryDep_supplement.tex
Spell-checked, corrected version with respect to velocity computation.

@@
 The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3.
-Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is neglicible.}
+Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is negligible.}
 \end{figure}
@@
   \caption{Resulting $R_b$ for different species.
-Shown are results of one day model integrations for both, Southerland's law and $\mu(T) = m\cdot T$.
-Shown are also the percentages above / below the given threashold.}
+Shown are results of one day model integrations for both, Sutherland's law and $\mu(T) = m\cdot T$.
+Shown are also the percentages above / below the given threshold.}
 \end{figure}
@@
   \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
-  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global describtion.
-Example heights are shown for coniferous, decideous, needleleaf, and broadleaf forests as well as Mediterranean shrub.
+  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global description.
+Example heights are shown for coniferous, deciduous, needleleaf, and broadleaf forests as well as Mediterranean shrub.
 As reference the original range of EMEP is added.}

### 2018-12-10 DryDep.tex
Added more references and reorganized some text.

@@
 With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
-With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs, especially in regions covered by tropical or coniferous forest.
+With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical or coniferous forest \citep{GCB:Anav2017}.
 %
@@
 For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
-Although we do not assess the impacts of differing resolution in models on compatibility with observation, we can assume that both, high and low biases, may occur.
 
@@
 Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
-In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
+Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both, high and low biases, may exist depending on the location of the observation side and the distribution of local sources \citep{AE:Schaap2015}.
+Since we use a rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the respective model gridbox.
 
@@
 Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
-The model performance could also be improved by allowing for more plant functional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
+The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
 The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.

### 2018-12-12 DryDep.tex
Updated citations regarding greening season evaluation.

@@
 SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses leaprates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
-As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.
+As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.\\
 
 In agriculture, there are different empirical rules to estimate the SGS and EGS.
-The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above $5\,\unit{^\circ C}$ and vice versa for EGS.
-Based on this rule of thumb, we have pre-processed our meteorological input data offline.
-If, within these 5 days, the average temperature drops below / rises above $5\,\unit{^\circ C}$, the counters are reset, respectively.
+The simplest assumption is that greening starts after five consecutive days with a daily average temperature above $5\,\unit{^\circ C}$ and vice versa for EGS.
+Other estimates use growing degree days \citep{JC:Levis2004,PO:Fu2014}, include soil moisture \citep{GCB:Fu2014}, or rely on satellite observations.
+A comprehensive evaluation of different tecniques is given by \citet{GCB:Anav2017}.
+
+Based on the empirical rule ($5\,\unit{^\circ C}$-days), we have pre-processed our meteorological input data offline.
+We added some additional criteria to prevent for \emph{false spring}: If, within these five days, the average temperature drops below / rises above $5\,\unit{^\circ C}$, the counter is reset, respectively.
 First, we used the $5\,\unit{^\circ C}$-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
-In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we use the SMOKE-BEIS parameterization.
+In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we still use the SMOKE-BEIS parameterization.
 The described algorithm written in python~2.7 has been included as Supplement~S.5.

### 2018-12-13 DryDep.tex
More comments from Frode - part 1.

@@
 
-For each gridbox, average dry deposition velocities have been reconstructed from the ratio between the monthly averaged  surface ozone concentration and total ozone dry deposition 
+As dry deposition velocities are not directly available from the model output, monthly average dry deposition velocities $v^\chem{O_3}_\text{DD}$ have been reconstructed from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged surface ozone amount $\chem{O_3}(z_0)$
 \begin{equation}
@@
 Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
-In comparison, the available two month of the multi-model-mean display a different pattern.
+In comparison, the available two months of the multi-model-mean display a somewhat different pattern.
 Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
@@
 
-Finally, we take a look at the different global as well as hemispheric sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
+Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
+Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink due to dry deposition in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
 The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process-modeling.
@@
 \end{equation}
-with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
+with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number of degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
 The closer to $1$ this test scores, the better does the simulation represent the observation.
@@
 The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
-With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed.
+With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
 Further more, the observational averages comprise at most nine years worth of data.
@@
 Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
-Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both, high and low biases, may exist depending on the location of the observation side and the distribution of local sources \citep{AE:Schaap2015}.
-Since we use a rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the respective model gridbox.
+Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both, high and low biases, exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
+Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox.
 
@@
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
-Total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increase.
+Total column ozone in the tropics is predicted to decrease due to changes in the atmospheric circulation \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increase.
 The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.

### 2018-12-14 DryDep.tex
Some more polishing of conclusion chapter.

@@
 \label{subsubsec:greening}
-In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use prescribed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with leaprates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
+In Eqs.~(\ref{eq:non-stomata-cond}--\ref{eq:SAI}), \citet{ACP:Simpson2012} use prescribed start of growing season (SGS) and end of growing season (EGS) at $50\,\unit{^\circ N}$ ($d_\text{SGS}$, $d_\text{EGS}$) together with lapse rates ($\nabla d_\text{SGS}$, and $\nabla d_\text{EGS}$) to define phenology and dry deposition over agricultural areas.
 For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
@@
 First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).
-SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses leaprates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
+SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses lapse rates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
 As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.\\
@@
 Other estimates use growing degree days \citep{JC:Levis2004,PO:Fu2014}, include soil moisture \citep{GCB:Fu2014}, or rely on satellite observations.
-A comprehensive evaluation of different tecniques is given by \citet{GCB:Anav2017}.
+A comprehensive evaluation of different techniques is given by \citet{GCB:Anav2017}.
 
@@
 \end{equation}
-Herein $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$ with the average height of the lowermost model level in each gridbox $\Delta h_\text{month}$, which is changing with temperature and surface pressure, and the respective number of seconds in a month $s_\text{month}$.
+Herein $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$ with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
 %
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a), a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ applying the EMEP scheme compared to the Wesely scheme.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
@@
 
-There seems to be a discrepancy between the OsloCTM3 response and the multi-model-mean, since the Wesely scheme is closer to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
-Without knowing the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this discrepancy.
+There seems to be a discrepancy between the OsloCTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
+This could be a sign of differences in photo-chemistry and transport (e.g.
+convective, advective, STE) between the OsloCTM3 and the average TF~HTAP model, but without looking at the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
 
@@
 The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme.
-From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
+From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
 This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme.
 Similarly, the dry deposition velocities over water differ.
-From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}.
+From measurements during ship-campaigns a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ over the ocean has been deduced \citep{JGR:Helmig2012}.
 In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
@@
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
-The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology.
-Due to changes in dry deposition velocities $v_\text{DD}^\chem{O_3}$, the annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ especially over large desert regions and the oceans.
-Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average.
-However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
-We found the response of the OsloCTM3 to the changes in dry deposition velocities consistent between the old and the new dry deposition scheme.
+We have presented an update of the dry deposition scheme in the OsloCTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
+In our implementation based on \citet{ACP:Simpson2012}, we follow a classical, resistance analogous approach in computing contributions to dry deposition.
+The quasi-laminar resistance in the dry deposition scheme is capable of adjusting to surface wind stress but falls back to its prescribed limits in most of the tested cases.
+The surface resistance computation is divided into stomatal and non-stomatal resistance.
+The stomatal resistance is sensitive to parameters associated with the state of vegetation such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology, while the non-stomatal resistance is affected by the existence or absence of vegetation.
+
+The annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ changing from the old dry deposition scheme to the new one.
+Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below average.
+However, there seems to be a tendency, that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+
+We found the response of the OsloCTM3 to the changes in dry deposition velocities from the old and the new dry deposition scheme to be consistent.
 A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in surface ozone concentration $[\chem{O_3}(z_0)]$.
-However, there is an apparent discrepancy when comparing with the multi-model-mean.
-Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved.
-
-Most of the decrease in ozone dry deposition in the OsloCTM3 is due to changes in dry deposition velocities over the ocean and deserts.
-In case of desert and grasslands this change is at the order of one magnitude.
-Over the ocean it is due to the accumulative effect of a small absolute change in dry deposition which is especially amplified in the southern hemisphere.
+As the new scheme is more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy when comparing with the multi-model-mean.
+Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may even hint to differences in photo-chemistry as well as stratosphere-troposphere-exchange.
+
+Most of the decrease in ozone dry deposition in the OsloCTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
+This is mainly due to updates of the respective prescribed dry deposition velocities $v_\text{DD}^\chem{O_3}$.
+In case of desert and grasslands the difference between the old and new prescribed value is at the order of one magnitude.
+Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
 With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
@@
 We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants.
+The difference between our most extreme test cases amounts to $15\,\unit{\%}$with respect to total surface ozone, while the more realistic test case using differing years of emission amounts to about $5\,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.

### 2019-01-08 DryDep.tex
Some small asjustments to the text.

@@
 Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
-The resulting atomic oxygen reacts then with \chem{O_2} (and any other available co-reactant $M$) to form an ozone molecule.
+The resulting atomic oxygen reacts then with \chem{O_2} (and also under the presence of available co-reactants) to form an ozone molecule.
 %\begin{reaction}
@@
 \end{equation}
-with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and a its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
+with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
 The used ratios $D_\chem{H_2O}/D_i$ are tabulated in Supplement~S.1.
@@
 There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.).
-Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}.
+Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation.
 To reflect part of the underlying mechanism, the stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
@@
 All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or tabulated in Supplement~S.3.
-Since $f_T$ becomes negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+Since $f_T$ turns negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
 
@@
 
-The phenology of a plant typically describes its life-cycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
+The phenology of a plant typically describes its life-cycle throughout a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
 In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
@@
 
-Light in the wavelength band $(400-700)\,\unit{nm}$ to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
+Light in the wavelength band $400-700\,\unit{nm}$ to which the plant chlorophyll is sensitive is called photosynthetic active radiation (PAR).
 The integral of PAR over these wavelengths is the photosynthetic photon flux density (PPFD).
@@
 \end{equation}
-As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs.
+As initially mentioned, the necessary depth of snow to cover a certain type of vegetation differs.
 Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.
@@
 \subsubsection*{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between $60$ and $74\,\unit{^\circ north}$.
+The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
 To adapt this to a global model, we need to make a few additional assumptions.
@@
 Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (S.4: Fig.4).
+For four example PFTs, results are shown in the Supplement (S.4, Fig.4).
 
@@
 In the following, we present the necessary pre-processing of the variables.
+It is planed to implement an online computation of these variables into the OsloCTM3 later on.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Based on the empirical rule ($5\,\unit{^\circ C}$-days), we have pre-processed our meteorological input data offline.
-We added some additional criteria to prevent for \emph{false spring}: If, within these five days, the average temperature drops below / rises above $5\,\unit{^\circ C}$, the counter is reset, respectively.
+We added some additional criteria to prevent for \emph{false spring}: If, within these five days, the average temperature drops below or rises above $5\,\unit{^\circ C}$, the counter is reset, respectively.
 First, we used the $5\,\unit{^\circ C}$-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
@@
 From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be de-accumulated.
-Since this field is already integrated presumably over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD.
+Since this field is already integrated presumably over the spectrum $400-700\,\unit{nm}$, it is not PAR but PPFD.
 Therefore, we will refer to the de-accumulated field as PPFD.
@@
 \label{subsec:sens}
-Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change.
+Due to significant differences between the EMEP scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every difference in results to a respective change.
 Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
@@
 Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia.
-The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
+The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details): Mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
 Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}).
@@
 We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-A possible explanation to the divergence especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme.
+A possible explanation to this divergence especially in desert regions is the difference between the prescribed dry deposition velocities in the Wesely scheme in comparison to those used in the EMEP scheme.
 We cover this in more detail in the following section.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a), a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ applying the EMEP scheme compared to the Wesely scheme.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a), a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the EMEP scheme compared to the Wesely scheme.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
@@
 
-In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown.
+In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($30\,\unit{^\circ}-90\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($30\,\unit{^\circ}-90\,\unit{^\circ S}$) is shown.
 We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference.
@@
 Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
+Small adjustments to the lower limits in our quasi-laminar layer resistance formulation may help improve the OsloCTM3 performance in this regard.
 With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
@@
 
-In the northern hemisphere mid and high latitudes ($(50-75)\,\unit{^\circ N}$), total surface ozone increases by about $7.7\,\unit{\%}$ if the beginning and end of the vegetation period is estimated based on a $5\,\unit{^\circ C}$-days criteria instead of prescribed.
+In the northern hemisphere mid and high latitudes ($50\,\unit{^\circ}-75\,\unit{^\circ N}$), total surface ozone increases by about $7.7\,\unit{\%}$ if the beginning and end of the vegetation period is estimated based on a $5\,\unit{^\circ C}$-days criteria instead of prescribed.
 This is very important for any study focusing on ozone in the boreal and subarctic regions.

### 2019-01-21 DryDep.tex
Second round of Amunds comments and update of copernicus style files.

@@
 A publication focusing on this is planed.
-Parallelization of the OsloCTM3 has not been optimized as of now.
-While the meteorological data driving the OsloCTM3 is given in a resolution of T159N80L60, with the highest model level at $0.1\,\unit{hPa}$, it is very time and memory consuming to run the OsloCTM3 with full chemistry at this resolution.
+While the meteorological data driving the OsloCTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the OsloCTM3 with full chemistry at this resolution.
 Therefore, we reduced the horizontal resolution to $2.25^\circ\times2.25^\circ$ in our experiments.
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated surface PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be de-accumulated.
-Since this field is already integrated presumably over the spectrum $400-700\,\unit{nm}$, it is not PAR but PPFD.
-Therefore, we will refer to the de-accumulated field as PPFD.
-An example PAR for January 2nd 2005 is shown in Supplement~S.6 (Fig.~5).
-De-Accumulation of hour 3 till 21, simply means computing
+From OpenIFS an accumulated surface PAR is available, however, it is already integrated presumably over the spectrum $400-700\,\unit{nm}$.
+Hence, it is not PAR but PPFD.
+The field has not been de-accumulated in the post-processing of the OpenIFS output and it is not feasible to redo all of it for this work.
+Thus, this field needs to be de-accumulated for practical use in Eq.~(\ref{eq:flight}).
+The main obstacle is that PAR has been accumulated since model start, so that the first field kept from the original OpenIFS simulation ($00\,\unit{UTC}$) is $12\,\unit{h}$ after model start ($12\,\unit{UTC}$ on the previous day).
+In other words, the first time step of each day in the OsloCTM3 has already accumulated PAR from $12\,\unit{UTC}$ on the previous day.
+De-Accumulation of $03\,\unit{UTC}$ till $21\,\unit{UTC}$, simply means computing the difference
 \begin{equation}
@@
 \end{equation}
-If everything was right in the OpenIFS output, we would not need to do any further processing, however the first time step of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (Supplement~S.6: Fig.~6), so it is impossible to accurately de-accumulate this time step.
-The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.6: Fig.~7)
-%
-\begin{equation}
-  \text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right].
-\end{equation}
-%
+For de-accumulation of the remaining timestep, the best choice is subtracting the difference between $21\,\unit{UTC}$ and $12\,\unit{UTC}$ of the previous day
+%
+\begin{equation}
+  \text{PPFD}(t=00\,\unit{UTC}) = \text{PAR}(t=00\,\unit{UTC}) - \left[\text{PAR}(t=21\,\unit{UTC}-1\,\unit{day})-\text{PAR}(t=12\,\unit{UTC}-1\,\unit{day})\right]
+\end{equation}
+%
+and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.6 (Fig.~5--7).
 The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$.
@@
 We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
-The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1nc4}.
+The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}.
 CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
-Biogenic emissions are taken from MEGAN output.
+Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}.
 In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
@@
 This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years.
-After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}.
+After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} is the baseline simulation for the sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}.
 As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a), a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the EMEP scheme compared to the Wesely scheme.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the EMEP scheme compared to the Wesely scheme.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
@@
 
-The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b).
+The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b.
 The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ lower than in the Wesely scheme.
@@
 
-The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c).
+The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c.
 In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}.
@@
 %
-In Appendix Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month.
+In Appendix Fig.~\ref{fig:mmm_drydep_season}a, the average zonal ozone dry deposition is shown separated by month.
 Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
@@
 In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ.
-The corresponding total monthly ozone dry deposition is shown in Appendix Fig.~\ref{fig:mmm_drydep_season}b).
+The corresponding total monthly ozone dry deposition is shown in Appendix Fig.~\ref{fig:mmm_drydep_season}b.
 In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
@@
 As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
 Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
@@
 Results from \citep{ACP:Hardacre2015} are shown as a reference.
-We refrain from showing the full extend of \emph{EMEP\_offLight} here, since it is an extreme scenario and has been discussed already.}
+We refrain from showing the full extent of \emph{EMEP\_offLight} here, since it is an extreme scenario and has been discussed already.}
   \label{fig:mmm_drydep_season_pft}
@@
 
-In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown.
+In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown.
 We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band.
@@
 A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
-The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b).
+The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
 The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
-Further more, the observational averages comprise at most nine years worth of data.
+Furthermore, the observational averages comprise at most nine years worth of data.
 Statistically, these data may still subject to interannual variability.
 Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
-Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both, high and low biases, exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
+Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both high and low biases exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
 Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox.
@@
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below average.
-However, there seems to be a tendency, that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+However, there seems to be a tendency that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
 
@@
 As the new scheme is more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy when comparing with the multi-model-mean.
-Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may even hint to differences in photo-chemistry as well as stratosphere-troposphere-exchange.
+Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may hint to, e.g., differences in photo-chemistry, stratosphere-troposphere-exchange as well as to differences in aerosol loadings among many other aspects.
 
@@
 This may be due to several reasons.
-The rather low model resolution in both horizontal ($2.25^\circ\times 2.25^\circ$) and vertical (L60, $P_\text{max}=0.1\,\unit{hPa}$) affects the distribution and transport (e.g., long-range, convection, and stratosphere--troposphere exchange) of ozone and its precursors.
+The model resolution in both horizontal ($2.25^\circ\times 2.25^\circ$) and vertical (L60, $P_\text{max}=0.02\,\unit{hPa}$) does not capture all details in transport, thus affecting the distribution and transport (e.g., long-range, convection, and stratosphere--troposphere exchange) of ozone and its precursors.
 Depending on the location of the observation site and its respective representativeness for a larger area, ozone dry deposition and ozone concentrations are expected to be over- or underestimated in the model.
@@
 
-\codedataavailability{The OsloCTM3 is made publicly available on git-hub under a MIT license.
+\codedataavailability{The OsloCTM3 is publicly available on git-hub under a MIT license.
 Model results can be made available under request.} %% use this section when having data sets and software code available
@@
 %\sampleavailability{TEXT} %% use this section when having geoscientific samples available
+%\videosupplement{TEXT} %% use this section when having video supplements available
 
@@
 Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the documentation.
-Both authors contributed to the writing of the paper.} %% it is strongly recommended to make use of this section
+Both authors contributed to the writing and discussion of the paper.} %% this section is mandatory for the journals ACP and GMD.
+For all other journals it is strongly recommended to make use of this section
 
@@
 https://doi.org/10.3334/ORNLDAAC/1293
+  The authors would like to thank Prof.
+Frode Stordal (Institut for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institut) for technical support.
 \end{acknowledgements}

### 2019-01-21 DryDep.tex
Deleted obsolete file. Started to conform to copernicus picture composition.

@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/phenology.pdf}
+  \includegraphics[width=8.3cm]{fig01}
   \caption{Sketch of the five different phases described in $f_\text{phen}$.}
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{pictures/pft_landuse_mapping.pdf}
+  \includegraphics[width=12cm]{fig02}
   \caption{Mapping of land surface categories.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005_she.pdf}
+  \includegraphics[width=8.3cm, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{fig03}
   \caption{Pre-processing of greening season from meteorological surface temperature fields.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
+  \includegraphics[width=8.3cm]{fig04}
   \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
-  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
+  \includegraphics[width=8.3cm]{fig05}
   \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
+  \includegraphics[width=12cm]{fig06}
   \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH).
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
+  \includegraphics[width=12cm]{fig07}
   \caption{Average seasonal cycles of ozone dry deposition velocities separated by land use type.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{pictures/final-monthly_ozone_drydep_stations_True}\\
-  \includegraphics[width=8.3cm]{pictures/final-monthly_drydep_divergence_True}
+  \includegraphics[width=8.3cm]{fig08}
   \caption{Ozone dry deposition fluxes at different observation sites.

### 2019-01-22 DryDep.tex
Fine-tuning to GMD requirements.

@@
 %\usepackage{rotating}
+\graphicspath{{final_figures/}}
 
@@
 We also do not consider the contribution of very short-lived ozone depleting substances (VSLS), that affect stratospheric ozone \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017}, to the tropospheric ozone abundance.\\
-Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to one year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
+Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
 The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Land}
 Over land, we use \citep[Eq.~(53),][]{ACP:Simpson2012}
@@
 %
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Ocean}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Over ocean, we use \citep[Eq.~(54),][]{ACP:Simpson2012}
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.2: Figs.~1--2).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.2: Figs.~S1--S2).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 The allowed maximum roughness length in both cases is set to $2\,\unit{mm}$.
-Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.2: Fig.~3).
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.2: Fig.~S3).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Stomatal conductance}
+%\subsubsection*{Stomatal conductance}
 The stomatal conductance is a measure of the rate of \chem{CO_2} exchange and evapotranspiration through the stomata of a leaf.
@@
   \includegraphics[width=8.3cm]{fig01}
-  \caption{Sketch of the five different phases described in $f_\text{phen}$.}
+  \caption{Sketch of the five different phases in plant phenology $f_\text{phen}$ in accordance to Eq.~(\ref{eq:fphen}).}
   \label{fig:phenology}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Non-stomatal conductance}
+%\subsubsection*{Non-stomatal conductance}
 In the EMEP scheme, non-stomatal conductances are explicitly calculated for \chem{O_3}, \chem{SO_2}, \chem{HNO_3}, and \chem{NH_3}.
@@
 
-\subsubsection*{Latitude dependent vegetation height}
+\subsubsection{Latitude dependent vegetation height}
 The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
@@
 Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (S.4, Fig.4).
-
-\subsubsection*{Mapping of land surface types}
+For four example PFTs, results are shown in the Supplement (S.4, Fig.~S4).
+
+\subsubsection{Mapping of land surface types}
 The OsloCTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
@@
 In agriculture, there are different empirical rules to estimate the SGS and EGS.
-The simplest assumption is that greening starts after five consecutive days with a daily average temperature above $5\,\unit{^\circ C}$ and vice versa for EGS.
+The simplest assumption is that greening starts after $5$ consecutive days with a daily average temperature above $5\,\unit{^\circ C}$ and vice versa for EGS.
 Other estimates use growing degree days \citep{JC:Levis2004,PO:Fu2014}, include soil moisture \citep{GCB:Fu2014}, or rely on satellite observations.
@@
 Based on the empirical rule ($5\,\unit{^\circ C}$-days), we have pre-processed our meteorological input data offline.
-We added some additional criteria to prevent for \emph{false spring}: If, within these five days, the average temperature drops below or rises above $5\,\unit{^\circ C}$, the counter is reset, respectively.
+We added some additional criteria to prevent for \emph{false spring}: If, within these $5$ days, the average temperature drops below or rises above $5\,\unit{^\circ C}$, the counter is reset, respectively.
 First, we used the $5\,\unit{^\circ C}$-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH.
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated surface PAR is available, however, it is already integrated presumably over the spectrum $400-700\,\unit{nm}$.
-Hence, it is not PAR but PPFD.
-The field has not been de-accumulated in the post-processing of the OpenIFS output and it is not feasible to redo all of it for this work.
-Thus, this field needs to be de-accumulated for practical use in Eq.~(\ref{eq:flight}).
-The main obstacle is that PAR has been accumulated since model start, so that the first field kept from the original OpenIFS simulation ($00\,\unit{UTC}$) is $12\,\unit{h}$ after model start ($12\,\unit{UTC}$ on the previous day).
+From OpenIFS an accumulated surface PAR is available which is already integrated presumably over the spectrum $400-700\,\unit{nm}$.
+For this reason, it will be refered to as PPFD instead of PAR.
+This PPFD field has not been de-accumulated in the post-processing of the OpenIFS output and it is not feasible to redo all of the post-processing for this work.
+But for practical use in Eq.~(\ref{eq:flight}), this field needs to be de-accumulated.
+The main obstacle is that PAR has been accumulated since model start, so that the first field kept from the original OpenIFS simulation ($00\,\unit{UTC}$) is $12$ hours after model start ($12\,\unit{UTC}$ on the previous day).
 In other words, the first time step of each day in the OsloCTM3 has already accumulated PAR from $12\,\unit{UTC}$ on the previous day.
-De-Accumulation of $03\,\unit{UTC}$ till $21\,\unit{UTC}$, simply means computing the difference
+De-accumulation of times $03\,\unit{UTC}$ to $21\,\unit{UTC}$, simply means computing the difference
 \begin{equation}
@@
 %
-and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.6 (Fig.~5--7).
-The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$.
+and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.6 (Figs.~S5--S7).
+The resulting PPFD fields are still accumulated over a time period of $3$ hours and should be divided by $3$.
 A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations.
@@
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.7: Fig.~8).
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.7: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
@@
 The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c.
-In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}.
+In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one-half of the amount given by \emph{Wesely\_type}.
 The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a change in $v_\text{desert}^\chem{O_3}$.
@@
 Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
-In comparison, the available two months of the multi-model-mean display a somewhat different pattern.
+In comparison, the available $2$ months of the multi-model-mean display a somewhat different pattern.
 Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
@@
 In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
-The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme.
+The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
 From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
@@
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
-Furthermore, the observational averages comprise at most nine years worth of data.
+Furthermore, the observational averages comprise at most $9$ years worth of data.
 Statistically, these data may still subject to interannual variability.
@@
 This is mainly due to updates of the respective prescribed dry deposition velocities $v_\text{DD}^\chem{O_3}$.
-In case of desert and grasslands the difference between the old and new prescribed value is at the order of one magnitude.
+In case of desert and grasslands the difference between the old and new prescribed value is at the order of $1$ magnitude.
 Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
@@
 Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
-The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE).
+The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
 The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{pictures/pft_categories.pdf}\\
-  \includegraphics[width=0.8\textwidth]{pictures/pft_categories_zonal.pdf}
+  \includegraphics[width=0.8\textwidth]{fig09}
   \caption{Partitioning of land surface types. (a) CLM~2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=0.8\textwidth]{pictures/final-total_ozone_drydepvelo_monthly}\\
-  \includegraphics[width=0.8\textwidth]{pictures/final-total_ozone_drydep_monthly}\\
-  \includegraphics[width=0.7\textwidth]{pictures/legend}
+  \includegraphics[width=0.8\textwidth]{fig10}
   \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
@@
 
-\authorcontribution{Stefanie Falk has finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the OsloCTM3, conducted the tests, and evaluated the results.
-Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the documentation.
+\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the OsloCTM3, conducted the simulations, and analysed and evaluated the results.
+Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the respective documentation.
 Both authors contributed to the writing and discussion of the paper.} %% this section is mandatory for the journals ACP and GMD.
@@
 \begin{acknowledgements}
-  This work was supported by the Norwegian Research Council (NRC) through the project The double punch: Ozone and climate stresses on vegetation (OzoNorClim).\\
+  This work was supported by the Norwegian Research Council (NRC) through the project The double punch: Ozone and climate stresses on vegetation (268073).\\
   %We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
@@
 https://doi.org/10.3334/ORNLDAAC/1293
-  The authors would like to thank Prof.
-Frode Stordal (Institut for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institut) for technical support.
+  We would like to thank Prof.
+Frode Stordal (Section for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institut) for technical support.
 \end{acknowledgements}

### 2019-01-23 DryDep.tex
Second round f fine tuning to GMD requirements, some minor adaptions of text and typo corrections. Ready to submit.

@@
 
-\title{Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0}
+\title{Update and evaluation of the ozone dry deposition in the Oslo~CTM3 v1.0}
 
@@
 
-\runningtitle{Update of ozone dry deposition in OsloCTM3}
+\runningtitle{Update of ozone dry deposition in Oslo~CTM3}
 
@@
 The conceptual dry deposition velocities are calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these are hard to measure explicitly.
-In this paper, we present an update of the dry deposition scheme implemented in the OsloCTM3.
-We change from an empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere and vegetation into account.
+In this paper, we present an update of the dry deposition scheme implemented in the Oslo~CTM3.
+We change from a purely empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere and vegetation into account.
 Examining the sensitivity of the scheme to various parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
@@
 In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
-%\begin{reaction}
-%  ^\bullet\chem{OH} + \chem{CO} \rightarrow {^\bullet\chem{HOCO}}\\
-%  ^\bullet\chem{HOCO} + \chem{O_2} \rightarrow \chem{HO_2^\bullet} + \chem{CO_2}.
-%\end{reaction}
 Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
 The resulting atomic oxygen reacts then with \chem{O_2} (and also under the presence of available co-reactants) to form an ozone molecule.
-%\begin{reaction}
-%  \chem{HO_2^\bullet} + \chem{NO} \rightarrow {^\bullet\chem{OH}} + \chem{NO_2}\\
-%  \chem{NO_2} + h\nu \rightarrow \chem{NO} + \chem{O(^3{P})}\\
-%^\bullet\chem{O(^3P)} + \chem{O_2} + M \rightarrow \chem{O_3} + M.
-%\end{reaction}
 Such a cycle leads to a net production via:
@@
 Major sinks are photolysis followed by a reaction with water vapor to from \chem{OH},
-%\begin{reaction}
-%  \chem{O_3} + h\nu \rightarrow \chem{O(^1D)} + \chem{O_2})\\
-%  \chem{O(^1D)} + \chem{H_2O} \rightarrow 2\chem{OH},
-%\end{reaction}
-reactions with \chem{HO_2},
-%\begin{reaction}
-%  {\bf TODO: Reaction},
-%\end{reaction}
-de-nitration reactions,
-%\begin{reaction}
-%  \chem{NO_x} + \chem{O_3} \rightarrow \chem{NO_{x+1}} + \chem{O_2},
-%\end{reaction}
-and dry deposition. We will come back to the latter later in this section and cover the implemented scheme in more detail in Section~\ref{subsec:DryDep}.
+reactions with \chem{HO_2}, titration reactions, and dry deposition.
+We will come back to the latter later in this section and cover the implemented scheme in more detail in Section~\ref{subsec:DryDep}.
 
@@
 Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.
-Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the OsloCTM3 as of now.
+Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the Oslo~CTM3 as of now.
 We also do not consider the contribution of very short-lived ozone depleting substances (VSLS), that affect stratospheric ozone \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017}, to the tropospheric ozone abundance.\\
@@
 
-In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
+In Section~\ref{sec:model_des}, we will briefly describe the Oslo~CTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
 In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations at the surface (Section~\ref{subsec:obs}).
@@
 \label{sec:model_des}
-The OsloCTM3 is a three dimensional global chemistry transport model (CTM).
-The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}.
-A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
-The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}.
+The Oslo~CTM3 is a three dimensional global chemistry transport model (CTM).
+The key components of the Oslo~CTM3 have been described and evaluated by \citet{GMD:Sovde2012}.
+A detailed account of the capabilities of the Oslo~CTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}.
+The Oslo~CTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}.
 A publication focusing on this is planed.
-While the meteorological data driving the OsloCTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the OsloCTM3 with full chemistry at this resolution.
+While the meteorological data driving the Oslo~CTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
 Therefore, we reduced the horizontal resolution to $2.25^\circ\times2.25^\circ$ in our experiments.
@@
 \label{subsec:DryDep}
-In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account.
+In the original dry deposition scheme of the Oslo~CTM3, the state of the atmosphere was not taken into account.
 Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}.
@@
 
-The phenology of a plant typically describes its life-cycle throughout a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
+The phenology of a plant typically describes its life-cycle throughout a year, e.g., at mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
 In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
@@
 \begin{table}[t]
-  \caption{Definition of growing season for crops used in the OsloCTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
+  \caption{Definition of growing season for crops used in the Oslo~CTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
   \begin{tabular}{lrr}
@@
 To adapt this to a global model, we need to make a few additional assumptions.
-The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes ($40-60\,\unit{^\circ}$).
+The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
 Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
@@
 \subsubsection{Mapping of land surface types}
-The OsloCTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
+The Oslo~CTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
@@
 %
-The total aerosol surface deposition is then
+The total aerosol surface deposition is defined as
 \begin{equation}
@@
 As mentioned in the previous section, there are two variables needed for computing the stomatal conductance which are not directly available from the meteorological input data.
-The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the amount of light plants need to photosynthesize.
+The greening season, as the time of the year in the mid and high latitudes when it is most likely for plants to grow, and the photosynthetic photon flux density, as the amount of light that plants need to photosynthesize.
 In the following, we present the necessary pre-processing of the variables.
-It is planed to implement an online computation of these variables into the OsloCTM3 later on.
+It is planed to implement an online computation of these variables into the Oslo~CTM3 later on.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The main obstacle is that PAR has been accumulated since model start, so that the first field kept from the original OpenIFS simulation ($00\,\unit{UTC}$) is $12$ hours after model start ($12\,\unit{UTC}$ on the previous day).
-In other words, the first time step of each day in the OsloCTM3 has already accumulated PAR from $12\,\unit{UTC}$ on the previous day.
+In other words, the first time step of each day in the Oslo~CTM3 has already accumulated PAR from $12\,\unit{UTC}$ on the previous day.
 De-accumulation of times $03\,\unit{UTC}$ to $21\,\unit{UTC}$, simply means computing the difference
@@
 \label{sec:eval}
-In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme.
+In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the EMEP scheme.
 We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively.
 We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
-The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}.
+The Oslo~CTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}.
 CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
@@
 Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
-A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+A reference simulation featuring the Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
 For all model integrations, the meteorological reference year is 2005.
@@
 During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
-Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia.
+Especially in Europe, this basically results in a suppression of canopy resistance in northern Scandinavia.
 The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details): Mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
-Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}).
+Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the Oslo~CTM3 (\emph{EMEP\_ppgssh\_ice}).
 Accidentally, we have used emissions for the year 2014 instead of 2005.
@@
 \end{equation}
-Herein $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$ with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
+Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
 %
@@
 Among the sensitivity studies, there is only a low absolute variance.
-Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
+Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
 Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
@@
 In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean.
-This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3.
+This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
+This may, however, lead to an overcompensation of the current low-bias in surface ozone in the Oslo~CTM3.
 The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
@@
 
-There seems to be a discrepancy between the OsloCTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
+There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
 This could be a sign of differences in photo-chemistry and transport (e.g.
-convective, advective, STE) between the OsloCTM3 and the average TF~HTAP model, but without looking at the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
+convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
 
@@
   \includegraphics[width=8.3cm]{fig05}
-  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
+  \caption{Comparison of the manifold of Oslo~CTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
 The different colors indicate sets of simulation with similar baselines.
@@
 Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink due to dry deposition in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
 The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process-modeling.
-The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}.
-But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
+The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
+But also the results of \citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
 %
@@
 In this section, we compare our results to observations.
-For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
+For all comparisons, we use the original resolution of the Oslo~CTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
 
@@
 The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
-The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the Oslo~CTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
@@
 (a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
-The model uncertainty of the OsloCTM3 is given as $1\,\sigma$ error band.
+The model uncertainty of the Oslo~CTM3 is given as $1\,\sigma$ error band.
 The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
@@
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the OsloCTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
+We have presented an update of the dry deposition scheme in the Oslo~CTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
 In our implementation based on \citet{ACP:Simpson2012}, we follow a classical, resistance analogous approach in computing contributions to dry deposition.
@@
 The surface resistance computation is divided into stomatal and non-stomatal resistance.
-The stomatal resistance is sensitive to parameters associated with the state of vegetation such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology, while the non-stomatal resistance is affected by the existence or absence of vegetation.
+The stomatal resistance is sensitive to parameters associated with the vegetation such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology, while the non-stomatal resistance is affected by the existence or absence of vegetation.
 
 The annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ changing from the old dry deposition scheme to the new one.
-Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below average.
+Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is about $33\,\unit{\%}$ below average.
 However, there seems to be a tendency that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
 
-We found the response of the OsloCTM3 to the changes in dry deposition velocities from the old and the new dry deposition scheme to be consistent.
+We found the response of the Oslo~CTM3 to the changes in dry deposition velocities from the old and the new dry deposition scheme to be consistent.
 A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in surface ozone concentration $[\chem{O_3}(z_0)]$.
@@
 
-Most of the decrease in ozone dry deposition in the OsloCTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
+Most of the decrease in ozone dry deposition in the Oslo~CTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
 This is mainly due to updates of the respective prescribed dry deposition velocities $v_\text{DD}^\chem{O_3}$.
@@
 Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
-Small adjustments to the lower limits in our quasi-laminar layer resistance formulation may help improve the OsloCTM3 performance in this regard.
-With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
+Small adjustments to the lower limits in our quasi-laminar layer resistance formulation may help improve the Oslo~CTM3 performance in this regard.
+With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new Oslo~CTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
 Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}.
@@
 Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
-By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model.
+By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model, however, this could lead to an overcompansation (high-bias).
 
 We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants.
-The difference between our most extreme test cases amounts to $15\,\unit{\%}$with respect to total surface ozone, while the more realistic test case using differing years of emission amounts to about $5\,\unit{\%}$.
+The difference between our most extreme test cases amounts to $15\,\unit{\%}$ with respect to total surface ozone, while the more realistic test case using differing years of emission amounts to about $5\,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
@@
 
-Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
+Future work on the Oslo~CTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
 The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
@@
 
-\codedataavailability{The OsloCTM3 is publicly available on git-hub under a MIT license.
+\codedataavailability{The Oslo~CTM3 is publicly available on git-hub under a MIT license.
 Model results can be made available under request.} %% use this section when having data sets and software code available
@@
 
-\clearpage
+%\clearpage
 
@@
 \section{Figures}    %% Appendix A
+
+\appendixfigures
+\begin{figure}[!htbp]
+  \includegraphics[width=0.8\textwidth]{fig09}
+  \caption{Comparison of the manifold of Oslo~CTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
+The multi-model mean from the evaluation of TF~HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
+  \label{fig:mmm_drydep_season}
+\end{figure}
+
 \appendixfigures
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{fig09}
+  \includegraphics[width=0.8\textwidth]{fig10}
   \caption{Partitioning of land surface types. (a) CLM~2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}
@@
 
-\appendixfigures
-\begin{figure}[!htbp]
-  \includegraphics[width=0.8\textwidth]{fig10}
-  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
-The multi-model mean from the evaluation of TF~HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
-  \label{fig:mmm_drydep_season}
-\end{figure}
 %\subsection{}     %% Appendix A1, A2, etc.
@@
 
-\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the OsloCTM3, conducted the simulations, and analysed and evaluated the results.
+\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the Oslo~CTM3, conducted the simulations, and analysed and evaluated the results.
 Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the respective documentation.

### 2019-01-24 DryDep.tex
Finished acknowledgments.

@@
 Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
-By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model, however, this could lead to an overcompansation (high-bias).
+By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model, however, this could lead to an overcompensation (high-bias).
 
@@
   This work was supported by the Norwegian Research Council (NRC) through the project The double punch: Ozone and climate stresses on vegetation (268073).\\
-  %We would like to thank Greg Bodeker, Stefanie Kremser (Bodeker Scientific) and Birgit Hassler (DLR) for providing the combined vertical ozone profile database (\url{http://www.bodekerscientific.com}.
+  The simulations were performed on resources provided by UNINETT Sigma2 -- the National Infrastructure for High Performance Computing and Data Storage in Norway (project nn2806k).\\
   The used Leaf Area Index (LAI) and roughness length ($z_0$) are available online from Oak Ridge National Laboratory Distributed Active Archive Center, Oak Ridge, Tennessee, U.S.A.
@@
   We would like to thank Prof.
-Frode Stordal (Section for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institut) for technical support.
+Frode Stordal (Section for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institute) for technical support.
+  We would also like to thank the Center for International Climate Research (CICERO) for their support of this work.
 \end{acknowledgements}

### 2019-04-03 version_2/response_version_1.tex
Created the directories to store the comments from referees and the corresponding answers.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To gmd-2019-21 Anonymous Referee \#1 (02 Apr 2019):
We thank the referee for his/her useful comments and we will take them into account in the revised version of this paper.
In the following, we will answer the questions in detail.
\begin{itemize}
\item {\color{blue}  Abstract L15-17: \emph{“While high sensitivity to changes in dry deposition to
    vegetation is found in the tropics, the largest impact on global scales is associated to
    changes in dry deposition to the ocean and deserts.”} The authors do not provide details
  in the paper as to what has changed in the updated scheme for such an impact.
  \begin{itemize}
  \item Is it the surface resistance ($R_c$) value? Or the other two resistances ($R_a$ and $R_b$)?
  \item What are the typical values?
  \item What value of $R_c$ for water has been used and
  \item how it compares with the value used in the Wesely scheme?
  \end{itemize}
}
  \emph{}

\item {\color{blue}  P22 Table 3:}
  \begin{itemize}
  \item {\color{blue}  Why the deposition values for ocean + ice + land do to add up to the total
    values reported, for all simulations?} \emph{}
  \item {\color{blue}  The new land-based deposition values are much lower than what has
    been reported in previous studies (e.g. Hardacre et al., 2015) and the authors largely
    attribute this to the changes in the updated scheme for the desert surface type. However,
    the paper does not provide any observational support to back this up. Are there
    any relevant deposition measurements (velocity or flux) that can be used for this purpose?
    At least, some comparison with ozone measurements (or even $\mathrm{O_3}$ reanalyses)
    should be provided for this surface type (and perhaps others) to see if the model is
    heading in the right direction with the updated deposition scheme.} \emph{}
  \item {\color{blue}  It will also be useful to report the global ozone burden from the various
    simulations.}
    \emph{}
  \end{itemize}

\item {\color{blue}  Section 2.1.1, Eq. (2): The statement \emph{“For certain values of z, z0, and L, this
    may result in nonphysical (negative) values for Ra.”} I do not comprehend as to why
  this would occur since this equation is simply based on the well-used Monin-Obukhov
  similarity theory (MOST) for the surface layer. This occurrence would also imply negative wind speeds.
  Actually Eq. (2) is incorrect: the term $\psi_m((z-d)/z_0)$ should be
  $\psi_m((z-d)/L)$, and the sign of the third term on the right-hand side should be positive
  (not negative). Given that $(z-d) > z_0$ (assuming the model is formulated correctly), Eq. (2) should always yield positive values.
  Eqs. (3--5): I am not sure why Monteith (1973) needs to be invoked here. Given that
  the term in the square brackets on right hand side of Eq. (2) is equal to $k\cdot u(z)/u^{*}$ as per
  MOST, substituting this into Eq. (2) results in Eq. (5).
  Define $z$, $z_0$ in Eq. (2). The parameter $d$ is the so-called displacement height, and is
  not a constant (depends on the surface type).}
  \emph{}

\item {\color{blue}  P2 L25-33: The first reference to the Oslo~CTM3 in the body of the paper is
  made here as \emph{“...we have not implemented any parameterization of these processes
    in the Oslo CTM3 as of now.”} Some brief introductory text is required here (or better at
  the start of the paragraph) to introduce the model properly. Also, the text between lines
  25 -- 33 on what is not considered in the model is too detailed to be here, so shorten
  and move it to Section 2.}
  Since we intent to properly introduce our model (Oslo CTM3) in Section~2, we rather
  move the sentence in L30-32 about polar boundary layer ozone depletion to Section~2.
  Regarding that the influence of VSLS on tropospheric ozone is indirect, this reference has not been
  well placed in this section and will be moved to the discussions (Section~4).

\item {\color{blue}  P3 L19/L28 and P21 L34: There is a newer ozone dry
  deposition study by Luhar et al. (2018, ACP, 18, 4329-4348) which, using global ozone
  reanalyses and a more realistic process-based oceanic deposition scheme, estimates
  the total global deposition at $722.8 \pm 87.3\,\mathrm{Tg O_3 yr^{-1}}$, which includes an oceanic
  component of $98.4 \pm 30.0\,\mathrm{Tg O_3 yr^{-1}}$. These figures should be cited for comparison.}
  Thank you for pointing this out. We were not aware of this study and will compare our results and refer
  to it at the given places and within our discussion.
  \emph{
    \begin{itemize}
    \item A newer study by Luhar et al. (2018), however, indicates much lower amounts ($722.8 \pm 87.3\,\mathrm{Tg\,a^{-1}}$).
    \item Based on the global atmospheric composition reanalysis performed in the\\ ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, Luhar et al. (2018) found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\mathrm{Tg\,a^{-1}}$.
    \item But also the results of  Luhar et al. (2017, 2018) yield a $(19-27)\,\mathrm{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\mathrm{\%}$ of the total annual ozone dry deposition.
    \end{itemize}
  }

\item {\color{blue}  Section 2.2: Since the present paper is about ozone dry deposition, this section
seems like a distraction and hence should be omitted.} \emph{}
\item {\color{blue}  Page 14, lines 15 -- 18: Anthropogenic, biomass burning, and biogenic emis-
sions are included in the model. How are other emissions such as soil NOx, wetland
methane, and oceanic methane and CO specified?} \emph{}
\item {\color{blue}  Page 15, line 4: The statement “Accidentally, we have used emissions for the year
2014 instead of 2005.” It is not clear what the consequences on the results are of this? } \emph{}
\item {\color{blue}  Section 3.2.1: In the Fig 5 discussion, although snow and ice is discussed, there is
no discussion on the oceanic differences between the present study and Hardacre et
al. (2015). This is particularly important for the Southern latitudes.} \emph{}
\item {\color{blue}  Section 3.2.1: The Hardacre et al. (2015) simulations were for the year 2001,
whereas the present study is mostly for the year 2015 emissions (see Table 1) driven
by the year 2005 meteorology. In addition, the observational averages used in Fig. 8
are based on multi-year data. The authors should discuss the implications of these
differences about different years on the deposition results presented (e.g. uncertainty).} \emph{}
\item {\color{blue}  Page 24, lines 3 -- 4: “The annual amount of ozone dry deposition decreases by up
to 100\% changing from the old dry deposition scheme to the new one.” Table 3 does
not support this, but this may be true for some surface types. So please qualify the
statement.} \emph{}
\item {\color{blue}  Page 24, line 15: “Most of the decrease in ozone dry deposition in the Oslo CTM3
can be attributed to changes in dry deposition velocities over the ocean and deserts.”
What are the dominant factors in these changes? For example, is it mostly the surface
resistance (Rc) term? For the ocean, it is likely to be Rc. For deserts, maybe Rb? Is it
possible to quantify these differences in the resistance terms?} \emph{}
\item {\color{blue}  Page 24, line 24: “2-layer gas exchange with ocean waters (Luhar et al., 2017).”
As mentioned earlier, Luhar et al. (2018) has derived a more realistic process-based
deposition scheme for the ocean, but the results for deposition velocity do not seem to
be too different from those in Luhar et al. (2017).} \emph{}
\item {\color{blue}  Page 25, lines 11 -- 12: The comment “This is most likely reflecting the ongoing
industrialization process of countries in the southern hemisphere and the commitment
and implementation of air quality regulations of industrialized nations in the northern
hemisphere” is quite speculative and may be omitted.} \emph{}
\item {\color{blue}  Eq. (13) cf. Eq. (14): $g_STO or G_sto$ -- use consistency with notation.} \emph{}
\item {\color{blue}  The first half of the abstract, the text before “In this paper...,” is introductory
material and can be deleted.} \emph{}
\item {\color{blue}  In the abstract (lines 15-16), it is better to say “...leading to an increase in surface
ozone of up to 100\% in some regions.”} \emph{}
\item {\color{blue}  Page 22, line 7: “At about 4 of 6 sites.” About? Not sure?} \emph{}
\end{itemize}
\newpage

\end{document}

### 2019-04-03 DryDep.tex
Started including changes in response to RC1.

@@
 They occur in spring-time in the polar boundary layer where an outburst of bromine monoxide \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
-Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.
-Although these ozone depleting events are important to understand surface ozone abundance in Arctic regions, we have not implemented any parameterization of these processes in the Oslo~CTM3 as of now.
-We also do not consider the contribution of very short-lived ozone depleting substances (VSLS), that affect stratospheric ozone \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017}, to the tropospheric ozone abundance.\\
+Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.\\
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
@@
 It is estimated that about $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
+A newer study by \citet{ACPLuhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
@@
 \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
+Based on the global atmospheric composition reanalysis performed in the ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, \citet{ACP:Luhar2018} found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\unit{Tg\,a^{-1}}$.
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.
@@
 A publication focusing on this is planed.
+Although the above mentioned ozone depleting events are important to understand surface ozone abundance in Arctic regions, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.
 While the meteorological data driving the Oslo~CTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
@@
 The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
-But also the results of \citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
+But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition.
 %
@@
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is about $33\,\unit{\%}$ below average.
-However, there seems to be a tendency that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
+However, there seems to be a tendency that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,ACP:Luhar2018,AE:Hu2017}.
 
@@
 Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may hint to, e.g., differences in photo-chemistry, stratosphere-troposphere-exchange as well as to differences in aerosol loadings among many other aspects.
+Especially the STE depends on the stratospheric ozone abundance which is, e.g.
+strongly affected by the contribution of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and not taken into account in the Oslo~CTM3.
 

### 2019-04-04 version_2/response_version_1.tex
Continue to answer referee questions. Revise paper and recompute ozone dry dep to different surfaces.

@@
 \usepackage{xcolor}
+\usepackage{url}
 \thispagestyle{empty}
@@
 \begin{itemize}
-\item {\color{blue}  Abstract L15-17: \emph{“While high sensitivity to changes in dry deposition to
+\item {\color{blue}  Abstract L15--17: \emph{“While high sensitivity to changes in dry deposition to
     vegetation is found in the tropics, the largest impact on global scales is associated to
@@
   \item {\color{blue}  Why the deposition values for ocean + ice + land do to add up to the total
-    values reported, for all simulations?} \emph{}
+    values reported, for all simulations?}
+    Are we right to assume that the referee is trying to state that the values for ocean, ice, and land do
+    not add up to the total global ozone dry deposition? If that is the case, he is right, for we exclusively
+    selected gridboxes associated to more then 98\% with the three surface types (ocean, ice, and land), while
+    \emph{total} comprises all gridboxes. We admit that this is not clear.
+    We shall clarify this in the table caption and the revised manuscript, repeat the computation of the values
+    with the proper weighting, and adapt the numbers.
+    We may also add that \emph{ice} in this analyis only
+    referes to regions at high latitudes that are permanently covered by ice and snow and hence does not take
+    sea ice and mountain glaciers into account. In the Oslo~CTM3, as written elsewhere in the manuscript, we
+    compute dry deposition on ice and snow based on meteorological data. Therefore, the given values in Table~3
+    comprise ozone dry deposition on sea ice in the case of \emph{ocean} and on snow covered land during winter
+    in case of \emph{land}.
+
   \item {\color{blue}  The new land-based deposition values are much lower than what has
@@
 
-\item {\color{blue}  Section 2.2: Since the present paper is about ozone dry deposition, this section
-seems like a distraction and hence should be omitted.} \emph{}
-\item {\color{blue}  Page 14, lines 15 -- 18: Anthropogenic, biomass burning, and biogenic emis-
+\item {\color{blue}  P11 Section 2.2: Since the present paper is about ozone dry deposition, this section
+  seems like a distraction and hence should be omitted.}
+  The referee is right in his/her assessment. We therefore omit this section in the revised version of our manuscript.
+
+\item {\color{blue}  P14 L15-18: Anthropogenic, biomass burning, and biogenic emis-
 sions are included in the model. How are other emissions such as soil NOx, wetland
-methane, and oceanic methane and CO specified?} \emph{}
-\item {\color{blue}  Page 15, line 4: The statement “Accidentally, we have used emissions for the year
-2014 instead of 2005.” It is not clear what the consequences on the results are of this? } \emph{}
-\item {\color{blue}  Section 3.2.1: In the Fig 5 discussion, although snow and ice is discussed, there is
+methane, and oceanic methane and CO specified?}
+  Emissions from soil and wetlands are computed by MEGAN. $\mathrm{NO_x}$ emissions therein are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
+  For oceanic emissions of $\mathrm{CO}$, we use predifined global fields of POET (available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
+$\mathrm{CH_4}$ from oceans is taken from surface data from HYMN (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}), given for years 2003--2005.
+  We shall include this information in the revised manuscript.
+
+\item {\color{blue}  P15 L4: The statement “Accidentally, we have used emissions for the year
+  2014 instead of 2005.” It is not clear what the consequences on the results are of this? }
+  \emph{}
+
+\item {\color{blue}  Section 3.2.1: }
+  \begin{itemize}
+  \item {\color{blue}  Section 3.2.1:In the Fig 5 discussion, although snow and ice is discussed, there is
     no discussion on the oceanic differences between the present study and Hardacre et
-al. (2015). This is particularly important for the Southern latitudes.} \emph{}
-\item {\color{blue}  Section 3.2.1: The Hardacre et al. (2015) simulations were for the year 2001,
+    al. (2015). This is particularly important for the Southern latitudes.}
+    \emph{}
+
+  \item {\color{blue}  The Hardacre et al. (2015) simulations were for the year 2001,
     whereas the present study is mostly for the year 2015 emissions (see Table 1) driven
@@
     are based on multi-year data. The authors should discuss the implications of these
-differences about different years on the deposition results presented (e.g. uncertainty).} \emph{}
-\item {\color{blue}  Page 24, lines 3 -- 4: “The annual amount of ozone dry deposition decreases by up
-to 100\% changing from the old dry deposition scheme to the new one.” Table 3 does
+    differences about different years on the deposition results presented (e.g. uncertainty).}
+    \emph{}
+  \end{itemize}
+
+\item {\color{blue}  P24 L3--4: \emph{“The annual amount of ozone dry deposition decreases by up
+to 100\% changing from the old dry deposition scheme to the new one.”} Table 3 does
 not support this, but this may be true for some surface types. So please qualify the
-statement.} \emph{}
-\item {\color{blue}  Page 24, line 15: “Most of the decrease in ozone dry deposition in the Oslo CTM3
-can be attributed to changes in dry deposition velocities over the ocean and deserts.”
+statement.}
+  \emph{}
+
+\item {\color{blue}  P24 L15: \emph{“Most of the decrease in ozone dry deposition in the Oslo CTM3
+can be attributed to changes in dry deposition velocities over the ocean and deserts.”}
 What are the dominant factors in these changes? For example, is it mostly the surface
-resistance (Rc) term? For the ocean, it is likely to be Rc. For deserts, maybe Rb? Is it
-possible to quantify these differences in the resistance terms?} \emph{}
-\item {\color{blue}  Page 24, line 24: “2-layer gas exchange with ocean waters (Luhar et al., 2017).”
+resistance ($R_c$) term? For the ocean, it is likely to be $R_c$. For deserts, maybe $R_b$? Is it
+possible to quantify these differences in the resistance terms?}
+  \emph{}
+
+\item {\color{blue}  P24 L24: \emph{“2-layer gas exchange with ocean waters (Luhar et al., 2017).”}
 As mentioned earlier, Luhar et al. (2018) has derived a more realistic process-based
 deposition scheme for the ocean, but the results for deposition velocity do not seem to
-be too different from those in Luhar et al. (2017).} \emph{}
-\item {\color{blue}  Page 25, lines 11 -- 12: The comment “This is most likely reflecting the ongoing
+be too different from those in Luhar et al. (2017).}
+  \emph{}
+
+\item {\color{blue}  P25 L11--12: The comment \emph{“This is most likely reflecting the ongoing
 industrialization process of countries in the southern hemisphere and the commitment
 and implementation of air quality regulations of industrialized nations in the northern
-hemisphere” is quite speculative and may be omitted.} \emph{}
-\item {\color{blue}  Eq. (13) cf. Eq. (14): $g_STO or G_sto$ -- use consistency with notation.} \emph{}
-\item {\color{blue}  The first half of the abstract, the text before “In this paper...,” is introductory
-material and can be deleted.} \emph{}
-\item {\color{blue}  In the abstract (lines 15-16), it is better to say “...leading to an increase in surface
-ozone of up to 100\% in some regions.”} \emph{}
-\item {\color{blue}  Page 22, line 7: “At about 4 of 6 sites.” About? Not sure?} \emph{}
+hemisphere”} is quite speculative and may be omitted.}
+  \emph{}
+
+\item {\color{blue}  Eq. (13) cf. Eq. (14): $g_\mathrm{STO}$ or $G_\mathrm{sto}$ -- use consistency with notation.}
+  Thanks for pointing this out. We will change this in the revised version of the manuscript.
+
+\item {\color{blue}  The first half of the abstract, the text before \emph{“In this paper...,”} is introductory
+  material and can be deleted.}
+  Thanks for pointing this out. We will remove it in the revised version.
+
+\item {\color{blue}  Abstract L15--16: it is better to say “...leading to an increase in surface
+  ozone of up to 100\% in some regions.”}
+  We follow the advice and change the sentence accordingly.
+
+\item {\color{blue}  P22 L7: \emph{“At about 4 of 6 sites.” About? Not sure?}}
+  Thanks for pointing out the missplaced \emph{"about"} in this context. We are certain about the number.
+
 \end{itemize}

### 2019-04-04 DryDep.tex
Continue to answer referee questions. Revise paper and recompute ozone dry dep to different surfaces.

@@
 The conceptual dry deposition velocities are calculated following a resistance-analogous approach wherein aerodynamic, quasi laminar, and canopy resistances are key components, but these are hard to measure explicitly.
-In this paper, we present an update of the dry deposition scheme implemented in the Oslo~CTM3.
+We present an update of the dry deposition scheme implemented in the Oslo~CTM3.
 We change from a purely empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere and vegetation into account.
@@
 We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies and also estimate the impact on the modeled ozone concentrations at the surface.
-We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-47\,\unit{\%}$), leading to an increase in surface ozone of up to $100\,\unit{\%}$.
+We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-47\,\unit{\%}$), leading to an increase in surface ozone of up to $100\,\unit{\%}$ in some regions.
 While high sensitivity to changes in dry deposition to vegetation is found in the tropics, the largest impact on global scales is associated to changes in dry deposition to the ocean and deserts.
@@
 It is estimated that about $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
-A newer study by \citet{ACPLuhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
+A newer study by \citet{ACP:Luhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
 Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
@@
 Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation.
-To reflect part of the underlying mechanism, the stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
-%
-\begin{equation}
-  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
+To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
+%
+\begin{equation}
+  G_\text{sto} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
   \label{eq:stomatal}
@@
 
-\subsection{Dry deposition scheme for other gases and aerosols}
-For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}.
-Since the dry deposition velocity for \chem{CO} is very small, we use a prescribed value for $v_\text{DD}^\chem{CO}$ of $0.03\,\unit{cm\,s^{-1}}$.
-
-The EMEP scheme for aerosols (BC, OC, \chem{SO_4}, \chem{MSA}, SOA) defines the surface deposition velocity $V_\text{ds}$ as
-\begin{equation}
-  \frac{V_\text{ds}}{u_*} =
-  \begin{cases}
-    a_1 &  \emph{if} \quad L \ge 0\,\unit{m}, \\
-    a_1 \cdot F_N \left[ 1 + \left(\frac{a_2}{25}\right)^{2/3}\right] & \emph{if} \quad -25\,\unit{m} < L < 0\,\unit{m}, \\
-    a_1 \cdot F_N \left[ 1 + \left(\frac{-a_2}{L}\right)^{2/3}\right] & \emph{if} \quad L \le -25\,\unit{m}, \\
-  \end{cases}
-\end{equation}
-wherein $F_N = 3$ for fine nitrate and ammonium and $F_N = 1$ for all other species.
-$L$ is again the Obukhov length.
-$a_2 = 300\,\unit{m}$, while $a_1$ differs for forest and non-forest.
-To account for hydrophilic and hydrophobic BC/OC aerosols on dry and wet surfaces, we diverge slightly from the EMAP scheme in the definition of the parameter $a_1$.
-From pre-studies with an aerosol microphysic model, we find
-\begin{equation}
-  a_1^\text{hydrophob.} = 0.025\,\unit{cm\,s^{-1}} \cdot \overline{u_*}
-\end{equation}
-and
-\begin{equation} 
-  a_1^\text{hydrophil.} = 0.2\,\unit{cm\,s^{-1}} \cdot \overline{u_*},
-\end{equation}
-with the annual mean friction velocity $\overline{u_*}$. Our definition of $a_1$ is then:
-\begin{equation}
-  a_1 =
-  \begin{cases}
-    \emph{if} \quad \emph{land} &
-    \begin{cases}
-      \emph{if} \quad \emph{non-forest} &  a_1^\text{hydrophob.}\\
-      \emph{if} \quad \emph{forest} &
-      \begin{cases}
-        \emph{if} \quad \emph{dry} & \max{\left\{a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10}\right\}}\\
-        \emph{if} \quad \emph{wet} & 1.5\cdot\max{\left\{a_1^\text{hydrophob.}, 0.008\cdot\frac{\text{SAI}}{10}\right\}}\\
-    \end{cases}\\
-    \end{cases}\\
-    \emph{if} \quad \emph{ocean} &
-    \begin{cases}
-      a_1^\text{hydrophil.}\\
-      a_1^\text{hydrophob.}\\
-    \end{cases}\\
-    \emph{if} \quad \emph{ice} &
-    \begin{cases}
-      \emph{if} \quad \theta \le 0\,\unit{^\circ C} & a_1^\text{hydrophob.}\\
-      \emph{if} \quad \theta > 0\,\unit{^\circ C} &
-      \begin{cases}
-        a_1^\text{hydrophil.}\\
-        a_1^\text{hydrophob.}\\
-      \end{cases}\\
-    \end{cases}\\
-  \end{cases}
-  \label{eq:dd_aerosol_a1}
-\end{equation}
-%
-The total aerosol surface deposition is defined as
-\begin{equation}
- V_\text{ds}^\text{tot} = \sum_{N=0}^{N_\text{max}} V_\text{ds}(N) \cdot f_L(N).
-\end{equation}
-%
 \begin{figure*}[t]
@@
 Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$.
-The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process-modeling.
+Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$.
+The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling.
 The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
 But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition.
+In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analyisis.
+This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
 %
 \begin{table*}[t]
-\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
-\begin{tabular}{lccccccccccccr}
+\caption{Total ozone dry deposition for the respective model experiment in \unit{Tg\,a^{-1}}.
+The global ozone dry deposition has been weighted by ocean, ice and, land fraction in each gridbox, respectively.
+\emph{Ice} herein refers to regions at high latitudes that are permanently covered by ice and snow.}
+\begin{tabular}{lccccccccc|cccr}
 \tophline
@@
 \middlehline
-Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
-EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
-EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
-EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
-EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
-EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
-EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
-EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
-EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
+Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
+EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
+EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
+EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
+EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
+EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
+EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
+EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
+EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
 \bottomhline
@@
 The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}.
-At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
+At $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
 We use a $\chi^2$-test
@@
 With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new Oslo~CTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
-Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to revise the dry deposition scheme for these regimes and add more process-oriented formulations, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017, ACP:Luhar2018}, wave braking and spray \citep{ACP:Pozzer2006}.
 

### 2019-04-05 version_2/response_version_1.tex
Continue to write the response. Added Torp ozone burden table.

@@
     attribute this to the changes in the updated scheme for the desert surface type. However,
-    the paper does not provide any observational support to back this up. Are there
-    any relevant deposition measurements (velocity or flux) that can be used for this purpose?
-    At least, some comparison with ozone measurements (or even $\mathrm{O_3}$ reanalyses)
+    the paper does not provide any observational support to back this up. }
+    \begin{itemize}
+    \item {\color{blue}  Are there any relevant deposition measurements (velocity or flux) that can be used for this purpose?}
+      The only paper we know of is a study by G\"{u}sten~et~al.~(1996) in which measurements of ozone concentrations
+      and fluxes onto the Sahara desert are described and dry deposition velocities deduced. We can include a more
+      thorough discussion of our results with respect to the observed fluxes by G\"{u}sten~et~al.~(1996) in the
+      revision of our manuscript. 
+    \item {\color{blue}  At least, some comparison with ozone measurements (or even $\mathrm{O_3}$ reanalyses)
       should be provided for this surface type (and perhaps others) to see if the model is
-    heading in the right direction with the updated deposition scheme.} \emph{}
+      heading in the right direction with the updated deposition scheme.}
+      Thank you for the advise. We will look into this.
+    \end{itemize}
   \item {\color{blue}  It will also be useful to report the global ozone burden from the various
     simulations.}
-    \emph{}
+    Since Table~3 is already at maximum width with respect to the the page width, we shall show the global
+    tropospheric ozone burden in a seperate table (\ref{tab:trop_ozone_burden}). If we compare our results
+    with Stevenson~et~al.~(2006)
+    ($344\pm 39\,\mathrm{Tg}$) and the number given in IPPC AR5 (2013) ($337\pm 23\,\mathrm{Tg}$), we find
+    that the ozone burden in the Oslo~CTM3 is higher then the model average from the start (\emph{Wesely scheme}).
+    The implementation of the \emph{EMEP scheme}
+    increases the tropospheric burden by roughly 8\,\% (compare \emph{Wesely type} and EMEP\_full).
+    \begin{table*}[h]
+        \caption{Annual mean tropospheric ozone burden for all experiments and $1 \sigma$ standart deviation.}
+        \centering
+        \begin{tabular}{lrcl}
+          \hline
+          Experiment & \multicolumn{3}{c}{Trop. $\mathrm{O_3}$}\\
+          &  \multicolumn{3}{c}{(Tg)}\\
+          \hline
+          Wesely type & 361 & $\pm$ & 21\\
+          EMEP\_full & 392 & $\pm$ & 28\\
+          EMEP\_offLight & 388 & $\pm$ & 26\\
+          EMEP\_offPhen & 392 & $\pm$ & 27\\
+          EMEP\_SWVL4 & 402 & $\pm$ & 31\\
+          EMEP\_ppgs & 392 & $\pm$ & 28\\
+          EMEP\_ppgssh & 391 & $\pm$ & 27\\
+          EMEP\_ppgssh\_ice & 403 & $\pm$ & 31\\
+          EMEP\_ppgs\_2005 & 386 & $\pm$ & 26\\
+          \hline
+        \end{tabular}
+        \label{tab:trop_ozone_burden}
+    \end{table*}
   \end{itemize}

### 2019-04-05 DryDep.tex
Continue to write the response. Added Torp ozone burden table.

@@
     \tophline
-\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
+    \multirow{3}{*}{Experiment} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
     & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\

### 2019-04-08 version_2/response_version_1.tex
Continue to respond to referee questions.

@@
 We thank the referee for his/her useful comments and we will take them into account in the revised version of this paper.
-In the following, we will answer the questions in detail.
+In the following, we will respond to the questions in detail.
 \begin{itemize}
@@
     changes in dry deposition to the ocean and deserts.”} The authors do not provide details
-  in the paper as to what has changed in the updated scheme for such an impact.
+  in the paper as to what has changed in the updated scheme for such an impact.}
+  We elaborate on the source of the \emph{"largest impact on global scales"} in Section~3.2.2 (P18 L4--11).
+But we have to admit that it might not be clear to which resistance term the changes in prescribed dry deposition velocities apply.
+In our formulation $R_c$ includes both stomatal and non-stomatal conductance.
   \begin{itemize}
-  \item Is it the surface resistance ($R_c$) value? Or the other two resistances ($R_a$ and $R_b$)?
-  \item What are the typical values?
-  \item What value of $R_c$ for water has been used and
-  \item how it compares with the value used in the Wesely scheme?
+  \item {\color{blue} Is it the surface resistance ($R_c$) value? Or the other two resistances ($R_a$ and $R_b$)? What are the typical values?}
+    It would be indeed interesting to look at the resistence terms seperately. However, they are not available in our output files.
+    Technically, it would be possible to force the model output but such would involve redoing the experiments and run for at least a couple of (model) weeks. 
+  \item {\color{blue} What value of $R_c$ for water has been used and how it compares with the value used in the Wesely scheme?}
+    We have given the prescribed dry deposition velocities in Section~3.2.2 for the \emph{Wesely scheme}:
+    \begin{itemize}
+    \item $v_\mathrm{desert}^\mathrm{O_3}\equiv v_\mathrm{tundra}^\mathrm{O_3} = 0.26\,\mathrm{cm\,s^{-1}}$,
+    \item $v_\mathrm{ice/snow}^\mathrm{O_3}\equiv v_\mathrm{water}^\mathrm{O_3}=0.07\,\mathrm{cm\,s^{-1}}$
     \end{itemize}
-}
-  \emph{}
+    For the \emph{EMEP scheme}:
+    \begin{itemize}
+    \item $v_\mathrm{desert}^\mathrm{O_3}=0.05\,\mathrm{cm\,s^{-1}}$,
+    \item $v^\mathrm{O_3}_\mathrm{tundra} = 0.24\,\mathrm{cm\,s^{-1}}$,
+    \item $v_\mathrm{ice/snow}^\mathrm{O_3}\equiv v_\mathrm{water}^\mathrm{O_3}=0.05\,\mathrm{cm\,s^{-1}}$.
+    \end{itemize}
+    They are dominated by the prescribed values of TODO...
+  \end{itemize}
 
@@
     Since Table~3 is already at maximum width with respect to the the page width, we shall show the global
-    tropospheric ozone burden in a seperate table (\ref{tab:trop_ozone_burden}). If we compare our results
+    tropospheric ozone burden in a seperate table (see the following Table~\ref{tab:trop_ozone_burden}). If we compare our results
     with Stevenson~et~al.~(2006)
@@
   $\psi_m((z-d)/L)$, and the sign of the third term on the right-hand side should be positive
-  (not negative). Given that $(z-d) > z_0$ (assuming the model is formulated correctly), Eq. (2) should always yield positive values.
+  (not negative). Given that $(z-d) > z_0$ (assuming the model is formulated correctly), Eq. (2) should always yield positive values.\\
   Eqs. (3--5): I am not sure why Monteith (1973) needs to be invoked here. Given that
@@
   not a constant (depends on the surface type).}
-  \emph{}
+  The reviewer is indeed correct that this equation is wrong. In fact, we originally
+  used Eq.~(2) with $L$ in denominator for the second term. The sign error was likely the reason why the \emph{Monteith method}
+  was chosen. Certainly, an update shall be considered in the future, but it is not feasible to
+  redo all simulations now. In the revised version of the manuscript, we change the text from \emph{"In Simpson et al.
+  (2003,2012) it is described as [...] fall back to the [...]"} to \emph{"For technical reasons, we
+have used the [...]"}.
 
@@
   and move it to Section 2.}
-  Since we intent to properly introduce our model (Oslo CTM3) in Section~2, we rather
-  move the sentence in L30-32 about polar boundary layer ozone depletion to Section~2.
-  Regarding that the influence of VSLS on tropospheric ozone is indirect, this reference has not been
-  well placed in this section and will be moved to the discussions (Section~4).
+  \begin{itemize}
+  \item The Oslo CTM3 is properly introduced in Section~2. Hence, we move the sentence in L30-32 about \emph{polar boundary layer ozone depletion} to Section~2:
+    \emph{"Although the ozone depleting events in the polar boundary layer (Section~1) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now."}
+  \item Given that the influence of VSLS on tropospheric ozone is indirect (through depletion of ozone in the upper troposphere -- lower stratosphere and STE), this reference (L32--33) rather belongs to the discussion in Section~4 and will be moved:
+    \emph{"In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
+affected by very short-lived ozone depleting substances (VSLS) (Warwick et al., 2006; Ziska et al., 2013; Hossaini et al., 2016; Falk et al., 2017) and not taken into account in the Oslo~CTM3."}
+  \end{itemize}
 
@@
     \begin{itemize}
-    \item A newer study by Luhar et al. (2018), however, indicates much lower amounts ($722.8 \pm 87.3\,\mathrm{Tg\,a^{-1}}$).
-    \item Based on the global atmospheric composition reanalysis performed in the\\ ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, Luhar et al.
-(2018) found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\mathrm{Tg\,a^{-1}}$.
-    \item But also the results of  Luhar et al.
-(2017, 2018) yield a $(19-27)\,\mathrm{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\mathrm{\%}$ of the total annual ozone dry deposition.
+    \item "A newer study by Luhar et al. (2018), however, indicates much lower amounts ($722.8 \pm 87.3\,\mathrm{Tg\,a^{-1}}$)."
+    \item "Based on the global atmospheric composition reanalysis performed in the\\ ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, Luhar et al.
+(2018) found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\mathrm{Tg\,a^{-1}}$."
+    \item "But also the results of  Luhar et al.
+(2017, 2018) yield a $(19-27)\,\mathrm{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\mathrm{\%}$ of the total annual ozone dry deposition."
     \end{itemize}
@@
 methane, and oceanic methane and CO specified?}
-  Emissions from soil and wetlands are computed by MEGAN. $\mathrm{NO_x}$ emissions therein are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
-  For oceanic emissions of $\mathrm{CO}$, we use predifined global fields of POET (available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
-$\mathrm{CH_4}$ from oceans is taken from surface data from HYMN (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}), given for years 2003--2005.
+  Emissions from soil and wetlands are computed by MEGAN. Resultant $\mathrm{NO_x}$ emissions are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
+  For oceanic emissions of $\mathrm{CO}$, we use predifined global fields (POET, available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
+$\mathrm{CH_4}$ for oceans is taken from surface data from NASA's HYMN project (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}) given for the years 2000--2004.
 We shall include this information in the revised manuscript.
@@
 statement.}
-  \emph{}
+  We have indeed not specified our statement in the mentioned sentence, while we had done so elsewhere in the manuscript (P15 L12). We complete our statement in the revised version of the manuscript:
+  \emph{"[...] ozone dry deposition decreases by up to $100\,\%$ over all major desert areas [...]. At the same time, it increases over tropical forest.}
 
@@
 be too different from those in Luhar et al. (2017).}
-  \emph{}
+  We acknowledge Luhar et al. (2018) and write:
+  \emph{“2-layer gas exchange with ocean waters (Luhar et al., 2017, 2018).”}
 
@@
 hemisphere”} is quite speculative and may be omitted.}
-  \emph{}
+  We follow the kind advise of the referee and remove the sentence in the revised version of the manuscript.
 
@@
   material and can be deleted.}
-  Thanks for pointing this out. We will remove it in the revised version.
+  This is indeed the case and we will remove it in the revised version.
 
@@
   ozone of up to 100\% in some regions.”}
-  We follow the advice and change the sentence accordingly.
+  We follow the advice of the referee and change the sentence accordingly.
 
 \item {\color{blue}  P22 L7: \emph{“At about 4 of 6 sites.” About? Not sure?}}
-  Thanks for pointing out the missplaced \emph{"about"} in this context. We are certain about the number.
+  Thanks for pointing out the misplaced \emph{"about"} in this sentence. We are certain regarding that number.
 

### 2019-04-08 DryDep.tex
Continue to respond to referee questions.

@@
 \introduction  %% \introduction[modified heading if necessary]
+\label{sec:intro}
 Ozone is an important trace gas for all lifeforms on Earth.
@@
 A publication focusing on this is planed.
-Although the above mentioned ozone depleting events are important to understand surface ozone abundance in Arctic regions, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.
+Although the ozone depleting events in the polar boundary layer (Section~\ref{sec:intro}) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.\\
 While the meteorological data driving the Oslo~CTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
@@
 \label{subsubsec:Ra}
-The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. In \citet{WASP:Simpson2003,ACP:Simpson2012} it is described as
-\begin{equation}
-  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{z_0}\right)-\Psi_m\left(\frac{z_0}{L}\right)}\right],
-\end{equation}
-with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a constant $d$ (typically $0.7\,\unit{m}$), and the Obukhov length $L$.
+The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. %In \citet{WASP:Simpson2003,ACP:Simpson2012} it is described as
+%\begin{equation}
+%  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{z_0}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
+%\end{equation}
+%with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\,\unit{m}$), and the Obukhov length $L$.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
-For this reason, we diverge from the EMEP scheme at this point and fall back to the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
+For this reason, we diverge from the EMEP scheme at this point and fall back to
+For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
 \begin{equation}
@@
 CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
-Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}.
+Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}, while emissions from soil and wetlands are computed by MEGAN.
+Resultant $\mathrm{NO_x}$ emissions are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
+For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}.
+Emissions of \chem{CH_4} from oceans are taken from surface data from \citet{HYMN}.
 In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
@@
 
-The annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ changing from the old dry deposition scheme to the new one.
+The annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ over all major desert areas changing from the old dry deposition scheme to the new one.
+At the same time, it increases over tropical forest.
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is about $33\,\unit{\%}$ below average.
@@
 Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may hint to, e.g., differences in photo-chemistry, stratosphere-troposphere-exchange as well as to differences in aerosol loadings among many other aspects.
-Especially the STE depends on the stratospheric ozone abundance which is, e.g.
-strongly affected by the contribution of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and not taken into account in the Oslo~CTM3.
+In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
+affected by very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and not taken into account in the Oslo~CTM3.
 
@@
 In 2014, surface ozone burden is higher in the southern hemisphere and in the tropics ($\sim 5\,\unit{\%}$) compared to 2005, while it is lower in the northern hemisphere ($\sim 2\,\unit{\%}$).
-This is most likely reflecting the ongoing industrialization process of countries in the southern hemisphere and the commitment and implementation of air quality regulations of industrialized nations in the northern hemisphere.
 

### 2019-04-09 version_2/response_version_1.tex
2 questions to go...

@@
   We elaborate on the source of the \emph{"largest impact on global scales"} in Section~3.2.2 (P18 L4--11).
-But we have to admit that it might not be clear to which resistance term the changes in prescribed dry deposition velocities apply.
-In our formulation $R_c$ includes both stomatal and non-stomatal conductance.
+But we have to admit that it might not be clear to which resistance term the changes in the prescribed dry deposition velocities apply.
   \begin{itemize}
   \item {\color{blue} Is it the surface resistance ($R_c$) value? Or the other two resistances ($R_a$ and $R_b$)? What are the typical values?}
-    It would be indeed interesting to look at the resistence terms seperately. However, they are not available in our output files.
+    It would be indeed interesting to look at the resistance terms separately. However, they are not available in our output files.
     Technically, it would be possible to force the model output but such would involve redoing the experiments and run for at least a couple of (model) weeks.
+In our formulation, the surface resistance $R_c$ includes both stomatal and non-stomatal conductance.
+In case, of non-vegetated surfaces, such as \emph{desert}, \emph{ocean}, and \emph{snow/ice}, the non-stomatal conductance, $G_\mathrm{ns}^\mathrm{O_3}$, is dominant.
+For water, $R_b = 10\,\mathrm{s\,m^{-1}}$ in most cases.
+Thus, $R_c^\mathrm{O_3} \approx R_\mathrm{gs}^\mathrm{O_3} \propto (v_\mathrm{DD}^\mathrm{O_3})^{-1}$ (for $R_c$ values see Table~S2 in the manuscript Supplement~S.3).
   \item {\color{blue} What value of $R_c$ for water has been used and how it compares with the value used in the Wesely scheme?}
-    We have given the prescribed dry deposition velocities in Section~3.2.2 for the \emph{Wesely scheme}:
-    \begin{itemize}
-    \item $v_\mathrm{desert}^\mathrm{O_3}\equiv v_\mathrm{tundra}^\mathrm{O_3} = 0.26\,\mathrm{cm\,s^{-1}}$,
-    \item $v_\mathrm{ice/snow}^\mathrm{O_3}\equiv v_\mathrm{water}^\mathrm{O_3}=0.07\,\mathrm{cm\,s^{-1}}$
-    \end{itemize}
-    For the \emph{EMEP scheme}:
-    \begin{itemize}
-    \item $v_\mathrm{desert}^\mathrm{O_3}=0.05\,\mathrm{cm\,s^{-1}}$,
-    \item $v^\mathrm{O_3}_\mathrm{tundra} = 0.24\,\mathrm{cm\,s^{-1}}$,
-    \item $v_\mathrm{ice/snow}^\mathrm{O_3}\equiv v_\mathrm{water}^\mathrm{O_3}=0.05\,\mathrm{cm\,s^{-1}}$.
-    \end{itemize}
-    They are dominated by the prescribed values of TODO...
+    We have given the prescribed dry deposition velocities in Section~3.2.2 for the \emph{Wesely scheme}
+    $v_\mathrm{water}^\mathrm{O_3} = 0.07\,\mathrm{cm\,s^{-1}}$ and for the \emph{EMEP scheme}
+    $v_\mathrm{water}^\mathrm{O_3} = 0.05\,\mathrm{cm\,s^{-1}}$.
+Surface resistances are thus $R_c \approx 1429\,\mathrm{s\,m^{-1}}$ (\emph{Wesely scheme}) and $R_c = 2000\,\mathrm{s\,m^{-1}}$ (\emph{EMEP scheme}), respectively.
   \end{itemize}
@@
     with the proper weighting, and adapt the numbers.
-    We may also add that \emph{ice} in this analyis only
-    referes to regions at high latitudes that are permanently covered by ice and snow and hence does not take
+    We may also add that \emph{ice} in this analysis only
+    refers to regions at high latitudes that are permanently covered by ice and snow and hence does not take
     sea ice and mountain glaciers into account. In the Oslo~CTM3, as written elsewhere in the manuscript, we
@@
     Since Table~3 is already at maximum width with respect to the the page width, we shall show the global
-    tropospheric ozone burden in a seperate table (see the following Table~\ref{tab:trop_ozone_burden}). If we compare our results
+    tropospheric ozone burden in a separate table (see the following Table~\ref{tab:trop_ozone_burden}). If we compare our results
     with Stevenson~et~al.~(2006)
@@
     \begin{table*}[h]
-        \caption{Annual mean tropospheric ozone burden for all experiments and $1 \sigma$ standart deviation.}
+        \caption{Annual mean tropospheric ozone burden for all experiments and $1 \sigma$ standard deviation.}
         \centering
@@
   The reviewer is indeed correct that this equation is wrong. In fact, we originally
-  used Eq.~(2) with $L$ in denominator for the second term. The sign error was likely the reason why the \emph{Monteith method}
+  used Eq.~(2) with $L$ in the denominator for the second term. The sign error was likely the reason why the \emph{Monteith method}
   was chosen. Certainly, an update shall be considered in the future, but it is not feasible to
@@
     \emph{"Although the ozone depleting events in the polar boundary layer (Section~1) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now."}
-  \item Given that the influence of VSLS on tropospheric ozone is indirect (through depletion of ozone in the upper troposphere -- lower stratosphere and STE), this reference (L32--33) rather belongs to the discussion in Section~4 and will be moved:
+  \item Given that the influence of VSLS on tropospheric ozone is indirect (through depletion of ozone in the upper troposphere -- lower stratosphere and subsequent STE), this reference (L32--33) rather belongs to the discussion in Section~4 and will be moved:
     \emph{"In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
@@
     \item "A newer study by Luhar et al. (2018), however, indicates much lower amounts ($722.8 \pm 87.3\,\mathrm{Tg\,a^{-1}}$)."
-    \item "Based on the global atmospheric composition reanalysis performed in the\\ ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, Luhar et al.
+    \item "Based on the global atmospheric composition reanalysis performed in the\\ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, Luhar et al.
 (2018) found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\mathrm{Tg\,a^{-1}}$."
@@
 
-\item {\color{blue}  P14 L15-18: Anthropogenic, biomass burning, and biogenic emis-
-sions are included in the model. How are other emissions such as soil NOx, wetland
-methane, and oceanic methane and CO specified?}
+\item {\color{blue}  P14 L15-18: Anthropogenic, biomass burning, and biogenic emissions are included in the model.
+How are other emissions such as soil $\mathrm{NO_x}$, wetland methane, and oceanic methane and CO specified?}
   Emissions from soil and wetlands are computed by MEGAN. Resultant $\mathrm{NO_x}$ emissions are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
-  For oceanic emissions of $\mathrm{CO}$, we use predifined global fields (POET, available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
+  For oceanic emissions of $\mathrm{CO}$, we use predefined global fields (POET, available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
 $\mathrm{CH_4}$ for oceans is taken from surface data from NASA's HYMN project (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}) given for the years 2000--2004.
@@
 
-\item {\color{blue}  P15 L4: The statement “Accidentally, we have used emissions for the year
-  2014 instead of 2005.” It is not clear what the consequences on the results are of this? }
+\item {\color{blue}  P15 L4: The statement \emph{“Accidentally, we have used emissions for the year
+  2014 instead of 2005.”} It is not clear what the consequences on the results are of this?}
+  In the introduction to Section~3.1, we wrote \emph{"For all model integrations, the meteorological reference year is 2005.
+This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years."}.
+We shall elaborate on the discussion on the implications based on our model results (EMEP\_ppgs vs EMEP\_ppgs\_2005) in the revision of the manuscript.
+Though, the major consequence of this is that, for the majority of our model experiments, one can neither directly compare observations for the years 2005 nor for 2014 directly to the model results.
+Surface ozone observations, in fact, show a strong interannual veriability in ozone dry deposition and ozone concentrations at the sites, but studying these in detail may be well beyond the scope of this manuscript.
   \emph{}
@@
   \begin{itemize}
-  \item {\color{blue}  Section 3.2.1:In the Fig 5 discussion, although snow and ice is discussed, there is
+  \item {\color{blue}  Section 3.2.1: In the Fig 5 discussion, although snow and ice is discussed, there is
     no discussion on the oceanic differences between the present study and Hardacre et
@@
 possible to quantify these differences in the resistance terms?}
+  We have already answered the question with respect to ocean (see first bulletpoint).
+In summary, since $R_b$ is quite small in most of the cases, the dominant factor for the ozone dry deposition onto ocean is the surface resistance $R_c$ which is tabulated in Table~S2.
+Regarding ozone dry deposition onto deserts, we use Eqs.~(7--8) to deduce
+  \begin{equation}
+    R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{D_\mathrm{H_2O}}{D_i} \cdot \frac{\mathrm{Sc}_\mathrm{H_2O}}{\mathrm{Pr}}\right)^{\frac{2}{3}},
+  \end{equation}
+  with $\mathrm{Pr}=0.72$, $\kappa=0.4$, $\mathrm{Sc}_\mathrm{H_2O} = 0.6$, $D_\mathrm{H_2O}/D_i = 1.6$.
+  We estimate $u_*$ from Eq. (16.67) in Seinfeld~and~Pandis (2006)
+  \begin{equation}
+    u_*(h) = \frac{\kappa\cdot \overline{u_x}(h)}{\mathrm{ln}(h/z_0)},
+  \end{equation}
+  with $h = 8\,\mathrm{m}$, $z_0^\mathrm{desert}\approx 10^{-3}\,\mathrm{m}$, and windspeeds not exceeding a gentle breeze ($1.8\,\mathrm{km\,h^{-1}} \leq \overline{u_x}(h) \leq 28\,\mathrm{km\,h^{-1}}$), we find $272\,\mathrm{s\,m^{-1}} \geq R_b \geq 17\,\mathrm{s\,m^{-1}}$.
+This is $1-2$ orders of magnitude smaller than $R_c = 2000\,\mathrm{s\,m^{-1}}$ and thus not negligible for low windspeeds.
+In summary, $R_c$ is dominant in our formulation of dry deposition of ozone to deserts (unless we have calm wind conditions).
   \emph{}
@@
 be too different from those in Luhar et al. (2017).}
-  We acknowledge Luhar et al. (2018) and write:
-  \emph{“2-layer gas exchange with ocean waters (Luhar et al., 2017, 2018).”}
+  We acknowledge Luhar et al. (2018) an update the sentence:
+  \emph{“[...] 2-layer gas exchange with ocean waters (Luhar et al., 2017, 2018).”}
 

### 2019-04-09 DryDep.tex
2 questions to go...

@@
 with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
-The used ratios $D_\chem{H_2O}/D_i$ are tabulated in Supplement~S.1.
+The used ratios $D_\chem{H_2O}/D_i$ are tabulated in Supplement~S.1.\\
 %
@@
 For all model integrations, the meteorological reference year is 2005.
-This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years.
+This choice affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years.
 After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} is the baseline simulation for the sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}.
@@
 Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the Oslo~CTM3 (\emph{EMEP\_ppgssh\_ice}).
-Accidentally, we have used emissions for the year 2014 instead of 2005.
-Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}).
-As mentioned previously, the simulations were initially conducted with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+Accidentally, we have used emissions for the year 2014 instead of 2005 in all model experiments except for \emph{EMEP\_ppgs\_2005}.
+{\bf TODO: Actually assess the impact of the differing emissions!} As mentioned previously, the simulations were initially conducted with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
 A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).
@@
 This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
+{\bf TODO include reference to Table~\ref{tab:trop_ozone_burden} in the text.}
 %
@@
 %
-
+\begin{table*}[h]
+  \caption{Annual mean tropospheric ozone burden for all experiments and $1 \sigma$ standard deviation.}
+  \centering
+  \begin{tabular}{lrcl}
+    \tophline
+    Experiment & \multicolumn{3}{c}{Trop. \chem{O_3}}\\
+    &  \multicolumn{3}{c}{(Tg)}\\
+    \middlehline
+    Wesely type & 361 & $\pm$ & 21\\
+    EMEP\_full & 392 & $\pm$ & 28\\
+    EMEP\_offLight & 388 & $\pm$ & 26\\
+    EMEP\_offPhen & 392 & $\pm$ & 27\\
+    EMEP\_SWVL4 & 402 & $\pm$ & 31\\
+    EMEP\_ppgs & 392 & $\pm$ & 28\\
+    EMEP\_ppgssh & 391 & $\pm$ & 27\\
+    EMEP\_ppgssh\_ice & 403 & $\pm$ & 31\\
+    EMEP\_ppgs\_2005 & 386 & $\pm$ & 26\\
+    \bottomhline
+  \end{tabular}
+  \label{tab:trop_ozone_burden}
+\end{table*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-04-09 version_2/response_version_1.tex
Some corrections and added table (needs to be properly formated, though).

@@
 \usepackage{url}
+\usepackage{multirow}
 \thispagestyle{empty}
@@
 For water, $R_b = 10\,\mathrm{s\,m^{-1}}$ in most cases.
-Thus, $R_c^\mathrm{O_3} \approx R_\mathrm{gs}^\mathrm{O_3} \propto (v_\mathrm{DD}^\mathrm{O_3})^{-1}$ (for $R_c$ values see Table~S2 in the manuscript Supplement~S.3).
+Thus, $R_\mathrm{gs}^\mathrm{O_3} \approx R_c^\mathrm{O_3} \propto (v_\mathrm{DD}^\mathrm{O_3})^{-1}$ (for $R_c$ values see Table~S2 in the manuscript Supplement~S.3).
   \item {\color{blue} What value of $R_c$ for water has been used and how it compares with the value used in the Wesely scheme?}
-    We have given the prescribed dry deposition velocities in Section~3.2.2 for the \emph{Wesely scheme}
-    $v_\mathrm{water}^\mathrm{O_3} = 0.07\,\mathrm{cm\,s^{-1}}$ and for the \emph{EMEP scheme}
+    We have given the prescribed dry deposition velocities in Section~3.2.2:
+    \begin{itemize}
+      \item For the \emph{Wesely scheme}
+        $v_\mathrm{water}^\mathrm{O_3} = 0.07\,\mathrm{cm\,s^{-1}}$ and
+        \item for the \emph{EMEP scheme}
           $v_\mathrm{water}^\mathrm{O_3} = 0.05\,\mathrm{cm\,s^{-1}}$.
-Surface resistances are thus $R_c \approx 1429\,\mathrm{s\,m^{-1}}$ (\emph{Wesely scheme}) and $R_c = 2000\,\mathrm{s\,m^{-1}}$ (\emph{EMEP scheme}), respectively.
+    \end{itemize}
+    Surface resistances for water are thus $R_c \approx 1429\,\mathrm{s\,m^{-1}}$ (\emph{Wesely scheme}) and $R_c = 2000\,\mathrm{s\,m^{-1}}$ (\emph{EMEP scheme}), respectively.
   \end{itemize}
@@
     not add up to the total global ozone dry deposition? If that is the case, he is right, for we exclusively
-    selected gridboxes associated to more then 98\% with the three surface types (ocean, ice, and land), while
+    selected gridboxes associated to more then 98\% to the three surface types (ocean, ice, and land), while
     \emph{total} comprises all gridboxes. We admit that this is not clear.
-    We shall clarify this in the table caption and the revised manuscript, repeat the computation of the values
-    with the proper weighting, and adapt the numbers.
+    We shall repeat the computation of the values using the proper weighting and adapt the numbers (Table~\ref{tab:ozone_sinks}).
     We may also add that \emph{ice} in this analysis only
@@
     in case of \emph{land}.
+    \begin{table*}[t]
+  \caption{Total ozone dry deposition for the respective model experiment in $\mathrm{Tg\,a^{-1}}$.
+The global ozone dry deposition has been weighted by ocean, ice and, land fraction in each gridbox, respectively.
+\emph{Ice} herein refers to regions at high latitudes that are permanently covered by ice and snow.}
+  \begin{tabular}{lccccccccc|cccr}
+    \hline
+    \multirow{3}{*}{Experiment} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
+    & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
+    & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & $(\mathrm{\%})$\\
+    \hline
+    Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
+    EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
+    EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
+    EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
+    EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
+    EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
+    EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
+    EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
+    EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
+    \hline
+  \end{tabular}
+  {$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
+  \label{tab:ozone_sinks}
+\end{table*}
 

### 2019-04-10 version_2/response_version_1.tex
Finished first draft of comments, sent to co-author. Need to fix all the ToDos in the manuscripts.

@@
 In case, of non-vegetated surfaces, such as \emph{desert}, \emph{ocean}, and \emph{snow/ice}, the non-stomatal conductance, $G_\mathrm{ns}^\mathrm{O_3}$, is dominant.
-For water, $R_b = 10\,\mathrm{s\,m^{-1}}$ in most cases.
+For water, $R_b = 10\,\mathrm{s\,m^{-1}}$ in most cases which is $2$ orders of magnitude smaller than $R_c$.
 Thus, $R_\mathrm{gs}^\mathrm{O_3} \approx R_c^\mathrm{O_3} \propto (v_\mathrm{DD}^\mathrm{O_3})^{-1}$ (for $R_c$ values see Table~S2 in the manuscript Supplement~S.3).
@@
     values reported, for all simulations?}
-    Are we right to assume that the referee is trying to state that the values for ocean, ice, and land do
+    Are we right to assume that the referee asks why the values for ocean, ice, and land do
     not add up to the total global ozone dry deposition? If that is the case, he is right, for we exclusively
@@
 \emph{Ice} herein refers to regions at high latitudes that are permanently covered by ice and snow.}
-  \begin{tabular}{lccccccccc|cccr}
+  \begin{tabular}{lccccccccc}%|cccr}
     \hline
-    \multirow{3}{*}{Experiment} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
-    & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
-    & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & $(\mathrm{\%})$\\
+    \multirow{3}{*}{Experiment} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land}\\% & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
+    & NH & SH & Global & NH & SH & Global & NH & SH & Global\\% & NH & SH & Global\\
+    & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$}\\ %& \multicolumn{3}{c}{$(\mathrm{Tg\,a^{-1}})$} & $(\mathrm{\%})$\\
     \hline
-    Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
-    EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
-    EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
-    EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
-    EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
-    EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
-    EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
-    EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
-    EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
+    Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3\\% & 612.4 & 347.9 & 960.2 & 46.8\\
+    EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6\\% & 408.6 & 245.6 & 654.2 & 0.0\\
+    EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0\\% & 451.3 & 267.2 & 718.6 & 9.8\\
+    EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7\\% & 413.9 & 249.9 & 663.7 & 1.5\\
+    EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8\\% & 417.2 & 250.7 & 667.9 & 2.1\\
+    EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9\\% & 411.4 & 246.0 & 657.4 & 0.5\\
+    EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6\\% & 419.4 & 253.6 & 673.0 & 2.9\\
+    EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2\\% & 413.0 & 251.1 & 664.1 & 1.5\\
+    EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7\\% & 409.2 & 240.7 & 649.9 & -0.7\\
     \hline
   \end{tabular}
-  {$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
+  %{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
   \label{tab:ozone_sinks}
@@
     \item {\color{blue}  Are there any relevant deposition measurements (velocity or flux) that can be used for this purpose?}
-      The only paper we know of is a study by G\"{u}sten~et~al.~(1996) in which measurements of ozone concentrations
-      and fluxes onto the Sahara desert are described and dry deposition velocities deduced. We can include a more
-      thorough discussion of our results with respect to the observed fluxes by G\"{u}sten~et~al.~(1996) in the
-      revision of our manuscript. 
+      G\"{u}sten~et~al.~(1996) conducted measurements of ozone concentrations
+      and fluxes onto the Sahara desert and deduced dry deposition velocities.
+      We have not found any recent paper conducting field experiments in desert regions. We can include a more
+      thorough discussion of our results with respect to the observed fluxes by G\"{u}sten~et~al.~(1996) and the
+      references therein the revision of our manuscript. 
     \item {\color{blue}  At least, some comparison with ozone measurements (or even $\mathrm{O_3}$ reanalyses)
@@
   made here as \emph{“...we have not implemented any parameterization of these processes
-    in the Oslo CTM3 as of now.”} Some brief introductory text is required here (or better at
-  the start of the paragraph) to introduce the model properly. Also, the text between lines
+    in the Oslo CTM3 as of now.”} }
+  \begin{itemize}
+  \item {\color{blue} Some brief introductory text is required here (or better at
+    the start of the paragraph) to introduce the model properly.}
+    The Oslo CTM3 is properly introduced in Section~2. Hence, we move the sentence in L30-32 about \emph{polar boundary layer ozone depletion} to Section~2:
+    \emph{"Although the ozone depleting events in the polar boundary layer (Section~1) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now."}
+  \item {\color{blue} Also, the text between lines
     25 -- 33 on what is not considered in the model is too detailed to be here, so shorten
     and move it to Section 2.}
-  \begin{itemize}
-  \item The Oslo CTM3 is properly introduced in Section~2. Hence, we move the sentence in L30-32 about \emph{polar boundary layer ozone depletion} to Section~2:
-    \emph{"Although the ozone depleting events in the polar boundary layer (Section~1) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now."}
-  \item Given that the influence of VSLS on tropospheric ozone is indirect (through depletion of ozone in the upper troposphere -- lower stratosphere and subsequent STE), this reference (L32--33) rather belongs to the discussion in Section~4 and will be moved:
+    Given that the influence of VSLS on tropospheric ozone is indirect (through depletion of ozone in the upper troposphere -- lower stratosphere and subsequent STE), this reference (L32--33) rather belongs to the discussion in Section~4 and will be moved:
     \emph{"In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
-affected by very short-lived ozone depleting substances (VSLS) (Warwick et al., 2006; Ziska et al., 2013; Hossaini et al., 2016; Falk et al., 2017) and not taken into account in the Oslo~CTM3."}
+affected by very short-lived ozone depleting substances (VSLS) (Warwick et al., 2006; Ziska et al., 2013; Hossaini et al., 2016; Falk et al., 2017) and is not taken into account in the Oslo~CTM3."}
   \end{itemize}
@@
 This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years."}.
-We shall elaborate on the discussion on the implications based on our model results (EMEP\_ppgs vs EMEP\_ppgs\_2005) in the revision of the manuscript.
-Though, the major consequence of this is that, for the majority of our model experiments, one can neither directly compare observations for the years 2005 nor for 2014 directly to the model results.
-Surface ozone observations, in fact, show a strong interannual veriability in ozone dry deposition and ozone concentrations at the sites, but studying these in detail may be well beyond the scope of this manuscript.
+We will clarify the sentence with respect to probable consequences.
   \emph{}
@@
     al. (2015). This is particularly important for the Southern latitudes.}
+    We agree that ozone dry deposition to oceans is of high importance for the southern hemisphere, last but not least in the zonal band $(50-70)^\circ\mathrm{S}$.
+In fact, we have mainly discussed out results with respect to seasonal cycles of dry deposition velocities onto ocean found in Hardacre et al.
+(2015) (Fig.
+7; P21 L1--10: \emph{"Similarly, the dry deposition velocities over water differ.
+[...]"}).
+We will include a discussion based on Fig.~5 in the revision of the manuscript.
     \emph{}
@@
     differences about different years on the deposition results presented (e.g. uncertainty).}
+    We regard this remark as a follow-up of the question raised regarding P15 L4.
+    We will elaborate on the discussion regarding the implications based on our model results (EMEP\_ppgs vs EMEP\_ppgs\_2005) in the revision of the manuscript.
+Though, the major consequence of this is that, for the majority of our model experiments, one can neither directly compare to observations for the years 2005 and 2014 nor to other the model results.
+Emission inventories may always be lacking in certain chemical species and the non-linearities in ozone formation and destruction make ozone concentrations sensitive to both emissions of precursors and meteorological conditions.
+Surface ozone observations, in fact, show a strong variability in ozone dry deposition and ozone concentrations at the sites.
+But studying these in detail may be well beyond the scope of this manuscript.
     \emph{}
@@
 possible to quantify these differences in the resistance terms?}
-  We have already answered the question with respect to ocean (see first bulletpoint).
+  We have already answered the question with respect to ocean (see first bullet point).
 In summary, since $R_b$ is quite small in most of the cases, the dominant factor for the ozone dry deposition onto ocean is the surface resistance $R_c$ which is tabulated in Table~S2.
@@
   \end{equation}
-  with $h = 8\,\mathrm{m}$, $z_0^\mathrm{desert}\approx 10^{-3}\,\mathrm{m}$, and windspeeds not exceeding a gentle breeze ($1.8\,\mathrm{km\,h^{-1}} \leq \overline{u_x}(h) \leq 28\,\mathrm{km\,h^{-1}}$), we find $272\,\mathrm{s\,m^{-1}} \geq R_b \geq 17\,\mathrm{s\,m^{-1}}$.
-This is $1-2$ orders of magnitude smaller than $R_c = 2000\,\mathrm{s\,m^{-1}}$ and thus not negligible for low windspeeds.
+  with $h = 8\,\mathrm{m}$, $z_0^\mathrm{desert}\approx 10^{-3}\,\mathrm{m}$, and wind speeds not exceeding a gentle breeze ($1.8\,\mathrm{km\,h^{-1}} \leq \overline{u_x}(h) \leq 28\,\mathrm{km\,h^{-1}}$), we find $272\,\mathrm{s\,m^{-1}} \geq R_b \geq 17\,\mathrm{s\,m^{-1}}$.
+This is $1-2$ orders of magnitude smaller than $R_c = 2000\,\mathrm{s\,m^{-1}}$ and thus not negligible for low wind speeds.
 In summary, $R_c$ is dominant in our formulation of dry deposition of ozone to deserts (unless we have calm wind conditions).
@@
 
-\item {\color{blue}  P22 L7: \emph{“At about 4 of 6 sites.” About? Not sure?}}
+\item {\color{blue}  P22 L7: \emph{“At about 4 of 6 sites.”} About? Not sure?}
   Thanks for pointing out the misplaced \emph{"about"} in this sentence. We are certain regarding that number.

### 2019-04-10 DryDep.tex
Finished first draft of comments, sent to co-author. Need to fix all the ToDos in the manuscripts.

@@
 In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
-affected by very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and not taken into account in the Oslo~CTM3.
+affected by very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and is not taken into account in the Oslo~CTM3.
 

### 2019-04-24 version_2/response_version_1.tex
Responds to referee 1.

@@
     values reported, for all simulations?}
-    Are we right to assume that the referee asks why the values for ocean, ice, and land do
-    not add up to the total global ozone dry deposition? If that is the case, he is right, for we exclusively
+    Yes, that is, for we exclusively
     selected gridboxes associated to more then 98\% to the three surface types (ocean, ice, and land), while
@@
       and fluxes onto the Sahara desert and deduced dry deposition velocities.
-      We have not found any recent paper conducting field experiments in desert regions. We can include a more
+      We have not found any recent paper conducting field experiments in desert regions. We include a more
       thorough discussion of our results with respect to the observed fluxes by G\"{u}sten~et~al.~(1996) and the
@@
   For oceanic emissions of $\mathrm{CO}$, we use predefined global fields (POET, available through ACCENT/GEIA, \url{http://accent.aero.jussieu.fr/database_table_inventories.php}).
-$\mathrm{CH_4}$ for oceans is taken from surface data from NASA's HYMN project (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}) given for the years 2000--2004.
+$\mathrm{CH_4}$ is taken from surface data from the EU project Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN; EU GOCE 037048) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA (\url{https://www.esrl.noaa.gov/gmd/ccgg/trends_ch4/}) given for the years 2000--2004.
 We shall include this information in the revised manuscript.

### 2019-04-24 DryDep.tex
Responds to referee 1.

@@
 For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}.
-Emissions of \chem{CH_4} from oceans are taken from surface data from \citet{HYMN}.
+Emissions of \chem{CH_4} from surface data from the EU project (EU GOCE 037048) Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA.
 In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.

### 2019-05-13 version_2/response_version_1_rc2.tex
Added draft for answers to ref2.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{url}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To gmd-2019-21 David Simpson (10 Apr 2019):
We appologize for the late response regarding the comments on our paper. We had to evaluate the major concerns raised in the review and found that we have to revise our model and repeat the model experiments to address all concerns. Nevertheless, we will respond to the questions in more detail in the following.

\section{Major points}
\begin{itemize}
\item {\color{blue} Mosaic formulation: [...] In any case, I think this approach has serious problems. Why average first for $G_s$ and then for $G_{ns}$ , when it is the fluxes ($F_k$ , or $V^i_{g,k}(z_\mathrm{ref}) \times \chi^i_\mathrm{avg}(z_\mathrm{ref})$) which need to be averaged? I also do not understand why they would use the same $u_*$ and $R_a$ for all land-covers. I think the authors need to make a case for their approach, or change it.}
  Thank you very much for your detailed account of the \emph{mosaic approach}. We have discussed the concerns raised within this comment and found that we may have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
\item {\color{blue} Calculation of $R_a$:}
  \begin{itemize}
  \item {\color{blue} $R_a$ in CTM3 appears to be calculated just once, and from a height of 8\,m. This means
that there is no consistency between the $R_a$ term and the underlying surface, which is
clearly wrong. The similarity equation for $R_a$ given in their Eqn~(2) is very standard and has been
used for decades (Garratt 1992). As pointed out by Ref~1, the equation as given has
errors. The correct equation will not give negative values unless presented with wrong
inputs, and I suspect that that is what has happened. It is actually difficult to tell what
was tested from the manuscript though, since they state simply that $d$ is ’typically 0.7\,m’.
Did they use $d$ properly, consistent with the underlying land-cover and its $z_0$ ? Did
they assume that their 8\,m meteorology was at a physical height of 8\,m, or at $d + 8\,\mathrm{m}$.
If the latter, which $d$?}
    We have to object slightly, because we have not presented "wrong inputs" to the algorithm. As pointed out by Ref \#1, we have made a mistake early on in the derivation of the formular to fit it into the framework of the Oslo~CTM3 (sign flip in fact). This let to negative values. 
    Since we are going to revise the model, we will address this issue and fix the calculation of $R_a$.

    \item {\color{blue} Lines 19--30 of this section explain the Monteith alternative, but in a rather confusing
way. For example, when is $z_0$ ever zero, as stated on line 23, or why does $\partial_z R_a \rightarrow R_a $
for finite $z$? (I know what they intend to say, but it isn’t at all clear.) In any case, here
the authors end up with a stability-independent equation for $R_a$ , without mentioning or
discussing that fact. This very shallow layer is also very problematic for deposition calculations in gen-
eral, since the model cell seems to be run here with horizontal dimensions of $2.25 \times 2.25^\circ$ ,
or about $250 \times 250\,\mathrm{km}$ near the equator, but a vertical mid-level (CTM3’s $z_\mathrm{ref}$ ) of
just 8\,m. Now, profiles of wind and depositing gases are very sensitive to the underlying
surface, and should be very different for forests or lakes for example. Any wind-speed
or friction velocity calculated from a model of such large horizontal resolution will nec-
essarily give values at 8\,m which reflect the whole grid. Deposition rates for a specific
land-cover will vary enormously depending on what else is in the grid-square. (Al-
though not strictly comparable, we showed in Schwede et al. (2018) that differences
between the grid-average and forest specific deposition rates of N-compounds could be
as much as a factor of two and up to more than a factor of five in extreme cases. These
differences were largely dependent on how much forest occupied each grid cell.)}
\end{itemize}      
\item {\color{blue} Why so much focus on sea areas?: The text seems rather unbalanced with regard to the different land-covers. Sect. 2 uses 1/2 page on various $z_0$ corrections for oceans, but say nothing about the ecosystem
where ozone is expected to deposit at high rates: forests, crops, and other terrestrial
ecosystems. The supplementary has three Figs related to this oceanic deposition. Why?}
\item {\color{blue} Use of the term 'EMEP scheme'?: Sect. 3 discusses the comparisons of $V_g$ in terms of ’EMEP scheme’ versus ’We-
sely scheme’, and sensitivity tests are named e.g. ’EMEP\_offlight’. As noted above
the scheme implemented in CTM3 is very different to that implemented in the EMEP
model, so this is very misleading. Please rename your scheme to something else.
I am worried that readers might get the impression that it is the EMEP scheme which
is being tested here, but it certainly is not. [...]}
  We appologize for the missleading naming of the new dry deposition scheme in the Oslo~CTM3 which is (partly) based on the formulations in the publication the referee refers to as \emph{S2012}. By no means have we meant to offend any of the original authors of the EMEP/MSC-W model nor intented to misguide the readers. We will rename the revised scheme apropriately. (\emph{mOSaic} maybe?) 
\item {\color{blue} Reproduction of material from S2012:}
  \begin{itemize}
  \item {\color{blue} As far as I can see, Table S1, S2 and S3 are taken directly from S2012, with no change to parameters. It is not usual to copy tables from the work of other authors in this way.
    Just refer to S2012 (and give Table number as help).}
  \item {\color{blue} Many of the equations are from S2012, and many not. I would like the authors to
    make this very explicit, so that readers are not confused as to what comes from EMEP,
    and what has changed for CTM3.}
  \end{itemize}
\end{itemize}
\section{Minor points}
\begin{itemize}
\item {\color{blue}P1 L22. Isn’t H 2 O the most important greenhouse gas? (Say anthropogenic
GHG perhaps?)}
\item {\color{blue}P2 L3. The Wilson ref only concerns Europe, and its focus on the 95th per-
centile can hide trends found at higher percentiles (e.g. Simpson et al. 2014). A
better ref would be Fleming et al. (2018) or Mills et al. (2018a). By the way,
the most recent calculation on food security (using flux approaches) is now Mills
et al. (2018b).}
\item {\color{blue}P2 L11. What does in situ mean here? Ozone production can take place over
days of transport.}
\item {\color{blue}P2 L25--35. This text about halogens is not really relevant to a dry deposition
paper. Reactions with bromine can be important sinks, but are not usually counted
as deposition.}
\item {\color{blue}P3 L2--3. Why specify mid-June maximum for ozone. Monks (2000) might
take issue with that, as would for example Sinha et al. (2015).}
\item {\color{blue}P3 L4--5. There are plenty of ozone measurements made outside Europe. The
authors appear to be unaware of the massive ozone collections made under the
TOAR project (see e.g. Flemming, Mills refs below), or the high quality data
available from GAW (inha 2015).}
\item {\color{blue}P3 L16. One also has dry deposition to water, as this paper makes clear later
on.}
\item {\color{blue}P3 L20. One usually refers to dry deposition as something between a near-
surface height (e.g. z = 1m, 10m, or 50m) and the surface, not from $z_0$. In fact,
at $z = z_0$ one has $u_z = 0$, and hence the author’s $R_a$ just below should be zero.}
\item {\color{blue}P4 L20. I would remove the term textbook knowledge, since there are many
different approaches to nearly all these equations. It is thus good that the equations
as used in CTM3 are spelled out explicitly.}
\item {\color{blue}P5 L5. I would add Emberson et al. (2000a) and Tuovinen et al. (2004) to
the list of EMEP refs here, since this was the first publication of the methods that
have more or less been used until today.}
\item {\color{blue}P7, notation. In S2012 and EMEP generally, we use upper-case G and R to refer
to canopy-scale (bulk) variables, and lower-case for leaf-scale. Thus, in EMEP
we would have $G_\mathrm{sto} = LAI g_\mathrm{sto}$. Here the authors seem to mix upper and lower
case between their equations (13) and (14).}
\item {\color{blue}P7 Eqn (13). Is LAI one-sided, 2-sided, projected .... define.}
\item {\color{blue}P8 L1. S2012 do not suggest using depths lower than 1\,m. We use SMI3 which
is from 28-100 cm.}
\item {\color{blue}P8 L2 - why did you choose to use the surface (0--7\,cm) soil moisture?}
\item {\color{blue}P8 L18. This is wrong. Nothing in the EMEP model is used to ’mimic the
time lag..’. We use the light function to modify stomatal conductance, as with the
other $f$ factors.}
\item {\color{blue}P9 L20, and Table 1. The consequence of Table 1 is that vegetation at $0.5^\circ\,\mathrm{N}$
will start growing at day 90, whereas those at $0.5^\circ\,\mathrm{S}$ will start on day 272. (By the
way, in EMEP now we use monthly factors from the LPJ-GUESS model to derive
phenology for non-European areas, because of such difficulties with tabulations.)}
\item {\color{blue}P9 L18. what do you mean by "surface or 2\,m"? Surface might refer to skin or
leaf temperature?}
\item {\color{blue}P9 Eqn (26). Say 1st and 2nd, not I. and II.}
\item {\color{blue}P9, Eqn (27). This equation is a modification of Erisman’s original (1994) formu-
lation, so explain that.}
\item {\color{blue}P10 L12. Be explicit that this statement refers to S2012. The current EMEP
model uses different heights for e.g. tropical vegetation.}
\item {\color{blue}P11 Sect 2.2. I also found this aerosol section confusing. Eqn (30) is from S2012,
and so is the factor 0.008.SAI/10 used in Eqn (33), but here new a 1 coefficients
are defined. Did the ’aerosol microphysic model’ referred to also mix equations
in this way? Is there any publication as to the reliability of this method?}
\item {\color{blue}P12 Fig.2. I didn’t understand what is being done here. The Figure suggests that
the EMEP scheme has one category for ’Forests, Med. scrub’, whereas S2012
lists 4 types of forest, as well as Mediterranean scrub as a separate ecosystem.
This figure also suggests that EMEP has savanna, which it doesn’t, but do have
many other categories (Table 3 of S2012 lists 16 main categories. The current
EMEP model has 32.)}
\item {\color{blue}P12 L13. Again, the current EMEP model is not eurocentric, and uses global
phenology calculations.}
\item {\color{blue}P13 Sect 2.3.2. The initial lines (14-16) are hard to understand and only by
reading further do I see what they mean by ’de-accumulated’. If working with
IFS PPFD is so hard, why didn’t the authors just calculate hourly (or minute-by-
minute) PPFD using cloud-cover and zenith angles?}

\end{itemize}
\newpage

\end{document}

### 2019-05-16 version_2/response_version_1_rc2.tex
Edited responses. Needs probably one more week to answer them all...

@@
 If the latter, which $d$?}
-    We have to object slightly, because we have not presented "wrong inputs" to the algorithm.
-As pointed out by Ref \#1, we have made a mistake early on in the derivation of the formular to fit it into the framework of the Oslo~CTM3 (sign flip in fact).
-This let to negative values.
+    We have, in fact, not presented "wrong inputs" to the algorithm.
+    As pointed out by Ref \#1, we have made a mistake when deriving the formulation for $R_a$, as it is presented in S2012, for the framework of the Oslo~CTM3.
+This sign-flip in the derived formulation let to negative values.
+Because we oversaw this flaw, we concluded that we would no be able to use the S2012 formulation of $R_a$ in our model and decided to use the Monteith alternative.
     Since we are going to revise the model, we will address this issue and fix the calculation of $R_a$.
@@
 for finite $z$? (I know what they intend to say, but it isn’t at all clear.) In any case, here
-the authors end up with a stability-independent equation for $R_a$ , without mentioning or
-discussing that fact. This very shallow layer is also very problematic for deposition calculations in gen-
-eral, since the model cell seems to be run here with horizontal dimensions of $2.25 \times 2.25^\circ$ ,
+the authors end up with a stability-independent equation for $R_a$, without mentioning or
+discussing that fact. This very shallow layer is also very problematic for deposition calculations in general,
+since the model cell seems to be run here with horizontal dimensions of $2.25 \times 2.25^\circ$ ,
 or about $250 \times 250\,\mathrm{km}$ near the equator, but a vertical mid-level (CTM3’s $z_\mathrm{ref}$ ) of
@@
 surface, and should be very different for forests or lakes for example. Any wind-speed
-or friction velocity calculated from a model of such large horizontal resolution will nec-
-essarily give values at 8\,m which reflect the whole grid. Deposition rates for a specific
-land-cover will vary enormously depending on what else is in the grid-square. (Al-
-though not strictly comparable, we showed in Schwede et al. (2018) that differences
+or friction velocity calculated from a model of such large horizontal resolution will necessarily give values
+at 8\,m which reflect the whole grid. Deposition rates for a specific
+land-cover will vary enormously depending on what else is in the grid-square. (Although not strictly comparable,
+we showed in Schwede et al. (2018) that differences
 between the grid-average and forest specific deposition rates of N-compounds could be
@@
 ecosystems. The supplementary has three Figs related to this oceanic deposition. Why?}
-\item {\color{blue} Use of the term 'EMEP scheme'?: Sect. 3 discusses the comparisons of $V_g$ in terms of ’EMEP scheme’ versus ’We-
-sely scheme’, and sensitivity tests are named e.g. ’EMEP\_offlight’. As noted above
+\item {\color{blue} Use of the term 'EMEP scheme'?: Sect. 3 discusses the comparisons of $V_g$ in terms of ’EMEP scheme’ versus
+  ’Wesely scheme’, and sensitivity tests are named e.g. ’EMEP\_offlight’. As noted above
 the scheme implemented in CTM3 is very different to that implemented in the EMEP
@@
     Just refer to S2012 (and give Table number as help).}
+    We will follow this advice and refer the tables in S2012 accordingly.
   \item {\color{blue} Many of the equations are from S2012, and many not. I would like the authors to
@@
     and what has changed for CTM3.}
+    We are going to ellaborate on this matter in a revised version and point out our equations' sources more clearly.
   \end{itemize}
@@
 GHG perhaps?)}
-\item {\color{blue}P2 L3. The Wilson ref only concerns Europe, and its focus on the 95th per-
-centile can hide trends found at higher percentiles (e.g. Simpson et al. 2014). A
+\item {\color{blue}P2 L3. The Wilson ref only concerns Europe, and its focus on the 95th percentile can hide trends
+  found at higher percentiles (e.g. Simpson et al. 2014). A
 better ref would be Fleming et al. (2018) or Mills et al. (2018a). By the way,
@@
 \item {\color{blue}P9 Eqn (26). Say 1st and 2nd, not I. and II.}
-\item {\color{blue}P9, Eqn (27). This equation is a modification of Erisman’s original (1994) formu-
-lation, so explain that.}
+\item {\color{blue}P9, Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}
 \item {\color{blue}P10 L12. Be explicit that this statement refers to S2012. The current EMEP

### 2019-05-16 DryDep.tex
Edited responses. Needs probably one more week to answer them all...

@@
 with the molecular diffusivity for any gas $D_i$, the Schmidt number of water ($\text{Sc}_\chem{H_2O}=0.6$) and its molecular diffusivity ($D_\chem{H_2O}=0.21\cdot10^{-4}\,\unit{m^2\,s^{-1}}$).
-The used ratios $D_\chem{H_2O}/D_i$ are tabulated in Supplement~S.1.\\
+The used ratios $D_\chem{H_2O}/D_i$ are taken from \citet[][Table S18]{ACP:Simpson2012}.\\
 %
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.2: Figs.~S1--S2).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.1: Figs.~S1--S2).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 The allowed maximum roughness length in both cases is set to $2\,\unit{mm}$.
-Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.2: Fig.~S3).
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.1: Fig.~S3).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various land surface types.
-All parameters used in the EMEP scheme can be found in the supplementary to \citet{ACP:Simpson2012} or tabulated in Supplement~S.3.
+All parameters used in the EMEP scheme can be found in the supplementary to \citet[][Tables S16, S19]{ACP:Simpson2012}.
 Since $f_T$ turns negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
@@
 Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (S.4, Fig.~S4).
+For four example PFTs, results are shown in the Supplement (S.2, Fig.~S4).
 
@@
 In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we still use the SMOKE-BEIS parameterization.
-The described algorithm written in python~2.7 has been included as Supplement~S.5.
+The described algorithm written in python~2.7 has been included as Supplement~S.3.
 An example map of the computed GLEN using the $5\,\unit{^\circ C}$-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
@@
 %
-and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.6 (Figs.~S5--S7).
+and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.4 (Figs.~S5--S7).
 The resulting PPFD fields are still accumulated over a time period of $3$ hours and should be divided by $3$.
@@
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.7: Fig.~S8).
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.5: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.

### 2019-05-16 DryDep_supplement.tex
Second round of Amunds comments and update of copernicus style files.
Fine-tuning to GMD requirements.
Second round f fine tuning to GMD requirements, some minor adaptions of text and typo corrections. Ready to submit.
Edited responses. Needs probably one more week to answer them all...

@@
 \graphicspath{{pictures/}}          % graphics
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
 \begin{document}
@@
 \section*{Supplement}
-\subsection*{S.1 Aerodynamical resistance}
-\begin{table}[!htbp]
-  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
-  \begin{tabular}{lcrr@{$\cdot$}l}
-    \tophline
-    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
-    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
-    \middlehline
-    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
-    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
-    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
-    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
-    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
-    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
-    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
-    \bottomhline
-  \end{tabular}
+%\subsection*{S.1 Aerodynamical resistance}
+%\begin{table}[!htbp]
+%  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
+%  \begin{tabular}{lcrr@{$\cdot$}l}
+%    \tophline
+%    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
+%    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
+%    \middlehline
+%    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
+%    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
+%    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
+%    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
+%    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
+%    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
+%    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
+%    \bottomhline
+%  \end{tabular}
   %\belowtable{} % Table Footnotes
-\end{table}
+%\end{table}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
+\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
 \begin{figure}[!htbp]
@@
   \caption{Sutherland's law and divergence linear fits with one or two degrees of freedom.
-The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3.
+The fit $\mu(T) = m\cdot T$ is implemented in the Oslo~CTM3.
 Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is negligible.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{S.3 Stomatal conductance}
-\begin{table*}[!htbp]
-  \caption{Tabulated parameters for stomatal conductance computation in the EMEP scheme}
-  \begin{tabular}{lcccccccccccccc}
-    \tophline
-    Code$^*$ & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $\alpha_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
-    & (\unit{mmol\,s^{-1}\,m^{-2}}) &&&&&&(\unit{days})&(\unit{days})&(\unit{days})&(\unit{days})&&(\unit{^\circ C})&(\unit{^\circ C})&(\unit{^\circ C})\\
-    \middlehline
-    CF & 140 & 0.1 & 0.8 & 0.8 & 0.8 & 0.8 & 1 & 1 & 0 & 0 & 0.006 & 0 & 18 & 36  \\
-    DF & 150 & 0.1 & 0 & 0 & 1 & 0 & 20 & 30 & 0 & 0 & 0.006 & 0 & 20 & 35  \\
-    NF & 200 & 0.1 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.013 & 8 & 25 & 38 \\
-    BF & 200 & 0.02 & 1 & 1 & 0.3 & 1 & 130 & 60 & 80 & 35 & 0.009 & 1 & 23 & 39 \\
-    TC & 300 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0105 & 0.01 & 12 & 26 & 40 \\
-    MC & 300 & 0.019 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0048 & 0 & 25 & 51  \\
-    RC & 360 & 0.02 & 0.2 & 0.2 & 1 & 0.2 & 20 & 45 & 0 & 0 & 0.0023 & 8 & 24 & 50  \\
-    SNL & 60 & 0.01 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0.009 & 1 & 18 & 36 \\
-    GR & 270 & 0.01 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0.009 & 12 & 26 & 40 \\
-    MS & 200 & 0.01 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.012 & 4 & 20 & 37 \\
-    WE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
-    TU & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
-    DE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
-    W & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
-    ICE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
-    U & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
-    \bottomhline
-  \end{tabular}
-  \\
-  \vspace{2\baselineskip}
-  \begin{tabular}{lcccccccccc}
-    \tophline
-    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$\\% & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
-    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})\\%&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\
-    \middlehline
-    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0\\% & 5 & 5 & 1 & 1\\
-    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 \\%& 0 & 4 & 20 & 30\\
-    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0\\% & 4 & 4 & 1 & 1\\
-    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 \\%& 4 & 4 & 1 & 1\\
-    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57\\% & 0 & 3.5 & 70 & 20\\
-    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57\\% & 0 & 3 & 70 & 44\\
-    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0\\% & 0 & 4.2 & 35 & 65\\
-    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 2 & 3 & 192 & 96\\
-    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0\\% & 2 & 3.5 & 140 & 135\\
-    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0\\% & 2.5 & 2.5 & 1 & 1\\
-    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
-    \bottomhline
-  \end{tabular}
-  \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean scrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
-\end{table*}
+%\subsection*{S.3 Stomatal conductance}
+%\begin{table*}[!htbp]
+%  \caption{Tabulated parameters for stomatal conductance computation in the EMEP scheme}
+%  \begin{tabular}{lcccccccccccccc}
+%    \tophline
+%    Code$^*$ & $g_\text{max}$ & $f_\text{min}$ & $\phi_a$ & $\phi_b$ & $\phi_c$ & $\phi_d$ & $\phi_e$ & $\phi_f$ & $\phi_\text{AS}$ & $\phi_\text{AE}$ & $\alpha_\text{light}$ & $T_\text{min}$ & $T_\text{opt}$ & $T_\text{max}$ \\
+%    & (\unit{mmol\,s^{-1}\,m^{-2}}) &&&&&&(\unit{days})&(\unit{days})&(\unit{days})&(\unit{days})&&(\unit{^\circ C})&(\unit{^\circ C})&(\unit{^\circ C})\\
+%    \middlehline
+%    CF & 140 & 0.1 & 0.8 & 0.8 & 0.8 & 0.8 & 1 & 1 & 0 & 0 & 0.006 & 0 & 18 & 36  \\
+%    DF & 150 & 0.1 & 0 & 0 & 1 & 0 & 20 & 30 & 0 & 0 & 0.006 & 0 & 20 & 35  \\
+%    NF & 200 & 0.1 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.013 & 8 & 25 & 38 \\
+%    BF & 200 & 0.02 & 1 & 1 & 0.3 & 1 & 130 & 60 & 80 & 35 & 0.009 & 1 & 23 & 39 \\
+%    TC & 300 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0105 & 0.01 & 12 & 26 & 40 \\
+%    MC & 300 & 0.019 & 0.1 & 0.1 & 1 & 0.1 & 0 & 45 & 0 & 0 & 0.0048 & 0 & 25 & 51  \\
+%    RC & 360 & 0.02 & 0.2 & 0.2 & 1 & 0.2 & 20 & 45 & 0 & 0 & 0.0023 & 8 & 24 & 50  \\
+%    SNL & 60 & 0.01 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0.009 & 1 & 18 & 36 \\
+%    GR & 270 & 0.01 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0.009 & 12 & 26 & 40 \\
+%    MS & 200 & 0.01 & 1 & 1 & 0.2 & 1 & 130 & 60 & 80 & 35 & 0.012 & 4 & 20 & 37 \\
+%    WE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+%    TU & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
+%    DE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0  \\
+%    W & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+%    ICE & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+%    U & 0 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 1 & 0 \\
+%    \bottomhline
+%  \end{tabular}
+%  \\
+%  \vspace{2\baselineskip}
+%  \begin{tabular}{lcccccccccc}
+%    \tophline
+%    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$\\% & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
+%    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})\\%&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\
+%    \middlehline
+%    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0\\% & 5 & 5 & 1 & 1\\
+%    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 \\%& 0 & 4 & 20 & 30\\
+%    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0\\% & 4 & 4 & 1 & 1\\
+%    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 \\%& 4 & 4 & 1 & 1\\
+%    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57\\% & 0 & 3.5 & 70 & 20\\
+%    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57\\% & 0 & 3 & 70 & 44\\
+%    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0\\% & 0 & 4.2 & 35 & 65\\
+%    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 2 & 3 & 192 & 96\\
+%    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0\\% & 2 & 3.5 & 140 & 135\\
+%    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0\\% & 2.5 & 2.5 & 1 & 1\\
+%    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
+%    \bottomhline
+%  \end{tabular}
+%  \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean scrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
+%\end{table*}
 
-\clearpage
-\subsection*{S.4 Latitude dependent vegetation height}
+%\clearpage
+\subsection*{S.2 Latitude dependent vegetation height}
 \begin{figure}[!htbp]
@@
 
-\subsection*{S.5 Temperature dependent greening season (python 2.7)}
+\subsection*{S.3 Temperature dependent greening season (python 2.7)}
 {\scriptsize
@@
 
-\subsection*{S.6 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
+\subsection*{S.4 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
 \begin{figure}[!htbp]
@@
   \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.65cm}]{pictures/PAR_deaccumulating_00h.png}
-  \caption{De-accumulation of hour 00 in the partly de-accumulated fields.
-$\text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right]$.}
+  \caption{De-accumulation of $t=00\,\unit{UTC}$ in the partly de-accumulated fields.
+$\text{PPFD}(t=00\,\unit{UTC}) = \text{PAR}(t=00\,\unit{UTC}) - \left[\text{PAR}(t=21\,\unit{UTC}-1\,\unit{day})-\text{PAR}(t=12\,\unit{UTC}-1\,\unit{day})\right]$.}
 \end{figure}
@@
 
-\subsection*{S.7 Normalized zonal average ozone dry deposition velocity}
+\subsection*{S.5 Normalized zonal average ozone dry deposition velocity}
 \begin{figure}[!htbp]

### 2019-05-21 version_2/response_version_1_rc2.tex
Working on the response to referee comments.

--- 2019-05-13 version_2/response_version_1_rc2.tex
@@
 \usepackage{multirow}
+\usepackage{graphicx}
 \thispagestyle{empty}
@@
 To gmd-2019-21 David Simpson (10 Apr 2019):
-We appologize for the late response regarding the comments on our paper.
+We apologize for the late response regarding the comments on our paper.
 We had to evaluate the major concerns raised in the review and found that we have to revise our model and repeat the model experiments to address all concerns.
@@
   Thank you very much for your detailed account of the \emph{mosaic approach}.
-We have discussed the concerns raised within this comment and found that we may have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
+We have discussed the concerns raised within this comment and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
 \item {\color{blue} Calculation of $R_a$:}
@@
 errors. The correct equation will not give negative values unless presented with wrong
-inputs, and I suspect that that is what has happened. It is actually difficult to tell what
+inputs, and I suspect that that is what has happened.}
+We have, in fact, not presented "wrong inputs" to the algorithm.
+As pointed out by Ref \#1, Eq.
+(2) in our manuscript is erroneous.
+We have not only made a mistake when deriving the formulation for $R_a$, from $u_*$ as presented in S2012 (sign-flip preceding the last term in the series) but also mistype the denominator of the second term.
+The corrected equation:
+\begin{equation}
+  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right].
+  \label{eq:aero_res}
+\end{equation}
+
+\item {\color{blue} It is actually difficult to tell what
 was tested from the manuscript though, since they state simply that $d$ is ’typically 0.7\,m’.
-Did they use $d$ properly, consistent with the underlying land-cover and its $z_0$ ? Did
+Did they use $d$ properly, consistent with the underlying land-cover and its $z_0$? Did
 they assume that their 8\,m meteorology was at a physical height of 8\,m, or at $d + 8\,\mathrm{m}$.
 If the latter, which $d$?}
-    We have to object slightly, because we have not presented "wrong inputs" to the algorithm.
-As pointed out by Ref \#1, we have made a mistake early on in the derivation of the formular to fit it into the framework of the Oslo~CTM3 (sign flip in fact).
-This let to negative values.
-    Since we are going to revise the model, we will address this issue and fix the calculation of $R_a$.
+  It is indeed not clear from the manuscript that we never actually implemented the formula (Eq.~(2) in our manuscript) in the model.
+The reason for this were simple, mathematical considerations.
+Given $z = z_\mathrm{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$, we looked at certain parameter spaces for $L$ and $z_0$ (e.g., $0 \le z_0 \le 2.5$ for vegetated land surfaces from a climatology mean derived from ISLSCP2 FASIR) but have not taken the consistency "with the underlying land-cover and its $z_0$" into account (see Fig.~\ref{fig:test_ra}).
+This might have been not unsophisticated enough, though.
+The statement "$d$ is ’typically 0.7\,m’" is a typo and should of cause be corrected to \emph{typically $d = 0.7\cdot h(N,\mathrm{lat})\,\mathrm{m}$ for vegetation other than forests}.
+Note: The average height of the lowermost model level is actually $20\,\mathrm{m}$ ($10\,\mathrm{m}$ for mid-level).
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{test_Ra_wrong_form.pdf}
+    \caption{Test cases for $R_a$ with $z = z_{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$.}
+    \label{fig:test_ra}
+  \end{figure}
 
@@
 way. For example, when is $z_0$ ever zero, as stated on line 23, or why does $\partial_z R_a \rightarrow R_a $
-for finite $z$? (I know what they intend to say, but it isn’t at all clear.) In any case, here
-the authors end up with a stability-independent equation for $R_a$ , without mentioning or
-discussing that fact. This very shallow layer is also very problematic for deposition calculations in gen-
-eral, since the model cell seems to be run here with horizontal dimensions of $2.25 \times 2.25^\circ$ ,
-or about $250 \times 250\,\mathrm{km}$ near the equator, but a vertical mid-level (CTM3’s $z_\mathrm{ref}$ ) of
+for finite $z$? (I know what they intend to say, but it isn’t at all clear.)}
+  The referee is right.
+The text and derivation of the Monteith relation is confusing.
+We will elaborate on this in a revised version of the manuscript in case we are not going to use Eq.~(\ref{eq:aero_res}).
+
+\item {\color{blue} In any case, here
+the authors end up with a stability-independent equation for $R_a$, without mentioning or
+discussing that fact.}
+  We shall elaborate on this.
+\item {\color{blue} This very shallow layer is also very problematic for deposition calculations in general,
+since the model cell seems to be run here with horizontal dimensions of $2.25 \times 2.25^\circ$,
+or about $250 \times 250\,\mathrm{km}$ near the equator, but a vertical mid-level (CTM3’s $z_\mathrm{ref}$) of
 just 8\,m. Now, profiles of wind and depositing gases are very sensitive to the underlying
 surface, and should be very different for forests or lakes for example. Any wind-speed
-or friction velocity calculated from a model of such large horizontal resolution will nec-
-essarily give values at 8\,m which reflect the whole grid. Deposition rates for a specific
-land-cover will vary enormously depending on what else is in the grid-square. (Al-
-though not strictly comparable, we showed in Schwede et al. (2018) that differences
+or friction velocity calculated from a model of such large horizontal resolution will necessarily give values
+at 8\,m which reflect the whole grid. Deposition rates for a specific
+land-cover will vary enormously depending on what else is in the grid-square. (Although not strictly comparable,
+we showed in Schwede et al. (2018) that differences
 between the grid-average and forest specific deposition rates of N-compounds could be
@@
 differences were largely dependent on how much forest occupied each grid cell.)}
+  In S2012, $R_a$ is evaluated "at around $45\,\mathrm{m}$" which is similar to the center of our second lowest model level. We can easily correct for this in a revised model version. 
 \end{itemize}      
@@
 ecosystems. The supplementary has three Figs related to this oceanic deposition. Why?}
-\item {\color{blue} Use of the term 'EMEP scheme'?: Sect. 3 discusses the comparisons of $V_g$ in terms of ’EMEP scheme’ versus ’We-
-sely scheme’, and sensitivity tests are named e.g. ’EMEP\_offlight’. As noted above
+  The referee is right that ozone is predominately deposited to terrestrial ecosystems, which cover about 2/9 of the Earth's surface, while 2/3 is covered by oceans.
+Nevertheless, due to the ocean's vastness, relatively small changes in modeled dry deposition easily accumulate and influence the atmospheric ozone concentration.
+Luhar et al.
+(2017, 2018) show that the current models may overestimate the oceanic ozone sink by a factor of three -- rendering the ocean a even smaller sink for ozone.
+
+  This said, Section~2 as a whole consists of about 9 pages, of which 1/2 page is dedicated to oceans ($R_b$ computation) and 7 pages are dedicated to vegetation especially stomatal and non-stomatal conductance ($R_c$ computation).
+Hence, we do not see a significant unbalance in favor of oceans herein.
+
+  Regarding the three figures with respect to ocean in the supplementary material: The reason for these is that we found the "legacy code" in the model (linear fit dynamic viscosity of air $\mu(T)$) after we had finished the implementation of the new dry deposition scheme, run the experiments, and done the analysis on these.
+We had to verify that this had no influence on the results.
+
+\item {\color{blue} Use of the term 'EMEP scheme'?: Sect. 3 discusses the comparisons of $V_g$ in terms of ’EMEP scheme’ versus
+  ’Wesely scheme’, and sensitivity tests are named e.g. ’EMEP\_offlight’. As noted above
 the scheme implemented in CTM3 is very different to that implemented in the EMEP
@@
 is being tested here, but it certainly is not. [...]}
-  We appologize for the missleading naming of the new dry deposition scheme in the Oslo~CTM3 which is (partly) based on the formulations in the publication the referee refers to as \emph{S2012}.
-By no means have we meant to offend any of the original authors of the EMEP/MSC-W model nor intented to misguide the readers.
-We will rename the revised scheme apropriately.
-(\emph{mOSaic} maybe?)
+  We apologize for the misleading naming of the new dry deposition scheme in the Oslo~CTM3 which is (partly) based on the formulations in the publication the referee refers to as \emph{S2012}.
+By no means have we meant to offend any of the original authors of the EMEP/MSC-W model nor intended to misguide the readers.
+We will rename the revised scheme appropriately ($\rightarrow$\emph{mOSaic scheme}).
 \item {\color{blue} Reproduction of material from S2012:}
@@
     Just refer to S2012 (and give Table number as help).}
+    We will follow this advice and refer to the tables in S2012 accordingly.
   \item {\color{blue} Many of the equations are from S2012, and many not. I would like the authors to
@@
     and what has changed for CTM3.}
+    We are going to elaborate on this matter in a revised version and point out our equations' sources more clearly.
   \end{itemize}
@@
 \begin{itemize}
-\item {\color{blue}P1 L22. Isn’t H 2 O the most important greenhouse gas? (Say anthropogenic
+\item {\color{blue}P1 L22. Isn’t $\mathrm{H_2O}$ the most important greenhouse gas? (Say anthropogenic
   GHG perhaps?)}
-\item {\color{blue}P2 L3. The Wilson ref only concerns Europe, and its focus on the 95th per-
-centile can hide trends found at higher percentiles (e.g. Simpson et al. 2014). A
-better ref would be Fleming et al. (2018) or Mills et al. (2018a). By the way,
-the most recent calculation on food security (using flux approaches) is now Mills
-et al. (2018b).}
+  Thank you for pointing this out. We follow the advice and write: \emph{"[...] ozone is a potent anthropogenic greenhouse gas [...]"}
+\item {\color{blue}P2 L3. The Wilson ref only concerns Europe, and its focus on the 95th percentile can hide trends
+  found at higher percentiles (e.g.
+Simpson et al.
+2014).
+A better ref would be Fleming et al.
+(2018) or Mills et al.
+(2018a).
+By the way, the most recent calculation on food security (using flux approaches) is now Mills et al.
+(2018b).}
+  Thank you for pointing out these new publications. We take them into consideration, when revising the manuscript.
 \item {\color{blue}P2 L11. What does in situ mean here? Ozone production can take place over
   days of transport.}
+  \emph{In situ} typically means \emph{on site, locally}.
+In this context, we use it exactly in this meaning.
+Elevated ozone levels at a site originate mainly from the local production of ozone from its precurors that are transported than from advection of ozone itself.
+Although long-range ozone transport occures and might be most important in regions that else lack precursors.
+We elaborate on this in a revised version of the paper: \emph{Tropospheric ozone is mainly produced locally [...] involving advected precursor gases [...]}
 \item {\color{blue}P2 L25--35. This text about halogens is not really relevant to a dry deposition
@@
 as deposition.}
+  The removal of ozone from the Arctic boundary layer in the precence of halogens is indeed not directly related to ozone dry deposition, but one of the discussed mechanisms to trigger the bromine explosions involves, e.g., a "initialisation" by ozone dry deposition.
+But we see that this is of no relevance to ozone dry deposition itself, but to the observed ozone abundances in the Arctics.
+We may remove these reference in the revised version of the manuscript.
+
 \item {\color{blue}P3 L2--3. Why specify mid-June maximum for ozone. Monks (2000) might
   take issue with that, as would for example Sinha et al. (2015).}
+
 \item {\color{blue}P3 L4--5. There are plenty of ozone measurements made outside Europe. The

### 2019-05-21 DryDep.tex
Working on the response to referee comments.

@@
 Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
-In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ACP:Wilson2012}.
+In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018} {\bf TODO: Read through papers and confirm numbers.}.
 At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
@@
 It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} as well as $7\,\unit{\%}$ loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
-The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-Tropospheric ozone is mainly produced in situ in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
+The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} {\bf TODO: Read through papers and confirm numbers.} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
+{\bf TODO: Think about how to balance the wording of were ozone originates from transport VS local production.}Tropospheric ozone is mainly produced locally in complex photochemical cycles involving advected precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
 A typical reaction mechanism for \chem{CO} is sketched in the following.
@@
 
-A dry deposition related sink limited to Arctic regions are so called ozone depleting events.
+%A dry deposition related sink limited to Arctic regions are so called ozone depleting events.
 They occur in spring-time in the polar boundary layer where an outburst of bromine monoxide \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
@@
 Regarding the new dry deposition scheme, we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{WASP:Simpson2003,ACP:Simpson2012}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
-We will refer to this as \emph{EMEP scheme} throughout the rest of the paper.
-The EMEP scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature)  of the atmosphere as well as dynamics (e.g., wind stress) of the boundary layer into account.
+We will refer to this as \emph{mOSaic scheme} throughout the rest of the paper.
+The mOSaic scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature)  of the atmosphere as well as dynamics (e.g., wind stress) of the boundary layer into account.
 To a certain degree, the global variety of plants and their variability throughout the seasons is also acknowledged.
-The EMEP scheme is implemented for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
+The mOSaic scheme is implemented for the gaseous species \chem{O_3}, \chem{H_2O_2}, \chem{NO_2}, \chem{PAN}, \chem{SO_2}, \chem{NH_3}, \chem{HCHO}, and \chem{CH_3CHO}.
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
@@
 %\begin{equation}
-%  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{z_0}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
+%  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
 %\end{equation}
-%with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\,\unit{m}$), and the Obukhov length $L$.
+%with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
-For this reason, we diverge from the EMEP scheme at this point and fall back to
+For this reason, we diverge from the mOSaic scheme at this point and fall back to
 For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
 \begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial_z T}{\partial_z R_a}.
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
   \label{eq:sens_heat}
 \end{equation}
-Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $\partial_z R_a$ is the diffusion resistance to sensible heat between surface ($z_0 = 0$) and the reference height $z$, and $\partial_z T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
+Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
 From the eddy diffusion theorem, a similar formulation can be derived
 \begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial_z T}{\partial_z u} \cdot u_*^2.
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
   \label{eq:eddy_theo}
 \end{equation}
-Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $\partial_z R_a \rightarrow R_a $ and $\partial_z u \rightarrow u_z $ for a finite $z$, we find 
+For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
 \begin{equation}
@@
 \end{equation}
-As reference height $z$, we use the height at mid-level of the lowermost model level (roughly $8\,\unit{m}$).
+As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$).
 From the meteorological input fields $u_z$ is directly available, while $u_*$ can be computed via
@@
 \end{equation}
-with the turbulent surface stress $\tau_0^2 = \text{NSSS}^2 + \text{EWSS}^2$, and northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS).
+with the turbulent surface stress $\tau_0^2 = \tau_v^2 + \tau_u^2$, and northward turbulent surface stress $\tau_v$, eastward turbulent surface stress $\tau_u$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The used ratios $D_\chem{H_2O}/D_i$ are taken from \citet[][Table S18]{ACP:Simpson2012}.\\
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various land surface types.
-All parameters used in the EMEP scheme can be found in the supplementary to \citet[][Tables S16, S19]{ACP:Simpson2012}.
+All parameters used in the mOSaic scheme can be found in the supplementary to \citet[][Tables S16, S19]{ACP:Simpson2012}.
 Since $f_T$ turns negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
@@
 The phenology of a plant typically describes its life-cycle throughout a year, e.g., at mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
-In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
+In the mOSaic scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
 Details about our treatment of these are given in Section~\ref{subsubsec:greening}.
@@
 Herein, we use the SGS and EGS derived parameters day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
-The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}).
+The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the mOSaic scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}).
 If GLEN is zero we are, e.g., in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
 Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$.
-Since the EMEP scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
+Since the mOSaic scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
 %
@@
 %\subsubsection*{Non-stomatal conductance}
-In the EMEP scheme, non-stomatal conductances are explicitly calculated for \chem{O_3}, \chem{SO_2}, \chem{HNO_3}, and \chem{NH_3}.
+In the mOSaic scheme, non-stomatal conductances are explicitly calculated for \chem{O_3}, \chem{SO_2}, \chem{HNO_3}, and \chem{NH_3}.
 For all other species, an interpolation between \chem{O_3} and \chem{SO_2} values is carried out.
@@
 %
-Extending the EMEP scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
+Extending the mOSaic scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
 %
@@
 \subsubsection{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
+The vegetation height $h(N, \text{lat})$ in the mOSaic scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
 To adapt this to a global model, we need to make a few additional assumptions.
-The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
+The tabulated height for each vegetation type $h(N)$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
 Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
@@
 \subsubsection{Mapping of land surface types}
-The Oslo~CTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}).
+The Oslo~CTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the mOSaic scheme (Fig.~\ref{fig:pft_mapping}).
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
@@
 \label{sec:eval}
-In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the EMEP scheme.
+In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the mOSaic scheme.
 We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively.
@@
 \label{subsec:sens}
-Due to significant differences between the EMEP scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every difference in results to a respective change.
-Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme.
+Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every difference in results to a respective change.
+Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the mOSaic scheme.
 A reference simulation featuring the Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
@@
     \tophline
-    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Emissions\\
+    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Emissions\\
     & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
@@
 We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-A possible explanation to this divergence especially in desert regions is the difference between the prescribed dry deposition velocities in the Wesely scheme in comparison to those used in the EMEP scheme.
+A possible explanation to this divergence especially in desert regions is the difference between the prescribed dry deposition velocities in the Wesely scheme in comparison to those used in the mOSaic scheme.
 We cover this in more detail in the following section.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the EMEP scheme compared to the Wesely scheme.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the mOSaic scheme compared to the Wesely scheme.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
 In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
-We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
+We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
 Among the sensitivity studies, there is only a low absolute variance.
@@
 The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b.
-The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ lower than in the Wesely scheme.
-In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean.
+The dry deposition velocities in the mOSaic scheme are on average $43\,\unit{\%}$ lower than in the Wesely scheme.
+In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the mOSaic scheme are constantly above the multi-model-mean.
 This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
@@
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.5: Fig.~S8).
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are similar (Supplement~S.5: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
@@
 The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a change in $v_\text{desert}^\chem{O_3}$.
-Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
-Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
-These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP scheme.
+Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
+Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original mOSaic scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
+These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the mOSaic scheme.
 We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).
 
-There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the EMEP scheme match better.
+There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better.
 This could be a sign of differences in photo-chemistry and transport (e.g.
@@
 Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity and a slight increase of $v_\text{DD}^\chem{O_3}$ in the region $(20-40)\,\unit{^\circ S}$.
-In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ.
+In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the mOSaic scheme, indicating that the seasonal cycles differ.
 The corresponding total monthly ozone dry deposition is shown in Appendix Fig.~\ref{fig:mmm_drydep_season}b.
-In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
+In general, the seasonal patterns are quite similar in the Wesely scheme and the mOSaic scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively.
 What differs most is the molding and intensity of the NH peak dry deposition.
@@
 In comparison, the available $2$ months of the multi-model-mean display a somewhat different pattern.
-Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
+Especially the NH/SH asymmetry is much more pronounced in February, than it is the mOSaic scheme.
 %
@@
 As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface).
-The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme.
+The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and mOSaic scheme.
 In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum.
-The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH.
-The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
+The seasonal cycle in the mOSaic scheme appears to be shifted towards later in the year in the NH.
+The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the mOSaic scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
 This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in long-wave radiative transfer (e.g., due changes in cloud cover or aerosol).
@@
 In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories.
-The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean.
-The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$).
-In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
-The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
+The patterns and absolute numbers differ substantially between the Wesely scheme and the mOSaic scheme and the multi-model-mean.
+The divergence of the average dry deposition velocities between the Wesely scheme and the mOSaic scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$).
+In contrast to the Wesely scheme and the mOSaic scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
+The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
 From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
-This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme.
+This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the mOSaic scheme.
 Similarly, the dry deposition velocities over water differ.
@@
 In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
-With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
+With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the mOSaic scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
 With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical or coniferous forest \citep{GCB:Anav2017}.
@@
 Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative mOSaic scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$.
 The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling.
-The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
+The total annual dry deposition in the mOSaic scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
 But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition.
-In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analyisis.
+In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analysis.
 This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
@@
 The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}.
-At $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
+At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
 We use a $\chi^2$-test
@@
 The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
-The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the Oslo~CTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.

### 2019-05-21 version_2/response_version_1_rc2.tex
Some more response.

@@
   take issue with that, as would for example Sinha et al. (2015).}
-
+  The referee has got a point here. Of cause the maximum of ozone in the annual cycle depends on the location. We remove this specification.
 \item {\color{blue}P3 L4--5. There are plenty of ozone measurements made outside Europe. The
@@
 available from GAW (inha 2015).}
-\item {\color{blue}P3 L16. One also has dry deposition to water, as this paper makes clear later
-on.}
+  Thank you for reminding us of the data collected under the TOAR project.
+We knew about this dataset but have to admitt that we have not made much use of it, yet.
+We have, however, not stated that there are, in general, no sites outside of Europe, but that long-term obersvations (started before the 1950s) are only found in Europe.
+Maybe the term "long-term" is not clear enough.
+We add: \emph{"[...] a limited number of long-term ozone observations (started before the 1950s)[...]"}
+
+\item {\color{blue}P3 L16. One also has dry deposition to water, as this paper makes clear later on.}
+  That is indeed true, but we didn't want to name all possible surfaces onto which dry deposition takes place.
+We, hence, shall write: \emph{"Removal of any substance from the atmosphere which is not involving rain, e.g., [...]}
+
 \item {\color{blue}P3 L20. One usually refers to dry deposition as something between a near-

### 2019-05-21 DryDep.tex
Some more response.

@@
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
-The abundance of tropospheric ozone therefore varies, e.g., with time of the day (maximum $\sim$15:00 local time), season (mid-June maximum), altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
+The abundance of tropospheric ozone therefore varies, e.g., with time of the day, season, altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
 Typical concentrations of surface ozone range from $10\,\unit{ppb}$ over the tropical Pacific to $100\,\unit{ppb}$ in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
 This variability poses a challenge on both, trend analysis from observation as well as validation and intercomparison of models.
-At the observational side, there is only a limited number of long-term ozone observations, mainly restricted to European sites.
+At the observational side, there is only a limited number of long-term ozone observations (started before the 1950s), mainly restricted to European sites.
 Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
@@
 A key to fathom these slightly contradicting results may lie in the used dry deposition schemes.\\
-Removal of any substance through gravitational settling or by uptake by plants and soil is referred to as dry deposition.
+Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants and soil, is referred to as dry deposition.
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.

### 2019-05-22 version_2/response_version_1_rc2.tex
Continue with thereferee comments.

@@
 \usepackage[english]{babel}
+\usepackage{amsmath}
 \usepackage{xcolor}
@@
 \begin{itemize}
-\item {\color{blue} Mosaic formulation: [...] In any case, I think this approach has serious problems.
+\item {\color{blue} Mosaic formulation:}
+  \begin{itemize}
+  \item {\color{blue} CTM3 claims to implement a mosaic approach, but instead of calculating deposition
+rates over each land-cover, and then aggregating using Eqn.3, CTM3 seems to perform the following steps:
+\begin{align}
+  \label{eq:ra}
+  R_a &= \frac{u_z}{u^2_*}\\
+  \label{eq:Gsto}
+  \overline{Gs^i_g(z)} &= \Sigma^N_{k=1} f_k \times Gs^i_{g,k}(z)\\
+  \label{eq:Gns}
+  \overline{Gns^i_g(z)} &= \Sigma^N_{k=1} f_k \times Gns^i_{g,k}(z)
+\end{align}
+As far as I can tell from the text, $R_a$ is calculated once, with the same value of $u_*$ for all land-covers.
+The CTM3 $R_b$ calculation seems to also use the same $u_*$ over different land-covers, except over sea where a more sophisticated scheme is used.
+Equations 5-6 above are equivalent to Eqns (22) and (26) from the manuscript.
+Now I am puzzled however as to how all this is put together.
+Do they use:
+\begin{equation}
+  G_c = \mathrm{LAI} \cdot \overline{Gs^i_g(z)} + \overline{Gns^i_g(z)}
+\end{equation}
+  }
+    The referee is correct regarding the usage of Eq.~(\ref{eq:ra}) and Eq.~(\ref{eq:Gns}). In case of stomatal conductance, we first compute:
+    \begin{equation}
+       \overline{gs^i_g(z)} = \Sigma^N_{k=1} f_k \times gs^i_{g,k}(z).
+    \end{equation}
+    This is put together as follows:
+    \begin{equation}
+      \label{eq:Gsto_corr}
+      G_c = \mathrm{LAI} \cdot \overline{gs^i_g(z)} + \overline{Gns^i_g(z)}
+    \end{equation}
+  \item {\color{blue} In any case, I think this approach has serious problems.
 Why average first for $G_s$ and then for $G_{ns}$ , when it is the fluxes ($F_k$ , or $V^i_{g,k}(z_\mathrm{ref}) \times \chi^i_\mathrm{avg}(z_\mathrm{ref})$) which need to be averaged?
@@
 We have discussed the concerns raised within this comment and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
+  \end{itemize}
 \item {\color{blue} Calculation of $R_a$:}
@@
   That is indeed true, but we didn't want to name all possible surfaces onto which dry deposition takes place.
-We, hence, shall write: \emph{"Removal of any substance from the atmosphere which is not involving rain, e.g., [...]}
+We, hence, shall write: \emph{"Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants and soil, is referred to as dry deposition.}
 
@@
 at $z = z_0$ one has $u_z = 0$, and hence the author’s $R_a$ just below should be zero.}
+Thank you for pointing out this inaccuracy.
+We, of cause, meant "near-surface" height or lowermost model level (which is indeed not at $z_0=0$).
+We change the text accordingly: \emph{"[...] dry deposition is a product between near-surface ozone concentration $\mathrm{[O_3](z_0)}$ (e.g.
+the lowermost model level) [...] "}
+
 \item {\color{blue}P4 L20. I would remove the term textbook knowledge, since there are many
@@
 as used in CTM3 are spelled out explicitly.}
+  We follow the advice and remove the sentence in which the term occurs and write: \emph{"[...] we will give a detailed account of the new dry deposition scheme and the equations that we use."}
+
 \item {\color{blue}P5 L5. I would add Emberson et al. (2000a) and Tuovinen et al. (2004) to
@@
 have more or less been used until today.}
+  We have added the references regarding the EMEP MSC-W model.
+\emph{"[...] we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model (Emberson et al., 2000; Simpson et al., 2003; Tuovinen et al., 2004; Simpson et al., 2012), [...]"}
+
 \item {\color{blue}P7, notation. In S2012 and EMEP generally, we use upper-case G and R to refer
@@
 case between their equations (13) and (14).}
+  This is probably related to the referee's major point 1.
+Since our approach, averaging the resistances/conductances instead of the fluxes, is incorrect, this part of the manuscript has to be revised in any case.
+But to answer the question: Eq.~(13) should have been placed at the end of the section, after the definition of $G_\mathrm{sto}$ and $G_\mathrm{ns}$.
+A better notation is found Eq.~(\ref{eq:Gsto_corr}).
+We shall also go through all our notations in that section to make it more comprehensible.
+
 \item {\color{blue}P7 Eqn (13). Is LAI one-sided, 2-sided, projected .... define.}
+  LAI is one-sided taken from ISLSCP2 FASIR. We add this to the sentence: \emph{"[...] $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR [...]"}
+
 \item {\color{blue}P8 L1. S2012 do not suggest using depths lower than 1\,m. We use SMI3 which
   is from 28-100 cm.}
-\item {\color{blue}P8 L2 - why did you choose to use the surface (0--7\,cm) soil moisture?}
+Thank you for the clarification. We might have mixed this up. Since we have to repeat our simulations, we will choose SWLV3 from the start.
+
+\item {\color{blue}P8 L2. Why did you choose to use the surface (0--7\,cm) soil moisture?}
+  The reason for the initial choice of 0--7\,cm came from the project this work is related to.
+In our project, we are interested in ozone in boreal regions where the soil column is generally more shallow.
+
 \item {\color{blue}P8 L18. This is wrong. Nothing in the EMEP model is used to ’mimic the
@@
 other $f$ factors.}
+  We change this inadequate formulation: \emph{"[...] this integrated photon flux is used to modify the stomatal conductance in response to light."}
+
 \item {\color{blue}P9 L20, and Table 1. The consequence of Table 1 is that vegetation at $0.5^\circ\,\mathrm{N}$
@@
 phenology for non-European areas, because of such difficulties with tabulations.)}
+  Thank you for the remark. We see this problem. Using an actual land model would solve this, but at that point it is not feasible to integrate such in the Oslo~CTM3.
+
 \item {\color{blue}P9 L18. what do you mean by "surface or 2\,m"? Surface might refer to skin or
   leaf temperature?}
+
 \item {\color{blue}P9 Eqn (26). Say 1st and 2nd, not I. and II.}
+  We may change Table~1 accordingly.
+
 \item {\color{blue}P9, Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}

### 2019-05-22 DryDep.tex
Continue with thereferee comments.

@@
 A newer study by \citet{ACP:Luhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
-Conceptually, dry deposition is a product between surface ozone concentration \chem{[O_3](z_0)} and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
+Conceptually, dry deposition is a product between near-surface ozone concentration \chem{[O_3](z_0)} (e.g.
+the lowermost model level) and a dry deposition velocity $v^\chem{O_3}_\text{DD}$.
 Species dependent dry deposition velocities $v^i_\text{DD}$, which are synonymously referred to as conductance $G^i$, for any gaseous species $i$, are typically calculated following a resistance-analogous approach
@@
 Therefore, we reduced the horizontal resolution to $2.25^\circ\times2.25^\circ$ in our experiments.
-In the following, we will give a detailed account of the new dry deposition scheme.
-Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, nonetheless.
+In the following, we will give a detailed account of the new dry deposition scheme and the equations that we use.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Regarding the new dry deposition scheme, we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{WASP:Simpson2003,ACP:Simpson2012}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
+Regarding the new dry deposition scheme, we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{Emberson2000, WASP:Simpson2003, AE:Tuovinen2004, ACP:Simpson2012}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
 We will refer to this as \emph{mOSaic scheme} throughout the rest of the paper.
@@
 \end{equation}
-wherein $\text{LAI}$ is the leaf area index (zero for non-vegetated surfaces), $G_\text{sto}$ the stomatal conductance, and $G_\text{ns}$ the non-stomatal conductance.
+wherein $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR (zero for non-vegetated surfaces), $G_\text{sto}$ the stomatal conductance, and $G_\text{ns}$ the non-stomatal conductance.
 
@@
 The integral of PAR over these wavelengths is the photosynthetic photon flux density (PPFD).
-In the EMEP parameterization, this integrated photon flux is used to mimic the time lag of the opening and closing of stomata with respect to light instead of a simple day-on/night-off treatment:
+In the EMEP parameterization, this integrated photon flux is used to modify the stomatal conductance in response to light:
 \begin{equation}
@@
     \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
-    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, I. part of growing season},\\
-    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, II. part of growing season},\\
+    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, 1. part of growing season},\\
+    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, 2. part of growing season},\\
     0 & \emph{if} \quad \text{\emph{cropland}, winter}.
@@
     \tophline
-    & I. part & II. part \\
+    & 1. part & 2. part \\
     & (days) & (days) \\

### 2019-05-23 version_2/response_version_1_rc2.tex
More work on the response. Almost there to have the first version ready and to start with the actual coading.

@@
 \usepackage[english]{babel}
+%\usepackage[T1]{fontenc}
 \usepackage{amsmath}
@@
 We have, in fact, not presented "wrong inputs" to the algorithm.
-As pointed out by Ref \#1, Eq.
-(2) in our manuscript is erroneous.
-We have not only made a mistake when deriving the formulation for $R_a$, from $u_*$ as presented in S2012 (sign-flip preceding the last term in the series) but also mistype the denominator of the second term.
-The corrected equation:
+As pointed out by Ref \#1, Eq. (2) in our manuscript is erroneous. The corrected equation (Erisman, 1994) for stable atmospheric conditions:
 \begin{equation}
@@
 \item {\color{blue}P9 Eqn (26). Say 1st and 2nd, not I. and II.}
-  We may change Table~1 accordingly.
+  We follow the advise regarding Eq.~(26) and also change Table~1 accordingly.
 
-\item {\color{blue}P9, Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}
+\item {\color{blue}P9 Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}
+  We follow the advise and write: \emph{"The in-canopy resistance $R_\text{inc}$ (Erisman et al., 1994) is then modified with respect to each (vegetated) land type $N$ [...]"} {\bfseries ToDo -- To Amund: Were does this modification come from?}
+
 \item {\color{blue}P10 L12. Be explicit that this statement refers to S2012. The current EMEP
   model uses different heights for e.g. tropical vegetation.}
+  We follow this advice and write: \emph{"The vegetation height $h(N, \text{lat})$ as decribed by Simpson et al. (2012) [...]"}
+
 \item {\color{blue}P11 Sect 2.2. I also found this aerosol section confusing. Eqn (30) is from S2012,
-and so is the factor 0.008.SAI/10 used in Eqn (33), but here new a 1 coefficients
+and so is the factor 0.008.SAI/10 used in Eqn (33), but here new $a_1$ coefficients
 are defined. Did the ’aerosol microphysic model’ referred to also mix equations
 in this way? Is there any publication as to the reliability of this method?}
+  The $a_1$ coefficients were actually calculated from deposition velocities taken from the Oslo~CTM2 which has been socumented {\bfseries ToDo: Citation of CTM2?}.
+
 \item {\color{blue}P12 Fig.2. I didn’t understand what is being done here. The Figure suggests that
@@
 EMEP model has 32.)}
+  The caption may become clearer after we changed the name of our scheme, so that it can no longer be confused with the actual EMEP model.
+The purpose of the mapping is to project the more detailed categories for stomatal conductance to the 10 ground-surface resistance categories (Table~S19 in S2012 supplement).
+
 \item {\color{blue}P12 L13. Again, the current EMEP model is not eurocentric, and uses global
   phenology calculations.}
+  We follow the advice and change the text accordingly: \emph{"Since the eurocentric parameterization of SGS and EGS in Simpson et al. (2012) is not applicable [...]"}
+
 \item {\color{blue}P13 Sect 2.3.2. The initial lines (14-16) are hard to understand and only by
@@
 minute) PPFD using cloud-cover and zenith angles?}
-
+We have not used that ansatz, because we do not have the radiation fields in the relevant wavelengths available from OpenIFS output.
 \end{itemize}

### 2019-05-23 DryDep.tex
More work on the response. Almost there to have the first version ready and to start with the actual coading.

@@
 \label{subsubsec:Ra}
-The aerodynamic resistance is describing the turbulent transport of any substance down to the surface. %In \citet{WASP:Simpson2003,ACP:Simpson2012} it is described as
-%\begin{equation}
-%  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
-%\end{equation}
-%with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$.
+{\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and $\Psi_m$ from \citet{ACP:Seinfeld2006} and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
+Following \citet{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
+\begin{equation}
+  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
+\end{equation}
+with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$.
 For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
@@
 Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation.
-To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
+To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the mOSaic scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
 %
@@
     \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
-    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, 1. part of growing season},\\
-    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, 2. part of growing season},\\
+    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, 1st part of growing season},\\
+    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, 2nd part of growing season},\\
     0 & \emph{if} \quad \text{\emph{cropland}, winter}.
@@
     \tophline
-    & 1. part & 2. part \\
+    & 1st part & 2nd part \\
     & (days) & (days) \\
@@
 
-In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata. The in-canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
+In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata.
+The in-canopy resistance $R_\text{inc}$ \citep{AE:Erisman1994} is then modified with respect to each (vegetated) land type $N$
 \begin{equation}
@@
 \end{equation}
-where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant.
+where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{m^{-1}}$ is an empirical constant.
 According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
@@
 \subsubsection{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ in the mOSaic scheme is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
-To adapt this to a global model, we need to make a few additional assumptions.
+The vegetation height $h(N, \text{lat})$ as decribed by \citet{ACP:Simpson2012} is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
+To adapt this to a global model, we made a few additional assumptions.
 The tabulated height for each vegetation type $h(N)$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
@@
   \caption{Mapping of land surface categories.
-Either land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types used in the EMAP scheme.
+Either land surface categories from ISLSCP2 product of MODIS or the Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types usedwe use in the mOSaic scheme.
 Water bodies and snow and ice categories of MODIS are actually not mapped.
@@
 For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
-Since the eurocentric EMEP parameterization of SGS and EGS is not applicable in a global model, another latitude dependent parameterization is needed.
+Since the eurocentric parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed.
 First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).

### 2019-05-24 version_2/response_version_1_rc2.tex
Finished first draft of responses.

@@
 \section*{Authors' response}
-To gmd-2019-21 David Simpson (10 Apr 2019):
-We apologize for the late response regarding the comments on our paper.
-We had to evaluate the major concerns raised in the review and found that we have to revise our model and repeat the model experiments to address all concerns.
-Nevertheless, we will respond to the questions in more detail in the following.
+To gmd-2019-21 referee\#2 David Simpson (10 Apr 2019):
+We apologize for the confusion the misuse of the terms \emph{EMEP} and \emph{EMEP scheme} may have caused.
+We evaluated the major concerns raised in the review and found them severe enough that we have to revise our model and repeat the model experiments to address all concerns.
+Nevertheless, we will respond to all of the questions and concerns in more detail in the following.
 
@@
 The CTM3 $R_b$ calculation seems to also use the same $u_*$ over different land-covers, except over sea where a more sophisticated scheme is used.
-Equations 5-6 above are equivalent to Eqns (22) and (26) from the manuscript.
+Equations 2--3 above are equivalent to Eqns (22) and (26) from the manuscript.
 Now I am puzzled however as to how all this is put together.
@@
   }
-    The referee is correct regarding the usage of Eq.~(\ref{eq:ra}) and Eq.~(\ref{eq:Gns}). In case of stomatal conductance, we first compute:
+    Thank you very much for your detailed account of the \emph{mosaic approach}.
+It is correct that we use Eq.~(\ref{eq:Gsto}) and Eq.~(\ref{eq:Gns}).
+As pointed out, we have not properly indicated the weighted mean in Eq.~(22) and Eq.~(29) in the manuscript.
+Additionally, there is also a typo in Eq.~(13) and Eq.~(22) ($G_\mathrm{sto}$ should have been $g_\mathrm{sto}$).
+Eq.~(22) should have read as follows:
     \begin{equation}
-       \overline{gs^i_g(z)} = \Sigma^N_{k=1} f_k \times gs^i_{g,k}(z).
+      \overline{g^\mathrm{m}_\text{sto}} = \sum_{n=0}^{N} f_{\text{L}, n}\cdot g^\text{m}_{\text{sto}, n}
     \end{equation}    
-    This is put together as follows:
+    All is put together in Eq.~(13):
     \begin{equation}
       \label{eq:Gsto_corr}
-      G_c = \mathrm{LAI} \cdot \overline{gs^i_g(z)} + \overline{Gns^i_g(z)}
+      G_c = \mathrm{LAI} \cdot \overline{g^\mathrm{m}_\mathrm{sto}} + \overline{G_\mathrm{ns}}
     \end{equation}
+    We acknowledge that Eq.~(13) should not procede the other equations and that this might have caused additional confusion.
+
   \item {\color{blue} In any case, I think this approach has serious problems.
@@
 I think the authors need to make a case for their approach, or change it.}
-    Thank you very much for your detailed account of the \emph{mosaic approach}.
     We have discussed the concerns raised within this comment and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
   \end{itemize}
+
 \item {\color{blue} Calculation of $R_a$:}
@@
 used for decades (Garratt 1992). As pointed out by Ref~1, the equation as given has
-errors. The correct equation will not give negative values unless presented with wrong
-inputs, and I suspect that that is what has happened.}
-We have, in fact, not presented "wrong inputs" to the algorithm.
-As pointed out by Ref \#1, Eq. (2) in our manuscript is erroneous. The corrected equation (Erisman, 1994) for stable atmospheric conditions:
+errors.}
+Eq.~(2) in our manuscript is indeed erroneous. The correct equation is (e.g., Erisman, 1994):
 \begin{equation}
@@
 
-\item {\color{blue} It is actually difficult to tell what
+\item {\color{blue} The correct equation will not give negative values unless presented with wrong
+inputs, and I suspect that that is what has happened. It is actually difficult to tell what
 was tested from the manuscript though, since they state simply that $d$ is ’typically 0.7\,m’.
@@
 If the latter, which $d$?}
-  It is indeed not clear from the manuscript that we never actually implemented the formula (Eq.~(2) in our manuscript) in the model.
-The reason for this were simple, mathematical considerations.
-Given $z = z_\mathrm{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$, we looked at certain parameter spaces for $L$ and $z_0$ (e.g., $0 \le z_0 \le 2.5$ for vegetated land surfaces from a climatology mean derived from ISLSCP2 FASIR) but have not taken the consistency "with the underlying land-cover and its $z_0$" into account (see Fig.~\ref{fig:test_ra}).
-This might have been not unsophisticated enough, though.
-The statement "$d$ is ’typically 0.7\,m’" is a typo and should of cause be corrected to \emph{typically $d = 0.7\cdot h(N,\mathrm{lat})\,\mathrm{m}$ for vegetation other than forests}.
-Note: The average height of the lowermost model level is actually $20\,\mathrm{m}$ ($10\,\mathrm{m}$ for mid-level).
+  At that point, it is indeed neither clear from the manuscript nor from our internal documentation.
+The assumption that Eq.~(2) (in the manuscript) would become negative, was most likly based on a wrong form of the integrated stability function $\Psi_m$.
 
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{test_Ra_wrong_form.pdf}
-    \caption{Test cases for $R_a$ with $z = z_{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$.}
-    \label{fig:test_ra}
-  \end{figure}
+  Given $z = z_\mathrm{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$, it is right that neither the correct Eq.~(\ref{eq:aero_res}) nor the erroneous equation will result in negative $R_a$.
+The statement "$d$ is ’typically 0.7\,m’" is a typo.
+We correct: \emph{"[...] typically $d = 0.7\cdot h(N,\mathrm{lat})\,\mathrm{m}$ for vegetation other than forests [...]"}.
+
+  Correction note: The average height of the lowermost model level is actually $20\,\mathrm{m}$ ($10\,\mathrm{m}$ for mid-level).
 
@@
   The referee is right.
-The text and derivation of the Monteith relation is confusing.
-We will elaborate on this in a revised version of the manuscript in case we are not going to use Eq.~(\ref{eq:aero_res}).
+The text and derivation of the Monteith relation is confusing and also erroneous.
+If we do not implement Eq.~(\ref{eq:aero_res}), we will elaborate on this in a revised version of the manuscript: \emph{"
+\begin{equation}
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
+  \label{eq:sens_heat}
+\end{equation}
+Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
+From the eddy diffusion theorem, a similar formulation can be derived
+\begin{equation}
+  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
+  \label{eq:eddy_theo}
+\end{equation}
+For finite $z$, $\partial T \rightarrow \Delta T$.
+  "}
 
@@
 discussing that fact.}
-  We shall elaborate on this.
+  In the revison of the model, we are going to implement the stability dependent $R_a$ following Eq.~(\ref{eq:aero_res}).
+
 \item {\color{blue} This very shallow layer is also very problematic for deposition calculations in general,
@@
 differences were largely dependent on how much forest occupied each grid cell.)}
-  In S2012, $R_a$ is evaluated "at around $45\,\mathrm{m}$" which is similar to the center of our second lowest model level. We can easily correct for this in a revised model version. 
+  Thank you for pointing this out.
+In S2012 we find, $R_a$ evaluated "at around $45\,\mathrm{m}$" which is similar to the center of our second lowest model level.
+We can easily correct for this in a revised model version.
   \end{itemize}
+
 \item {\color{blue} Why so much focus on sea areas?: The text seems rather unbalanced with regard to the different land-covers.
@@
 
-  Regarding the three figures with respect to ocean in the supplementary material: The reason for these is that we found the "legacy code" in the model (linear fit dynamic viscosity of air $\mu(T)$) after we had finished the implementation of the new dry deposition scheme, run the experiments, and done the analysis on these.
-We had to verify that this had no influence on the results.
+  Regarding the three figures with respect to ocean in the supplementary material: The reason for these is that we found some "legacy code" in the model (linear fit to dynamic viscosity of air $\mu(T)$) after we had finished the implementation of the new dry deposition scheme, run the experiments, and done the analysis on these.
+We had to verify that this had no significant influence on the results.
 
@@
 We will rename the revised scheme appropriately ($\rightarrow$\emph{mOSaic scheme}).
+
 \item {\color{blue} Reproduction of material from S2012:}
@@
     We will follow this advice and refer to the tables in S2012 accordingly.
+
   \item {\color{blue} Many of the equations are from S2012, and many not. I would like the authors to
@@
     and what has changed for CTM3.}
-    We are going to elaborate on this matter in a revised version and point out our equations' sources more clearly.
+    We are going to elaborate on this matter in a revised version and refere to our equations' sources more properly.
   \end{itemize}
@@
 (2018b).}
-  Thank you for pointing out these new publications. We take them into consideration, when revising the manuscript.
+
+  Thank you for pointing out these publications. We take them into consideration, when revising the manuscript.
 \item {\color{blue}P2 L11. What does in situ mean here? Ozone production can take place over
@@
 In this context, we use it exactly in this meaning.
-Elevated ozone levels at a site originate mainly from the local production of ozone from its precurors that are transported than from advection of ozone itself.
-Although long-range ozone transport occures and might be most important in regions that else lack precursors.
-We elaborate on this in a revised version of the paper: \emph{Tropospheric ozone is mainly produced locally [...] involving advected precursor gases [...]}
+We elaborate on this in a revised version of the paper: \emph{Elevated ozone levels at a site originate rather from the local production of ozone from its precursors, which are transported, than from advection of ozone itself.
+Although long-range ozone transport occurs and might be most important in regions that else lack precursors.
+Tropospheric ozone is produced in complex photochemical cycles involving precursor gases [...]}
+
 \item {\color{blue}P2 L25--35. This text about halogens is not really relevant to a dry deposition
@@
 as deposition.}
-  The removal of ozone from the Arctic boundary layer in the precence of halogens is indeed not directly related to ozone dry deposition, but one of the discussed mechanisms to trigger the bromine explosions involves, e.g., a "initialisation" by ozone dry deposition.
-But we see that this is of no relevance to ozone dry deposition itself, but to the observed ozone abundances in the Arctics.
-We may remove these reference in the revised version of the manuscript.
+  The removal of ozone from the Arctic boundary layer in the presence of halogens is indeed not directly related to ozone dry deposition, but one of the discussed mechanisms to trigger the bromine explosions involves, e.g., an "initialization" by ozone dry deposition.
+We see that this is of no relevance to ozone dry deposition itself, but to the observed ozone abundances in the Arctics.
+We remove this text in the revised version of the manuscript.
 
@@
   Thank you for reminding us of the data collected under the TOAR project.
-We knew about this dataset but have to admitt that we have not made much use of it, yet.
-We have, however, not stated that there are, in general, no sites outside of Europe, but that long-term obersvations (started before the 1950s) are only found in Europe.
+We knew about this data set but have to admit that we have not made much use of it, yet.
+We have, however, not stated that there are, in general, no sites outside of Europe, but that long-term observations (started before the 1950s) are only found in Europe.
 Maybe the term "long-term" is not clear enough.
@@
 case between their equations (13) and (14).}
-  This is probably related to the referee's major point 1.
-Since our approach, averaging the resistances/conductances instead of the fluxes, is incorrect, this part of the manuscript has to be revised in any case.
-But to answer the question: Eq.~(13) should have been placed at the end of the section, after the definition of $G_\mathrm{sto}$ and $G_\mathrm{ns}$.
-A better notation is found Eq.~(\ref{eq:Gsto_corr}).
-We shall also go through all our notations in that section to make it more comprehensible.
+  We have indeed mixed up the cases here.
 
@@
 
-\item {\color{blue}P9 L18. what do you mean by "surface or 2\,m"? Surface might refer to skin or
-  leaf temperature?}
+\item {\color{blue}P9 L18. What do you mean by "surface or 2\,m"? Surface might refer to skin or leaf temperature?}
+  This may have been indeed unclear. It is the $\mathrm{2\,m}$ temperature. We change the text accordingly: \emph{"[...] $\theta_0$ is the $2\,\mathrm{m}$ temperature in $\mathrm{^\circ C}$."}
 
@@
   model uses different heights for e.g. tropical vegetation.}
-  We follow this advice and write: \emph{"The vegetation height $h(N, \text{lat})$ as decribed by Simpson et al. (2012) [...]"}
+  We follow this advice and write: \emph{"The vegetation height $h(N, \text{lat})$ as described by Simpson et al. (2012) [...]"}
 
@@
 in this way? Is there any publication as to the reliability of this method?}
-  The $a_1$ coefficients were actually calculated from deposition velocities taken from the Oslo~CTM2 which has been socumented {\bfseries ToDo: Citation of CTM2?}.
+  The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 which has been documented {\bfseries ToDo: Citation of CTM2?}.
+We have actually not used this new scheme for aerosols in the Oslo~CTM3, yet.
+Hence, we will remove this section from the paper.
 
@@
   phenology calculations.}
-  We follow the advice and change the text accordingly: \emph{"Since the eurocentric parameterization of SGS and EGS in Simpson et al. (2012) is not applicable [...]"}
+  We follow the advice and drop the term "eurocentric" the text: \emph{"Since the parameterization of SGS and EGS in Simpson et al. (2012) is not applicable [...]"}
 

### 2019-05-24 DryDep.tex
Finished first draft of responses.

@@
 The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} {\bf TODO: Read through papers and confirm numbers.} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-{\bf TODO: Think about how to balance the wording of were ozone originates from transport VS local production.}Tropospheric ozone is mainly produced locally in complex photochemical cycles involving advected precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
+Elevated ozone levels at a site originate rather from the local production of ozone from its precursors, which are transported, than from advection of ozone itself.
+Although long-range ozone transport occurs and might be most important in regions that else lack precursors.
+Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
 A typical reaction mechanism for \chem{CO} is sketched in the following.
@@
 \label{subsubsec:Ra}
-{\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and $\Psi_m$ from \citet{ACP:Seinfeld2006} and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
+{\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and note down the right $\Psi_m$ and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
 Following \citet{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
@@
 with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$.
-For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$.
-For this reason, we diverge from the mOSaic scheme at this point and fall back to
-For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
-\begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
-  \label{eq:sens_heat}
-\end{equation}
-Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
+
+%For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$. For this reason, we diverge from the mOSaic scheme at this point and fall back to
+%For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
+%\begin{equation}
+%  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
+%  \label{eq:sens_heat}
+%\end{equation}
+%Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
 From the eddy diffusion theorem, a similar formulation can be derived
-\begin{equation}
-  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
-  \label{eq:eddy_theo}
-\end{equation}
-For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
-\begin{equation}
-  R_a = \frac{u_z}{u_*^2}.
-\end{equation}
+%\begin{equation}
+%  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
+%  \label{eq:eddy_theo}
+%\end{equation}
+%For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
+%\begin{equation}
+%  R_a = \frac{u_z}{u_*^2}.
+%\end{equation}
 As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$).
@@
 \begin{equation}
-  G_c = \text{LAI} \cdot G_\text{sto} + G_\text{ns}, 
+  G_c = \text{LAI} \cdot g_\text{sto} + G_\text{ns}, 
 \end{equation}
@@
 \begin{equation}
-  G_\text{sto} = \sum_{N=0}^{N_\text{max}} f_L(N)\cdot g^\text{m}_{\text{sto}}(N)
+  g_\text{sto} = \sum_{N=0}^{N_\text{max}} f_L(N)\cdot g^\text{m}_{\text{sto}}(N)
 \end{equation}
@@
 \end{equation}
-$\theta_0$ is the surface or $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
+$\theta_0$ is the $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
 %
@@
 For the growing season of crops in the computation of non-stomatal conductances, we use also prescribed values (Table~\ref{tab:growing_season}), while for the stomatal conductances, as shown in Eq.~(\ref{eq:fphen}), we use the SGS and EGS derived parameters: day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS.
-Since the eurocentric parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed.
+Since the parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed.
 First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).

### 2019-06-21 version_2/response_version_1_rc2.tex
Begin with finalizing the statements.

@@
   \emph{In situ} typically means \emph{on site, locally}.
-In this context, we use it exactly in this meaning.
-We elaborate on this in a revised version of the paper: \emph{Elevated ozone levels at a site originate rather from the local production of ozone from its precursors, which are transported, than from advection of ozone itself.
-Although long-range ozone transport occurs and might be most important in regions that else lack precursors.
+In this context, we used it exactly in this meaning.
+We elaborate on this in a revised version of the paper: \emph{Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
+Long-range ozone transport occurs and might be most important in regions that else lack precursors.
 Tropospheric ozone is produced in complex photochemical cycles involving precursor gases [...]}
@@
 as deposition.}
-  The removal of ozone from the Arctic boundary layer in the presence of halogens is indeed not directly related to ozone dry deposition, but one of the discussed mechanisms to trigger the bromine explosions involves, e.g., an "initialization" by ozone dry deposition.
-We see that this is of no relevance to ozone dry deposition itself, but to the observed ozone abundances in the Arctics.
-We remove this text in the revised version of the manuscript.
-
+  The removal of ozone from the Arctic boundary layer in the presence of halogens is indeed not part of ozone dry deposition and we remove this reference in this context.
+Although these processes are not relevant with respect to ozone dry deposition, they matter regarding the observed ozone abundances in the Arctics.
+One of the proposed mechanisms to trigger bromine explosions involves, e.g., an "initialization" by ozone dry deposition.
 \item {\color{blue}P3 L2--3. Why specify mid-June maximum for ozone. Monks (2000) might
@@
 Maybe the term "long-term" is not clear enough.
-We add: \emph{"[...] a limited number of long-term ozone observations (started before the 1950s)[...]"}
+We add: \emph{"From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sides."}
 
@@
   We have added the references regarding the EMEP MSC-W model.
-\emph{"[...] we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model (Emberson et al., 2000; Simpson et al., 2003; Tuovinen et al., 2004; Simpson et al., 2012), [...]"}
+\emph{"[...] we mainly follow Simpson et al.
+(2012) in their description of dry deposition used in the European Monitoring and Evaluation Programme (EMEP) MSC-W model (see also, Emberson et al., 2000; Simpson et al., 2003; Tuovinen et al., 2004;), [...]"}
 
@@
   The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 which has been documented {\bfseries ToDo: Citation of CTM2?}.
-We have actually not used this new scheme for aerosols in the Oslo~CTM3, yet.
-Hence, we will remove this section from the paper.
+We do actually not used this new scheme for aerosols in the Oslo~CTM3, yet.
+Hence, we will remove this section from the paper and add: \emph{"In addition to the gaseous species, Simpson at al.
+(2012) also has some of the aerosol deposition velocities modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we do not update our model with respect to these."}.
 

### 2019-06-21 DryDep.tex
Begin with finalizing the statements.

@@
 Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
-In recent years, the number of episodes of peak concentrations has been decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018} {\bf TODO: Read through papers and confirm numbers.}.
+In recent years, the number of episodes of peak concentrations has been, in general, decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
 At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
@@
 It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} as well as $7\,\unit{\%}$ loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
-The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} {\bf TODO: Read through papers and confirm numbers.} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-Elevated ozone levels at a site originate rather from the local production of ozone from its precursors, which are transported, than from advection of ozone itself.
-Although long-range ozone transport occurs and might be most important in regions that else lack precursors.
+The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
+Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
+Long-range ozone transport occurs and might be most important in regions that else lack precursors.
 Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
@@
 
-%A dry deposition related sink limited to Arctic regions are so called ozone depleting events.
-They occur in spring-time in the polar boundary layer where an outburst of bromine monoxide \chem{BrO} (so called bromine explosion) leads to a rapid depletion of surface ozone \citep{JGR:Oltmans1981,GRL:Bottenheim1986,Nat:Barrie1988,JGR:Bottenheim2006}.
-Various schemes ranging from bulk-snow parameterization \citep{ACP:Toyota2011,GMD:Falk2018} to detailed in-snow \citep{ACP:Toyota2014a}, and aerosol chemistry \citep{ACP:Yang2010} have been successfully applied to different types of atmospheric models but do not yet cover the full range of observed events.\\
 Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
@@
 This variability poses a challenge on both, trend analysis from observation as well as validation and intercomparison of models.
-At the observational side, there is only a limited number of long-term ozone observations (started before the 1950s), mainly restricted to European sites.
+From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sides.
 Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
@@
 In Section~\ref{sec:model_des}, we will briefly describe the Oslo~CTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
-In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations at the surface (Section~\ref{subsec:obs}).
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to surface ozone observations (Section~\ref{subsec:obs}).
 In Section~\ref{sec:conc}, we will summarize and discuss our results and draw conclusions for further development of the model.
@@
 A publication focusing on this is planed.
-Although the ozone depleting events in the polar boundary layer (Section~\ref{sec:intro}) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.\\
+
+%Although the ozone depleting events in the polar boundary layer (Section~\ref{sec:intro}) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.\\
 While the meteorological data driving the Oslo~CTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
@@
 
-Regarding the new dry deposition scheme, we follow the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep{Emberson2000, WASP:Simpson2003, AE:Tuovinen2004, ACP:Simpson2012}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
+Regarding the new dry deposition scheme, we mainly follow \citet{ACP:Simpson2012} in their description of dry deposition used in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep[see also,][]{Emberson2000, WASP:Simpson2003, AE:Tuovinen2004}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
 We will refer to this as \emph{mOSaic scheme} throughout the rest of the paper.
@@
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
-In addition to the gaseous species, some of the aerosol deposition velocities are also modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA).
+In addition to the gaseous species, \citet{ACP:Simpson2012} also has some of the aerosol deposition velocities modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we do not update our model with respect to these.
 
@@
 {\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and note down the right $\Psi_m$ and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
-Following \citet{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
-\begin{equation}
-  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_m\left(\frac{z-d}{L}\right)+\Psi_m\left(\frac{z_0}{L}\right)}\right],
+Following \citet[]{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
+\begin{equation}
+  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_h\left(\frac{z-d}{L}\right)+\Psi_h\left(\frac{z_0}{L}\right)}\right],
 \end{equation}
@@
 
-%For certain values of $z$, $z_0$, and $L$, this may result in nonphysical (negative) values for $R_a$. For this reason, we diverge from the mOSaic scheme at this point and fall back to
-%For technical reasons, we used the method of \citet{Monteith1973}, wherein the sensible heat flux $\Phi_{Q_\text{sensible}}$ is written as
-%\begin{equation}
-%  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\Delta T}{R_{a, H}}.
-%  \label{eq:sens_heat}
-%\end{equation}
-%Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
-From the eddy diffusion theorem, a similar formulation can be derived
-%\begin{equation}
-%  \Phi_{Q_\text{sensible}} = \rho_\text{air}\,c_P \cdot \frac{\partial T}{\partial u} \cdot u_*^2.
-%  \label{eq:eddy_theo}
-%\end{equation}
-%For finite $z$, $\partial T \rightarrow \Delta T$. Comparing Eq.~(\ref{eq:sens_heat}) and Eq.~(\ref{eq:eddy_theo}) and assuming $R_{a, H} \equiv R_a$, we find the stability independent equation
-%\begin{equation}
-%  R_a = \frac{u_z}{u_*^2}.
-%\end{equation}
 As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$).

### 2019-06-25 version_2/response_version_1_rc2.tex
Rewrote aerodynamical and quasi-laminar resistance text. Changed mapping picture.

@@
 Hence, we will remove this section from the paper and add: \emph{"In addition to the gaseous species, Simpson at al.
-(2012) also has some of the aerosol deposition velocities modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we do not update our model with respect to these."}.
+(2012) also modify aerosol deposition velocities, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we have not updated our model with respect to these."}.
 

### 2019-06-25 DryDep.tex
Rewrote aerodynamical and quasi-laminar resistance text. Changed mapping picture.

@@
 \label{subsec:DryDep}
-In the original dry deposition scheme of the Oslo~CTM3, the state of the atmosphere was not taken into account.
+In the original dry deposition scheme, the state of the atmosphere was not taken into account.
 Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}.
@@
 Regarding the new dry deposition scheme, we mainly follow \citet{ACP:Simpson2012} in their description of dry deposition used in the European Monitoring and Evaluation Programme (EMEP) MSC-W model \citep[see also,][]{Emberson2000, WASP:Simpson2003, AE:Tuovinen2004}, which is used for air quality modeling implementing the Convention on Long-Range Transboundary Air Pollution (CLRTAP).
-We will refer to this as \emph{mOSaic scheme} throughout the rest of the paper.
+We will refer to the new scheme as \emph{mOSaic scheme} throughout the rest of the paper.
 The mOSaic scheme is a more physical approach compared to the previously used Wesely scheme, because it takes state (e.g., pressure, temperature)  of the atmosphere as well as dynamics (e.g., wind stress) of the boundary layer into account.
@@
 Since \chem{CO} has a very small uptake and is not included in \citet{WASP:Simpson2003,ACP:Simpson2012}, the Wesely parameterization is kept.
-In addition to the gaseous species, \citet{ACP:Simpson2012} also has some of the aerosol deposition velocities modified, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we do not update our model with respect to these.
-
-As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is typically split into three different resistance contributions (aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy $R^i_c$) which we will recap in the following.
+In addition to the gaseous species, \citet{ACP:Simpson2012} also modify aerosol deposition velocities, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we have not updated our model with respect to these.
+
+As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is subdivided into contributions from three different resistances.
+The main idea of a mosaic approach is to calculate these resistences dependent on the fraction of land surface types $k$ in each grid cell: $R_a^k$,$R^{i,k}_b$, and $R^{i,k}_c$.
 
@@
 \label{subsubsec:Ra}
-{\bf ToDO: Implement \citet{AE:Erisman1994} formulation for Ra with d's from \citet{ACP:Simpson2012} and note down the right $\Psi_m$ and remove text about Monteith alternative!} The aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
-Following \citet[]{AE:Garland1978}, \citet{WASP:Simpson2003,ACP:Simpson2012} has
-\begin{equation}
-  R_a = \frac{1}{\kappa u_*}\left[{\ln{\left(\frac{z-d}{z_0}\right)}-\Psi_h\left(\frac{z-d}{L}\right)+\Psi_h\left(\frac{z_0}{L}\right)}\right],
-\end{equation}
-with the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the friction velocity $u_*$, integrated stability equations for momentum $\Psi_m$,  a the deplacement height $d$ (typically $0.7\cdot h(N,\mathrm{lat})\,\unit{m}$ for vegetation other than forests), and the Obukhov length $L$.
-
-As reference height $z$, we use the height at mid-level of the lowermost model level (on average $10\,\unit{m}$).
-From the meteorological input fields $u_z$ is directly available, while $u_*$ can be computed via
-\begin{equation}
-  u_* = \sqrt{\frac{\tau_0}{\rho}},
-\end{equation}
-with the turbulent surface stress $\tau_0^2 = \tau_v^2 + \tau_u^2$, and northward turbulent surface stress $\tau_v$, eastward turbulent surface stress $\tau_u$.
+In general, the aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
+To derive $R_a^k$, \citet{WASP:Simpson2003,ACP:Simpson2012} compute a local friction velocity at reference height $z_\mathrm{ref}$ \citep[Eq.~(52),][]{ACP:Simpson2012}
+\begin{equation}
+  u_*^k = \frac{\overline{u}(z_\mathrm{ref})\cdot\kappa}{\ln{(\frac{z_\mathrm{ref}-d_k}{z_0^k})}-\Psi_m(\frac{z_\mathrm{ref}-d_k}{L})+\Psi_m(\frac{z_0^k}{L})},
+  \label{eq:local_fric_vel}
+\end{equation}
+with an average wind speed $\overline{u}(z_\mathrm{ref})$, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and surface height $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
+We have chosen a model level where $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
+From Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ is found \citep[Eq.~(8),][]{ACP:Simpson2012}:
+\begin{equation}
+  L_k = - \frac{\rho c_p T_\mathrm{2m} u_*^k}{\kappa g H}.
+\end{equation}
+Herein, $H$ is the sensible heat flux, $g$ is the standard gravitational acceleration, $c_p$ the specific heat capacity, and $T_{2m}$ the $2\,\unit{m}$ temperature.
+With this we get \citep[Eq.~(8.8),][]{WASP:Simpson2003}
+\begin{equation}
+  R_a^k = \frac{1}{\kappa u_*^k}\left[{\ln{\left(\frac{z_\mathrm{ref}-d_k}{z_0^k}\right)}-\Psi_h\left(\frac{z_\mathrm{ref}-d_k}{L_k}\right)+\Psi_h\left(\frac{z_0^k}{L_k}\right)}\right],
+\end{equation}
+with the integrated stability equation for heat $\Psi_h$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsubsec:Rb}
-The quasi-laminar layer resistance $R_b^i$ is species specific and differs over land and ocean surfaces. In case a gridbox contains both, a weighted mean is calculated.
+The quasi-laminar layer resistance $R_b^{i,k}$ is species specific and differs only over land and ocean surfaces.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{equation}
-  R_b^i = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
+  R_b^{i,k} = \frac{2}{\kappa u_*} \cdot \left(\frac{\text{Sc}_i}{\text{Pr}}\right)^{\frac{2}{3}},
   \label{eq:Rb}
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement~S.1: Figs.~S1--S2).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b^{i,k}$ (Supplement~S.1: Figs.~S1--S2).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.

### 2019-06-26 version_2/response_version_1_rc2.tex
First version of revised model description.

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \thispagestyle{empty}
@@
     \end{equation}
-    We acknowledge that Eq.~(13) should not procede the other equations and that this might have caused additional confusion.
+    We acknowledge that Eq.~(13) should not proceed the other equations and that this might have caused additional confusion.
 
@@
   At that point, it is indeed neither clear from the manuscript nor from our internal documentation.
-The assumption that Eq.~(2) (in the manuscript) would become negative, was most likly based on a wrong form of the integrated stability function $\Psi_m$.
+The assumption that Eq.~(2) (in the manuscript) would become negative, was most likely based on a wrong form of the integrated stability function $\Psi_m$.
 
@@
 discussing that fact.}
-  In the revison of the model, we are going to implement the stability dependent $R_a$ following Eq.~(\ref{eq:aero_res}).
+  In the revision of the model, we are going to implement the stability dependent $R_a$ following Eq.~(\ref{eq:aero_res}).
 
@@
     and what has changed for CTM3.}
-    We are going to elaborate on this matter in a revised version and refere to our equations' sources more properly.
+    We are going to elaborate on this matter in a revised version and refer to our equations' sources more properly.
   \end{itemize}
@@
   is from 28-100 cm.}
-Thank you for the clarification. We might have mixed this up. Since we have to repeat our simulations, we will choose SWLV3 from the start.
+Thank you for the clarification. This may have been a misunderstanding on our side. Since we have to repeat our simulations, we will choose the appropriate SWVL from the start.
 
 \item {\color{blue}P8 L2. Why did you choose to use the surface (0--7\,cm) soil moisture?}
-  The reason for the initial choice of 0--7\,cm came from the project this work is related to.
-In our project, we are interested in ozone in boreal regions where the soil column is generally more shallow.
+  Initially we choose 0--7\,cm due to the original scope of this work within our research project, in which we study ozone concentrations and uptake by plants in boreal regions where the soil columns are expected to be more shallow then, e.g., at mid-latitudes.
+We expect the stomatal conductance in these environments to be more sensitive to precipitation.
+Using the actual water availability to vegetation, would be better, but this is not feasible without a proper land surface model.
 
@@
 in this way? Is there any publication as to the reliability of this method?}
-  The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 which has been documented {\bfseries ToDo: Citation of CTM2?}.
-We do actually not used this new scheme for aerosols in the Oslo~CTM3, yet.
+  The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 (\href{http://folk.uio.no/mariantl/osloctm3/history.html}{Oslo~CTM2 list of publications}).
+We meant to use this new scheme for aerosols in the Oslo~CTM3, but as of now it has not been evaluated and is actually not used.
 Hence, we will remove this section from the paper and add: \emph{"In addition to the gaseous species, Simpson at al.
-(2012) also modify aerosol deposition velocities, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols (\chem{SO_4}, \chem{MSA}) and secondary organic aerosols (SOA), but we have not updated our model with respect to these."}.
+(2012) also modify aerosol deposition velocities, namely black carbon (BC) and organic carbon (OC), sulfuric aerosols ($\mathrm{SO_4}$, $\mathrm{MSA}$) and secondary organic aerosols (SOA), but we have not updated our model with respect to these."}.
 

### 2019-06-26 DryDep.tex
First version of revised model description.

@@
 As displayed in Eq.~(\ref{eq:drydep_velo}), the dry deposition computation is subdivided into contributions from three different resistances.
-The main idea of a mosaic approach is to calculate these resistences dependent on the fraction of land surface types $k$ in each grid cell: $R_a^k$,$R^{i,k}_b$, and $R^{i,k}_c$.
+The main idea of a mosaic approach is to calculate these resistances separately for each land surface type $k$ in each grid cell: $R_a^k$,$R^{i,k}_b$, and $R^{i,k}_c$.
+The grid cell average dry deposition velocity $\overline{v}_\mathrm{DD}^{i}$ is then defined by weighting each individual $v_\mathrm{DD}^{i,k}$ by the corresponding land fraction factor $f_k$:
+\begin{equation}
+  \overline{v}_\mathrm{DD}^{i} = \sum_k{f_k v_\mathrm{DD}^{i,k}}
+  \label{eq:ddep_velo_wa}
+\end{equation}
 
@@
 In general, the aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
-To derive $R_a^k$, \citet{WASP:Simpson2003,ACP:Simpson2012} compute a local friction velocity at reference height $z_\mathrm{ref}$ \citep[Eq.~(52),][]{ACP:Simpson2012}
+To derive $R_a^k$, we follow \citet{WASP:Simpson2003,ACP:Simpson2012} and compute a local friction velocity at reference height $z_\mathrm{ref}$ \citep[Eq.~(52),][]{ACP:Simpson2012}
 \begin{equation}
@@
 \end{equation}
-with an average wind speed $\overline{u}(z_\mathrm{ref})$, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and surface height $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
-We have chosen a model level where $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
-From Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ is found \citep[Eq.~(8),][]{ACP:Simpson2012}:
+with the average wind speed $\overline{u}(z_\mathrm{ref})$ at reference height, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and roughness length $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
+We have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
+Using the derived $u_*^k$ from Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ can by obtained with \citep[Eq.~(8),][]{ACP:Simpson2012}:
 \begin{equation}
@@
 Herein, $H$ is the sensible heat flux, $g$ is the standard gravitational acceleration, $c_p$ the specific heat capacity, and $T_{2m}$ the $2\,\unit{m}$ temperature.
-With this we get \citep[Eq.~(8.8),][]{WASP:Simpson2003}
+With these, we get \citep[Eq.~(8.8),][]{WASP:Simpson2003}
 \begin{equation}
@@
 \end{equation}
-with the integrated stability equation for heat $\Psi_h$.
+with the integrated stability equation for heat $\Psi_h$ \citep[e.g.,][]{Garratt1992}.
+Both integrated stability functions ($\Psi_m$, $\Psi_h$) and corresponding parameters are listed in supplement~S.1.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsubsec:Rb}
-The quasi-laminar layer resistance $R_b^{i,k}$ is species specific and differs only over land and ocean surfaces.
+The quasi-laminar layer resistance $R_b^{i,k}$ is species specific and differs over land and ocean surfaces.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$.
-But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b^{i,k}$ (Supplement~S.1: Figs.~S1--S2).
+But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b^{i,k}$ (Supplement~S.2: Figs.~S1--S2).
 In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law.
@@
 The allowed maximum roughness length in both cases is set to $2\,\unit{mm}$.
-Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.1: Fig.~S3).
+Since the $z_0$ computed with this parameterization are rather small ($0 < z_0^\text{calm} < 1\cdot 10^{-4}\,\unit{m}$, $0 < z_0^\text{rough} < 2 \cdot 10^{-3}\,\unit{m}$), $R^i_b$ is set to its lower limit of $10\,\unit{s\,m^{-1}}$ in about $91\,\unit{\%}$ of all cases (see Supplement~S.2: Fig.~S3).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{subsubsec:Rc}
-The surface resistance consists of both, stomatal and non-stomatal resistances. When computing gridbox averages, we use the sum of conductances (deposition velocities)
-\begin{equation}
-  G_c = \text{LAI} \cdot g_\text{sto} + G_\text{ns}, 
-\end{equation}
-wherein $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR (zero for non-vegetated surfaces), $G_\text{sto}$ the stomatal conductance, and $G_\text{ns}$ the non-stomatal conductance.
-
+The surface resistance consists of both, stomatal and non-stomatal resistances.\\ 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The stomatal conductance is a measure of the rate of \chem{CO_2} exchange and evapotranspiration through the stomata of a leaf.
-There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.).
+There are several environmental conditions affecting the opening and closing of the stomata and hence the capability of respiration (e.g., light, available water, etc.).
 Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation.
-To reflect part of the underlying mechanism, the stomatal conductance $G_\text{sto}$ in the mOSaic scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
-%
-\begin{equation}
-  G_\text{sto} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
+To reflect part of the underlying mechanism, the leaf-level stomatal conductance in the mOSaic scheme is computed using a common multiplicative ansatz \citep{BallBerry1987, ICP:MappingManual2017}:
+%
+\begin{equation}
+  g^k_\text{sto, m} = g^k_\text{max, m} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_T \cdot f^k_D \cdot f^k_\text{SW}\right\}}.
   \label{eq:stomatal}
@@
 %
-The factors herein are normalized and vary within the range $0-1$.
-They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
-
-The temperature adjustment $f_T$ of plants is computed from
+The factors herein are normalized and vary within the range $0-1$ and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
+All factors differ with land use type $k$.
+For clarity reasons, we drop this index in the following, as long is it not necessary for the equation.
+
+The temperature adjustment $f_T$ is computed from
 %
@@
 with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$.
-The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various land surface types.
-All parameters used in the mOSaic scheme can be found in the supplementary to \citet[][Tables S16, S19]{ACP:Simpson2012}.
-Since $f_T$ turns negative outside its range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various plant functional types.
+All parameters are taken from Supplementary \citet[][Tables S16, S19]{ACP:Simpson2012}.
+Since $f_T$ turns negative outside the range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
 
@@
 \end{equation}
-\citet{ACP:Simpson2012} suggest using SW of soil depths below $1\,\unit{m}$ (SWVL4 in OpenIFS).
-Throughout most of our simulations (Section~\ref{sec:eval}), we have, however, used the soil water contained in the top layer (SWVL1, $0-7\,\unit{cm}$), which makes the canopy conductance more sensitive to precipitation than it would be otherwise.
+SW is evaluated at a soil depths of $0.28-1\,\unit{m}$, which corresponds to SWVL3 in OpenIFS.
 
 The phenology of a plant typically describes its life-cycle throughout a year, e.g., at mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
-In the mOSaic scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
+In the mOSaic scheme, phenology is parameterized with respect to the start of greening season (SGS) and its end (EGS).
 Details about our treatment of these are given in Section~\ref{subsubsec:greening}.
@@
   \begin{cases}
+    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding tropics)} \\
     \emph{if} \quad \text{GDAY = 0}  & \quad 0 \\
-    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding tropics)} \\
     \emph{else}  & \quad 
@@
 Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$.
-Since the mOSaic scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
+Since the mOSaic scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics.
+We therefore decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
 %
@@
 The integral of PAR over these wavelengths is the photosynthetic photon flux density (PPFD).
-In the EMEP parameterization, this integrated photon flux is used to modify the stomatal conductance in response to light:
+The correction factor $f_\mathrm{light}$ in response to varying PPFD is:
 \begin{equation}
@@
 
-Since $g_\text{max}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,m^{-2}}, a unit conversion is done:
-\begin{equation}
-  g^\text{m}_{\text{sto}}(N) = g_{\text{sto}}(N) \cdot R \cdot \frac{T_0}{P_0}.
-\end{equation}
-Herein, $R$ is the universal gas constant. With these, the total gridbox stomatal conductance $G_\text{sto}$ is computed weighted by the land surface type fraction $f_L(N)$:
-\begin{equation}
-  g_\text{sto} = \sum_{N=0}^{N_\text{max}} f_L(N)\cdot g^\text{m}_{\text{sto}}(N)
-\end{equation}
-
+Since $g_\text{max, m}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,m^{-2}}, a unit conversion to \unit{m\,s^{-1}} is necessary in our model:
+\begin{equation}
+  g^k_{\text{sto}} = g_{\text{sto, m}}^k \cdot R \cdot \frac{T_0}{P_0}.
+\end{equation}
+Herein, $R$ is the universal gas constant.\\
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For all other species, an interpolation between \chem{O_3} and \chem{SO_2} values is carried out.
-
-The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
-\begin{equation}
-  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
+The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For land surface types $k$, we can write
+\begin{equation}
+  G_\text{ns}^{\chem{O_3}, k} = \frac{\text{SAI}_k}{r_\text{ext}} + \frac{1}{R_\text{inc}^k + R_\text{gs}^{\chem{O_3},k}}.
   \label{eq:non-stomata-cond}
 \end{equation}
-$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
+$\text{SAI}_k$ is the surface area index for vegetation type $k$, which is $\text{LAI}$ plus some value that represents cuticles and others surfaces. The external leaf resistance is defined by
 \begin{equation}
@@
 \begin{equation}
-  F_T = \exp{(-0.2\cdot(1+\theta_0))}.
-\end{equation}
-$\theta_0$ is the $2\,\unit{m}$ temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
+  F_T = \exp{(-0.2\cdot(1+\theta_\text{2m}))}.
+\end{equation}
+$\theta_\text{2m}$ is the $2\,\unit{m}$ temperature in \unit{^\circ C}. For most land surface types, $\text{SAI} \equiv \text{LAI}$. Some exceptions are: 
 %
@@
 %
-Extending the mOSaic scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
+Extending the mOSaic scheme to the southern hemisphere, we use the growing season for crops defined in Table~\ref{tab:growing_season}.
 %
@@
 In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata.
-The in-canopy resistance $R_\text{inc}$ \citep{AE:Erisman1994} is then modified with respect to each (vegetated) land type $N$
-\begin{equation}
-  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
-\end{equation}
-where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{m^{-1}}$ is an empirical constant.
-According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
-$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
-\begin{equation}
-  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
-\end{equation}
+The in-canopy resistance $R_\text{inc}$ \citep{AE:Erisman1994} is then modified with respect to each (vegetated) land surface type in $k$
+\begin{equation}
+  R_\text{inc} = b \cdot \text{SAI}_k \cdot \frac{h_k(\text{lat})}{u_*},
+\end{equation}
+where $h_k(\text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{m^{-1}}$ is an empirical constant.
+The canopy resistance described in \citet{ACP:Simpson2012} does not take temperature and snow into account and is zero for non-vegetated surfaces, but we will adopt the correction previously used in the Oslo~CTM3 Wesely scheme.\\
 As initially mentioned, the necessary depth of snow to cover a certain type of vegetation differs.
 Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.
-
-The total conductance $G_\text{ns}^\chem{O_3}$ in each gridbox is the sum of all $G_\text{ns}^\chem{O_3}(N)$ weighted by the land surface type fraction $f_L(N)$
-\begin{equation}
-  G_\text{ns}^\chem{O_3} = \sum_{N=0}^{N_\text{max}}{G_\text{ns}^\chem{O_3}(N) \cdot f_L(N)}.
-\end{equation}
-
+$R_\text{gs}^{\chem{O_3},k}$ is tabulated and we correct for temperature by $F_T$ and for snow cover fraction:
+\begin{equation}
+  \frac{1}{R_\text{gs}^{\chem{O_3},k}} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^{\chem{O_3},k}} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
+\end{equation}
+
+The bulk canopy conductance is then defined as:
+\begin{equation}
+  G_c^k = \text{LAI} \cdot g^k_\text{sto, m} + G^k_\text{ns}, 
+\end{equation}
+wherein $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR (zero for non-vegetated surfaces), $g_\text{sto}$ the leaf-level stomatal conductance, and $G_\text{ns}$ the bulk non-stomatal conductance.\\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Latitude dependent vegetation height}
-The vegetation height $h(N, \text{lat})$ as decribed by \citet{ACP:Simpson2012} is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
+The vegetation height $h_k(\text{lat})$ as described by \citet{ACP:Simpson2012} is linearly decreasing with latitude between $60\,\unit{^\circ}$ and $74\,\unit{^\circ N}$.
 To adapt this to a global model, we made a few additional assumptions.
-The tabulated height for each vegetation type $h(N)$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
+The tabulated height for each vegetation type $h_k$ in the mOSaic scheme is regarded as constant at mid latitudes ($40\,\unit{^\circ}-60\,\unit{^\circ}$).
 Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}.
-At a latitude of $74\,\unit{^\circ}$ a minimum height of $3/10 \cdot h(N)$ is reached and kept constant.
-Towards the equator, we increase the height linearly so that at a latitude of $10\,\unit{^\circ}$ a maximum height of $2 \cdot h(N)$ is achieved which is then held constant.
+At a latitude of $74\,\unit{^\circ}$ a minimum height of $3/10 \cdot h_k$ is reached and kept constant.
+Towards the equator, we increase the height linearly so that at a latitude of $10\,\unit{^\circ}$ a maximum height of $2 \cdot h_k$ is reached which is then held constant.
 We also assume symmetry in both hemispheres.
 Presuming a typical tree height of $20\,\unit{m}$ at mid latitudes, this step-wise function yields a height of $8\,\unit{m}$ at high latitudes and $40\,\unit{m}$ in the tropics which is not unrealistic.
-For four example PFTs, results are shown in the Supplement (S.2, Fig.~S4).
+For four example PFTs, results are shown in the Supplement (S.3, Fig.~S4).
 
 \subsubsection{Mapping of land surface types}
-The Oslo~CTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the nine land surface types used in the mOSaic scheme (Fig.~\ref{fig:pft_mapping}).
-For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+The Oslo~CTM3 is configured to read land surface types from, either \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, which have to be mapped to the land surface types used in the mOSaic scheme (Fig.~\ref{fig:pft_mapping}).
+For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while $f_L^\text{water}$ is defined as $1-\sum_{k} f_L^k$.
 From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.
@@
   \caption{Mapping of land surface categories.
-Either land surface categories from ISLSCP2 product of MODIS or the Community Land Model (CLM)~2 can be chosen for mapping to the nine land surface types usedwe use in the mOSaic scheme.
+Either land surface categories from ISLSCP2 product of MODIS or the Community Land Model (CLM)~2 can be chosen for mapping to the land surface types we use in the mOSaic scheme.
 Water bodies and snow and ice categories of MODIS are actually not mapped.
-For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$.
+For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{k} f_L^k$.
 From MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}
@@
 In all other cases and where the $5\,\unit{^\circ C}$-days criteria fails, we still use the SMOKE-BEIS parameterization.
-The described algorithm written in python~2.7 has been included as Supplement~S.3.
+The described algorithm written in python~2.7 has been included as Supplement~S.4.
 An example map of the computed GLEN using the $5\,\unit{^\circ C}$-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
@@
 From OpenIFS an accumulated surface PAR is available which is already integrated presumably over the spectrum $400-700\,\unit{nm}$.
-For this reason, it will be refered to as PPFD instead of PAR.
+For this reason, it will be referred to as PPFD instead of PAR.
 This PPFD field has not been de-accumulated in the post-processing of the OpenIFS output and it is not feasible to redo all of the post-processing for this work.
@@
 \end{equation}
-For de-accumulation of the remaining timestep, the best choice is subtracting the difference between $21\,\unit{UTC}$ and $12\,\unit{UTC}$ of the previous day
+For de-accumulation of the remaining time step, the best choice is subtracting the difference between $21\,\unit{UTC}$ and $12\,\unit{UTC}$ of the previous day
 %
@@
 %
-and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.4 (Figs.~S5--S7).
+and limit the result to positive values only. An example PAR de-accumulation for January 2nd 2005 is shown in Supplement~S.5 (Figs.~S5--S7).
 The resulting PPFD fields are still accumulated over a time period of $3$ hours and should be divided by $3$.
@@
 Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}, while emissions from soil and wetlands are computed by MEGAN.
-Resultant $\mathrm{NO_x}$ emissions are upscaled to match Global Emissions InitiAtive (GEIA) inventory.
+Resultant $\mathrm{NO_x}$ emissions are up-scaled to match Global Emissions InitiAtive (GEIA) inventory.
 For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}.
@@
 That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are similar (Supplement~S.5: Fig.~S8).
+Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are similar (Supplement~S.6: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
@@
 
-\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the Oslo~CTM3, conducted the simulations, and analysed and evaluated the results.
+\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the Oslo~CTM3, conducted the simulations, and analyzed and evaluated the results.
 Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the respective documentation.

### 2019-06-26 DryDep_supplement.tex
First version of revised model description.

@@
 \section*{Supplement}
-%\subsection*{S.1 Aerodynamical resistance}
+\subsection*{S.1 Aerodynamical resistance}
+\begin{equation}
+  \Psi_m(\zeta) =
+  \begin{cases}
+   \log(\frac{1+x^2}{2}\cdot(\frac{1+x}{2})^2) - 2\cdot \arctan(x+\frac{\pi}{2}) & \quad \emph{if} \quad -2 < \zeta < 0,\\
+   - \beta \cdot \zeta & \quad \emph{if} \quad \zeta \ge 0.
+  \end{cases}
+  \label{eq:psim}
+\end{equation}
+\begin{equation}
+  \Psi_h(\zeta) =
+  \begin{cases}
+   2\cdot\log(\frac{1+x^2}{2}) & \quad \emph{if} \quad -2 < \zeta < 0,\\
+   - \beta \cdot \zeta & \quad \emph{if} \quad \zeta \ge 0.
+  \end{cases}
+  \label{eq:psim}
+\end{equation}
+$x = (1-\gamma\cdot\zeta)^\frac{1}{4}$, $\beta = 4.7$, $\gamma = 15$ (from Cansas experiment).
 %\begin{table}[!htbp]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
+\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
 \begin{figure}[!htbp]
@@
 %  \end{tabular}
-%  \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean scrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
+%  \belowtable{$^*$: CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean shrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
 %\end{table*}
@@
 %\clearpage
-\subsection*{S.2 Latitude dependent vegetation height}
+\subsection*{S.3 Latitude dependent vegetation height}
 \begin{figure}[!htbp]
@@
   \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global description.
-Example heights are shown for coniferous, deciduous, needleleaf, and broadleaf forests as well as Mediterranean shrub.
+Example heights are shown for coniferous (CF), deciduous (DF), needleleaf (NF), and broadleaf (BF) forests as well as Mediterranean shrub (MS).
 As reference the original range of EMEP is added.}
@@
 
-\subsection*{S.3 Temperature dependent greening season (python 2.7)}
+\subsection*{S.4 Temperature dependent greening season (python 2.7)}
 {\scriptsize
@@
 
-\subsection*{S.4 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
+\subsection*{S.5 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
 \begin{figure}[!htbp]
@@
 
-\subsection*{S.5 Normalized zonal average ozone dry deposition velocity}
+\subsection*{S.6 Normalized zonal average ozone dry deposition velocity}
 \begin{figure}[!htbp]

### 2019-07-02 DryDep.tex
Updated plots and started to update tables.

@@
 \label{sec:model_des}
-The Oslo~CTM3 is a three dimensional global chemistry transport model (CTM).
+The Oslo~CTM3 is an offline, three dimensional, global chemistry transport model (CTM).
 The key components of the Oslo~CTM3 have been described and evaluated by \citet{GMD:Sovde2012}.
@@
 Herein, $H$ is the sensible heat flux, $g$ is the standard gravitational acceleration, $c_p$ the specific heat capacity, and $T_{2m}$ the $2\,\unit{m}$ temperature.
-With these, we get \citep[Eq.~(8.8),][]{WASP:Simpson2003}
+With these, we can compute the aerodynamical resistance for each land surface type \citep[Eq.~(8.8),][]{WASP:Simpson2003}
 \begin{equation}
@@
 If GLEN is zero we are, e.g., in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$.
-Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$.
-Since the mOSaic scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics.
+Before the start of the greening season ($\text{GDAY}=0$) $f_\text{phen}=0$.
+Since this phenology is tuned to northern hemisphere (NH) mid latitudes, it does not apply to the tropics.
 We therefore decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
@@
 Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to $10\,\unit{\%}$ of the vegetation height.
-$R_\text{gs}^{\chem{O_3},k}$ is tabulated and we correct for temperature by $F_T$ and for snow cover fraction:
-\begin{equation}
-  \frac{1}{R_\text{gs}^{\chem{O_3},k}} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^{\chem{O_3},k}} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
+$R_\text{gs}^{\chem{O_3},k}$ is tabulated. We correct for temperature by $F_T$ and for snow cover fraction:
+\begin{equation}
+  \frac{1}{R_\text{gs}^{\chem{O_3},k}} = \frac{1-f_\text{snow}^k}{\hat{R}_\text{gs}^{\chem{O_3},k}} + \frac{f_\text{snow}^k}{R_\text{snow}^{\chem{O_3},k}}.
 \end{equation}
@@
 \end{equation}
-wherein $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR (zero for non-vegetated surfaces), $g_\text{sto}$ the leaf-level stomatal conductance, and $G_\text{ns}$ the bulk non-stomatal conductance.\\
+wherein $\text{LAI}$ is the one-sided leaf area index taken from ISLSCP2 FASIR, $g_\text{sto}$ the leaf-level stomatal conductance, and $G_\text{ns}$ the bulk non-stomatal conductance.\\
 
@@
     \tophline
-    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Emissions\\
-    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
+    \multirow{2}{*}{Simulation} & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $R^\chem{O_3}$ & Emissions\\
+    & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & & (year)\\
     \middlehline
-    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
-    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
-    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
-    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
-    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
-    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
-    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
-    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
-    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
+    Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
+    mOSaic            &   x & x & x                &   &  x & x  & 2005 \\
+    mOSaic\_offLight  &   x & 1 & x                &   &  x & x  & 2005 \\
+    mOSaic\_offPhen   &   1 & x & x                &   &  x & x  & 2005 \\
+    mOSaic\_SWVL1     &   x & x & SWLV1            &   &  x & x  & 2005 \\
+    mOSaic\_ice       &   x & x & x                &   &  x & x; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
+    mOSaic\_desert    &   x & x & x                &   &  x & x; \citet{AE:Gusten1995}$^*$  & 2005 \\
+    mOSaic\_emis2014  &   x & x & x                &   &  x & x  & 2014 \\
+    mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
     \bottomhline
@@
   \label{tab:simsum}
-  %\belowtable{} % Table Footnotes
+  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$Supplement~S.7} % Table Footnotes
 \end{table*}

### 2019-07-02 DryDep_supplement.tex
Updated plots and started to update tables.

@@
   \end{cases}
-  \label{eq:psim}
+  \label{eq:psih}
 \end{equation}
-$x = (1-\gamma\cdot\zeta)^\frac{1}{4}$, $\beta = 4.7$, $\gamma = 15$ (from Cansas experiment).
+$x = (1-\gamma\cdot\zeta)^\frac{1}{4}$, $\beta = 4.7$, $\gamma = 15$ (from Kansas experiment).
 %\begin{table}[!htbp]
@@
 
+\subsection*{S.7 Wesely-Hough $R^\chem{O_3}$}
+\begin{table*}[!htbp]
+  \caption{Surface resistance of ozone: $^*$ CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean shrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
+  \begin{tabular}{lc}
+    \tophline
+    Code$^*$ & $R_\chem{O_3}$\\
+         & (\unit{s\,m^{-1}})\\
+    \middlehline
+    CF &  1000\\
+    DF &  1000 \\
+    NF &  1000 \\
+    BF &  1000 \\
+    TC &  200 \\
+    SNL & 400 \\
+    GR  & 1000 \\
+    MS  & 1000 \\
+    WE  & 400 \\
+    TU  & 385 \\
+    DE  & 385 \\
+    W   & 1430 \\
+    U   & 400 \\
+    ICE & 1430\\
+    \bottomhline
+ \end{tabular}
+  %\belowtable{}
+\end{table*}
 \end{document}

### 2019-07-03 DryDep.tex
Updated tables.

@@
 A comprehensive evaluation of different techniques is given by \citet{GCB:Anav2017}.
+Another solution would be the usage of a proper landsurface model, e.g.
+LPJ-GUESS, CLM, but the integration of such into the Oslo~CTM3 is not planed at the moment.
 
@@
 To counter this, we decided to refrain from the division at this stage, but need to bear this in mind for later OpenIFS cycles.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%      
@@
 In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the mOSaic scheme.
-We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively.
-We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}).
+We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, ozone concentrations in the lowermost model level $[\chem{O_3}](p_0)$, and tropospheric burden $\sum_\text{trop}\chem{O_3}$.
+We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and observations (Section~\ref{subsec:obs}).
 The Oslo~CTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}.
@@
     Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
-    mOSaic            &   x & x & x                &   &  x & x  & 2005 \\
-    mOSaic\_offLight  &   x & 1 & x                &   &  x & x  & 2005 \\
-    mOSaic\_offPhen   &   1 & x & x                &   &  x & x  & 2005 \\
-    mOSaic\_SWVL1     &   x & x & SWLV1            &   &  x & x  & 2005 \\
-    mOSaic\_ice       &   x & x & x                &   &  x & x; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
-    mOSaic\_desert    &   x & x & x                &   &  x & x; \citet{AE:Gusten1995}$^*$  & 2005 \\
-    mOSaic\_emis2014  &   x & x & x                &   &  x & x  & 2014 \\
+    mOSaic            &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
+    mOSaic\_offLight  &   x & 1 & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
+    mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
+    mOSaic\_SWVL1     &   x & x & LV1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
+    mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
+    mOSaic\_desert    &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{AE:Gusten1995}$^*$  & 2005 \\
+    mOSaic\_emis2014  &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2014 \\
     mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
@@
   \label{tab:simsum}
-  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$Supplement~S.7} % Table Footnotes
+  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.7} % Table Footnotes
 \end{table*}
@@
   \includegraphics[width=8.3cm]{fig04}
-  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between the sensitivity simulation \emph{mOSaic} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
   \label{fig:diff_maps}
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition.
+In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition.
 Surface ozone increases globally except for regions covered by tropical forest.
@@
     \middlehline
-    Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
-    EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
-    EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
-    EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
-    EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
-    EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
-    EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
-    EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
-    EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
+    Wesely\_type & 160.5 & 147.7 & 308.2 & 7.0 & 6.4 & 13.4 & 417.3 & 190.8 & 608.2 & 613.4 & 344.9 & 958.3 & 36.7\\
+    mOSaic & 108.1 & 105.3 & 213.4 & 4.3 & 3.1 & 7.4 & 236.2 & 130.3 & 366.4 & 368.3 & 238.6 & 606.9 & 0.0\\
+    mOSaic\_offLight & 110.5 & 106.3 & 216.8 & 4.3 & 3.0 & 7.4 & 263.1 & 145.3 & 408.4 & 399.9 & 254.7 & 654.6 & 7.3\\
+    mOSaic\_offPhen & 108.3 & 105.3 & 213.6 & 4.3 & 3.1 & 7.4 & 246.9 & 132.2 & 379.1 & 379.5 & 240.5 & 620.0 & 2.1\\
+    mOSaic\_SWVL1 & 108.1 & 105.2 & 213.3 & 4.3 & 3.1 & 7.4 & 234.6 & 128.8 & 363.4 & 366.6 & 237.1 & 603.7 & -0.5\\
+    mOSaic\_ice & 105.6 & 103.0 & 208.6 & 2.6 & 1.0 & 3.6 & 232.7 & 130.2 & 362.9 & 358.3 & 234.2 & 592.6 & -2.4\\
+    mOSaic\_desert & 109.1 & 105.5 & 214.6 & 4.3 & 3.1 & 7.4 & 250.4 & 132.6 & 383.0 & 383.4 & 241.2 & 624.6 & 2.8\\
+    mOSaic\_emis2014 & 108.9 & 107.0 & 215.9 & 4.3 & 3.1 & 7.3 & 238.3 & 133.8 & 372.1 & 370.8 & 243.9 & 614.7 & 1.3\\
+    mOSaic\_hough & 133.3 & 131.1 & 264.4 & 4.9 & 3.6 & 8.4 & 265.8 & 132.0 & 397.8 & 423.6 & 266.7 & 690.3 & 12.1\\
     \bottomhline
   \end{tabular}
-  \belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
+  \belowtable{$^\dagger$: Relative change of global annual total in comparison to \emph{mOSaic}.}% Table Footnotes
   \label{tab:ozone_sinks}
@@
     \middlehline
-    Wesely type & 361 & $\pm$ & 21\\
-    EMEP\_full & 392 & $\pm$ & 28\\
-    EMEP\_offLight & 388 & $\pm$ & 26\\
-    EMEP\_offPhen & 392 & $\pm$ & 27\\
-    EMEP\_SWVL4 & 402 & $\pm$ & 31\\
-    EMEP\_ppgs & 392 & $\pm$ & 28\\
-    EMEP\_ppgssh & 391 & $\pm$ & 27\\
-    EMEP\_ppgssh\_ice & 403 & $\pm$ & 31\\
-    EMEP\_ppgs\_2005 & 386 & $\pm$ & 26\\
+    Wesely\_type & 364 & $\pm$ & 23 \\
+    mOSaic & 399 & $\pm$ & 31 \\
+    mOSaic\_offLight & 395 & $\pm$ & 30 \\
+    mOSaic\_offPhen & 398 & $\pm$ & 31 \\
+    mOSaic\_SWVL1 & 399 & $\pm$ & 31 \\
+    mOSaic\_ice & 401 & $\pm$ & 31 \\
+    mOSaic\_desert & 398 & $\pm$ & 31 \\
+    mOSaic\_emis2014 & 405 & $\pm$ & 32 \\
+    mOSaic\_hough & 393 & $\pm$ & 30 \\
     \bottomhline
@@
 \end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparison with MACC-reanalysis}
+\label{subsec:macc}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-07-03 DryDep_supplement.tex
Updated tables.

--- 2019-06-26 DryDep_supplement.tex
@@
 
+\subsection*{S.7 Surface resistance of ozone}
+\begin{table*}[!htbp]
+  \caption{Surface resistance of ozone $R^\chem{O_3}$ adapted from  Wesely (1989); Hough (1991): $^*$ CF -- temperate/boreal coniferous; DF -- temperate/boreal deciduous; NF -- Mediterranean needleleaf; BF -- Mediterranean broadleaf; TC -- temperate crop; MC -- Mediterranean crop; RC -- root crop; SNL -- moorland; GR -- grass; MS -- Mediterranean shrub; WE -- wetlands; TU -- tundra; DE -- desert; W -- water; ICE -- ice.}
+  \begin{tabular}{lc}
+    \tophline
+    Code$^*$ & $R_\chem{O_3}$\\
+         & (\unit{s\,m^{-1}})\\
+    \middlehline
+    CF, DF, NF, BF &  1000\\
+    TC &  200 \\
+    SNL, WE, U & 400 \\
+    GR, MS  & 1000 \\
+    TU, DE  & 385 \\
+    W, ICE  & 1430 \\
+    \bottomhline
+ \end{tabular}
+  %\belowtable{}
+\end{table*}
 \end{document}

### 2019-07-04 DryDep.tex
Finished revision of Sec 3.1.

@@
 Since the parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed.
-First, we have used a parameterization as implemented in the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).
+First, we used a parameterization which was already implemented in the Oslo~CTM3 and which had been adopted from the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).
 SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses lapse rates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
-As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.\\
+As this parameterization is optimized for North America, it does not work well in Europe, e.g., most of northern Scandinavia has no allocated vegetation period.
+This basically results in a suppression of canopy resistance in northern Scandinavia.\\
 
@@
 \label{subsec:sens}
-Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every difference in results to a respective change.
-Therefore, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the mOSaic scheme.
-A reference simulation featuring the Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every single difference in results to a respective change.
+Therefore, we conducted one reference simulation denoted as \emph{mOSaic} implementing all parameters as tabulated in \citet{ACP:Simpson2012} and in total nine sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}).
+A reference simulation featuring the Oslo~CTM3 Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
+All model experiments discussed in the following are summarized in Table~\ref{tab:simsum}.
+An \emph{x} therein denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.
 For all model integrations, the meteorological reference year is 2005.
-This choice affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years.
-After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} is the baseline simulation for the sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}.
-As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively.
-During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions.
-Especially in Europe, this basically results in a suppression of canopy resistance in northern Scandinavia.
-The following sensitivity study therefore comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details): Mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}).
-Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the Oslo~CTM3 (\emph{EMEP\_ppgssh\_ice}).
-Accidentally, we have used emissions for the year 2014 instead of 2005 in all model experiments except for \emph{EMEP\_ppgs\_2005}.
-{\bf TODO: Actually assess the impact of the differing emissions!} As mentioned previously, the simulations were initially conducted with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
-A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).
-All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
-An x in the table denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.
+This choice affects the direct comparison with data and studies that either show results based on decadal averages or differing years.
+
+First, we have a closer look at the influence of certain parameters on the stomatal conductance.
+As indicated by the names, \emph{mOSaic\_offLight} and \emph{mOSaic\_offPhen} are rather extreme scenarios completely switching off the sensitivity to light and phenology in Eq.~(\ref{eq:stomatal}) by setting $f_\text{light}$ and $f_\text{phen}$ to a fixed value of 1, respectively.
+Because of the underlying research project that focuses on arctic and alpine ecosystems, where water might only be available from upper soil layers, an experiment was conducted using the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+After this, we want to confirm the importance of choice of $R^\chem{O_3}$ for different land surface types.
+We conducted three experiments looking at a $R^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics (\emph{mOSaic\_ice}), observed $R^\chem{O_3}_\text{desert}$ \citep{AE:Gusten1995} (\emph{mOSaic\_desert}), and the $R^\chem{O_3}$ originally used in \emph{Wesely\_type} \citep{AE:Wesely1989, JGR:Hough1991}.
+Finally, we run a simulation with emissions for the year 2014 instead of 2005 (\emph{EMEP\_emis2014}) to characterize the general influence of differing emissions on ozone.
 
@@
     mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
-    mOSaic\_SWVL1     &   x & x & LV1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
+    mOSaic\_SWVL1     &   x & x & VL1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
     mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
@@
   \includegraphics[width=8.3cm]{fig04}
-  \caption{Relative difference between the sensitivity simulation \emph{mOSaic} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
   \label{fig:diff_maps}
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition.
-Surface ozone increases globally except for regions covered by tropical forest.
-Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$.
+In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition.
+The ozone burden increases globally except for regions covered by tropical forest.
+Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$.
 Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest.
@@
 We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible.
-A possible explanation to this divergence especially in desert regions is the difference between the prescribed dry deposition velocities in the Wesely scheme in comparison to those used in the mOSaic scheme.
+A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}.
 We cover this in more detail in the following section.
@@
 
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig08}
+  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{mOSaic} (lowermost model level); (c) Relative difference.}
+  \label{fig:macc_o3conc}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig08}
+  \includegraphics[width=8.3cm]{fig09}
   \caption{Ozone dry deposition fluxes at different observation sites.
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=0.8\textwidth]{fig09}
+  \includegraphics[width=0.8\textwidth]{fig10}
   \caption{Comparison of the manifold of Oslo~CTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition.
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{fig10}
+  \includegraphics[width=0.8\textwidth]{fig11}
   \caption{Partitioning of land surface types. (a) CLM~2 dynamic land surface types in $(0.5\times0.5)\,\unit{^\circ}$ resolution; (b) Zonal distribution of land surface types.}

### 2019-07-04 DryDep.tex
Finished sec 3.2.1 and updated one plot.

@@
 %
-In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}.
+In the evaluation of our model, we closely follow suggestions by \citet{ACP:Hardacre2015}.
 For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
-In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}).
+In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](p_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types (Section~\ref{subsubsec:seasons}).
 From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
 
-As dry deposition velocities are not directly available from the model output, monthly average dry deposition velocities $v^\chem{O_3}_\text{DD}$ have been reconstructed from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged surface ozone amount $\chem{O_3}(z_0)$
+Dry deposition velocities are directly available only for the new model version.
+For \emph{Wesely\_type}, monthly average dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged surface ozone amount $\chem{O_3}(z_0)$
 \begin{equation}
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by up to $10\,\unit{ppb}$ when applying the mOSaic scheme compared to the Wesely scheme.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $14-15\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
-In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$).
+In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$).
 We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
 Among the sensitivity studies, there is only a low absolute variance.
-Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
-Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly.
-Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics).
-Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $(54-70)\,\unit{^\circ N}$.
-The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
-An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.
+Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly.
+Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}), amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$).
+Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
+Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH.
+The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.7).
+The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
+An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, where an increase in ozone precursors is seen in CEDS.
 
 The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b.
-The dry deposition velocities in the mOSaic scheme are on average $43\,\unit{\%}$ lower than in the Wesely scheme.
-In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the mOSaic scheme are constantly above the multi-model-mean.
-This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
-This may, however, lead to an overcompensation of the current low-bias in surface ozone in the Oslo~CTM3.
-The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$.
-The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}).
-That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation.
-Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are similar (Supplement~S.6: Fig.~S8).
+The dry deposition velocities in the mOSaic scheme are well below the Wesely scheme and in remarkable agreement with the results shown by \citet{ACP:Hardacre2015}.
+In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean.
+This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
+This may, however, lead to an overcompensation of the current low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
+The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$ ($\mathcal{O}$), $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
+The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are more similar than the \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
-However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.
+However, it is not clear whether diverging dry deposition velocities over ocean or deserts are responsible for this.
 
 The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c.
-In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one-half of the amount given by \emph{Wesely\_type}.
-The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a change in $v_\text{desert}^\chem{O_3}$.
+In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies.
+In the most extreme case (NH subtropics and mid latitudes), the total ozone dry deposition drops to one-half of the amount given by \emph{Wesely\_type}.
+The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a substantial difference in $v_\text{desert}^\chem{O_3}$.
 Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original mOSaic scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
-These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the mOSaic scheme.
+These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference.
 We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).
 
-There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the reconstructed $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better.
+There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better.
 This could be a sign of differences in photo-chemistry and transport (e.g.
-convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}(z_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
+convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}(p_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
+Another hint may be found in the routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this subject to further investigations.
 
@@
 Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
+As for the global annual comparisons above, the mOSaic scheme matches the multi-model-mean values remarkably well with respect to dry deposition velocities, while it strongly underestimates the total dry deposition.
 There are two major phases apparent: NH and SH greening season.
@@
 Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
-In comparison, the available $2$ months of the multi-model-mean display a somewhat different pattern.
-Especially the NH/SH asymmetry is much more pronounced in February, than it is the mOSaic scheme.
 %

### 2019-07-05 DryDep.tex
Edited section, added and updated plots using the actual drydep velos from output.

@@
 As indicated by the names, \emph{mOSaic\_offLight} and \emph{mOSaic\_offPhen} are rather extreme scenarios completely switching off the sensitivity to light and phenology in Eq.~(\ref{eq:stomatal}) by setting $f_\text{light}$ and $f_\text{phen}$ to a fixed value of 1, respectively.
-Because of the underlying research project that focuses on arctic and alpine ecosystems, where water might only be available from upper soil layers, an experiment was conducted using the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+Because of the underlying research project's focus on arctic and alpine ecosystems, where water might only be available from upper soil layers, an experiment was conducted using the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
 After this, we want to confirm the importance of choice of $R^\chem{O_3}$ for different land surface types.
-We conducted three experiments looking at a $R^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics (\emph{mOSaic\_ice}), observed $R^\chem{O_3}_\text{desert}$ \citep{AE:Gusten1995} (\emph{mOSaic\_desert}), and the $R^\chem{O_3}$ originally used in \emph{Wesely\_type} \citep{AE:Wesely1989, JGR:Hough1991}.
+We conducted three experiments looking at a $R^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} (\emph{mOSaic\_ice}), observed $R^\chem{O_3}_\text{desert}$ \citep{AE:Gusten1995} (\emph{mOSaic\_desert}), and an approximation of $R^\chem{O_3}$ originally used in \emph{Wesely\_type} \citep{AE:Wesely1989, JGR:Hough1991}.
 Finally, we run a simulation with emissions for the year 2014 instead of 2005 (\emph{EMEP\_emis2014}) to characterize the general influence of differing emissions on ozone.
@@
 In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition.
-The ozone burden increases globally except for regions covered by tropical forest.
+The ozone burden increases globally except for some regions covered by tropical forest.
 Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$.
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $14-15\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
@@
 We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
-Among the sensitivity studies, there is only a low absolute variance.
+Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance.
 Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly.
-Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}), amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$).
+Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}) amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$).
 Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
@@
 This may, however, lead to an overcompensation of the current low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
-The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$ ($\mathcal{O}$), $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
+The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
 The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are more similar than the \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
@@
 Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
-Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original mOSaic scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$).
-These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference.
+Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameterset \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}.
+These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}).
 We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).
@@
 convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}(p_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
-Another hint may be found in the routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this subject to further investigations.
+Another reason might be found within the (chemistry) routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this is subject to further investigations.
 
@@
 As for the global annual comparisons above, the mOSaic scheme matches the multi-model-mean values remarkably well with respect to dry deposition velocities, while it strongly underestimates the total dry deposition.
-There are two major phases apparent: NH and SH greening season.
+Qualitatively, there are two major phases apparent: NH and SH greening season.
 Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer).

### 2019-07-05 DryDep.tex
Finished first re-vision of sec. 3. Added subsection with comparison to MACC-reanalysis. But this might need to be revised.

@@
 A comprehensive evaluation of different techniques is given by \citet{GCB:Anav2017}.
-Another solution would be the usage of a proper landsurface model, e.g.
+Another solution would be the usage of a proper land surface model, e.g.
 LPJ-GUESS, CLM, but the integration of such into the Oslo~CTM3 is not planed at the moment.
@@
   v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month}.
+  \label{eq:retro_vdd}
 \end{equation}
@@
 Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively.
-Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameterset \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}.
+Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameter set \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}.
 These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}).
@@
   \includegraphics[width=8.3cm]{fig05}
-  \caption{Comparison of the manifold of Oslo~CTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
+  \caption{Comparison of the manifold of Oslo~CTM3 integrations with respect to (a) Ozone concentrations in the lowermost model level, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition.
 The different colors indicate sets of simulation with similar baselines.
@@
 %
-To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.
+To further disentangle the contributions of different regions to the global ozone budget, we will look at different projections of seasonal cycles.
 
@@
 We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference.
+While the total ozone dry deposition of \emph{Wesely\_type} agrees well with the multi-model-mean in any zonal band, the mOSaic scheme displays a much smaller total ozone dry deposition.
+This deviation appears to be almost the same for each zonal band ($6-7\,\unit{\%}$).
+
 As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface).
@@
 In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum.
-The seasonal cycle in the mOSaic scheme appears to be shifted towards later in the year in the NH.
-The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realizations of the mOSaic scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively.
-This indicates that surface ozone in the tropics may be very sensitive to changes in emissions of ozone precursors and changes in long-wave radiative transfer (e.g., due changes in cloud cover or aerosol).
-
-As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types.
-Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
-Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
-Since we have not performed a full unfolding on the data, the results should be treated with slight caution.
+Compared to the multi-model-mean, the seasonal cycle in the Oslo~CTM3 NH appears to be shifted towards later in the year.
+The seasonal cycle in the tropics and subtropics only differs by magnitude otherwise the shapes are identical for both the mOSaic scheme, the Wesely scheme, and the multi-model-mean.
+The total amount of dry deposition of ozone differs strongly between the different model experiments, with \emph{mOSaic\_SWVL1} and \emph{mOSaic\_hough} displaying the lowest and highest amount, respectively.
+This indicates that surface ozone is much more sensitive to the choice of parameters ($\mathcal{O}(5\,\unit{ppb})$ for \emph{mOSaic\_hough} in the tropics) than to slight changes in precursor emissions ($\mathcal{O}(1\,\unit{ppb})$ for \emph{mOSaic\_emis2014} in the tropics).
+
+As suggested by \citet{ACP:Hardacre2015}, we also look at ozone dry deposition velocities with respect to surface types separately.
+Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to calculate these.
+Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+Thus, it is not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
+Since we have not performed a full unfolding on the data in case of \emph{Wesely\_type}, the results should be treated with slight caution.
+In case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
 %
@@
 %
+
 In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories.
 The patterns and absolute numbers differ substantially between the Wesely scheme and the mOSaic scheme and the multi-model-mean.
-The divergence of the average dry deposition velocities between the Wesely scheme and the mOSaic scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$).
-In contrast to the Wesely scheme and the mOSaic scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$).
-The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
+The divergence of the average dry deposition velocities between \emph{Wesely\_type} and \emph{mOSaic} in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.20\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.65\,\unit{cm\,s^{-1}}$) is quite remarkable.
+The difference of \emph{mOSaic} and \emph{Wesely\_type} to the multi-model-mean in tropical forest regions is $\Delta_\text{mOSaic} \overline{v}_\text{tropical forest}^\chem{O_3} = 0.61\,\unit{cm\,s^{-1}}$ and $\Delta_\text{Wesely\_type} \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$, respectively.
+The multi-model-mean displays a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$), which cannot be reproduced with the mOSaic scheme.
+The dry deposition velocities over desert regions are consistent with the estimated values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
+The difference between \emph{Wesely\_type} and \emph{mOSaic\_hough} stems most likely from $R_a$ and $R_b$ contributions.
 From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
@@
 In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed).
-With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the mOSaic scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented.
-With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical or coniferous forest \citep{GCB:Anav2017}.
+With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the mOSaic scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean, but is in line with the multi-model-mean.
+This implies that ozone concentrations might even become larger and dry deposition even lower in the model if a more advanced dry deposition scheme to the ocean would be implemented.
+With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical forest \citep{GCB:Anav2017} or in boreal regions.
 %
@@
 Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative mOSaic scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to about $35\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative mOSaic scheme (\emph{mOSaic}), while permanently ice and snow covered regions account for $1.2\,\unit{\%}$.
 The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling.
@@
 This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
-{\bf TODO include reference to Table~\ref{tab:trop_ozone_burden} in the text.}
 %
@@
 \end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparison with MACC-reanalysis}
+\label{subsec:macc}
+In this section, we conclude the comparison of our results with modeled ozone concentrations by comparing with ECWMF's MACC-reanalysis \citep{MACC}.
+In Fig,~\ref{fig:macc_o3conc}, we contrast \emph{mOSaic} ozone concentrations in the lowermost model level with surface concentrations deduced from the MACC-reanalysis for the year 2005.
+The MACC-reanalysis displays low ozone concentrations above all land masses except for the Greenland ice sheet, while the highest values are found within $(25-60)\,\unit{^\circ N}$ over the oceans.
+These high values over the oceans are relatively well reproduced by the mOSaic scheme, however over land (e.g.
+central Africa, Tibetan plateau) the ozone concentrations are up to $200\,\unit{\%}$ higher.
+While on average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to central Africa and the Tibetan plateau (Supplement~S.8).
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig08}
+  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{mOSaic} (lowermost model level); (c) Relative difference.}
+  \label{fig:macc_o3conc}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparison with observations}
+\label{subsec:obs}
+In this section, we compare our results to observations.
+For all comparisons, we use the original resolution of the Oslo~CTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
+
+In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six selected observation sites are shown.
+We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ uncertainty band.
+The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual uncertainty band since such is not given in \citet{ACP:Hardacre2015}.
+At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and similar to or better than the multi-model-mean.
+We use a $\chi^2$-test
+\begin{equation}
+  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
+\end{equation}
+with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number of degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
+The closer to $1$ this test scores, the better does the simulation represent the observation.
+A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
+The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
+The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to observed ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
+Furthermore, the observational averages comprise at most $9$ years worth of data.
+Statistically, these data may still be subject to interannual variability.
+Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
+Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both high and low biases exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
+Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox and not substantially affected by differences in modeled and actual topography (e.g., major wind directions).
+
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig09}
+  \caption{Ozone dry deposition fluxes at different observation sites.
+(a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
+The model uncertainty of the Oslo~CTM3 is given as $1\,\sigma$ error band.
+The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
+(b) Model divergence from observation and $\chi^2$-test results.}
+  \label{fig:mmm_drydep_stations}
+\end{figure}
+
+Table~\ref{tab:trop_ozone_burden} displays the average tropospheric ozone burden for all model experiments.
+Consistently with the previous findings, the mOSaic scheme increases the tropospheric ozone burden by about $35\,\unit{Tg}$.
+From various satellite ozone retrieval products, \citet[][Tab.~5]{ESA:Gaudel2018} deduce a lower limit estimate for global tropospheric ozone burden for the years $2010-2014$ of $333-345\,\unit{Tg}$, but remark that this amount underestimates the actual tropospheric ozone burden, since it is only based on daytime retrievals.
+Nevertheless, the results of \emph{mOSaic} lie $17\,\unit{\%}$ above that estimate and also well above the typical modeling range of $302-378\,\unit{Tg}$ \citep{ACP:Young2013}.
+The difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), however, lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
+
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with MACC-reanalysis}
-\label{subsec:macc}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig08}
-  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{mOSaic} (lowermost model level); (c) Relative difference.}
-  \label{fig:macc_o3conc}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with observations}
-\label{subsec:obs}
-In this section, we compare our results to observations.
-For all comparisons, we use the original resolution of the Oslo~CTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
-
-In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown.
-We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band.
-The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}.
-At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean.
-We use a $\chi^2$-test
-\begin{equation}
-  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
-\end{equation}
-with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number of degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
-The closer to $1$ this test scores, the better does the simulation represent the observation.
-A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
-The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
-The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
-With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
-Furthermore, the observational averages comprise at most $9$ years worth of data.
-Statistically, these data may still subject to interannual variability.
-Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
-Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both high and low biases exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
-Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox.
-
-%
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig09}
-  \caption{Ozone dry deposition fluxes at different observation sites.
-(a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
-The model uncertainty of the Oslo~CTM3 is given as $1\,\sigma$ error band.
-The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
-(b) Model divergence from observation and $\chi^2$-test results.}
-  \label{fig:mmm_drydep_stations}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]

### 2019-07-05 DryDep_supplement.tex
Edited section, added and updated plots using the actual drydep velos from output.
Finished first re-vision of sec. 3. Added subsection with comparison to MACC-reanalysis. But this might need to be revised.

@@
   \centering
-  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
+  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo_v2}
   \caption{Normalized zonal average ozone dry deposition velocity.}
@@
 \end{table*}
+
+\subsection{S.8 Comparison with MACC-reanalysis}
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[height=0.5\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/annual_mean_surface_ozone_2005_Wesely_type}
+  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{Wesely\_type} (lowermost model level); (c) Relative difference.}
+\end{figure}
+
 \end{document}

### 2019-07-08 version_2/response_version_1_rc2.tex
Revised version, not spell-checked.

@@
   The referee is right that ozone is predominately deposited to terrestrial ecosystems, which cover about 2/9 of the Earth's surface, while 2/3 is covered by oceans.
-Nevertheless, due to the ocean's vastness, relatively small changes in modeled dry deposition easily accumulate and influence the atmospheric ozone concentration.
+Nevertheless, due to the ocean's vastness, relatively small changes in modeled dry deposition easily accumulate and influence the atmospheric ozone concentration as pointed out in Hardacre et al.
+(2015) based on Ganzeveld et al.
+(2009).
 Luhar et al.

### 2019-07-08 DryDep.tex
Revised version, not spell-checked.

@@
 We change from a purely empirical dry deposition parameterization to a more process-based one which is taking the state of the atmosphere and vegetation into account.
-Examining the sensitivity of the scheme to various parameters, our focus lies mainly on the stomatal conductance-based description of the canopy resistance.
-We evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies and also estimate the impact on the modeled ozone concentrations at the surface.
-We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-47\,\unit{\%}$), leading to an increase in surface ozone of up to $100\,\unit{\%}$ in some regions.
-While high sensitivity to changes in dry deposition to vegetation is found in the tropics, the largest impact on global scales is associated to changes in dry deposition to the ocean and deserts.
+We examine the sensitivity of the scheme to various parameters, e.g.
+the stomatal conductance-based description of the canopy resistance and the choice of ozone surface resistance, and evaluate the resulting modeled ozone dry deposition with respect to observations and multi-model studies.
+Individual dry deposition velocities are now available for each land surface type and agree generally well with observations.
+We also estimate the impact on the modeled ozone concentrations at the surface.
+We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-37\,\unit{\%}$), leading to an increase in surface ozone of more than $100\,\unit{\%}$ in some regions.
+While high sensitivity to changes in dry deposition to vegetation is found in the tropics and the northern hemisphere, the largest impact on global scales is associated to changes in dry deposition to the ocean and deserts.
 \end{abstract}
@@
 \label{subsubsec:Ra}
-In general, the aerodynamic resistance is describing the turbulent transport of any substance down to the surface.
+In general, the aerodynamic resistance describes the turbulent transport of any substance down to the surface.
 To derive $R_a^k$, we follow \citet{WASP:Simpson2003,ACP:Simpson2012} and compute a local friction velocity at reference height $z_\mathrm{ref}$ \citep[Eq.~(52),][]{ACP:Simpson2012}
@@
 with the average wind speed $\overline{u}(z_\mathrm{ref})$ at reference height, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and roughness length $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
-We have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
-Using the derived $u_*^k$ from Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ can by obtained with \citep[Eq.~(8),][]{ACP:Simpson2012}:
+Concidering the height of vegetation, we have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
+Using the derived $u_*^k$ from Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ can be obtained from \citep[Eq.~(8),][]{ACP:Simpson2012}:
 \begin{equation}
@@
 \end{equation}
-wherein $\text{Pr}$ is the Prandtl number (typically 0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$.
+wherein $\text{Pr}$ is the Prandtl number (0.72 for air and other gases) and $\text{Sc}_i$ is the Schmidt number for a gas $i$.
 Eq.~(\ref{eq:Rb}) differs from a similar formulation in \citet{ACP:Seinfeld2006} by a factor of roughly $1.25$.
@@
 All factors differ with land use type $k$.
-For clarity reasons, we drop this index in the following, as long is it not necessary for the equation.
+For clarity reasons, we drop this index in the following, as long as it is not necessary for the equation's completeness.
 
@@
 The parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various plant functional types.
-All parameters are taken from Supplementary \citet[][Tables S16, S19]{ACP:Simpson2012}.
-Since $f_T$ turns negative outside the range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$.
+All parameters are taken from \citet[][Tables S16, S19]{ACP:Simpson2012}.
+Since $f_T$ turns negative outside the range defined by $T_\text{min}$, $T_\text{max}$, we impose a lower limit of $0.01$ for numerical reasons.
 
@@
 For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}.
-Emissions of \chem{CH_4} from surface data from the EU project (EU GOCE 037048) Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA.
+Emissions of \chem{CH_4} are taken from the EU project (EU GOCE 037048) Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA.
 In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
@@
 Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every single difference in results to a respective change.
-Therefore, we conducted one reference simulation denoted as \emph{mOSaic} implementing all parameters as tabulated in \citet{ACP:Simpson2012} and in total nine sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}).
+Therefore, we conducted one reference simulation denoted as \emph{mOSaic} implementing all parameters as tabulated in \citet{ACP:Simpson2012} and in total seven sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}).
 A reference simulation featuring the Oslo~CTM3 Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
@@
 Dry deposition velocities are directly available only for the new model version.
-For \emph{Wesely\_type}, monthly average dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged surface ozone amount $\chem{O_3}(z_0)$
-\begin{equation}
-  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month}.
+For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged ozone amount in the lowermost modellevel $\chem{O_3}(p_0)$
+\begin{equation}
+  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
   \label{eq:retro_vdd}
@@
 Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
+In case of \emph{mOSaic}, Eq.~(\ref{eq:retro_vdd}) yields similar values for $v^\chem{O_3}_\text{DD}$ as directly available from model output.
 %
@@
 This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
-This may, however, lead to an overcompensation of the current low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
+This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
 The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
-The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme and the multi-model-mean are more similar than the \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
+The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
-However, it is not clear whether diverging dry deposition velocities over ocean or deserts are responsible for this.
 
@@
 This could be a sign of differences in photo-chemistry and transport (e.g.
-convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}(p_0)]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
+convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
 Another reason might be found within the (chemistry) routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this is subject to further investigations.
@@
 Thus, it is not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
-Since we have not performed a full unfolding on the data in case of \emph{Wesely\_type}, the results should be treated with slight caution.
+Since we have not performed a full unfolding on the data, the results should be treated with slight caution (e.g.
+over cropland).
 In case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
@@
 The multi-model-mean displays a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$), which cannot be reproduced with the mOSaic scheme.
-The dry deposition velocities over desert regions are consistent with the estimated values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
+The dry deposition velocities over desert regions are consistent with the average values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
 The difference between \emph{Wesely\_type} and \emph{mOSaic\_hough} stems most likely from $R_a$ and $R_b$ contributions.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with MACC-reanalysis}
-\label{subsec:macc}
-In this section, we conclude the comparison of our results with modeled ozone concentrations by comparing with ECWMF's MACC-reanalysis \citep{MACC}.
-In Fig,~\ref{fig:macc_o3conc}, we contrast \emph{mOSaic} ozone concentrations in the lowermost model level with surface concentrations deduced from the MACC-reanalysis for the year 2005.
-The MACC-reanalysis displays low ozone concentrations above all land masses except for the Greenland ice sheet, while the highest values are found within $(25-60)\,\unit{^\circ N}$ over the oceans.
-These high values over the oceans are relatively well reproduced by the mOSaic scheme, however over land (e.g.
-central Africa, Tibetan plateau) the ozone concentrations are up to $200\,\unit{\%}$ higher.
-While on average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to central Africa and the Tibetan plateau (Supplement~S.8).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig08}
-  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{mOSaic} (lowermost model level); (c) Relative difference.}
-  \label{fig:macc_o3conc}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison with observations}
-\label{subsec:obs}
-In this section, we compare our results to observations.
-For all comparisons, we use the original resolution of the Oslo~CTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).
-
-In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six selected observation sites are shown.
-We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ uncertainty band.
-The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual uncertainty band since such is not given in \citet{ACP:Hardacre2015}.
-At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and similar to or better than the multi-model-mean.
-We use a $\chi^2$-test
-\begin{equation}
-  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
-\end{equation}
-with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number of degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
-The closer to $1$ this test scores, the better does the simulation represent the observation.
-A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
-The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
-The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to observed ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
-With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
-Furthermore, the observational averages comprise at most $9$ years worth of data.
-Statistically, these data may still be subject to interannual variability.
-Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
-Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both high and low biases exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
-Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox and not substantially affected by differences in modeled and actual topography (e.g., major wind directions).
-
-%
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig09}
-  \caption{Ozone dry deposition fluxes at different observation sites.
-(a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
-The model uncertainty of the Oslo~CTM3 is given as $1\,\sigma$ error band.
-The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
-(b) Model divergence from observation and $\chi^2$-test results.}
-  \label{fig:mmm_drydep_stations}
-\end{figure}
-
 Table~\ref{tab:trop_ozone_burden} displays the average tropospheric ozone burden for all model experiments.
@@
 The difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), however, lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
-
+This indicates that the Oslo~CTM3 responses resonable to given changes in global emissions despite the strong positive bias in ozone concentrations and accordingly low bias in total dry deposition.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
+\subsection{Comparison with MACC-reanalysis}
+\label{subsec:macc}
+In this section, we conclude the comparison of our results with respect to global ozone by looking at ECWMF's MACC-reanalysis (obtained from \href{https://apps.ecmwf.int/datasets/data/macc-reanalysis/levtype=ml/}{ECWMF's data center}) \citep{MACC-II}.
+In Fig,~\ref{fig:macc_o3conc}, we compare \emph{mOSaic} ozone concentrations in the lowermost model level with surface concentrations deduced from the MACC-reanalysis for the year 2005.
+The MACC-reanalysis displays low ozone concentrations above all land masses except for the Greenland ice sheet.
+The lowest values are found in the deep tropics (e.g.
+northern South America and central Africa), while the highest values occur within $(25-60)\,\unit{^\circ N}$ over the oceans.
+These values over the oceans are relatively well reproduced by \emph{mOSaic} ($\pm 20\,\unit{\%}$).
+Over, e.g., South America, central and southern Africa, the Arabian penisula, and the north-western Indian subcontinent, the ozone concentrations are elevated by up to $200\,\unit{\%}$.
+While on global average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to elevated ozone over continents, especially the deep tropics and the north-western Indian subcontinent (Supplement~S.8) which coincides with regions of high intensity in incoming UV radiation.
+This may indicated an imbalance in the photo-chemical production and loss of \chem{O_3} in the Oslo~CTM3.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig08}
+  \caption{Mean ozone concentrations for the year 2005. (a) MACC-reanalysis (surface); (b) Oslo~CTM3 \emph{mOSaic} (lowermost model level); (c) Relative difference.}
+  \label{fig:macc_o3conc}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparison with ground-based observations}
+\label{subsec:obs}
+In this section, we compare our model results to observations at a selected number of sites which provide ozone flux measurements.
+For all comparisons, we use the original resolution of the Oslo~CTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded resolution ($3^\circ\times 3^\circ$).
+
+In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six selected observation sites are shown.
+We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ uncertainty band.
+The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual uncertainty band since such is not given in \citet{ACP:Hardacre2015}.
+At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and similar to or better than the multi-model-mean.
+We use a $\chi^2$-test
+\begin{equation}
+  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
+\end{equation}
+with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number of degrees of freedom (NDF) to assess this subjective analysis in a more objective way.
+The closer to $1$ this test scores, the better does the simulation represent the observation.
+A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small.
+The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
+The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to observed ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
+Furthermore, the observational averages comprise at most $9$ years worth of data.
+Statistically, these data may still be subject to interannual variability.
+Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance.
+Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both high and low biases exist due to dilution of sources and sinks in coarse resolution models \citep{AE:Schaap2015}.
+Good matches between observation and model are only to be expected if the station's location is representative for an area similar to the respective model gridbox and not substantially affected by differences in modeled and actual topography (e.g., major wind directions).
+
+%
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig09}
+  \caption{Ozone dry deposition fluxes at different observation sites.
+(a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}.
+The model uncertainty of the Oslo~CTM3 is given as $1\,\sigma$ error band.
+The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band.
+(b) Model divergence from observation and $\chi^2$-test results.}
+  \label{fig:mmm_drydep_stations}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conclusions[Summary and conclusions]  %% \conclusions[modified heading if necessary]
 \label{sec:conc}
-We have presented an update of the dry deposition scheme in the Oslo~CTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}.
-In our implementation based on \citet{ACP:Simpson2012}, we follow a classical, resistance analogous approach in computing contributions to dry deposition.
-The quasi-laminar resistance in the dry deposition scheme is capable of adjusting to surface wind stress but falls back to its prescribed limits in most of the tested cases.
-The surface resistance computation is divided into stomatal and non-stomatal resistance.
-The stomatal resistance is sensitive to parameters associated with the vegetation such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology, while the non-stomatal resistance is affected by the existence or absence of vegetation.
-
-The annual amount of ozone dry deposition decreases by up to $100\,\unit{\%}$ over all major desert areas changing from the old dry deposition scheme to the new one.
-At the same time, it increases over tropical forest.
+We have presented an update of the dry deposition scheme in the Oslo~CTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account.
+Based on the describtion of dry deposition in \citet{WASP:Simpson2003,ACP:Simpson2012}, we have implemented a moasic approach to compute contributions to dry deposition by individual sub-grid land surface types.
+Aerodynamic, quasi-laminar, and surface resistance (latter divided into stomatal and non-stomatal contributions) are calculated for each land surface category seperately.
+Based on these, a landfraction-weighted mean is deduced.
+In addition, the various dry deposition velocities are now directly available from model output for diagnostics and further studies.
+
+The new dry deposition scheme named \emph{mOSaic} improves the modeled dry deposition velocities which are now compatible with observation and model studies \citep[e.g.,][]{ACP:Hardacre2015,ACP:Luhar2018}.
+Dry deposition velocities are reduced by $6-60\,\unit{\%}$.
+At the same time, the annual amount of ozone dry deposition decreases by more than $100\,\unit{\%}$ over all major desert areas and increases over tropical forest.
 Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is about $33\,\unit{\%}$ below average.
@@
 
-We found the response of the Oslo~CTM3 to the changes in dry deposition velocities from the old and the new dry deposition scheme to be consistent.
-A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in surface ozone concentration $[\chem{O_3}(z_0)]$.
-As the new scheme is more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy when comparing with the multi-model-mean.
-Without knowing $[\chem{O_3}(z_0)]$ of the TF~HTAP models, this cannot be resolved and may hint to, e.g., differences in photo-chemistry, stratosphere-troposphere-exchange as well as to differences in aerosol loadings among many other aspects.
-In particular, the STE depends on the stratospheric ozone abundance which is, e.g.
-affected by very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} and is not taken into account in the Oslo~CTM3.
-
-Most of the decrease in ozone dry deposition in the Oslo~CTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
-This is mainly due to updates of the respective prescribed dry deposition velocities $v_\text{DD}^\chem{O_3}$.
+We found the response of the Oslo~CTM3 to the changes in dry deposition velocities from the old and the mOSaic scheme to be consistent.
+A decrease in $v_\text{DD}^\chem{O_3}$ leads to a decrease in total ozone dry deposition and an increase in ozone concentration $[\chem{O_3}]$.
+As the new scheme is quantitatively more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy.
+By means of tropospheric ozone burden \citep{ESA:Gaudel2018} and surface ozone concentrations deduced from the MACC-reanalysis \citep{MACC-II}, the Oslo~CTM3 with operational mOSaic scheme shows a pronounced high bias of tropospheric ozone.
+While the average bias is small or even reversed when using the old scheme, elevated ozone in comparison with the MACC-reanalysis is also found in continental regions with high incoming UV radiation.
+The reason behind this bias could not be resolved and may hint to, e.g., issues in photo-chemistry (ozone production and loss) or previously introduced optimization with respect to the old, less physical dry deposition velocities.
+
+Most of the qualitative decrease in ozone dry deposition in the Oslo~CTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
+This is mainly due to updates of the respective, prescribed ozone surface resistances $R^\chem{O_3}$.
 In case of desert and grasslands the difference between the old and new prescribed value is at the order of $1$ magnitude.
@@
 
-We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants.
-The difference between our most extreme test cases amounts to $15\,\unit{\%}$ with respect to total surface ozone, while the more realistic test case using differing years of emission amounts to about $5\,\unit{\%}$.
+We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics and the northern hemisphere are most sensitive to changes therein.
+In the most extreme test case, the increase in total dry deposition amounts to $7.3\,\unit{\%}$, while the more realistic test cases, e.g.
+using differing years of emission amount to changes in the order of about $\pm 2,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
@@
 The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.
+The largest impact, as mentioned above, is associated to the choice of surface resistance ($-2-12\,\unit{\%}$).
 
@@
 
-Future work on the Oslo~CTM3 should include a direct output of dry deposition velocities for diagnostic purposes.
+Future work on the Oslo~CTM3 should resolve the ozone high-bias which may invlove revising the photolysis- and chemical reaction computation as well as reaction rates should.
+For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone abundance and a scheme covering actic spring-time ozone depletion \citep[][e.g.,]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
 The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
-The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
 

### 2019-07-09 DryDep.tex
Spell-checked.

@@
 with the average wind speed $\overline{u}(z_\mathrm{ref})$ at reference height, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and roughness length $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
-Concidering the height of vegetation, we have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
+Considering the height of vegetation, we have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
 Using the derived $u_*^k$ from Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ can be obtained from \citep[Eq.~(8),][]{ACP:Simpson2012}:
@@
 Dry deposition velocities are directly available only for the new model version.
-For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged ozone amount in the lowermost modellevel $\chem{O_3}(p_0)$
+For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$
 \begin{equation}
@@
 The difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), however, lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
-This indicates that the Oslo~CTM3 responses resonable to given changes in global emissions despite the strong positive bias in ozone concentrations and accordingly low bias in total dry deposition.
+This indicates that the Oslo~CTM3 responses reasonable to given changes in global emissions despite the strong positive bias in ozone concentrations and accordingly low bias in total dry deposition.
 %
@@
 These values over the oceans are relatively well reproduced by \emph{mOSaic} ($\pm 20\,\unit{\%}$).
-Over, e.g., South America, central and southern Africa, the Arabian penisula, and the north-western Indian subcontinent, the ozone concentrations are elevated by up to $200\,\unit{\%}$.
+Over, e.g., South America, central and southern Africa, the Arabian peninsula, and the north-western Indian subcontinent, the ozone concentrations are elevated by up to $200\,\unit{\%}$.
 While on global average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to elevated ozone over continents, especially the deep tropics and the north-western Indian subcontinent (Supplement~S.8) which coincides with regions of high intensity in incoming UV radiation.
@@
 We have presented an update of the dry deposition scheme in the Oslo~CTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account.
-Based on the describtion of dry deposition in \citet{WASP:Simpson2003,ACP:Simpson2012}, we have implemented a moasic approach to compute contributions to dry deposition by individual sub-grid land surface types.
-Aerodynamic, quasi-laminar, and surface resistance (latter divided into stomatal and non-stomatal contributions) are calculated for each land surface category seperately.
-Based on these, a landfraction-weighted mean is deduced.
+Based on the description of dry deposition in \citet{WASP:Simpson2003,ACP:Simpson2012}, we have implemented a moasic approach to compute contributions to dry deposition by individual sub-grid land surface types.
+Aerodynamic, quasi-laminar, and surface resistance (latter divided into stomatal and non-stomatal contributions) are calculated for each land surface category separately.
+Based on these, a land fraction-weighted mean is deduced.
 In addition, the various dry deposition velocities are now directly available from model output for diagnostics and further studies.
@@
 
-Future work on the Oslo~CTM3 should resolve the ozone high-bias which may invlove revising the photolysis- and chemical reaction computation as well as reaction rates should.
-For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone abundance and a scheme covering actic spring-time ozone depletion \citep[][e.g.,]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
+Future work on the Oslo~CTM3 should resolve the ozone high-bias which may involve revising the photolysis- and chemical reaction computation as well as reaction rates should.
+For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone abundance and a scheme covering arctic spring-time ozone depletion \citep[][e.g.,]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
 The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
@@
   We would like to thank Prof.
-Frode Stordal (Section for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript as well as Anne Fouilloux (scientific programmer at the same institute) for technical support.
+Frode Stordal (Section for Meteorology and Oceanography, University of Oslo) for discussions regarding early drafts of the manuscript, Anne Fouilloux (scientific programmer at the same institute) for technical support as well as Olimpia Bruno (Karlsruhe Institute of Technology) and Franziska Hellmuth (University of Oslo) for valuable input regarding the aerodynamic resistance formulation.
   We would also like to thank the Center for International Climate Research (CICERO) for their support of this work.

### 2019-07-09 version_2/response_version_1_rc2.tex
Wrote intro to final response, revised summary and conclusion. ToDo - check if all issues have been addressed. Revise response RC2.

@@
 % Authors' response
-\documentclass{scrartcl}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
+%\documentclass{scrartcl}
+%\usepackage[utf8]{inputenc}
+%\usepackage[english]{babel}
 %\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{multirow}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\thispagestyle{empty}
+%\usepackage{amsmath}
+%\usepackage{xcolor}
+%\usepackage{url}
+%\usepackage{multirow}
+%\usepackage{graphicx}
+%\usepackage{hyperref}
+%\thispagestyle{empty}
 
-\begin{document}
+%\begin{document}
 \section*{Authors' response}
@@
 \item {\color{blue}P9 Eqn (27). This equation is a modification of Erisman’s original (1994) formulation, so explain that.}
-  We follow the advise and write: \emph{"The in-canopy resistance $R_\text{inc}$ (Erisman et al., 1994) is then modified with respect to each (vegetated) land type $N$ [...]"} {\bfseries ToDo -- To Amund: Were does this modification come from?}
+  We follow the advise and write: \emph{"The in-canopy resistance $R_\text{inc}$ (Erisman et al., 1994) is then modified with respect to each (vegetated) land type $N$ [...]"} {We have added the respective reference in the text.}
 
@@
 \end{itemize}
-\newpage
+%\newpage
 
-\end{document}
+%\end{document}

### 2019-07-09 DryDep.tex
Wrote intro to final response, revised summary and conclusion. ToDo - check if all issues have been addressed. Revise response RC2.

@@
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
-It is estimated that about $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
+It is estimated that $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
 A newer study by \citet{ACP:Luhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
@@
 \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances.
-Based on the global atmospheric composition reanalysis performed in the ECWMF project Monitoring Atmospheric Composition and Climate (MACC) and a more realistic process-based oceanic deposition scheme, \citet{ACP:Luhar2018} found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\unit{Tg\,a^{-1}}$.
+Based on the global atmospheric composition reanalysis performed in the ECWMF project Monitoring Atmospheric Composition and Climate (MACC) \citep{MACC-II} and a more realistic process-based oceanic deposition scheme, \citet{ACP:Luhar2018} found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\unit{Tg\,a^{-1}}$.
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.
@@
 The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and mOSaic scheme.
-In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum.
+In the Wesely scheme, the average tropical ozone dry deposition diverges by $5\,\unit{Tg}$ in comparison to its corresponding NH maximum.
 Compared to the multi-model-mean, the seasonal cycle in the Oslo~CTM3 NH appears to be shifted towards later in the year.
@@
 The multi-model-mean displays a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$), which cannot be reproduced with the mOSaic scheme.
-The dry deposition velocities over desert regions are consistent with the average values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are about $1$ order of magnitude lower than in the Wesely scheme.
+The dry deposition velocities over desert regions are consistent with the average values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are $1$ order of magnitude lower than in the Wesely scheme.
 The difference between \emph{Wesely\_type} and \emph{mOSaic\_hough} stems most likely from $R_a$ and $R_b$ contributions.
@@
 Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}).
-Despite its vastness, the ocean amounts only to about $35\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative mOSaic scheme (\emph{mOSaic}), while permanently ice and snow covered regions account for $1.2\,\unit{\%}$.
+Despite its vastness, the ocean amounts only to $35\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative mOSaic scheme (\emph{mOSaic}), while permanently ice and snow covered regions account for $1.2\,\unit{\%}$.
 The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling.
-The total annual dry deposition in the mOSaic scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
+The total annual dry deposition in the mOSaic scheme is one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}.
 But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition.
@@
 Table~\ref{tab:trop_ozone_burden} displays the average tropospheric ozone burden for all model experiments.
-Consistently with the previous findings, the mOSaic scheme increases the tropospheric ozone burden by about $35\,\unit{Tg}$.
+Consistently with the previous findings, the mOSaic scheme increases the tropospheric ozone burden by $35\,\unit{Tg}$.
 From various satellite ozone retrieval products, \citet[][Tab.~5]{ESA:Gaudel2018} deduce a lower limit estimate for global tropospheric ozone burden for the years $2010-2014$ of $333-345\,\unit{Tg}$, but remark that this amount underestimates the actual tropospheric ozone burden, since it is only based on daytime retrievals.
@@
 At the same time, the annual amount of ozone dry deposition decreases by more than $100\,\unit{\%}$ over all major desert areas and increases over tropical forest.
-Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is about $33\,\unit{\%}$ below average.
+Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the Oslo~CTM3 is $(38\substack{+~\,1 \\ -10})\,\unit{\%}$ below average.
 However, there seems to be a tendency that newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,ACP:Luhar2018,AE:Hu2017}.
@@
 As the new scheme is quantitatively more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy.
-By means of tropospheric ozone burden \citep{ESA:Gaudel2018} and surface ozone concentrations deduced from the MACC-reanalysis \citep{MACC-II}, the Oslo~CTM3 with operational mOSaic scheme shows a pronounced high bias of tropospheric ozone.
-While the average bias is small or even reversed when using the old scheme, elevated ozone in comparison with the MACC-reanalysis is also found in continental regions with high incoming UV radiation.
-The reason behind this bias could not be resolved and may hint to, e.g., issues in photo-chemistry (ozone production and loss) or previously introduced optimization with respect to the old, less physical dry deposition velocities.
-
-Most of the qualitative decrease in ozone dry deposition in the Oslo~CTM3 can be attributed to changes in dry deposition velocities over the ocean and deserts.
+By means of tropospheric ozone burden \citep{ESA:Gaudel2018} and surface ozone concentrations deduced from the MACC-reanalysis \citep{MACC-II}, the Oslo~CTM3 with operational mOSaic scheme shows a pronounced high-bias of tropospheric ozone.
+While the average bias is small or even reversed when using the old scheme, both display elevated ozone in comparison with the MACC-reanalysis in continental regions with high average incoming UV radiation (e.g., northern South America, central and southern Africa, the Himalayas).
+The reason behind this bias has not yet been resolved and may hint to, e.g., issues in photo-chemistry (ozone production and loss) or previously introduced optimization of ozone removal with respect to the old, less physical dry deposition velocities.
+
+Most of the qualitative change in ozone dry deposition in the Oslo~CTM3 ($-2-12\,\unit{\%}$) can be attributed to changes in dry deposition velocities over the ocean and deserts.
 This is mainly due to updates of the respective, prescribed ozone surface resistances $R^\chem{O_3}$.
@@
 
-Although dry deposition to ice and snow amounts to less than $1\,\unit{\%}$ of the total global annual ozone dry deposition, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres.
-By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in the model, however, this could lead to an overcompensation (high-bias).
+Although dry deposition to ice and snow amounts to only $1\,\unit{\%}$ of the total global annual ozone dry deposition in \emph{mOSaic}, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at $50\,\unit{^\circ}$ in both hemispheres.
+By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in models, however, this currently leads to an overcompensation (high-bias) in the Oslo~CTM3.
 
 We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics and the northern hemisphere are most sensitive to changes therein.
-In the most extreme test case, the increase in total dry deposition amounts to $7.3\,\unit{\%}$, while the more realistic test cases, e.g.
-using differing years of emission amount to changes in the order of about $\pm 2,\unit{\%}$.
+In the most extreme test case, the increase in global total dry deposition amounts to $7.3\,\unit{\%}$, while the more realistic test cases, e.g.
+using differing years of emission amount to changes in the order of $\pm 2,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
@@
 The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.
-The largest impact, as mentioned above, is associated to the choice of surface resistance ($-2-12\,\unit{\%}$).
-
-In the northern hemisphere mid and high latitudes ($50\,\unit{^\circ}-75\,\unit{^\circ N}$), total surface ozone increases by about $7.7\,\unit{\%}$ if the beginning and end of the vegetation period is estimated based on a $5\,\unit{^\circ C}$-days criteria instead of prescribed.
+
+%In the northern hemisphere mid and high latitudes ($50\,\unit{^\circ}-75\,\unit{^\circ N}$), total surface ozone increases by $7.7\,\unit{\%}$ if the beginning and end of the vegetation period is estimated based on a $5\,\unit{^\circ C}$-days criteria instead of prescribed.
 This is very important for any study focusing on ozone in the boreal and subarctic regions.
@@
 We chose the years 2005 and 2014 for our comparison.
-In 2014, surface ozone burden is higher in the southern hemisphere and in the tropics ($\sim 5\,\unit{\%}$) compared to 2005, while it is lower in the northern hemisphere ($\sim 2\,\unit{\%}$).
-
-We also evaluated the model with respect to observed dry deposition velocities at six sites in the northern hemisphere and found that the updated dry deposition scheme performs better than the old one, but is not able to reproduce the measurements at most sites quantitatively.
+Ozone precursor emissions in 2014 are slightly lower in the NH while enhanced in the tropics and the SH.
+In 2014, surface ozone burden is higher in the southern hemisphere and in the tropics ($5\,\unit{\%}$) compared to 2005, while it is lower in the northern hemisphere ($2\,\unit{\%}$).
+
+We also evaluated the model with respect to observed dry deposition fluxes at six sites in the northern hemisphere and found that the mOSaic scheme performs better than the old one, but is not able to reproduce the measurements at most sites quantitatively.
 This may be due to several reasons.
@@
 
-Future work on the Oslo~CTM3 should resolve the ozone high-bias which may involve revising the photolysis- and chemical reaction computation as well as reaction rates should.
-For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone abundance and a scheme covering arctic spring-time ozone depletion \citep[][e.g.,]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
+Future work on the Oslo~CTM3 should resolve the ozone high-bias which may involve revising the photolysis- and chemical reaction computation as well as reaction rates.
+For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone and a scheme covering arctic spring-time ozone depletion \citep[e.g.,][]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
 The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
+A more efficient parallelization of the code would enable computation on higher horizontal resolutions.
 
@@
 
-\codedataavailability{The Oslo~CTM3 is publicly available on git-hub under a MIT license.
+\codedataavailability{The Oslo~CTM3 shall be publicly available on git-hub under a MIT license in the future.
+Until then, access can be made granted under request.
 Model results can be made available under request.} %% use this section when having data sets and software code available

### 2019-07-09 version_2/final_response_version_1.tex
Wrote intro to final response, revised summary and conclusion. ToDo - check if all issues have been addressed. Revise response RC2.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
%\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{url}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\thispagestyle{empty}

\begin{document}
\section*{Final response}
We thank both referees for their thorough review of our paper and the evaluation of the formulation of the dry deposition scheme.
In consideration of all comments, it became necessary to revise a major part of the dry deposition implementation in our model. Particularly, we  
\begin{itemize}
  \item have addressed and corrected the used equations and the implementation of the mosaic approach in the model,
  \item have fixed all related issues in the manuscript,
  \item have included a diagnostic output of dry deposition velocities,
  \item now refer to our new dry deposition scheme as \emph{mOSaic scheme},
  \item have refined the definition of our model experiments to make them more consistent, especially
    \begin{itemize}
    \item using SWVL3 and
    \item CEDS emissions in accordance to the meteorological year.
    \end{itemize}
    This also includes replacement or removal of some previous experiments
    \begin{itemize}
    \item \emph{EMEP\_SWVL4} $\rightarrow$ \emph{mOSaic\_SWVL1},
    \item \sout{\emph{EMEP\_ppgs}},
    \item \sout{\emph{EMEP\_ppgssh}}, and
    \item \emph{EMEP\_ppgs\_2005} $\rightarrow$ \emph{mOSaic\_emis2014}.
    \end{itemize}
    We have added two experiments in which we explicitly change the prescribed ozone surface resistance $R^\mathrm{O_3}$:
    \begin{itemize}
    \item \emph{mOSaic\_desert},
    \item\emph{mOSaic\_hough}.
    \end{itemize}
  \item have elaborated on the evaluation and discussion of the results and included a section in which we compare our results to the MACC-reanalysis.
\end{itemize}
In the following, we include a detailed account of all referee comments and the track-change-version of the manuscript. We have not updated the responses already published in gmdd (RC1, RC3). Hence, the referenced numbers will be different in the final manuscript.
\newpage
\input{response_version_1}
\input{response_version_1_rc2}
\input{response_version_1_rc3}
\newpage

\end{document}

### 2019-07-10 version_2/response_version_1_rc2.tex
Updated final response.

@@
 I think the authors need to make a case for their approach, or change it.}
-    We have discussed the concerns raised within this comment and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
+    We have discussed the concerns raised and found that we have indeed made a mistake in our implementation of the \emph{mosaic approach} which forces us to revise our model and repeat the model experiments.
+We update the equations in the revised manuscript accordingly.
   \end{itemize}
@@
   Given $z = z_\mathrm{ref} = 8\,\mathrm{m}$ and $d = 0.7\cdot 1\,\mathrm{m}$, it is right that neither the correct Eq.~(\ref{eq:aero_res}) nor the erroneous equation will result in negative $R_a$.
-The statement "$d$ is ’typically 0.7\,m’" is a typo.
-We correct: \emph{"[...] typically $d = 0.7\cdot h(N,\mathrm{lat})\,\mathrm{m}$ for vegetation other than forests [...]"}.
+The statement "$d$ is ’typically 0.7\,m’" is incorrect, indeed.
+We correct: \emph{"[...] $d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests [...]"}.
 
-  Correction note: The average height of the lowermost model level is actually $20\,\mathrm{m}$ ($10\,\mathrm{m}$ for mid-level).
+  Side note: The average height of the lowermost model level is actually $20\,\mathrm{m}$ ($10\,\mathrm{m}$ for mid-level).
 
@@
 The text and derivation of the Monteith relation is confusing and also erroneous.
-If we do not implement Eq.~(\ref{eq:aero_res}), we will elaborate on this in a revised version of the manuscript: \emph{"
+In a revised version of the dry deposition scheme, we do no longer use these formulations, hence they are removed from the revised manuscript.
+Nevertheless, the correct form should have read as follows:
 \begin{equation}
@@
 \end{equation}
-Herein, $\rho_\text{air}$ is the air density, $c_P$ the specific heat at constant pressure $P$, $R_{a, H}$ is the diffusion resistance to sensible heat between surface and the reference height $z$, and $\Delta T$ is the difference between the surface temperature $T_0$ and the temperature at a reference height $T_z$.
-From the eddy diffusion theorem, a similar formulation can be derived
 \begin{equation}
@@
 For finite $z$, $\partial T \rightarrow \Delta T$.
-  "}
 
@@
 discussing that fact.}
-  In the revision of the model, we are going to implement the stability dependent $R_a$ following Eq.~(\ref{eq:aero_res}).
+  In the revision of the model, we are going to implement the stability dependent $R_a$ following Eq.~(\ref{eq:aero_res}) and update the manuscript accordingly.
 
@@
   Thank you for pointing this out.
-In S2012 we find, $R_a$ evaluated "at around $45\,\mathrm{m}$" which is similar to the center of our second lowest model level.
-We can easily correct for this in a revised model version.
+Indeed, an evaluation of winds at $8\,\mathrm{m}$ does not make much sense, given that forests in the tropics reach heights up to $40\,\mathrm{m}$.
+In S2012, $R_a$ is actually evaluated "at around $45\,\mathrm{m}$" which is similar to the center of our second lowest model level.
+We correct for this in the revised model version.
   \end{itemize}
@@
 in this way? Is there any publication as to the reliability of this method?}
-  The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 (\href{http://folk.uio.no/mariantl/osloctm3/history.html}{Oslo~CTM2 list of publications}).
+  The $a_1$ coefficients were calculated from deposition velocities taken from the Oslo~CTM2 (\href{http://folk.uio.no/mariantl/osloctm3/history.html}{Oslo~CTM2 list of publications, online}).
 We meant to use this new scheme for aerosols in the Oslo~CTM3, but as of now it has not been evaluated and is actually not used.
@@
   The caption may become clearer after we changed the name of our scheme, so that it can no longer be confused with the actual EMEP model.
-The purpose of the mapping is to project the more detailed categories for stomatal conductance to the 10 ground-surface resistance categories (Table~S19 in S2012 supplement).
+The purpose of the mapping is to project the more detailed categories for stomatal conductance to the ground-surface resistance categories (Table~S19 in S2012 supplement).
+We have updated figure and text in accordance to the corrections made in the model code.
+It should be clearer now.
 
@@
 minute) PPFD using cloud-cover and zenith angles?}
+  We elaborate on the comprehensibility of this paragraph: \emph{"From OpenIFS an accumulated surface PAR is available.
+It is integrated both, spectral (presumably $400-60\,\mathrm{nm}$) and temporally.
+For practical use in Eq.~(18), we de-accumulate this field with respect to time and refer to the result as PPFD."}
 We have not used that ansatz, because we do not have the radiation fields in the relevant wavelengths available from OpenIFS output.

### 2019-07-10 version_2/response_version_1_rc2.tex
Went through the first two sections in detail again and made some adjustments. The last two sections will be revised later on.

@@
 We elaborate on this in a revised version of the paper: \emph{Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
-Long-range ozone transport occurs and might be most important in regions that else lack precursors.
+Long-range ozone transport occurs regularly and might be most important in regions that else lack precursors.
 Tropospheric ozone is produced in complex photochemical cycles involving precursor gases [...]}
@@
   That is indeed true, but we didn't want to name all possible surfaces onto which dry deposition takes place.
-We, hence, shall write: \emph{"Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants and soil, is referred to as dry deposition.}
+We, hence, shall write: \emph{"Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants, soil, and water, is referred to as dry deposition.}
 
@@
 \item {\color{blue}P9 L18. What do you mean by "surface or 2\,m"? Surface might refer to skin or leaf temperature?}
-  This may have been indeed unclear. It is the $\mathrm{2\,m}$ temperature. We change the text accordingly: \emph{"[...] $\theta_0$ is the $2\,\mathrm{m}$ temperature in $\mathrm{^\circ C}$."}
+  This may have been indeed unclear.
+It is the $\mathrm{2\,m}$ temperature.
+We change the text accordingly: \emph{"[...] $\theta_\mathrm{2m}$ is the $2\,\mathrm{m}$ temperature in $\mathrm{^\circ C}$."}
 

### 2019-07-10 DryDep.tex
Went through the first two sections in detail again and made some adjustments. The last two sections will be revised later on.

@@
 We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-37\,\unit{\%}$), leading to an increase in surface ozone of more than $100\,\unit{\%}$ in some regions.
-While high sensitivity to changes in dry deposition to vegetation is found in the tropics and the northern hemisphere, the largest impact on global scales is associated to changes in dry deposition to the ocean and deserts.
+While high sensitivity to changes in dry deposition to vegetation is found in the tropics and the northern hemisphere, the largest impact on global scales is associated to the choice of prescribed ozone surface resistance the ocean and deserts.
 \end{abstract}
@@
 Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
-Long-range ozone transport occurs and might be most important in regions that else lack precursors.
+Long-range ozone transport occurs regularly and might be most important in regions that else lack precursors.
 Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
@@
 Another source of tropospheric ozone is downward transport from the stratosphere via stratosphere-troposphere exchange (STE) \citep{WMO2014}.
-Based on observations, STE might only amount to roughly $10\,\unit{\%}$ ($550 \pm 140\,\unit{Tg\,a^{-1}}$) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be $5000\,\unit{Tg\,a^{-1}}$ \citep{ACP:Monks2015}.
+Based on observations, STE might only amount to $10\,\unit{\%}$ ($550 \pm 140\,\unit{Tg\,a^{-1}}$) of the total global ozone budget in the troposphere, while ozone from chemical production is estimated to be $5000\,\unit{Tg\,a^{-1}}$ \citep{ACP:Monks2015}.
 Ozone is removed from the atmosphere by photochemical reactions or scavenging processes.
@@
 A key to fathom these slightly contradicting results may lie in the used dry deposition schemes.\\
-Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants and soil, is referred to as dry deposition.
+Removal of any substance from the atmosphere which is not involving rain, e.g., through gravitational settling or by uptake by plants, soil, and water, is referred to as dry deposition.
 The process of dry deposition is important for predicting surface ozone concentrations and understanding the observed amount and increase of tropospheric background ozone.
@@
 In Section~\ref{sec:model_des}, we will briefly describe the Oslo~CTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
-In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to surface ozone observations (Section~\ref{subsec:obs}).
+In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and to surface ozone observations (Section~\ref{subsec:obs}).
 In Section~\ref{sec:conc}, we will summarize and discuss our results and draw conclusions for further development of the model.
@@
 Winter was defined by temperatures below $273.15\,\unit{K}$ for gridboxes containing land masses.
-For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes.
+For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment was needed for ocean gridboxes.
 In addition, a reduced uptake due to snow cover above $1\,\unit{m}$ for forest and $10\,\unit{cm}$ for grass/tundra, respectively, was taken into account.
@@
 with the average wind speed $\overline{u}(z_\mathrm{ref})$ at reference height, the K\'{a}rm\'{a}n constant $\kappa = 0.40$, the integrated stability equation for momentum $\Psi_m$ \citep[e.g.,][]{Garratt1992}, a grid average Obukhov length $L$, deplacement height $d_k$, and roughness length $z_0^k$ ($d_k=0.78\cdot h_k(\mathrm{lat})$, $z_0^k=0.07\cdot h_k(\mathrm{lat})$ for forests, $d_k=0.7\cdot h_k(\mathrm{lat})$, $z_0^k=0.1\cdot h_k(\mathrm{lat})$ for vegetation other than forests).
-Considering the height of vegetation, we have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
+Taking the height of vegetation in to consideration, we have chosen the model level such that $\overline{z}_\mathrm{ref}\approx 45\,\unit{m}$.
 Using the derived $u_*^k$ from Eq.~(\ref{eq:local_fric_vel}), a local Obukhov length $L_k$ can be obtained from \citep[Eq.~(8),][]{ACP:Simpson2012}:
@@
 %
-The factors herein are normalized and vary within the range $0-1$ and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
+The factors herein are normalized and vary within the range $0-1$.
+They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
 All factors differ with land use type $k$.
@@
 For all other species, an interpolation between \chem{O_3} and \chem{SO_2} values is carried out.
-The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For land surface types $k$, we can write
+The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For each land surface types $k$, we can write
 \begin{equation}
@@
 \end{equation}
-$\text{SAI}_k$ is the surface area index for vegetation type $k$, which is $\text{LAI}$ plus some value that represents cuticles and others surfaces. The external leaf resistance is defined by
+$\text{SAI}_k$ is the surface area index for vegetation type $k$, which is $\text{LAI}$ plus a value that represents cuticles and others surfaces. The external leaf resistance is defined by
 \begin{equation}
@@
 Either land surface categories from ISLSCP2 product of MODIS or the Community Land Model (CLM)~2 can be chosen for mapping to the land surface types we use in the mOSaic scheme.
-Water bodies and snow and ice categories of MODIS are actually not mapped.
+Water bodies of MODIS are actually not mapped.
 For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{k} f_L^k$.
@@
 \subsubsection{Photosynthetic photon flux density}
-From OpenIFS an accumulated surface PAR is available which is already integrated presumably over the spectrum $400-700\,\unit{nm}$.
-For this reason, it will be referred to as PPFD instead of PAR.
-This PPFD field has not been de-accumulated in the post-processing of the OpenIFS output and it is not feasible to redo all of the post-processing for this work.
-But for practical use in Eq.~(\ref{eq:flight}), this field needs to be de-accumulated.
+From OpenIFS an accumulated surface PAR is available.
+It is integrated both, spactrally (presumably $400-700\,\unit{nm}$) and temporally.
+For practical use in Eq.~(\ref{eq:flight}), we de-accumulate this field with respect to time and refer to the result as PPFD.
+
 The main obstacle is that PAR has been accumulated since model start, so that the first field kept from the original OpenIFS simulation ($00\,\unit{UTC}$) is $12$ hours after model start ($12\,\unit{UTC}$ on the previous day).

### 2019-07-10 DryDep.tex
Finished final revision of text. ToDo - diff and new abstract pdf.

@@
 In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the mOSaic scheme.
-We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, ozone concentrations in the lowermost model level $[\chem{O_3}](p_0)$, and tropospheric burden $\sum_\text{trop}\chem{O_3}$.
+We focus on changes in ozone total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, concentrations in the lowermost model level $[\chem{O_3}](p_0)$, and tropospheric burden $\sum_\text{trop}\chem{O_3}$.
 We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and observations (Section~\ref{subsec:obs}).
@@
 Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every single difference in results to a respective change.
-Therefore, we conducted one reference simulation denoted as \emph{mOSaic} implementing all parameters as tabulated in \citet{ACP:Simpson2012} and in total seven sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}).
+Therefore, we conducted one reference simulation denoted as \emph{mOSaic} and in total seven sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}).
 A reference simulation featuring the Oslo~CTM3 Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
 All model experiments discussed in the following are summarized in Table~\ref{tab:simsum}.
-An \emph{x} therein denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.
+An \emph{x} therein denotes that the model was run exactly in the configuration and with parameters as has been described in Section~\ref{sec:model_des}.
 For all model integrations, the meteorological reference year is 2005.
-This choice affects the direct comparison with data and studies that either show results based on decadal averages or differing years.
+This choice affects the direct comparison with data and studies that either show results based on decadal averages or differing years, because non-linearities in ozone formation and destruction make ozone concentrations sensitive to both, differences in local concentration of precursors and meteorological conditions \citep{JGR:Jin2013}.
 
@@
 A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}.
-We cover this in more detail in the following section.
+We come back to this in the following sections.
 
@@
 Dry deposition velocities are directly available only for the new model version.
-For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$
-\begin{equation}
-  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
+For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(p_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$
+\begin{equation}
+  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(p_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
   \label{eq:retro_vdd}
@@
 Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
-In case of \emph{mOSaic}, Eq.~(\ref{eq:retro_vdd}) yields similar values for $v^\chem{O_3}_\text{DD}$ as directly available from model output.
+In case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly ailable from model output.
 %
@@
 In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean.
-This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3.
+This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as new standard for the Oslo~CTM3.
 This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
@@
 The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
-The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
+The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$ (almost entirely covered by ocean), $(12-30)\,\unit{^\circ S}$ (coinciding with the location of Australia and its desert regions), as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
 
@@
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameter set \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}.
-These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}).
-We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).
+These differences in surface resistances over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}).
+We further elaborate on this in the following (Section~\ref{subsubsec:seasons}).
 
@@
 convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further.
-Another reason might be found within the (chemistry) routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this is subject to further investigations.
+This may also hint to issues in the Oslo~CTM3 photo-chemistry, which may have a too high ozone production, or the actual removal of ozone from the atmosphere, which might have been adjusted to the less physical dry deposition velocities in the past, but this is subject to further investigations.
 
@@
 As suggested by \citet{ACP:Hardacre2015}, we also look at ozone dry deposition velocities with respect to surface types separately.
-Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to calculate these.
+Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to estimate these.
 Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
@@
 The dry deposition velocities over desert regions are consistent with the average values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are $1$ order of magnitude lower than in the Wesely scheme.
-The difference between \emph{Wesely\_type} and \emph{mOSaic\_hough} stems most likely from $R_a$ and $R_b$ contributions.
+In the mOSaic scheme, dry deposition to deserts is dominated by contribution from $R_b$.
 From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$.
@@
 Nevertheless, the results of \emph{mOSaic} lie $17\,\unit{\%}$ above that estimate and also well above the typical modeling range of $302-378\,\unit{Tg}$ \citep{ACP:Young2013}.
-The difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), however, lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
-This indicates that the Oslo~CTM3 responses reasonable to given changes in global emissions despite the strong positive bias in ozone concentrations and accordingly low bias in total dry deposition.
+Despite the strong positive bias in ozone concentrations and accordingly low-bias in total dry deposition, the difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
+This indicates that the Oslo~CTM3 responses well to given changes in global emissions.
 %
@@
 \label{subsec:macc}
-In this section, we conclude the comparison of our results with respect to global ozone by looking at ECWMF's MACC-reanalysis (obtained from \href{https://apps.ecmwf.int/datasets/data/macc-reanalysis/levtype=ml/}{ECWMF's data center}) \citep{MACC-II}.
+In this section, we conclude the comparison of our results with respect to global ozone by looking at ECWMF's MACC-reanalysis \citep[][data obtained from \href{https://apps.ecmwf.int/datasets/data/macc-reanalysis/levtype=ml/}{ECWMF's data center}]{MACC-II}.
 In Fig,~\ref{fig:macc_o3conc}, we compare \emph{mOSaic} ozone concentrations in the lowermost model level with surface concentrations deduced from the MACC-reanalysis for the year 2005.
@@
 northern South America and central Africa), while the highest values occur within $(25-60)\,\unit{^\circ N}$ over the oceans.
-These values over the oceans are relatively well reproduced by \emph{mOSaic} ($\pm 20\,\unit{\%}$).
+These low values over the oceans are relatively well reproduced by \emph{mOSaic} ($\pm 20\,\unit{\%}$).
 Over, e.g., South America, central and southern Africa, the Arabian peninsula, and the north-western Indian subcontinent, the ozone concentrations are elevated by up to $200\,\unit{\%}$.
-While on global average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to elevated ozone over continents, especially the deep tropics and the north-western Indian subcontinent (Supplement~S.8) which coincides with regions of high intensity in incoming UV radiation.
+While on global average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to elevated ozone over continents.
+Enhanced ozone compared to MACC-reanalysis is found in the deep tropics and the north-western Indian subcontinent (Supplement~S.8) which coincides with regions of high intensity in incoming UV radiation.
 This may indicated an imbalance in the photo-chemical production and loss of \chem{O_3} in the Oslo~CTM3.
@@
 
-In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for six selected observation sites are shown.
+In Fig.~\ref{fig:mmm_drydep_stations}a, seasonal cycles of average ozone dry deposition fluxes for the six selected observation sites are shown.
 We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ uncertainty band.
@@
 While the average bias is small or even reversed when using the old scheme, both display elevated ozone in comparison with the MACC-reanalysis in continental regions with high average incoming UV radiation (e.g., northern South America, central and southern Africa, the Himalayas).
-The reason behind this bias has not yet been resolved and may hint to, e.g., issues in photo-chemistry (ozone production and loss) or previously introduced optimization of ozone removal with respect to the old, less physical dry deposition velocities.
-
-Most of the qualitative change in ozone dry deposition in the Oslo~CTM3 ($-2-12\,\unit{\%}$) can be attributed to changes in dry deposition velocities over the ocean and deserts.
+The reason behind this bias has not yet been resolved and may hint to, e.g., issues in photo-chemistry ($[\chem{OH}]$ related ozone production and loss) or previously introduced optimization of ozone removal with respect to the old, less physical dry deposition velocities.
+
+Most of the qualitative change in ozone dry deposition in the Oslo~CTM3 ($-2-12\,\unit{\%}$) can be attributed to changes in dry deposition over the ocean and deserts.
 This is mainly due to updates of the respective, prescribed ozone surface resistances $R^\chem{O_3}$.
-In case of desert and grasslands the difference between the old and new prescribed value is at the order of $1$ magnitude.
+In case of desert and grasslands the difference between the old and new prescribed value is in the order of $1$ magnitude.
 Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
@@
 Although dry deposition to ice and snow amounts to only $1\,\unit{\%}$ of the total global annual ozone dry deposition in \emph{mOSaic}, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at $50\,\unit{^\circ}$ in both hemispheres.
-By comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in models, however, this currently leads to an overcompensation (high-bias) in the Oslo~CTM3.
-
-We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics and the northern hemisphere are most sensitive to changes therein.
+Comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in models, however, this currently leads to an overcompensation (high-bias) in the Oslo~CTM3.
+
+We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics and the northern hemisphere is most sensitive to changes therein.
 In the most extreme test case, the increase in global total dry deposition amounts to $7.3\,\unit{\%}$, while the more realistic test cases, e.g.
-using differing years of emission amount to changes in the order of $\pm 2,\unit{\%}$.
+using differing years of emission amount to changes in the order of $\pm 2\,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
@@
 Depending on the location of the observation site and its respective representativeness for a larger area, ozone dry deposition and ozone concentrations are expected to be over- or underestimated in the model.
+Because of non-linearities in ozone formation and destruction, ozone concentrations are sensitive to both, differences in local concentration of precursors and meteorological conditions \citep{JGR:Jin2013}.
 In addition, a comparison of very few years of measurement to only one specific year of simulation may reflect the year to year variability more than the actual model performance.
@@
 Future work on the Oslo~CTM3 should resolve the ozone high-bias which may involve revising the photolysis- and chemical reaction computation as well as reaction rates.
-For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013, ACP:Hossaini2016, ACP:Falk2017} which affect the stratospheric ozone and a scheme covering arctic spring-time ozone depletion \citep[e.g.,][]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
-The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
+For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013} which affect the stratospheric ozone \citep{ACP:Hossaini2016, ACP:Falk2017} and a scheme covering arctic spring-time ozone depletion \citep[e.g.,][]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
+The general model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
 A more efficient parallelization of the code would enable computation on higher horizontal resolutions.

### 2019-07-11 DryDep.tex
Corrected order in supplement and manuscrit accordingly.

@@
   \label{tab:simsum}
-  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.7} % Table Footnotes
+  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.6} % Table Footnotes
 \end{table*}
@@
 Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH.
-The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.7).
+The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.6).
 The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years.
@@
 The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
-The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8).
+The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.7: Fig.~S8).
 The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$ (almost entirely covered by ocean), $(12-30)\,\unit{^\circ S}$ (coinciding with the location of Australia and its desert regions), as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.

### 2019-07-11 version_2/final_response_version_1.tex
Small adjustments to response. Added Diff.

@@
 % Authors' response
-\documentclass{scrartcl}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
+%\documentclass{scrartcl}
+%\usepackage[utf8]{inputenc}
+%\usepackage[english]{babel}
 %\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{multirow}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage[normalem]{ulem}
-\thispagestyle{empty}
+%\usepackage{amsmath}
+%\usepackage{xcolor}
+%\usepackage{url}
+%\usepackage{multirow}
+%\usepackage{graphicx}
+%\usepackage{hyperref}
+%\usepackage[normalem]{ulem}
+%\thispagestyle{empty}
 
-\begin{document}
+%\begin{document}
 \section*{Final response}
@@
 \begin{itemize}
-  \item have addressed and corrected the used equations and the implementation of the mosaic approach in the model,
-  \item have fixed all related issues in the manuscript,
+\item have addressed and corrected the equations in
+  \begin{itemize}
+  \item the implementation of the mosaic approach,
+  \item the computation of $R_a$.
+  \end{itemize}
+  \item have fixed the equations with respect to comprehensibility (e.g., indexing),
   \item have included a diagnostic output of dry deposition velocities,
@@
     \begin{itemize}
-    \item using SWVL3 and
+    \item using the soil moisture at a more appropriate level (SWVL3) and
     \item CEDS emissions in accordance to the meteorological year.
@@
     \end{itemize}
-  \item have elaborated on the evaluation and discussion of the results and included a section in which we compare our results to the MACC-reanalysis.
+  \item have elaborated on the evaluation and discussion of the results and
+  \item included a section in which we compare our results to the MACC-reanalysis.
 \end{itemize}
-In the following, we include a detailed account of all referee comments and the track-change-version of the manuscript.
-We have not updated the responses already published in gmdd (RC1, RC3).
-Hence, the referenced numbers will be different in the final manuscript.
+In the following, we account for all referee comments in detail and include the track-change-version of the manuscript.
+{\bf Remark:} We have not updated the responses that have been already published in gmdd (RC1, RC3).
+Hence, the referenced values therein differ from the final manuscript.
 \newpage
-\input{response_version_1}
-\input{response_version_1_rc2}
-\input{response_version_1_rc3}
+\input{version_2/response_version_1}
+\input{version_2/response_version_1_rc2}
+\input{version_2/response_version_1_rc3}
 \newpage
 
-\end{document}
+%\end{document}

### 2019-10-04 DryDep.tex
Added all remarks of editor.

@@
 
-\runningtitle{Update of ozone dry deposition in Oslo~CTM3}
+\runningtitle{Update of ozone dry deposition in the Oslo~CTM3}
 
@@
 We show that the global annual total ozone dry deposition decreases with respect to the previous model version ($-37\,\unit{\%}$), leading to an increase in surface ozone of more than $100\,\unit{\%}$ in some regions.
-While high sensitivity to changes in dry deposition to vegetation is found in the tropics and the northern hemisphere, the largest impact on global scales is associated to the choice of prescribed ozone surface resistance the ocean and deserts.
+While high sensitivity to changes in dry deposition to vegetation is found in the tropics and the northern hemisphere, the largest impact on global scales is associated with the choice of prescribed ozone surface resistance over the ocean and deserts.
 \end{abstract}
@@
 Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
-Long-range ozone transport occurs regularly and might be most important in regions that else lack precursors.
+Long-range ozone transport occurs regularly and might be most important in regions that otherwise lack precursors.
 Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
@@
 Typical concentrations of surface ozone range from $10\,\unit{ppb}$ over the tropical Pacific to $100\,\unit{ppb}$ in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
-This variability poses a challenge on both, trend analysis from observation as well as validation and intercomparison of models.
-From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sides.
+This variability poses a challenge on both, trend analysis from observations as well as validation and intercomparison of models.
+From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sites.
 Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
 But especially the very low pre-industrial ozone abundance cannot be reproduced by the likes of most models.
+These early observations, however, were subject to interference by other species, e.g., \chem{SO_2}.
 Among the participating models in the Atmospheric Chemistry and Climate Model Intercomparison Project (ACCMIP), there is a general tendency to underestimate tropospheric ozone burden (e.g., $10-20\,\unit{\%}$ negative bias at $250\,\unit{hPa}$ in the southern hemisphere (SH) tropical region) \citep[Chapter 8]{IPCC2013}.
@@
 It is estimated that $1000 \pm 200\,\unit{Tg\,a^{-1}}$ of ozone are removed from the atmosphere by dry deposition processes \citep{ACP:Monks2015}.
-A newer study by \citet{ACP:Luhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$).
+A newer study by \citet{ACP:Luhar2018}, however, indicates much lower amounts ($722.8 \pm 87.3\,\unit{Tg\,a^{-1}}$) due to lower dry deposition to the oceans.
 Conceptually, dry deposition is a product between near-surface ozone concentration \chem{[O_3](z_0)} (e.g.
@@
 Based on the global atmospheric composition reanalysis performed in the ECWMF project Monitoring Atmospheric Composition and Climate (MACC) \citep{MACC-II} and a more realistic process-based oceanic deposition scheme, \citet{ACP:Luhar2018} found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\unit{Tg\,a^{-1}}$.
+In particular, \citet{ACP:Luhar2018} find that the average surface resistance of ozone over ocean ($r_c=2200\,\unit{sm^{-1}}$) is highly overestimated in most models.
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.
@@
 
-In Section~\ref{sec:model_des}, we will briefly describe the Oslo~CTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
+In Section~\ref{sec:model_des}, we will briefly describe the Oslo~CTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as the begin and duration of the greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}).
 In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and to surface ozone observations (Section~\ref{subsec:obs}).
@@
 The Oslo~CTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}.
-A publication focusing on this is planed.
+A publication focussing on this is planned.
 
 %Although the ozone depleting events in the polar boundary layer (Section~\ref{sec:intro}) are important to understand surface ozone abundance in Arctic regions in spring-time, no parameterization of these processes is implemented in the Oslo~CTM3 as of now.\\
-While the meteorological data driving the Oslo~CTM3 is given in a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
+While the meteorological data driving the Oslo~CTM3 is given at a resolution of T159N80L60, with the highest model level at $0.02\,\unit{hPa}$, it is very time and memory consuming to run the Oslo~CTM3 with full chemistry at this resolution.
 Therefore, we reduced the horizontal resolution to $2.25^\circ\times2.25^\circ$ in our experiments.
@@
 \label{subsubsec:Rc}
-The surface resistance consists of both, stomatal and non-stomatal resistances.\\ 
+The surface resistance consists of both stomatal and non-stomatal resistances.\\ 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}, but is not taken into account in our formulation.
-To reflect part of the underlying mechanism, the leaf-level stomatal conductance in the mOSaic scheme is computed using a common multiplicative ansatz \citep{BallBerry1987, ICP:MappingManual2017}:
+To reflect part of the underlying mechanism, the leaf-level molar stomatal conductance in the mOSaic scheme is computed using a common multiplicative ansatz \citep{BallBerry1987, ICP:MappingManual2017}:
 %
@@
 For clarity reasons, we drop this index in the following, as long as it is not necessary for the equation's completeness.
+The maximum molar stomatal conductance is given by $g^k_\text{max, m}$ which is in units of \unit{mmol\,s^{-1}\,m^{-2}}.
+A unit conversion to \unit{m\,s^{-1}} is necessary in our model:
+
+\begin{equation}
+  g^k_{\text{sto}} = g_{\text{sto, m}}^k \cdot R \cdot \frac{T_0}{P_0}.
+\end{equation}
+Herein, $R$ is the universal gas constant. To annotate the differing units, we use the index $\text{m}$ in Eq.~(\ref{eq:stomatal}).\\
 
@@
 The phenology of a plant typically describes its life-cycle throughout a year, e.g., at mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall.
-In the mOSaic scheme, phenology is parameterized with respect to the start of greening season (SGS) and its end (EGS).
+In the mOSaic scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS).
 Details about our treatment of these are given in Section~\ref{subsubsec:greening}.
@@
 
-Since $g_\text{max, m}$ in Eq.~(\ref{eq:stomatal}) is in units of \unit{mmol\,s^{-1}\,m^{-2}}, a unit conversion to \unit{m\,s^{-1}} is necessary in our model:
-\begin{equation}
-  g^k_{\text{sto}} = g_{\text{sto, m}}^k \cdot R \cdot \frac{T_0}{P_0}.
-\end{equation}
-Herein, $R$ is the universal gas constant.\\
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For all other species, an interpolation between \chem{O_3} and \chem{SO_2} values is carried out.
-The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For each land surface types $k$, we can write
+The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For each land surface type $k$, we can write
 \begin{equation}
@@
 \end{equation}
-$\text{SAI}_k$ is the surface area index for vegetation type $k$, which is $\text{LAI}$ plus a value that represents cuticles and others surfaces. The external leaf resistance is defined by
+$\text{SAI}_k$ is the surface area index for vegetation type $k$, which is $\text{LAI}$ plus a value that represents cuticles and other surfaces. The external leaf resistance is defined by
 \begin{equation}
@@
 In the following, we present the necessary pre-processing of the variables.
-It is planed to implement an online computation of these variables into the Oslo~CTM3 later on.
+It is planned to implement an online computation of these variables into the Oslo~CTM3 later on.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Since the parameterization of SGS and EGS in \citet{ACP:Simpson2012} is not applicable in a global model, another latitude dependent parameterization is needed.
-First, we used a parameterization which was already implemented in the Oslo~CTM3 and which had been adopted from the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{model webpage}).
+First, we used a parameterization which was already implemented in the Oslo~CTM3 and which had been adopted from the Sparse Matrix Operational Kernel Emissions -- Biogenic Emission Inventory System (SMOKE-BEIS; \href{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}{https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis}).
 SMOKE-BEIS has fixed values for SGS and EGS for all regions but NH mid latitudes ($23\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$), where it uses lapse rates of $\nabla d_\text{SGS} = 4.5$ and $\nabla d_\text{EGS} = 3.3$.
@@
 Another solution would be the usage of a proper land surface model, e.g.
-LPJ-GUESS, CLM, but the integration of such into the Oslo~CTM3 is not planed at the moment.
+LPJ-GUESS, CLM, but the integration of such into the Oslo~CTM3 is not planned at the moment.
 
@@
 From OpenIFS an accumulated surface PAR is available.
-It is integrated both, spactrally (presumably $400-700\,\unit{nm}$) and temporally.
+It is integrated both, spectrally (presumably $400-700\,\unit{nm}$) and temporally.
 For practical use in Eq.~(\ref{eq:flight}), we de-accumulate this field with respect to time and refer to the result as PPFD.
@@
 The Oslo~CTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}.
-CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
+CEDS historical emission inventory \citep{GMD:Hoesly2018} is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4).
 Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}, while emissions from soil and wetlands are computed by MEGAN.
@@
 As indicated by the names, \emph{mOSaic\_offLight} and \emph{mOSaic\_offPhen} are rather extreme scenarios completely switching off the sensitivity to light and phenology in Eq.~(\ref{eq:stomatal}) by setting $f_\text{light}$ and $f_\text{phen}$ to a fixed value of 1, respectively.
-Because of the underlying research project's focus on arctic and alpine ecosystems, where water might only be available from upper soil layers, an experiment was conducted using the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
+Because of the underlying research project's focus on Arctic and alpine ecosystems, where water might only be available from upper soil layers, an experiment was conducted using the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$.
 After this, we want to confirm the importance of choice of $R^\chem{O_3}$ for different land surface types.
@@
   \includegraphics[width=8.3cm]{fig04}
-  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
+  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
   \label{fig:diff_maps}
@@
 %
-In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition.
-The ozone burden increases globally except for some regions covered by tropical forest.
-Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$.
+In Fig.~\ref{fig:diff_maps}, we show global distributions of the relative difference between \emph{mOSaic} and \emph{Wesely\_type} for surface ozone, dry deposition velocity, and total ozone dry deposition.
+The surface ozone increases globally except for some regions covered by tropical forest.
+Especially in desert regions in Africa, America, and Asia, the surface ozone increases by more than $100\,\unit{\%}$.
 Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest.
@@
 In the evaluation of our model, we closely follow suggestions by \citet{ACP:Hardacre2015}.
-For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
+For the purpose of comparison with the multi-model mean of the Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$.
 In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](p_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types (Section~\ref{subsubsec:seasons}).
@@
 Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$.
-In case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly ailable from model output.
+In the case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly available from model output.
 %
@@
 %
-The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}.
+The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, consistent with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}.
 This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts.
 In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$).
-We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon.
-Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance.
+We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemispheres as a consequence of the distribution of the continental land masses and vegetation thereon.
+Among the sensitivity studies focussing on the stomatal conductance, there is only a low absolute variance.
 Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly.
@@
 In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean.
-This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as new standard for the Oslo~CTM3.
-This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
-The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
+This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as the new standard for the Oslo~CTM3.
+This may, however, lead to an overcompensation of the current Arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation.
+The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity to $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$.
 The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.7: Fig.~S8).
@@
 Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameter set \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}.
-These differences in surface resistances over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}).
+These differences in surface resistances over huge parts of the unvegetated surface of the Earth account for most of the qualitative difference between the Wesely and the mOSaic scheme, but do not explain the quantitative difference (compare \emph{mOSaic\_hough}).
 We further elaborate on this in the following (Section~\ref{subsubsec:seasons}).
 
-There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better.
+There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme matches better.
 This could be a sign of differences in photo-chemistry and transport (e.g.
@@
 In Appendix Fig.~\ref{fig:mmm_drydep_season}a, the average zonal ozone dry deposition is shown separated by month.
-Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference.
+Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as a reference.
 As for the global annual comparisons above, the mOSaic scheme matches the multi-model-mean values remarkably well with respect to dry deposition velocities, while it strongly underestimates the total dry deposition.
@@
 In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition separated into mid and high latitudes in the northern hemisphere ($30\,\unit{^\circ}-90\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($30\,\unit{^\circ}-90\,\unit{^\circ S}$) is shown.
-We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference.
+We have added the multi-model-mean by \citet{ACP:Hardacre2015} as a reference.
 While the total ozone dry deposition of \emph{Wesely\_type} agrees well with the multi-model-mean in any zonal band, the mOSaic scheme displays a much smaller total ozone dry deposition.
@@
 Compared to the multi-model-mean, the seasonal cycle in the Oslo~CTM3 NH appears to be shifted towards later in the year.
-The seasonal cycle in the tropics and subtropics only differs by magnitude otherwise the shapes are identical for both the mOSaic scheme, the Wesely scheme, and the multi-model-mean.
+The seasonal cycle in the tropics and subtropics only differs by magnitude otherwise the shapes are identical for the mOSaic scheme, the Wesely scheme, and the multi-model-mean.
 The total amount of dry deposition of ozone differs strongly between the different model experiments, with \emph{mOSaic\_SWVL1} and \emph{mOSaic\_hough} displaying the lowest and highest amount, respectively.
@@
 Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to estimate these.
-Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest.
+Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ for desert, ocean, snow and ice, and tropical forest.
 Thus, it is not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type.
@@
 over cropland).
-In case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
+In the case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
 %
@@
 In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analysis.
-This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
+This implies that the ozone dry deposition to the ocean in the Oslo~CTM3 is too high as well.
 %
@@
 Despite the strong positive bias in ozone concentrations and accordingly low-bias in total dry deposition, the difference between \emph{mOSaic} and \emph{mOSaic\_emis2014} ($6\,\unit{Tg}$), lies well within the range given by satellites for the years 2005 and 2014 \citep[][Fig.~26]{ESA:Gaudel2018}.
-This indicates that the Oslo~CTM3 responses well to given changes in global emissions.
+This indicates that the Oslo~CTM3 responds well to given changes in global emissions.
 %
@@
 Over, e.g., South America, central and southern Africa, the Arabian peninsula, and the north-western Indian subcontinent, the ozone concentrations are elevated by up to $200\,\unit{\%}$.
-While on global average much more concise with the MACC-reanalysis, \emph{Wesely\_type} displays a similar tendency with respect to elevated ozone over continents.
-Enhanced ozone compared to MACC-reanalysis is found in the deep tropics and the north-western Indian subcontinent (Supplement~S.8) which coincides with regions of high intensity in incoming UV radiation.
+Regarding the global average surface ozone concentration, \emph{Wesely\_type} (Supplement S.8) is more consistent with the MACC-reanalysis than \emph{mOSaic}, but shows the same tendency of enhanced ozone over the continents as the latter.
+This enhancement is apparent mostly in the deep tropics and over the north-western Indian subcontinent which coincide with regions of high intensity in incoming UV radiation.
 This may indicated an imbalance in the photo-chemical production and loss of \chem{O_3} in the Oslo~CTM3.
@@
 The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual uncertainty band since such is not given in \citet{ACP:Hardacre2015}.
-At $4$ of $6$ sites, the mOSaic scheme performs better than the Wesely scheme and similar to or better than the multi-model-mean.
+At $4$ of the $6$ sites, the mOSaic scheme performs better than the Wesely scheme and similar to or better than the multi-model-mean.
 We use a $\chi^2$-test
@@
 The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b.
-The $\chi^2$-test reveals also that in $4$ of $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to observed ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
+The $\chi^2$-test reveals also that in $4$ of the $6$ cases the mOSaic scheme improves the performance of the Oslo~CTM3 with respect to observed ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
 With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be properly quantified.
@@
 We have presented an update of the dry deposition scheme in the Oslo~CTM3 from purely prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process-oriented parameterization taking the state of the atmosphere and vegetation into account.
-Based on the description of dry deposition in \citet{WASP:Simpson2003,ACP:Simpson2012}, we have implemented a moasic approach to compute contributions to dry deposition by individual sub-grid land surface types.
-Aerodynamic, quasi-laminar, and surface resistance (latter divided into stomatal and non-stomatal contributions) are calculated for each land surface category separately.
+Based on the description of dry deposition in \citet{WASP:Simpson2003,ACP:Simpson2012}, we have implemented a mosaic approach to compute contributions to dry deposition by individual sub-grid land surface types.
+Aerodynamic, quasi-laminar, and surface resistance (with the latter divided into stomatal and non-stomatal contributions) are calculated for each land surface category separately.
 Based on these, a land fraction-weighted mean is deduced.
@@
 As the new scheme is quantitatively more similar to the multi-model-mean \citep{ACP:Hardacre2015} with respect to dry deposition velocities, while the old scheme agrees better in terms of total dry deposition, there is an apparent discrepancy.
-By means of tropospheric ozone burden \citep{ESA:Gaudel2018} and surface ozone concentrations deduced from the MACC-reanalysis \citep{MACC-II}, the Oslo~CTM3 with operational mOSaic scheme shows a pronounced high-bias of tropospheric ozone.
+By means of tropospheric ozone burden \citep{ESA:Gaudel2018} and surface ozone concentrations deduced from the MACC-reanalysis \citep{MACC-II}, the Oslo~CTM3 with the operational mOSaic scheme shows a pronounced high-bias of tropospheric ozone.
 While the average bias is small or even reversed when using the old scheme, both display elevated ozone in comparison with the MACC-reanalysis in continental regions with high average incoming UV radiation (e.g., northern South America, central and southern Africa, the Himalayas).
@@
 This is mainly due to updates of the respective, prescribed ozone surface resistances $R^\chem{O_3}$.
-In case of desert and grasslands the difference between the old and new prescribed value is in the order of $1$ magnitude.
+In the case of desert and grasslands the difference between the old and new prescribed value is an order of magnitude.
 Over the ocean, the absolute change in dry deposition is small, but it is accumulated over a large area which is especially amplified in the southern hemisphere.
@@
 With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new Oslo~CTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate them over the latter.
-Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to revise the dry deposition scheme for these regimes and add more process-oriented formulations, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017, ACP:Luhar2018}, wave braking and spray \citep{ACP:Pozzer2006}.
-
-Although dry deposition to ice and snow amounts to only $1\,\unit{\%}$ of the total global annual ozone dry deposition in \emph{mOSaic}, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctics and affects surface ozone concentrations down to latitudes at $50\,\unit{^\circ}$ in both hemispheres.
+Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to revise the dry deposition scheme for these regimes and add more process-oriented formulations, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017, ACP:Luhar2018}, wave breaking and spray \citep{ACP:Pozzer2006}.
+
+Although dry deposition to ice and snow amounts to only $1\,\unit{\%}$ of the total global annual ozone dry deposition in \emph{mOSaic}, a decrease in prescribed dry deposition velocity in accordance to combined measurements and model studies \citep{ACP:Helmig2007} causes almost a doubling in the surface ozone concentrations in the high Arctic and affects surface ozone concentrations down to latitudes at $50\,\unit{^\circ}$ in both hemispheres.
 Comparing with results from the multi-model evaluation \citep{ACP:Hardacre2015}, we conclude that it is important to use this updated ozone dry deposition velocity to counter an Arctic surface ozone low-bias in models, however, this currently leads to an overcompensation (high-bias) in the Oslo~CTM3.
 
-We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics and the northern hemisphere is most sensitive to changes therein.
+We have studied the parameter space of the stomatal conductance parameterization and found that surface ozone in the tropics and the northern hemisphere is most sensitive to changes therein.
 In the most extreme test case, the increase in global total dry deposition amounts to $7.3\,\unit{\%}$, while the more realistic test cases, e.g.
-using differing years of emission amount to changes in the order of $\pm 2\,\unit{\%}$.
+using differing years of emission amount to changes of the order of $\pm 2\,\unit{\%}$.
 This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics.
@@
 Future work on the Oslo~CTM3 should resolve the ozone high-bias which may involve revising the photolysis- and chemical reaction computation as well as reaction rates.
-For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013} which affect the stratospheric ozone \citep{ACP:Hossaini2016, ACP:Falk2017} and a scheme covering arctic spring-time ozone depletion \citep[e.g.,][]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
+For a better modeling of ozone abundances, ocean emissions of very short-lived ozone depleting substances (VSLS) \citep{JGR:Warwick2006, ACP:Ziska2013} which affect stratospheric ozone \citep{ACP:Hossaini2016, ACP:Falk2017} and a scheme covering Arctic spring-time ozone depletion \citep[e.g.,][]{ACP:Yang2010, ACP:Toyota2011, GMD:Falk2018}, could be worthwhile implementing.
 The general model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants.
-A more efficient parallelization of the code would enable computation on higher horizontal resolutions.
+A more efficient parallelization of the code would enable computation at higher horizontal resolutions.
 
@@
 
-\authorcontribution{Stefanie Falk has compiled the manuscript, finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the Oslo~CTM3, conducted the simulations, and analyzed and evaluated the results.
-Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the respective documentation.
+\authorcontribution{Stefanie Falk has compiled the manuscript, finalized and revised the implementation of the stomatal conductance in the \emph{mOSaic} dry deposition scheme of the Oslo~CTM3, conducted the simulations, and analyzed and evaluated the results.
+Amund S{\o}vde Haslerud has implemented the \emph{mOSaic} dry deposition scheme and wrote the respective documentation.
 Both authors contributed to the writing and discussion of the paper.} %% this section is mandatory for the journals ACP and GMD.

### 2019-10-04 DryDep.tex
Small change in tense.

@@
 Based on the global atmospheric composition reanalysis performed in the ECWMF project Monitoring Atmospheric Composition and Climate (MACC) \citep{MACC-II} and a more realistic process-based oceanic deposition scheme, \citet{ACP:Luhar2018} found that the ozone dry deposition to oceans amounts to $98.4 \pm 30.0\,\unit{Tg\,a^{-1}}$.
-In particular, \citet{ACP:Luhar2018} find that the average surface resistance of ozone over ocean ($r_c=2200\,\unit{sm^{-1}}$) is highly overestimated in most models.
+In particular, \citet{ACP:Luhar2018} found that the average surface resistance of ozone over ocean ($r_c=2200\,\unit{sm^{-1}}$) is highly overestimated in most models.
 An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}.

### 2019-10-04 DryDep.tex
Corrected typo.

@@
 Regarding the global average surface ozone concentration, \emph{Wesely\_type} (Supplement S.8) is more consistent with the MACC-reanalysis than \emph{mOSaic}, but shows the same tendency of enhanced ozone over the continents as the latter.
-This enhancement is apparent mostly in the deep tropics and over the north-western Indian subcontinent which coincide with regions of high intensity in incoming UV radiation.
+This enhancement is apparent mostly in the deep tropics and over the north-western Indian subcontinent which coincides with regions of high intensity in incoming UV radiation.
 This may indicated an imbalance in the photo-chemical production and loss of \chem{O_3} in the Oslo~CTM3.
